### 2021-02-08 manuscript/manuscript.tex
New repository with squashed history

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{wrapfig}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}
\RequirePackage{wrapfig}

% line numbers:
\RequirePackage{lineno}
% \modulolinenumbers[5]
\linenumbers
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}

% spacing
\RequirePackage{setspace}
% \doublespacing

% \RequirePackage[osf]{mathpazo}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
                date=year,
                maxbibnames=9,maxnames=5,maxcitenames=2,
                backend=biber,uniquelist=false,uniquename=false,
                % style=apa,
                sorting=nyt,
                % sorting=,
                hyperref=true]{biblatex}
\RequirePackage[colorinlistoftodos]{todonotes}  %disable
\RequirePackage{color}
\RequirePackage{nicefrac}

\newcommand{\gc}[1]{{\it \color{red} #1 } }
\newcommand{\vb}[1]{{\it \color{blue} #1}}
\newcommand{\vbout}[1]{{\it \color{blue} \sout{#1}}}

% a /nonumber you can turn on/off
\newcommand{\nnn}{\nonumber}
%\newcommand{\nnn}{}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\newcommand{\Newnameref}[1]{\textit{\nameref{#1}}}

\renewbibmacro{in:}{}
\AtEveryBibitem{\clearlist{language}}

\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\flt}{\mathrm{flat}}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\logt}{\log_{10}}

\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }

\addbibresource{biblio.bib}

\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}

\author{Vince Buffalo}
\affil[]{\footnotesize University of Oregon, Institute of Ecology and Evolution\\Eugene, Oregon \\ \href{mailto:vsbuffalo@gmail.com}{vsbuffalo@gmail.com}}

\begin{document}
\maketitle

\begin{abstract}

  Under neutral theory, the level of polymorphism in an equilibrium population
  is expected to increase with population size. However, observed levels of
  diversity across metazoans vary only two orders of magnitude, while census
  population sizes ($N_c$) are expected to vary over several. This unexpectedly
  narrow range of diversity is a longstanding enigma in evolutionary genetics
  known as Lewontin's Paradox of Variation (1974). Since Lewontin's
  observation, it has been argued that selection constrains diversity across
  species, yet tests of this hypothesis seem to fall short of explaining the
  orders-of-magnitude reduction in diversity observed in nature. In this work,
  I revisit Lewontin's Paradox and assess whether current models of linked
  selection are likely to constrain diversity to this extent.  To quantify the
  discrepancy between pairwise diversity and census population sizes across
  species, I combine genetic data from 172 metazoan taxa with estimates of
  census sizes from geographic occurrence data and population densities
  estimated from body mass.  Next, I fit the relationship between
  previously-published estimates of genomic diversity and these approximate
  census sizes to quantify Lewontin's Paradox. While previous across-taxa
  population genetic studies have avoided accounting for phylogenetic
  non-independence, I use phylogenetic comparative methods to investigate the
  diversity census size relationship, estimate phylogenetic signal, and explore
  how diversity changes along the phylogeny. I consider whether the reduction
  in diversity predicted by models of recurrent hitchhiking and background
  selection could explain the observed pattern of diversity across species.
  Since the impact of linked selection is mediated by recombination map length,
  I also investigate how map lengths vary with census sizes. I find species
  with large census sizes have shorter map lengths, leading these species to
  experience greater reductions in diversity due to linked selection. Even
  after using high estimates of the strength of sweeps and background
  selection, I find linked selection likely cannot explain the shortfall
  between predicted and observed diversity levels across metazoan species.
  Furthermore, the predicted diversity under linked selection does not fit the
  observed diversity--census-size relationship, implying that processes other
  than background selection and recurrent hitchhiking must be limiting
  diversity.

\end{abstract}

A longstanding mystery in evolutionary genetics is that the observed levels of
genetic variation across sexual species are confined to an unexpectedly narrow
range. Under neutral theory, the average number of nucleotide differences
between lineages (pairwise diversity, $\pi$) is determined by the balance of
new mutations and their loss by genetic drift
\parencite{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the
expected diversity at neutral sites in a panmictic population of $N_c$ diploids
is expected to be $\pi \approx 4N_c \mu$, where $\mu$ is per generation
mutation rate. Given that metazoan germline mutation rates only differ 10-fold
($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}), and census sizes
vary over several orders of magnitude, one would expected under neutral theory
that heterozygosity should also vary over several orders of magnitude. However,
early allozyme surveys revealed that heterozygosity levels across a wide range
of species varied just an order of magnitude \parencite[][p.
208]{Lewontin1974-jb}; this anomaly is known as Lewontin's ``Paradox of
Variation". With modern sequencing-based estimates of $\pi$ across taxa ranging
over only three orders of magnitude ($0.01$--$10\%$, \cite{Leffler2012-zj}),
Lewontin's paradox has persisted unresolved through the genomics era.

From the beginning, explanations for Lewontin's Paradox have been framed in
terms of the neutralist--selectionist controversy
\parencite{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.
The neutralist view is that beneficial alleles are sufficiently rare and
deleterious alleles removed sufficiently quickly, that levels of genetic
diversity are shaped predominantly by genetic drift and mutation
\parencite{Kimura1984-ia}. Specifically, \emph{non-selective} processes
decouple the effective population size implied by observed levels of diversity
$\widehat{\pi}$, $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, from the
census size, $N_c$. By contrast, the selectionist view is that the direct and
indirect effects of linked selection suppress diversity levels across taxa,
specifically because the impact of linked selection is greater in large
populations. Undoubtedly, these opposing views represent a false dichotomy, as
population genomic studies have uncovered complex demographic histories that
impact diversity within a species (e.g. \cite{Zhao2013-vd,Palkopoulou2015-bg}),
as well as evidence that selection depresses genome-wide diversity (e.g.
\cite{Elyashiv2016-vt,Begun1992-ey,Aguade1989-jx,McVicker2009-ax}).

\subsection*{Possible Explanations of Lewontin's Paradox}

A resolution of Lewontin's Paradox would involve a mechanistic description and
quantification of the evolutionary processes that prevent diversity from
scaling with census sizes across species. This would necessarily connect to the
broader literature on the empirical relationship between diversity and
population size \parencite{Frankham1996-yb,Nei1984-zi,Soule1976-he}, and the
ecological and life history correlates of genetic diversity
\parencite{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of
processes stand out as potentially capable of decoupling census sizes from
diversity: non-equilibrium demography, variance and skew in reproductive
success, and selective processes.

It has long been appreciated that effective population sizes are typically less
than census population sizes, tracing back to early debates between R.A. Fisher
and Sewall Wright \parencite{Fisher1947-tf,Wright1948-fj}. Possible causes of
this divergence between effective and census population sizes include
demographic history (e.g. population bottlenecks), extinction and
recolonization dynamics, or the breeding structure of populations (e.g. the
variance in reproductive success and population substructure). Early
explanations for Lewontin's Paradox suggested bottlenecks during the last
glacial maximum severely reduced population sizes
\parencite{Kimura1984-ia,Ohta1973-pk,Nei1984-zi}, and emphasized that large
populations recover to equilibrium diversity levels more slowly
(\cite{Nei1984-zi}, \cite{Kimura1984-ia} p. 203-204). Another explanation is
that cosmopolitan species repeatedly endure extinction and recolonization
events, which reduces effective population size
\parencite{Maruyama1980-xz,Slatkin1977-kd}.

While chance demographic events like bottlenecks and recent expansions have
long-term impacts on diversity (equilibrium is reached on the order of size of
the population), characteristics of the breeding structure such as high
variance ($V_w$) or skew in reproductive success  also suppress diversity below
the levels predicted by the census size \parencite{Wright1938-tv}. In species
like marine animals, females are highly fecund, and dispersing larvae face
extremely low survivorship, leading to high variance in reproductive success
\parencite{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
``sweepstakes" reproductive systems can lead to remarkably small ratios of
effective population size to census population size (e.g. $\nicefrac{N_e}{N_c}$
can range from $10^{-6}$--$10^{-2}$), since $\nicefrac{N_e}{N} \approx
\nicefrac{1}{V_w}$
\parencite{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and
require multiple-merger coalescent processes  to describe their genealogies
\parencite{Eldon2006-ui}. Overall, these reproductive systems diminish the
diversity in many species, but seem unlikely to explain Lewontin's Paradox
broadly across metazoans.

Alternatively, selective processes, and in particular the indirect effects of
selection on linked neutral variation, could explain the observed narrow range
of diversity. The earliest mathematical model of hitchhiking was proffered as a
solution to Lewontin's Paradox \parencite{Maynard_Smith1974-zr}.  Since,
empirical observations have demonstrated that linked selection shapes patterns
of genome-wide diversity, as evidenced by the correlation between recombination
and diversity in a variety of species
\parencite{Aguade1989-jx,Begun1992-ey,Cutter2003-tl,Stephan1998-hh,Cai2009-by}.
Theoretic work to explain this pattern has considered diversity under a steady
influx of new beneficial mutations (recurrent hitchhiking;
\cite{Stephan1992-jc,Stephan1995-ry}), and purifying selection against new
deleterious mutations (background selection, BGS;
\cite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-xc}).
Indeed, empirical work indicates background selection diminishes diversity
around genic regions in a variety of species
\parencite{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now
efforts have shifted towards teasing apart the effects of positive and negative
selection on genomic diversity \parencite{Elyashiv2016-vt}.

An important class of theoretic selection models pertaining to Lewontin's
Paradox are recurrent hitchhiking models that decouple diversity from the
census population size. These models predict diversity when strongly selected
beneficial mutations regularly enter and sweep through the population, trapping
lineages and forcing them to coalesce
\parencite{Kaplan1989-sc,Gillespie2000-mh}. In general, decoupling occurs under
these hitchhiking models when the rate of coalescence due to selection is much
greater than the rate of neutral coalescence \parencite[equation
22]{Coop2012-cd}. Other selection models cannot alone decouple diversity from
population size, \emph{ceteris paribus}. For example, the reduction in
diversity predicted under background selection and polygenic fitness variation
is a proportion reduction in population size, mediated by the total
recombination map length
\parencite{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.

\subsection*{Recent Approaches Towards Solving Lewontin's Paradox}

Recently, \textcite{Corbett-Detig2015-gt} used population genomic data to
estimate the reduction in diversity due to background selection and hitchhiking
across 40 species, and showed the impact of selection increases with two
proxies of census population size, species range and with body size. These
authors argued this is evidence that selection could explain Lewontin's
Paradox; however, in a re-analysis, \textcite{Coop2016-gx} demonstrated that
the observed scale of these reductions is insufficient to explain the
orders-of-magnitude shortfall between observed and expected levels of diversity
across species. Other recent work has found that certain life history
characteristics related to parental investment, such as propagule size, are
good predictors diversity in animals \parencite{Romiguier2014-bp,Chen2017-nf}.
Nevertheless, while these diversity correlates are important clues, they do not
propose a mechanism by which these traits act to constrain diversity within a
few orders of magnitude.

% Furthermore, Coop's re-analysis finds no evidence
% of a remaining correlation between recombination and diversity among the
% highest recombination windows, as would be expected if these $\pi_0$ estimates
% were still impacted by linked selection.

Here, I revisit Lewontin's Paradox by integrating a variety of data sets and
assessing the predicted reductions in diversity under different selection
models. Prior surveys of genetic diversity either lacked census population size
estimates, used allozyme-based measures of heterozygosity, or included fewer
species. To address these shortcomings, I first estimate census sizes by
combining predictions of population density based on body size with ranges
estimated from geographic occurrence data. Using these estimates, I quantify
the relationship between census size and previously-published genomic diversity
estimates across 172 metazoan taxa within nine phyla, which provides a sense of
the scale of the divergence between $\pi$ and $N_c$ that leads to Lewontin's
Paradox.

Past work looking at the relationship between $\pi$ and $N_c$ has largely
ignored phylogenetic non-independence across taxa
\parencite{Felsenstein1985-an}. To address this shortcoming, I account for
phylogenetic non-independence across taxa using a synthetic time-calibrated
phylogeny and phylogenetic comparative methods (PCMs). Moreover,
\textcite{Lynch2011-qv} has argued that since coalescent times are much less
than divergence times, considering phylogenetic non-independence is unnecessary
for traits like effective population size. Using PCMs, I test this conjecture
by estimating phylogenetic signal in the diversity census size relationship,
and investigating how these traits evolve along the phylogeny.

Finally, I explore whether the predicted reductions of diversity under
background selection and recurrent hitchhiking are sufficiently strong enough
to resolve Lewontin's Paradox. These predicted reductions in diversity across
species are generously estimated using selection strength parameters from
\emph{Drosophila melanogaster}, a species known to be strongly affected by
linked selection. Given the effects of linked selection are mediated by
recombination map length, I investigate how recombination map lengths vary with
census population size using data from a previously-published survey
\parencite{Stapley2017-fs}. I find map lengths are typically shorter in
large--census-size species, increasing the effects of linked selection in these
species, which might further decouple diversity from census size. Still, I find
the combined impact of these selection models with available parameter
estimates falls short in explaining Lewontin's Paradox, and discuss future
avenues through which the Paradox of Variation could be fully resolved.

\section*{Results}

\subsection*{Estimates of Census Population Size}

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/Nc_figure.pdf}
  \caption{
     The distribution of approximate census population sizes estimated by this
     study. Some phyla containing few species were excluded for clarity.}
\label{fig:figure-nc}
\end{figure}

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range and body
size or qualitative categories as proxies for census size
\parencite{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (Figure
\ref{fig:figure-nc}). My approach predicts population densities from body sizes
using a previously-observed linear relationship that holds across metazoans
(Supplementary Figure \ref{suppfig:damuth};
\cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
estimate range sizes. Finally, I estimate population size as the product of
these predicted densities and range estimates (see Methods:
\Newnameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \parencite{Damuth1987-sg}; consequently, population
sizes for taxa like humans and their domesticated species are underestimated.
While these methods to estimate census size are crude and approximate, they can
be efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox.

\subsection*{Quantifying Lewontin's Paradox}

\begin{figure}[t!]
  \centering
  \includegraphics[width=\textwidth]{figures/diversity_popsize_full.pdf}

  \caption{An illustration of Lewontin's Paradox of Variation. Pairwise
    diversity (data from \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt},
    and \cite{Romiguier2014-bp}), which varies around three orders of
    magnitude, shows a weak relationship with approximate population size,
    which varies over 12 orders of magnitude.  The shaded curve shows the range
    of expected neutral diversity if $N_e$ were to equal $N_c$ under the
    four-alleles model, $\log_{10}(\pi) = \logt(\theta) - \logt{(1 +
    \nicefrac{4\theta}{3})}$ where $\theta = 4N_c \mu$, for two mutation rates,
    $\mu = 10^{-8}$ and $\mu = 10^{-9}$, and the light gray dashed line
    represents the maximum pairwise diversity under the four alleles model. The
    dark gray dashed line is the OLS regression fit, and the blue
    dashed line is the regression fit using a phylogenetic mixed-effects model.
    Points are colored by phylum.}

  \label{fig:figure-1}
\end{figure}

To determine which ecological or evolutionary processes could decouple
diversity from census population size, we first need to quantify this
relationship across a wide variety of taxa. Previous work has found there is a
significant relationship between heterozygosity and the logarithm of population
size, but these studies relied on heterozygosity measured from allozyme data
\parencite{Soule1976-he,Frankham1996-yb,Nei1984-zi}. Here, I confirm these
findings using pairwise diversity estimates from genomic sequence data and the
estimated census sizes (Figure \ref{fig:figure-1}). The pairwise diversity
estimates are from three sources: \textcite{Leffler2012-zj},
\textcite{Corbett-Detig2015-gt}, and \textcite{Romiguier2014-bp}, and are
predominantly from either synonymous or non-coding DNA (see Methods:
\Newnameref{sec:methods-diversity}). Overall, an ordinary least squares (OLS)
relationship on a log-log scale fits the data well (Figure \ref{fig:figure-1}).
The OLS slope estimate is significant and implies an increase in diversity of
0.09 differences per basepair for every order of magnitude census size grows
(95\% confidence interval $[0.08, 0.12]$; see also the OLS fit per-phyla,
Supplementary Figure \ref{suppfig:figure-1-ave}).

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\parencite{Nei1984-zi,Soule1976-he}; see Supplementary Figure
\ref{suppfig:figure-1-linear} for a version on a log-linear scale. However, it
is noteworthy that few taxa have diversity estimates below $10^{-3.5}$
differences per basepair. Those that do, lynx (\emph{Lynx Lynx}), wolverine
(\emph{Gulo gulo}), and Massasauga rattlesnake (\emph{Sistrurus catenatus})
face habitat fragmentation and declining population sizes. These three species
are all in the IUCN Red List, but are listed as least concern (though their
presence in the Red List indicates they are of conservation interest). I also
investigated the relationship between species' IUCN Red List categories (an
ordinal scale of how threatened a species is) and both diversity and population
size, finding that species categorized as more threatened have both smaller
population sizes and reduced diversity, compared to non-threatened species
(Figure \ref{fig:figure-1-redlist}) consistent with past work
\parencite{Spielman2004-mt}. A linear model of diversity regressed on
population size has lower AIC when the IUCN Red List categories are included,
and the estimates of the effect of IUCN status are all negative on diversity,
though not all are significant in part because some categories have three or
fewer species (Supplementary Table \ref{supptable:tbl-1}).

\begin{figure}[t!]
  \centering
  \includegraphics[width=\textwidth]{figures/diversity_popsize_redlist.pdf}

  \caption{A version of Figure \ref{fig:figure-1} with points colored by their
    IUCN Red List conservation status. Margin boxplots show the diversity and
    population size ranges (thin lines) and interquartile ranges (thick lines)
    for each category. NA/DD indicates no IUCN Red List entry, or Red List
  status Data Deficient; LC is Least Concern, NT is Near Threatened, VU is
Vulnerable, EN is Endangered, and CR is Critically Endangered.}

  \label{fig:figure-1-redlist}
\end{figure}

\subsection*{Phylogenetic Non-Independence and the Population Size Diversity Relationship}

In quantifying Lewontin's Paradox, I have additionally fit some simple trait
evolution models that account for phylogenetic non-independence, investigated
whether there is a signal of phylogenetic non-independence, estimated the
continuous trait values on the phylogeny, and assessed how diversity and
population size evolve. Prior population genetic comparative studies have
lacked time-calibrated phylogenies and assumed unit branch lengths
\parencite{Whitney2010-ud}, a shortcoming that has drawn criticism
\parencite{Lynch2011-qv}. Here, I use a synthetic time-calibrated phylogeny
created from the DateLife project \parencite{OMeara2020-ds} to account for
shared phylogenetic history (see Methods: \Newnameref{sec:methods-pcm}).

\begin{figure}[t!]
  \centering
  \includegraphics[width=\textwidth]{figures/diversity_pcm.pdf}

  \caption{(A) The ancestral continuous trait estimates for the population size
    and diversity (differences per bp, log scaled) across the phylogeny of 166
    taxa. The phyla of the tips are indicated by the color bar in the center.
    (B) The posterior distributions of the intercept, slope, and phylogenetic
    signal ($\lambda$, \cite{De_Villemereuil2014-kt}) of the phylogenetic
    mixed-effects model of diversity and population size (log scaled). Also
    shown are the 90\% credible interval (light blue shading), posterior mean
    (blue line), OLS estimate (gray solid line), and bootstrap OLS confidence
    intervals (light gray shading).  (C) The node-height tests of diversity,
    population size, and the two components of the population size estimates,
    body mass, and range (all traits on log scale before contrast was
    calculated). Each point shows the standardized phylogenetic independent
    contrast and branching time for a pair of lineages. Red lines are robust
    regression estimates (and are only shown for statistically significant
    relationships at the $\alpha = 0.05$ level). Note that some outlier pairs
    with very high phylogenetic independent contrasts were excluded (in all
    cases, these outliers were in the genus \emph{Drosophila}).}

  \label{fig:figure-2}
\end{figure}

Using a phylogenetic mixed-effects model
\parencite{Lynch1991-tp,Hadfield2010-ql,De_Villemereuil2014-kt} implemented in
Stan \parencite{Carpenter2017-xi,Stan_Development_Team2020-ea}, I estimated the
linear relationship between diversity and population size (on a log-log scale)
accounting for phylogeny, for the 166 taxa with non-missing data and present in
the synthetic chronogram. As with the non-phylogenetic regression, this
relationship was positive and significant (95\% credible interval $0.04,
0.11$), though somewhat attenuated compared to the OLS estimates (Figure
\ref{fig:figure-1}B). Since the population size estimates are based on range
and body mass, they are essentially a composite trait; fitting phylogenetic
mixed-effects models separately on body mass and range indicates these have
significant negative and positive effects, respectively (Supplementary Figure
\ref{suppfig:figure-div-range-bodymass}).

With the phylogenetic mixed-effects model, I also estimated the variance of the
phylogenetic effect ($\sigma_p^2$) and the residual variance ($\sigma_r^2$),
which can be used to estimate a measure of the phylogenetic signal, $\lambda =
\nicefrac{\sigma_p^2}{(\sigma_p^2 + \sigma_r^2)}$
(\cite{Lynch1991-tp,De_Villemereuil2014-kt}; see \cite{Freckleton2002-ly} for a
comparison to Pagel's $\lambda$).  If the relationship between diversity and
population size was free of shared phylogenetic history, $\lambda = 0$ and all
the variance could be explained by evolution on the tips; this is analogous to
Lynch's conjecture that coalescent times should be free of phylogenetic signal
(\citeyear{Lynch2011-qv}). In the relationship between population size and
diversity, the posterior mean of $\lambda = 0.67$ (90\% credible interval
$[0.59, 0.75]$) indicates that the majority of the variance perhaps might be
due to shared phylogenetic history (Figure \ref{fig:figure-2}B).

A closer visual inspection of the estimated ancestral continuous values for
diversity and population size on the phylogeny indicates the high phylogenetic
signal seems to be driven in part by chordates having low diversity and small
population sizes compared to non-chordates (Figure \ref{fig:figure-2}A). This
suggests \citeauthor{Gillespie1991-qa}'s (1991) earlier critique that the
$\pi$--$N_c$ relationship was driven by chordate-arthropod differences may be
valid. This problem resembles Felsenstein's worst-case scenario
\parencite{Felsenstein1985-an,Uyeda2018-wf}, where a singular event on a
lineage separating two clades generates a spurious association between two
traits. To further investigate whether clade-level differences dominated the
relationship between diversity and population size, I fit phylogenetic
mixed-effects models to phyla-level subsets of the data for clades with
sufficient sample sizes (see Methods: \Newnameref{sec:methods-pcm}). This
analysis shows a significant positive relationship between diversity and
population size in arthropods, and positive weak relationships in molluscs and
chordates (Supplementary Figure \ref{suppfig:div_popsize_post_phyla}). Each of
the 90\% credible intervals for slope overlap, indicating the relationship
between $\pi$ and $N_c$ is similar across these clades.

One limitation of the phylogenetic mixed-effects models employed here is that
they assume traits evolve under constant-rate Brownian motion. To test this
assumption, I performed node-height tests \parencite{Freckleton2006-ww}.
Node-height tests regress the absolute values of the standardized contrasts
between lineages against the branching time (since present) of these lineages.
Under Brownian Motion (BM), standardized contrasts are estimates of the rate of
character evolution \parencite{Felsenstein1985-an}; if a trait evolves under
constant rate BM, this relationship should be flat. For both diversity and
population size, node-height tests indicate a significant increase in the rate
of evolution towards the present (robust regression p-values 0.028 and 0.00070
respectively; Figure \ref{fig:figure-2}C).  Considering the constituents of the
population size estimate, range and body mass, separately, range but not body
mass shows a significant increase (p-value $1.9 \times 10^{-7}$) in rate
towards the present.

Interestingly, the diversity node-height test reveals two rate shifts at deeper
splits (Figure \ref{fig:figure-2}C, top left) around 570 Mya. These nodes
represent the branches between tunicates and vertebrates in chordates, and
cephalopods and pleistomollusca (bivalves and gastropods) in molluscs.  While
the cephalopod-pleistomollusca split outlier may be an artifact of having a
single cephalopod (\emph{Sepia officinalis}) in the phylogeny, the
tunicate-vertebrate split outlier is driven by the low diversity of vertebrates
and the previously-documented exceptionally high diversity of tunicates (sea
squirts; \cite{Nydam2010-kg,Small2007-mt}). This deep node representing a rate
shift in diversity could reflect a change in either effective population size
or mutation rate, and there is some evidence of both in this genus \emph{Ciona}
\parencite{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts
in diversity is mirrored in the population size node-height test (Figure
\ref{fig:figure-2}C, top right). Rather, it appears a trait impacting diversity
but not census size (e.g. mutation rate or offspring distributions) has
experienced a shift on the lineage separating tunicates and vertebrates. At
nearly 600 Mya, these deep nodes illustrate a counterexample to Lynch's claim
that expected coalescent times do not share phylogenetic history because they
are less than divergence times.

Finally, an important caveat is the increase in rate towards the tips could be
caused by measurement noise. Inspecting the lineage pairs that lead to this
increase in rate towards the tips indicates these represent plausible rate
shifts, e.g. between cosmopolitan and endemic sister species like
\emph{Drosophila simulans} and \emph{Drosophila sechellia}; however, ruling out
measurement noise entirely as an explanation would involve considering the
uncertainty of diversity and population size estimates.

\subsection*{Assessing the Impact of Linked Selection on Diversity Across Taxa}

The above analyses reemphasize the drastic shortfall of diversity levels as
compared to census sizes. Linked selection has been proposed as the mechanism
that acts to reduce diversity levels from what we would expect given census
sizes \parencite{Maynard_Smith1974-zr,Gillespie2000-mh,Corbett-Detig2015-gt}.
Here, I test this hypothesis by estimating the scale of diversity reductions
expected under background selection and recurrent hitchhiking, and compare
these to the observed relationship between $\pi$ and $N_c$.

I quantify the effect of linked selection on diversity as the ratio of observed
diversity ($\pi$) to the estimated diversity in the absence of linked selection
($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. There are two difficulties in
evaluating whether linked selection could resolve Lewontin's Paradox. The first
difficulty is that $\pi_0$ is unobserved. Previous work has estimated $\pi_0$
using methods that exploit the spatial heterogeneity in recombination and
functional density across the genome to fit linked selection models that
incorporate both hitchhiking and background selection
\parencite{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second difficulty is
understanding how $R$ varies across taxa, since we lack estimates of critical
model parameters for most species. Still, I can address a key question: if
diversity levels were determined by census sizes ($\pi_0 = 4N_c \mu$), are the
combined effects of background selection and recurrent hitchhiking sufficient
to reduce diversity to observed levels? Furthermore, does the relationship
between census size and predicted diversity under linked selection across
species, $\pi_{BGS+HH} = R \pi_0$, match the observed relationship in Figure
\ref{fig:figure-1}?

Since we lack estimates of key linked selection parameters across species, I
generously parameterize the hitchhiking and BGS models using estimates from
\emph{Drosophila melanogaster}, a species known to be strongly affected by
linked selection \parencite{Sella2009-nx}. Under a generalized model of
hitchhiking and background selection \parencite{Elyashiv2016-vt,Coop2012-cd}
and assuming $N_e = N_c$, expected diversity is

\begin{align} 
  \pi_\text{BGS+HH} \approx \frac{\theta}{\nicefrac{1}{B(U, L)} + 2N_c S(\gamma, L, J)} 
  \label{eqn:div} 
\end{align}
%
where $\theta = 4N_c\mu$, $B(U,L)$ is the effect of background selection, and
$S(\gamma, L, J)$ is the rate of coalescence caused by sweeps (c.f.
\cite{Elyashiv2016-vt}, equation 1, \cite{Coop2012-cd} equation 20). Under
background selection models with recombination, the reduction is $B(U,L) =
\exp(\nicefrac{-U}{L})$ where $U$ is the per diploid genome per generation
deleterious mutation rate, and $L$ is the recombination map length
\parencite{Hudson1994-oh,Hudson1995-xc,Nordborg1996-nq}. This BGS model is
similar to models of effective population size under polygenic fitness
variation, and can account for other modes of linked selection
(\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}, see Appendix Section
\ref{app:bgs}). The coalescent rate due to sweeps is $S(\gamma, L, J) =
\nicefrac{\gamma}{L} J$, where $\gamma$ is the number of adaptive substitutions
per generation, and $J$ is the probability a lineage is trapped by sweeps as
they occur across the genome (c.f. $J_{2,2}$ in equation 15 of
\cite{Coop2012-cd}).

Parameterizing the model this way, I then set the key parameters that
determine the impact of recurrent hitchhiking and background selection
($\gamma$, $J$, and $U$) to high values estimated from \emph{Drosophila
melanogaster} by \textcite{Elyashiv2016-vt}. My estimate of
$\gamma_\text{Dmel}$ based Elyashiv et al. implies $\nu_\text{BP,Dmel} \approx
2.34 \times 10^{-11}$, which is close to other estimates from \emph{D.
melanogaster} (see Supplementary Figure \ref{suppfig:linked-sel-params}A). The
rate of deleterious mutations per diploid genome, per generation is
parameterized using the estimate from Elyashiv et al., $U_\text{Dmel} = 1.6$,
which is a bit greater than previous estimates based on Bateman-Mukai
approaches \parencite{Mukai1985-bc,Mukai1988-vs,Charlesworth1987-ab}. Finally,
the probability that a lineage is trapped in a sweep, $J_\text{Dmel}$, is
calculated from the estimated genome-wide average coalescent rate due to sweeps
from Elyashiv et al. (see Supplementary Figured
\ref{suppfig:linked-sel-params}B and Methods:
\Newnameref{sec:methods-reduction} for more details on parameter estimates).
Using these \emph{Drosophila} parameters, I then explore how the predicted
range of diversity levels under background selection and recurrent hitchhiking
varies across species with recombination map length ($L$) and census population
size ($N_c$).

\begin{figure}[t!]
  \centering
  \includegraphics[width=\textwidth]{figures/figure_3.pdf}

  \caption{(A) The observed relationship between recombination map length ($L$)
    and census size ($N_c$) across 131 species with complete data and known
    phylogeny. Triangle points indicate social taxa excluded from the model
    fitting since these have adaptively higher recombination map lengths
    \parencite{Wilfert2007-dx}. The dark gray line is the estimated
    relationship under a phylogenetic mixed-effects model, and the gray
    interval is the 95\% posterior average.  (B) Points indicate the observed
    $\pi$--$N_c$ relationship across taxa shown in Figure \ref{fig:figure-1},
    and the blue ribbon is the range of predicted diversity were $N_e = N_c$
    for $\mu = 10^{-8}$--$10^{-9}$,  and after accounting for the expected
  reduction in diversity due to background selection and recurrent hitchhiking
under \emph{Drosophila melanogaster} parameters. In both plots, point color
indicates phylum.}

  \label{fig:figure-3}
\end{figure}

Previous work has found that the impact of linked selection increases with
$N_c$ (\cite{Corbett-Detig2015-gt}; see also Supplementary Figure
\ref{suppfig:corbett_detig}A), and it is often thought that this is driven by
higher rates of adaptive substitutions in larger populations, despite equivocal
evidence \parencite{Galtier2016-dq}. However, there is another mechanism by
which species with larger population sizes might experience a greater impact of
linked selection: recombinational map length, $L$, is known to correlate with
body mass \parencite{Burt1987-tq} and thus varies inversely with population
size. As this is a critical parameter that determines the genome-wide impact of
both hitchhiking and background selection, I examine the relationship between
recombination map length ($L$) and census population size ($N_c$) across taxa,
using available estimates of map lengths across species
\parencite{Stapley2017-fs,Corbett-Detig2015-gt}. I find a significant
non-linear relationship using phylogenetic mixed-effects models (Figure
\ref{fig:figure-3}A; see Methods: \Newnameref{sec:methods-pcm}). There is also
a correlation between map length and genome size (Supplementary Figure
\ref{supfig:genome_size_maplength}) and genome size and population size
(Supplementary Figure \ref{supfig:genome_size_popsize}). Overall, the negative
relationship between map length and census size indicates linked selection is
expected to be stronger in short map length, high-$N_c$ species.

Then, I predict the expected diversity ($\pi_{BGS+HH}$) under background
selection and hitchhiking, were $N_e= N_c$, and assuming all species had the
rate of sweeps and strength of BGS as \emph{D. melanogaster}. Since neutral
mutation rates $\mu$ are unknown and vary across species, I calculate the range
of predicted $\pi_{BGS+HH}$ estimates for $\mu = 10^{-8}$--$10^{-9}$ (using the
four-alleles model, \cite{Tajima1996-rb}), and compare this to the observed
relationship between $\pi$ and $N_c$ in Figured \ref{fig:figure-3}B.  Under
these parameters, linked selection begins to appreciably depress diversity
around $N_c \approx 10^9$, since $S \approx 10^{-8}$--$10^{-9}$ and linked
selection dominates drift when $S > \nicefrac{1}{2N}$.  Overall, this reveals
two problems for the hypothesis that linked selection could solve Lewontin's
Paradox. First, low to mid-$N_c$ species (census sizes between
$10^{6}$--$10^{14}$) have sufficiently long map lengths that their diversity
levels are only moderately reduced by linked selection, leading to a wide gap
between predicted and observed diversity levels. For this not to be the case,
the parameters that determine the strength of background selection and
recurrent hitchhiking would need to be \emph{higher} among these species than
in \emph{Drosophila melanogaster}. This would require that the rate of adaptive
mutations or the deleterious mutation rate be orders of magnitude higher for
species within this range than in \emph{Drosophila}, which is incompatible with
the rate of adaptive substitutions across species \parencite{Galtier2016-dq}
and mutation rates \parencite{Lynch2010-ki}. Furthermore, linked selection has
been quantified in humans, which fall in this census size range, and has been
found to be relatively weak
\parencite{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
Second, while hitchhiking and BGS can reduce predicted diversity levels for
high-$N_c$ species ($N_c > 10^{14}$) to observed levels, this would imply
available estimates of $\pi_0$ are underestimated by several orders of
magnitude in \emph{Drosophila} (Supplementary Figured
\ref{suppfig:corbett_detig}B). Overall, while linked selection could decouple
diversity from census size for high-$N_c$ species, recurrent hitchhiking and
background selection seem unlikely to explain the observed patterns of
diversity across species under our understanding of the range of parameter
estimates.

\section*{Discussion}

Nearly fifty years after Lewontin's description of the Paradox of Variation,
how evolutionary, life history, and ecological processes interact to constrain
diversity across taxa to a narrow range remains a mystery.  Since Wright
(\citeyear{Wright1931-fl,Wright1938-tv}), population geneticists have
appreciated that various demographic processes shrink effective population
sizes compared to census sizes, yet it has remained unclear whether these
neutral processes alone can explain Lewontin's Paradox and across-taxa
diversity patterns. Alternatively, selective processes that act more strongly
in larger populations could account for the observed narrow range of diversity.
A critical first step to discerning the processes that act to transform census
sizes to diversity levels across species is characterizing the observed
$\pi$--$N_c$ relationship.

Here, for 172 taxa, I estimated the relationship between genomic estimates of
pairwise diversity and approximate census population size. Previous surveys
have used allozyme-based estimates, fewer taxa, or qualitative measures of
population size. My estimates of census population sizes are quite approximate,
since they use body size to predict density. An improved estimate might
consider vagility (as \cite{Soule1976-he} did), though this is harder to do
systematically across many taxa. Future work might also use other ecological
information, such as total biomass and estimated numbers of species within
phyla, to improve census size estimates \parencite{Bar-On2018-kc, Mora2011-wm}.
Still, it seems more accurate estimates would be unlikely to change the
qualitative findings here, which resemble those of early surveys
\parencite{Nei1984-zi,Soule1976-he}.

One limitation of the dataset in this study is that diversity estimates are
collated from a variety of sources rather than estimated with a single
bioinformatic pipeline. This leads to technical noise across diversity
estimates; perhaps the relationship between $\pi$ and $N_c$ found here could be
tighter with a standardized bioinformatic pipeline. In addition to this
technical variation, there might be systematic bioinformatic sources of bias in
diversity estimates. For example high-diversity sequences may fail to align to
the reference genome and end up unaccounted for, leading to a downward bias.
Alternatively, high-diversity sequences might map to the reference genome, but
adjacent mis-matching SNPs might be mistaken for a short insertion or deletion.
While these issues might adversely affect the estimates in high-diversity
species, it is unlikely they will qualitatively change the observed
$\pi$--$N_c$ relationship.

\subsection*{Macroevolution and Across-Taxa Population Genomics}

% TODO -- talk about diversity as a trait. What are the trail evolution
% models?

Lewontin's Paradox arises from a comparison of diversity across species, yet
surprisingly, previous work on this problem has not considered the impact of
phylogenetic non-independence. I have addressed this limitation, showing that
diversity does have a significant positive relationship with census size, after
accounting for shared phylogenetic history among taxa. Additionally, I find a
high degree of phylogenetic signal, and that arthropods and chordates form
clusters, showing previous concern about phylogenetic non-independence was
warranted \parencite{Gillespie1991-qa}. Finally, this high degree of
phylogenetic signal, as well as evidence of shifts in the rate of evolution of
genetic diversity on deep timescales in molluscs and chordates, seem to
contradict Lynch's (2011) claim that since coalescent times are much less than
divergence times, they are not affected by shared phylogenetic history.

One can reconcile my findings with Lynch's claim by considering what
evolutionary, ecological, life history, and demographic causal factors
determine coalescent rates across species, and how these factors evolve across
deep timescales. Lynch's conjecture that coalescent times should be free of
phylogenetic signal may be true \emph{were we to condition} on these causal
factors that could be affected by shared phylogenetic history. In contrast, my
estimates of phylogenetic signal in diversity are not conditioned on these
factors. Importantly, even ``correcting for" phylogeny implicitly favors
certain causal interpretations over others
\parencite{Westoby1995-wz,Uyeda2018-wf}. Future work could try to untangle what
causal factors determine coalescent times across species, as well as how these
factors evolve across macroevolutionary timescales.

Furthermore, beyond just accounting for phylogenetic non-independence,
macroevolution and phylogenetic comparative methods are a promising way to
approach across-species population genomic questions.  For example, one could
imagine that diversification processes could contribute to Lewontin's Paradox.
If large-$N_c$ species were to have a rate of speciation that is greater than
the rate at which mutation and drift reach equilibrium (which is indeed slower
for large $N_c$ species), this could act to decouple diversity from census
population size. That is to say, even if the rate of random demographic
bottlenecks were constant across taxa, lineage-specific diversification
processes could lead certain clades to be systematically further from
demographic equilibrium, and thus have lower diversity than expected for their
census population size.

\subsection*{Spatial and Demographic Processes}

One limitation of this study is the inability to quantify the impact of spatial
population genetic processes on the relationship between diversity and census
population sizes across taxa. The genomic diversity estimates collated in this
study unfortunately lack details about the sampling process and spatial data,
which can have a profound impact on population genomic summary statistics
\parencite{Battey2020-lc}. These issues could systematically bias species-wide
diversity estimates; for example, if diversity estimates from a cosmopolitan
species were primarily from a single subpopulation, diversity would be an
underestimate relative to the entire population. However, biased spatial
sampling alone seems incapable of explaining the $\pi$-$N_c$ divergence in
high-$N_c$ taxa. In the extreme scenario in which only one subpopulation was
sampled, $F_\text{ST}$ would need to be close to one for population subdivision
alone to sufficiently reduce the total population heterozygosity to explain the
orders-of-magnitude shortfall between predicted and observed diversity levels.
Yet, across-taxa surveys indicate that $F_\text{ST}$ is almost never this high
within species \parencite{Roux2016-lm}.  Still, future work could quantify the
extent to which spatial processes contribute to Lewontin's Paradox. For
example, high-$N_c$ taxa usually experience range expansions, likely with
repeated founder effects and local extinction/recolonization dynamics that
doubtlessly depress diversity. In particular, with the appropriate data, one
could estimate the empirical relationship between dispersal distance, range
size, and coalescent effective population size across taxa.

In this study, I have focused entirely on assessing the role of linked
selection, rather than demography, in reducing diversity across taxa. In
contrast to demographic models, models of linked selection have comparatively
fewer parameters and more readily permit rough estimates of diversity
reductions across taxa. Still, a full resolution of Lewontin's Paradox would
require understanding how the demographic processes across taxa with incredibly
heterogeneous ecologies and life histories transform $N_c$ into $N_e$. With
population genomic data becoming available for more species, this could involve
systematically inferring the demographic histories of tens of species and
looking for correlations in the frequency and size of bottlenecks with $N_c$
across species.

\subsection*{How could selection still explain Lewontin's Paradox?}

In this study, my goal was not to accurately estimate the levels in diversity
across species, but rather to give linked selection the best possible chance to
solve Lewontin's Paradox. Still, I find that even after parameterizing
hitchhiking and background selection with strong selection parameter estimates
from \emph{Drosophila melanogaster}, the predicted patterns of diversity under
linked selection poorly fit observed patterns of diversity across species.
While this suggests these two common modes of linked selection seem unlikely to
explain across-taxa patterns of diversity, there are three major potential
limitations of my approach that need further evaluation.

First, I approximate the reduction in diversity using homogeneous background
selection and recurrent hitchhiking models
\parencite{Kaplan1989-sc,Hudson1995-xc,Coop2012-cd}, when in reality, there is
genome-wide heterogeneity in functional density, recombination rates, and the
adaptive substitutions across species. Each of these factors mediate how
strongly linked selection impacts diversity across the genome. Despite these
model simplifications, the predicted reduction in diversity in \emph{Drosophila
melanogaster} is $85\%$ (using $N_e$, not $N_c$), which is reasonably close to
the estimated $77\%$ from the more realistic model of Elyashiv et al. that
accounts for the actual position of substitutions, annotation features, and
recombination rate heterogeneity (though it should be noted that these both use
the same parameter estimates). Furthermore, even though my model fails to
capture the heterogeneity of functionality density and recombination rate in
real genomes, it is still extraordinary conservative, giving linked selection
the best possible chance to decouple diversity from census size and explain
Lewontin's Paradox. This is in part because the strong selection parameter
estimates from \emph{Drosophila melanogaster} used, but also because I assume
that the effective population size is the census size. Even then, this
decoupling only occurs in very high--census-size species, and implies that the
diversity in the absence of linked selection, $\pi_0$, is currently
underestimated by several orders of magnitude. Moreover, the study of
\textcite{Corbett-Detig2015-gt} did consider recombination rate and functional
density heterogeneity in estimating the reduction due to linked selection
across species, yet their predicted reductions are orders of magnitude weaker
than those considered here by assuming that $N_e = N_c$ (Supplementary Figure
\ref{suppfig:corbett_detig}B).  Overall, even with more realistic models of
linked selection, current models of linked selection seem fundamentally unable
to fit the diversity--census-size relationship.

Second, my model here only considers hard sweeps, and ignores the contribution
of soft sweeps (e.g. from standing variation or recurrent mutations;
\cite{Hermisson2005-hs,Pennings2006-lj}), partial sweeps (e.g those that do not
reach fixation), and the interaction of sweeps and spatial processes. While
future work exploring these alternative types of sweeps is needed, the
predicted reductions in diversity found here under the simplified sweep model
are likely relatively robust to these other modes of sweeps for a few reasons.
First, the shape of the diversity--recombination curve is equivalent under
models of partial sweeps and hard sweeps, though these imply different rates of
sweeps \parencite{Coop2012-cd}. Second, in the limit where most fitness
variation is due to weak soft sweeps from standing variation scattered across
the genome (i.e. due to polygenic fitness variation), levels of diversity are
well approximated by quantitative genetic linked selection models (QGLS
\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}). The reduction in
diversity under these models is nearly identical to that under background
selection models, in part because deleterious alleles at mutation-selection
balance constitute a considerable component of fitness variation (see Appendix
Section \ref{app:bgs}; \cite{Charlesworth2000-km,Charlesworth2015-am}).
Finally, I also disregarded the interaction of sweeps and spatial processes.
For populations spread over wide ranges, limited dispersal slows the spread of
sweeps, allowing for new beneficial alleles to arise, spread, and compete
against other segregating beneficial variants
\parencite{Ralph2015-kl,Ralph2010-ki}. Through limited dispersal should act to
``soften sweeps" and not impact my findings for the reasons described above,
future work could investigate how these processes impact diversity in ways not
captured by hard sweep models.

Third, other selective processes, such as fluctuating selection or hard
selective events, could reduce diversity in ways not captured by the background
selection and hitchhiking model. Since frequency-independent fluctuating
selection generally reduces diversity under most conditions
\parencite{Novak2017-yp}, this could lead seasonality and other sources of
temporal heterogeneity to reduce diversity in large-$N_c$ species with short
generation times more than longer-lived species with smaller population sizes.
Future work could consider the impact of fluctuating selection on diversity
under simple models \parencite{Barton2000-zg} if estimates of key parameters
governing the rate of such fluctuations were known across taxa. Additionally,
another mode of selection that could severely reduce diversity across taxa, yet
remains unaccounted for in this study, is periodic hard selective events. These
selective events could occur regularly in a species' history yet be
indistinguishable from demographic bottlenecks from just population genomic
data.

\subsection*{Measures of Effective Population Size, Timescales, and Lewontin's Paradox}

Lewontin's Paradox shows the extent to which the effective population sizes
implied by diversity, $\widetilde{N}_e$, diverge from census population sizes.
However, there are a variety other effective population size estimates
calculable from different data and summary statistics
\parencite{Wang2016-mi,Caballero1994-ao,Caballero2020-wm}. These include
estimators based on the observed decay in linkage disequilibrium or temporal
estimators that use the variance in allele frequency change. These alternate
estimators capture summaries of the effective population size on shorter
timescales than coalescent-based estimators \parencite{Wang2005-wy}, and thus
could be used to tease out processes that impact the $N_e$-$N_c$ relationship
in the more recent past.

Temporal $N_e$ estimators already play an important role in understanding
another summary of the $N_e$-$N_c$ relationship: the ratio
$\nicefrac{N_e}{N_c}$, which is an important quantity in conservation genetics
\parencite{Frankham1995-xz,Mace1991-ai} and in understanding evolution in
highly-fecund marine species. Surveys of the $\nicefrac{N_e}{N_c}$ relationship
across taxa indicate mean $\nicefrac{N_e}{N_c}$ is on order of $\approx 0.1$
\parencite{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though the
uncertainty in these estimates is high, and some species with sweepstakes
reproduction systems like Pacific Oyster (\emph{Crassostrea gigas}) can have
$\nicefrac{N_e}{N_c} \approx 10^{-6}$. Estimates of the $\nicefrac{N_e}{N_c}$
ratio are an important, and under appreciated piece of solving Lewontin's
Paradox. For example, if $N_e$ is estimated from the allele frequency change
across a single generation (i.e. \cite{Waples1989-sj}), $\nicefrac{N_e}{N_c}$
constrains the variance in reproductive success
\parencite{Wright1938-tv,Nunney1993-ef,Nunney1996-wy}. This implies that apart
from species with sweepstakes reproductive systems, the variance in
reproductives success each generation (whether heritable or non-heritable) is
likely insufficient to significantly contribute to constraining
$\widetilde{N}_e$ for most taxa. Still, further work is needed to characterize
(1) how $\nicefrac{N_e}{N_c}$ varies with $N_c$ across taxa (though see
\cite{Palstra2012-vy}, Figure 2), and (2) the variance of $\nicefrac{N_e}{N_c}$
over longer time spans (i.e. how periodic sweepstakes reproductive events act
to constrain $N_e$). Overall, characterizing how $\nicefrac{N_e}{N_c}$ varies
across taxa and correlates with ecology and life history traits could provide
clues into the mechanisms that leads propagule size and survivorship curves to
be predictive of diversity levels across taxa
\parencite{Romiguier2014-bp,Hallatschek2018-iu,Barry2020-pp}.

Finally, short-term temporal $N_e$ estimators may play an important role in
resolving Lewontin's Paradox. These estimators, along with short-term estimates
of the impact of linked selection \parencite{Buffalo2019-qs,Buffalo2020-my},
can inform us how much diversity is depressed across shorter timescales, free
from the rare strong selective events or severe bottlenecks that impact
pairwise diversity. It could be that in any one generation, selection
contributes more to the variance of allele frequency changes than drift, yet
across-taxa patterns in diversity are better explained processes acting
sporadically on longer timescales, such as colonization, founder effects, and
bottlenecks. Thus, the pairwise diversity may not give us the best picture of
the generation to generation evolutionary processes acting in a population to
change allele frequencies. Furthermore, certain observed adaptations are
inexplicable given implied long-term coalescent effective population sizes, and
are only possible if short-term effective population sizes are orders of
magnitude larger \parencite{Karasov2010-fb,Barton2010-pi}.

\subsection*{Conclusions}

In \emph{Building a Science of Population Biology}
(\citeyear{Lewontin2004-fb}), Lewontin laments the difficulty of uniting
population genetics and population ecology into a cohesive discipline of
population biology. Lewontin's Paradox of Variation remains a critical unsolved
problem at the nexus of these two different disciplines: across species, we
fail to understand the processes that connect a central parameter of population
ecology, census size, to a central parameter of population genetics, effective
population size. Given that selection seems to fall short in explaining
Lewontin's Paradox, a full resolution will require a mechanistic understanding
the ecological, life history, and macroevolutionary processes that connect
$N_c$ to $N_e$ across taxa. While I have focused exclusively on metazoan taxa
since their population densities are more readily approximated from body mass,
a full resolution must also include plant species (with the added difficulties
of variation in selfing rates, different dispersal strategies, pollination,
etc.).

Looking at Lewontin's Paradox through an macroecological and macroevolutionary
lens begets interesting questions outside of the realm of population genetics.
Here, I have found that diversity and $N_c$ have a surprisingly consistent
relationship without many outliers, despite the wildly disparate ecologies,
life histories, and evolutionary histories of the taxa included. Furthermore,
taxa with very large census sizes have surprisingly low diversity. Is this
explained by macroevolutionary processes, such as different rates of speciation
for large-$N_c$ taxa? Or, are the levels of diversity we observe today an
artifact of our timing relative to the last glacial maximum, or the last major
extinction? Did large-$N_c$ prehistoric animal populations living in other
geological eras have higher levels of diversity than our present taxa? Or, does
ecological competition occur on shorter timescales such that strong population
size contractions transpire and depress diversity, even if a species is
undisturbed by climatic shifts or mass extinctions? Overall, patterns of
diversity across taxa are determined by many overlaid evolutionary and
ecological processes occurring on vastly different timescales. Lewontin's
Paradox of Variation may persist unresolved for some time because the solution
requires synthesis and model building at the intersection of all these
disciplines.

\section*{Methods}

\subsection*{Diversity and Map Length Data}
\label{sec:methods-diversity}

The data used in this study are collated from a variety of previously published
surveys. Of the 172 taxa with diversity estimates, 14 are from
\textcite{Corbett-Detig2015-gt}, 96 are from \textcite{Leffler2012-zj}, and 62
are from \textcite{Romiguier2014-bp}. The Corbett-Detig et al. data is
estimated from four-fold degenerate sites, the Romiguier et al. data is
synonymous sites, and the Leffler et al. data is estimated predominantly from
silent, intronic, and non-coding sites. All types of diversity estimates from
\textcite{Leffler2012-zj} were included to maximize the taxa in the study,
since the variability of diversity across functional categories is much less
than the diversity across taxa. Multiple diversity estimates per taxa were
averaged. The total recombination map length data were from both
\citeauthor{Stapley2017-fs} (2017; 30 taxa), and
\citeauthor{Corbett-Detig2015-gt} (2015; 11 taxa). Both studies used
sex-averaged recombination maps estimated with cross-based approaches; in some
cases errors in the original data were found, documented, and corrected. These
studies also included genome size estimates used to create Supplementary
Figures \ref{supfig:genome_size_popsize} and
\ref{supfig:genome_size_maplength}.

\subsection*{Macroecological Estimates of Population Size}
\label{sec:methods-popsize}

A rough approximation for total population size (census size) is $N_c = D R$,
where $D$ is the population density in individuals per $\text{km}^2$ and $R$ is
the range size in $\text{km}^2$. Since population density estimates are not
available for many taxa included in this study, I used the macroecological
abundance-body size relationship to predict population density from body size.
Since body length measurements are more readily available than body mass, I
collated body length data from various sources (see
\url{https://github.com/vsbuffalo/paradox_variation/}); body lengths were
averaged across sexes for sexually dimorphic species, and if only a range of
lengths was available, the midpoint was used.

Then, I re-estimated the relationship between body mass and population density
using the data in the appendix table of \textcite{Damuth1987-sg}, which
includes 696 taxa with body mass and population density measurements across
mammals, fish, reptiles, amphibians, aquatic invertebrates, and terrestrial
arthropods. Though the abundance-body size relationship can be noisy at small
spatial or phylogenetic scales \parencite[Chapter 5,][]{Gaston2008-pn}, across
deeply diverged taxa such as those included in this study and
\textcite{Damuth1987-sg}, the relationship is linear and homoscedastic (see
Supplementary Figure \ref{suppfig:damuth}). Using Stan
\parencite{Stan_Development_Team2020-ea}, I jointly estimated the relationship
between body mass from body length using the \textcite{Romiguier2014-bp} taxa,
and used this relationship to predict body mass for the taxa in this study.
These body masses were then used to predict population density simultaneously,
using the \textcite{Damuth1981-st} relationship. The code of this routine
(\texttt{pred\_popsize\_missing\_centered.stan}) is available in the GitHub
repository (\url{https://github.com/vsbuffalo/paradox_variation/}).

To estimate range, I first downloaded occurrence records from Global
Biodiversity Information Facility \parencite{noauthor_2020-ey} using the
\texttt{rgbif} R package \parencite{Chamberlain2014-up,Chamberlain2017-uz}.
Using the occurrence locations, I inferred whether a species was marine or
terrestrial, based on whether the majority of their recorded occurrences
overlapped a continent using \texttt{rnaturalearth} and the \texttt{sf}
packages \parencite{South2017-db,Pebesma2018-bs}. For each taxon, I estimated
its range by finding the minimum $\alpha$-shape containing these occurrences.
The $\alpha$ parameters were set more permissive for marine species since
occurrence data for marine taxa were sparser. Then, I intersected the inferred
ranges for terrestrial taxa with continental polygons, so their ranges did not
overrun landmasses (and likewise with marine taxa and oceans). I inspected
diagnostic plots for each taxa for quality control (all of these plots are
available in \texttt{paradox\_variation} GitHub repository), and in some cases,
I manually adjusted the $\alpha$ parameter or manually corrected the range
based on known range maps (these changes are documented in the code
\texttt{data/species\_ranges.r} and \texttt{data/species\_range\_fixes.r}).
The range of \emph{C. elegans} was conservatively approximated as the area of
the Western US and Western Europe based on the map in \textcite{Frezal2015-iw}.
\emph{Drosophila} species ranges are from the Drosophila Speciation Patterns
website, \parencite{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate
these range estimates, I have compared these to the qualitative range
descriptions \textcite{Leffler2012-zj} (Supplementary Figure
\ref{suppfig:range-cats}) and compared my $\alpha$-shape method to a subset of
taxa with range estimates from IUCN Red List
(\cite{Chamberlain2020-tv,Iucn2020-ap}; Supplementary Figure
\ref{suppfig:redlist-ranges}). Each census population size is then estimated as
the product of range and density.

\subsection*{Phylogenetic Comparative Methods}
\label{sec:methods-pcm}

Of the full dataset of 172 taxa with diversity and population size estimates, a
synthetic calibrated phylogeny was created for 166 species that appear in
phylogenies in DateLife project \parencite{OMeara2020-ds,Sanchez-Reyes2019-io}.
This calibrated synthetic phylogeny was then subset for the analyses based on
what species had non-missing trait data. The diversity-population size
relationship assessed by a linear phylogenetic mixed-effects model implemented
in Stan \parencite{Stan_Development_Team2020-ea}, according to the methods
described in (\cite{De_Villemereuil2014-kt}, see
\texttt{stan/phylo\_mm\_regression.stan} in the GitHub repository). This same
Stan model was used to estimate the same relationship between arthropod,
chordate, and mollusc subsets of the data (Supplementary Figure
\ref{suppfig:div_popsize_post_phyla}).

The relationship between recombination map length and the logarithm of
population size is non-linear and heteroscedastic, and was fit using a
lognormal phylogenetic mixed-effects model. Since social insects have longer
recombination map lengths \parencite{Wilfert2007-dx}, social taxa were excluded
when fitting this model.  All \texttt{Rhat} \parencite{Vehtari2019-po} values
were below 1.01 and the effective number of samples was over 1,000, consistent
with good mixing; details about the model are available in the GitHub
repository (\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (Figure
\ref{fig:figure-2}A and Supplementary Material Figure \ref{suppfig:biconmap})
were created using \texttt{phytools} \parencite{Revell2012-zq}. Node-height
tests were implemented based on the methods in \texttt{Geiger}
\parencite{Pennell2014-nx,Harmon2008-tr}, and use robust regression to fit a
linear relationship between phylogenetic independent contrasts and branching
times.

\subsection*{Predicted Reductions in Diversity}
\label{sec:methods-reduction}

The predicted reductions in diversity due to linked selection are approximated
using selection and deleterious mutation parameters from \emph{Drosophila
melanogaster}, and the recombination map length estimates from
\textcite{Stapley2017-fs} and \textcite{Corbett-Detig2015-gt}. The mathematical
details of the simplified sweep model are explained in the Appendix Section
\ref{app:sweep}. I use estimates of the number of substitutions, $m$, in genic
regions between \emph{D.  melanogaster} and \emph{D. simulans} from
\textcite{Hu2013-pg}.  Following \textcite{Elyashiv2016-vt}, only substitutions
in UTRs and exons are included, since they found no evidence of sweeps in
introns. Then, I average over annotation classes to estimate the mean
proportion of substitutions that are beneficial, $\alpha_\text{Dmel}=0.42$,
which are consistent with the estimates of Elyashiv et al. and estimates from
MacDonald--Kreitman test approaches (see \cite{Eyre-Walker2006-li}, Table 1).
Then, I use divergence time estimates between \emph{D. melanogaster} and
\emph{D. simulans} of $4.2 \times 10^6$ and estimate of ten generations per
year \parencite{Obbard2012-on}, calculating there are $\gamma_\text{Dmel} =
\nicefrac{\alpha m}{2T} = 2.26 \times 10^{-3}$ substitutions per generation.
Given the length of the \emph{Drosophila} autosomes, $G$, this implies that the
rate of beneficial substitutions per basepair, per generation is $\nu_{BP,Dmel}
= \nicefrac{\gamma_\text{Dmel}}{G} = 2.34 \times 10^{-11}$. Finally, I estimate
$J_\text{Dmel}$ from the estimate of genome-wide average rate of sweeps from
Elyashiv et al. (Supplementary Table S6) and assuming \emph{Drosophila} $N_e =
10^6$. These \emph{Drosophila melanogaster} hitchhiking parameter estimates are
close to other previously-published estimates (Supplementary Figure
\ref{suppfig:linked-sel-params}). Finally, I use $U_\text{Dmel} = 1.6$, from
\textcite{Elyashiv2016-vt}. With these parameter estimates from \emph{D.
melanogaster}, the recombination map lengths across species, and Equation
\eqref{eqn:div}, I estimate $\pi_\text{BGS+HH}$ (assuming $N_c=N_c$) across all
species. This leads to a range of predicted diversity ranges across species
corresponding to $\mu=10^{-8}$--$10^{-9}$; to visualize these, I take a convex
hull of all diversity ranges and smooth this with R's \texttt{smooth.spline}
function.

\section*{Acknowledgments}

I would like to thank Andy Kern and Peter Ralph for helpful discussions and
supporting me during this work, and Graham Coop for inspiration and helpful
feedback during socially distanced nature walks at Yolo Basin. I thank Jessica
Stapley for kindly providing the recombination map length data, and Yaniv
Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Sally Otto, Jeff
Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli, and my
Kern-Ralph labmates for helpful discussions. Sarah Friedman, Katherine Corn,
and Josef Uyeda provided very useful advice about phylogenetic comparative
methods; yet I take full responsibility for any shortcomings of my analysis.  I
would like to also thank UO librarian Dean Walton for helping me track down
some rather difficult to find older papers. This work was supported by an NIH
Grant (1R01GM117241) awarded to Andrew Kern.

\section*{Appendix}

\setcounter{table}{0}
\renewcommand{\theequation}{A\arabic{equation}}
\setcounter{section}{0}
\renewcommand{\thesection}{A\arabic{section}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\section{Simplified Sweep Effects Model}
\label{app:sweep}

I use a simplified model of the effects of recurrent hitchhiking and background
selection (BGS) occurring uniformly along a genome. Expected diversity is given
by

\begin{align}
  \E(\pi) &= \frac{\theta}{\theta + \nicefrac{1}{B} + 2NS} \\
          &\approx \frac{\theta}{\nicefrac{1}{B} + 2NS} 
\end{align}
%
(cf. equation 1 \cite{Elyashiv2016-vt}, and equation 20 of \cite{Coop2012-cd}).
The BGS component is given by \textcite{Hudson1995-xc},

\begin{align}
  B(U, L) = N_e \exp\left(-\frac{U}{L}\right)
\end{align}
%
and the hitchhiking component is

\begin{align}
  S = \frac{\nu_\text{BP}}{r_\text{BP}} J
\end{align}
%
(cf. \cite{Coop2012-cd} equation 20) where $J$ is the probability that two
lineages coalesce down to one, given sweeps occur uniformly along the genome.
Under this homogeneous sweep model, $J$ is

\begin{align}
  J = \int_0^L q_f(r)^2 dr 
\end{align}
%
where $q_f(r)$ is the approximate probability that a lineage is trapped by a
sweep to frequency $f$ when it is $r$ recombination fraction away from this
sweep (cf. \cite{Coop2012-cd} equation 15).

Since I use \emph{Drosophila melanogaster} parameter estimates from
\textcite{Elyashiv2016-vt}, I now reconcile their model's $S$ term with the
simple model above. They estimate $S$ in \emph{Drosophila melanogaster} using a
composite likelihood model that considers hitchhiking and background selection
simultaneously, using substitutions and stratifying by annotation.  For a
neutral position at site $x$, the coalescent rate due to sweeps is given by
Elyashiv et al.'s equation 3,

\begin{align}
  S(x) = \frac{1}{T} \sum_{i_S} \alpha(i_S) \sum_{y \in a(i_S)} \int \exp(-r(x, y) \tau(s, N)) g(s | i_S) ds
\end{align}
%
where $T$ is the number of generations that substitutions accrue, $i_S = 1,
\ldots, I_S$ is the annotation class (e.g. exons, introns, UTRs), $\alpha(i_S)$
is the fraction of substitutions in annotation class $i_S$ that are beneficial,
$a(i_S)$ is the set of all substitutions in annotation class $i_S$, $\tau(s,
N)$ is the fixation time of a site with additive effect $s$, and $g(s | i_S)$
is the distribution of selection coefficients for annotation class $i_S$.

Note, that we can recover the model of \textcite{Coop2012-cd} from this
expression. Suppose there is only one annotation class, and $\alpha$ fraction
of substitutions are beneficial, and one selection coefficient $\bar{s}$, (i.e.
$g(s) = \delta_0(s - \bar{s})$), then

\begin{align}
  S(x) = \frac{\alpha}{T} \sum_{y \in {a}} \exp(-r(x,y)\tau(\bar{s}, N)).
\end{align}
%

Let the number of substitutions be $m := |a|$, and imagine their positions are
uniformly distributed on a segment of length $G$ basepairs with the focal site
is the middle at position $x=0$. Then, each substitution $y$ is a random
distance $l_y \sim U(-\nicefrac{G}{2}, \nicefrac{G}{2})$ away from the focal
site. Assuming the recombination rate is a constant $r_\text{BP}$ per
basepair, and approximating the sum with an integral, we have,

\begin{align}
  S &= \frac{\alpha}{T} \sum_{i=1}^m \E_{l_i} \left(\exp(-r_\text{BP} l_i \tau(\bar{s}, N)) \right) \\
    &= \frac{\alpha}{T G} \sum_{i=1}^m \int_0^{G} \exp(-r_\text{BP} \ell \tau(\bar{s}, N)) d\ell \\
    &= \frac{\alpha m}{T G} \int_0^{G} \exp(-r_\text{BP} \ell \tau(\bar{s}, N)) d\ell
\end{align}

Using $u$-substitution with $r = \ell r_\text{BP}$ this simplifies to

\begin{align}
  S &= \frac{\alpha m}{T G r_\text{BP}} \int_0^{L} \exp(-r \tau(\bar{s}, N)) dr
\end{align}
%
where $L = G r_\text{BP}$.

To simplify this notation, note that the rate of adaptive substitutions per
basepair per generation is $\nu_\text{BP} = \nicefrac{\alpha m}{GT}$, so

\begin{align}
  S &= \frac{\nu_\text{BP}}{r_\text{BP}} \int_0^{L} \exp(-r \tau(\bar{s}, N)) dr
\end{align}

This is analogous to the second term of \textcite{Coop2012-cd} equation 17,
with $k = i = 2$ and $x = 1$ (e.g.  conditioning on a sweep to fixation). Note
that there appears to be a factor of two error in \textcite{Elyashiv2016-vt}
compared to \textcite{Coop2012-cd}; here I include the factor of two.  Then,

\begin{align}
  S &= \frac{\nu_\text{BP}}{r_\text{BP}} \underbrace{\int_0^{L} \exp(-2r \tau(\bar{s}, N)) dr}_{J}
\end{align}
%

where the integral is equal to $J$ (c.f. $J_{2,2}$ of equation 15 in
\cite{Coop2012-cd}) since a simple model of $q_f(r) = f \exp(- 2r \tau(s, N))$
and if we condition on fixation, $f = 1$. This expression is useful to
generalize across species, since we know $N$ and $L$. Additionally, we have
estimates of $\alpha$ and $\nicefrac{m}{T}$ in \emph{Drosophila} and other
species. In Elyashiv et al, they consider the number of substitutions per
generation in genic regions only; it should be noted that the number of coding
basepairs varies little across species. For convenience, I define $\gamma =
\nicefrac{\alpha m}{T}$ as the number of adaptive substitutions per generation
per entire genome, such that $S(\gamma, L, J) = \nicefrac{\gamma}{L} \; J$ used
in the main text. Using the estimates of $m \approx 4.5 \times 10^{5}$, $\alpha
\approx 0.42$, and $T \approx 8.4 \times 10^{7}$ from the Supplementary
Material of Elyashiv et al., I arrive at $\gamma \approx 0.00226$ adaptive
substitutions per generation, per genome. For a $\approx 100$ megabase genome,
this translates to a $\nu_\text{BP} \approx 2.34 \times 10^{-11}$, which is
close to previous estimates (Supplementary Figure
\ref{suppfig:linked-sel-params}). For $J$, I use an empirical estimate
calculated from the genome-wide average of the rate of coalescent events due to
sweeps, from Supplementary Table S6 of Elyashiv et al. ($r_s = 2NS \approx
0.92$). This implies $J \approx 4.46 \times 10^{-4}$. Alternatively, I have
tried using the estimated distribution of selection coefficients from Elyashiv
et al., but this led to a weaker estimate of $J$, since the adaptive
substitutions considered tend to cluster around genic regions. Note that these
\emph{Drosophila} sweep parameters I have used are close to previous estimates
(Supplementary Figures \ref{suppfig:linked-sel-params} A and B).

%Given that \textcite{Elyashiv2016-vt} provides a distribution of binned
%selection coefficients $g(s|i_S)$ by annotation class $i_S$ (e.g. UTRs, exons),
%I average over $i_S$, weighting by the number of basepairs in each annotation
%category, to calculate the marginal distribution of selection coefficients $g(s)$. 
%Then, I calculate $J$ by marginalizing over $g(s)$,

%\begin{align}
%  J = \sum_{s} g(s) \int_{0}^{L} \exp(- 2 r \tau(s, N)) dr.
%\end{align}
%%
%Throughout, I use the expression for fixation time,

%\begin{align}
%  \tau(s, N) = \frac{2}{s} \left(\log(4 N s) + c - \nicefrac{1}{4Ns}\right)
%\end{align}
%%
%from equation A17 of \textcite{Hermisson2005-hs}, re-parameterized so the
%fitness of the heterozygote is $1+s$, and $c \approx 0.577$ is Euler's gamma.

%% We can take $L \to \infty$ as
%% in \textcite{Coop2012-cd}, leaving

%% \begin{align}
%%   S &= \frac{\alpha m}{T} \frac{1}{L \tau(\bar{s}, N)} 
%% \end{align}

\section{Background Selection and Polygenic Fitness Models}
\label{app:bgs}

Throughout the main text, I use recurrent hitchhiking and background selection
models to estimate the reduction in diversity due to linked selection. Another
class of linked selection models, which I refer to as quantitative genetic
linked selection models (QGLS;
\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}), can also depress
genome-wide diversity. Furthermore, these models may depress diversity at neutral sites unlinked to the regions containing fitness variation.  While I
did not explicitly incorporate these models into my estimates of the diversity
reductions, their effect is implicit in background selection models because
they are analytically nearly identical. Here, I briefly sketch out the
connection between BGS and QGLS models.

Under the \textcite{Santiago1998-bs} model, the effective population size is
$N_e^\text{SC98} = N \exp(\nicefrac{-C^2}{(1-Z)L})$, where $C^2$ is the
standardized heritable fitness variation, $1-Z$ is the decay of genetic
variance through time, and $L$ is the recombination map length. This model can
accommodate a variety of modes of selection such as selection on an
infinitesimal trait \parencite[p.  1016]{Santiago1995-hx}, and the flux of
either weakly advantageous or deleterious alleles \parencite[p.
2109]{Santiago1998-bs}. If the source of fitness variation is entirely the
input of new deleterious mutations with heterozygous effect $s h$ at rate $U$
per diploid genome per generation, then under mutation-selection balance, the
equilibrium relative variance in reproductive success $C^2 = U sh$
(\cite{Crow1970-wm}; \cite{Caballero2020-wm}, p. 167), and $Z = 1 - sh -
\nicefrac{1}{2N_c}$ \parencite{Santiago1998-bs}. Thus, if $\nicefrac{1}{2N_c}
<< sh << 1$, then $\nicefrac{C^2}{(1-Z)} \approx U$ and $N_e^\text{SC98}
\approx N \exp(\nicefrac{-U}{L})$, which is the BGS model used in the main text
and is a result of many background selection models with similar assumptions
(\cite{Hudson1994-oh} eqn.  15; \cite{Hudson1995-xc} eqn.  9;
\cite{Nordborg1996-nq} eqn. 4; \cite{Barton1995-pm} eqn. 22b). Intuitively, the
similarity of these models reflects the fact that a substantial proportion of
heritable fitness variation is caused by the continual flux of deleterious
alleles across the genome under mutation-selection balance
\parencite{Charlesworth2015-am,Charlesworth2000-km}.

\printbibliography

\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{equation}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{section}{0}
\renewcommand{\thesection}{S\arabic{section}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}

\section*{Supplementary Table and Figures}

\section{Diversity and IUCN Red List Status}
\begin{table}[ht]
  \centering
  \begin{tabular}{rrrr}
    \hline
 & mean & 2.5 \% & 97.5 \% \\ 
 \hline
    $\beta_0$ & -2.70 & -3.00 & -2.40 \\ 
    $\beta_{LC}$ & -0.37 & -0.55 & -0.19 \\ 
    $\beta_{NT}$ & -0.19 & -0.80 & 0.41 \\ 
    $\beta_{VU}$ & -0.31 & -0.81 & 0.19 \\ 
    $\beta_{EN}$ & -0.39 & -0.71 & -0.06 \\ 
    $\beta_{CR}$ & -0.03 & -0.65 & 0.59 \\ 
    $\beta_{N_c}$ & 0.06 & 0.04 & 0.09 \\ 
    \hline
  \end{tabular}

  \caption{The regression estimates of full IUCN Red List population size model
    for diversity, $\log_{10}(\pi) = \beta_0 + \beta_{LC} LC + \beta_{NT} NT +
    \beta_{VU} VU + \beta_{EN} EN + \beta_{CR} CR + \beta_{N_c} \log_{10}(N_c)$;
    $df = 165$.  Using AIC to compare this full model to a reduced model of
    $\log_{10}(\pi) = \beta_0 + \beta_{N_c} \log_{10}(N_c)$, $AIC_\text{full} =
  203.99$, $AIC_\text{reduced} =212.46$. }

\end{table}
\label{supptable:tbl-1} 
\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/diversity_popsize_linear.pdf}

  \caption{A version of Figure \ref{fig:figure-1} with diversity on a linear,
    rather than log, scale. Points are colored by phylum, and the shaded region
    is the predicted neutral level of diversity assuming $N_e = N_c$ with mutation
  range ranging between $10^{-10} \le \mu \le 10^{-8}$. }

  \label{suppfig:figure-1-linear}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/diversity_popsize_averages.pdf}

  \caption{Diversity and approximate population size for 172 taxa, colored by
    phylum; the dashed lines indicate the non-phylogenetic OLS estimates of the
  relationship between population size and diversity grouped by phyla.}

  \label{suppfig:figure-1-ave}
\end{figure}

\begin{figure}[!htb] \centering
  \includegraphics[width=\textwidth]{figures/diversity_range_bodymass.pdf}

  \caption{The relationship between diversity (differences per basepair) and
    body mass (left) and range (right) across 172 species. The top row are
    posterior distributions of parameters estimated using the phylogenetic
    mixed-effects model using 166 taxa in the synthetic phylogeny for the
    intercept, slope, and phylogenetic signal from the mixed-effects model. The
    bottom row contain each species as a point, colored by phyla. The gray
    dashed line is the non-phylogenetic standard regression estimate, and the
    blue dashed line is the relationship fit by the phylogenetic mixed-effects
  model. }

  \label{suppfig:figure-div-range-bodymass}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/genome_size_popsize.pdf}

  \caption{The relationship between genome size and approximate census
    population size. The dashed gray line indicates the OLS fit. Tiger
    salamander (\emph{Ambystoma tigrinum}) was excluded because of its
  exceptionally large genome size (~30Gbp).}

  \label{supfig:genome_size_popsize}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/genome_size_maplength.pdf}

  \caption{The relationship between genome size and recombination map length.
    The dashed gray line indicates the OLS fit for all taxa, and the dashed
    colored dashed lines indicate the linear relationship fit by phyla. Tiger
    salamander (\emph{Ambystoma tigrinum}) was excluded because of its
  exceptionally large genome size (~30Gbp).}

  \label{supfig:genome_size_maplength}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/iucn_redlist_ranges.pdf}

  \caption{The correspondence between the ranges estimated with the alpha hull
    method applied to GBIF data used in this paper and IUCN Red List's Extent
    of Occurrence for the subset of species in both datasets. Note that the
    IUCN Red List contains predominantly endangered species, which leads to
    ascertainment bias; still, the high correlation between the estimated
  ranges shows the alpha hull method works well.}

  \label{suppfig:redlist-ranges}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/range_categories.pdf}

  \caption{The estimated ranges using GBIF occurrence data, ordered within and
    colored by the original range category labels assigned in
  \textcite{Leffler2012-zj}.}

  \label{suppfig:range-cats}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/damuth.pdf}

  \caption{The appendix table of \textcite{Damuth1987-sg}; the color indicates
    Damuth's original group labels. The dashed line was estimated using a
    lognormal regression model in Stan. References to each measurement are
  available in \textcite{Damuth1987-sg}.}

  \label{suppfig:damuth}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/body_mass_length.pdf}

  \caption{The relationship between body length (meters) and body mass (grams)
    in the \textcite{Romiguier2014-bp} data set, used to infer body masses for
  taxa. The gray dashed line is the line of best fit inferred using Stan.}

  \label{suppfig:body-mass-length}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/corbett_detig.pdf}

  \caption{(A) The diversity data from \textcite{Corbett-Detig2015-gt} and the
    census population size estimated here for metazoan taxa. (B) The
    reductions in diversity, $R = \nicefrac{N_e}{N}$, plotted against census
    size across species. The red points are the reductions estimated by
    \textcite{Corbett-Detig2015-gt}. This confirms
    \citeauthor{Corbett-Detig2015-gt}'s (2015) finding that the impact of
    selection ($I = 1-R$) increases with census population size (though, in the
    original paper size body size and range were used as separate proxy
    variables for census population size). The green and red points are the
    predicted reduction in diversity under the recurrent hitchhiking (RHH) and
    background selection (BGS) model using the \emph{Drosophila melanogaster}
    parameters as described in the main text.  The reduction in the diversity
    due to sweeps, from Equation \eqref{eqn:div}, is determined by the term
    $2NS$.  Green points treat $N$ as the implied effective population size
    from diversity $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, assuming
    $\mu = 10^{-9}$. Yellow points treat $N$ as the census size, $N = N_c$.
    Overall, using the census size, e.g.  $2N_cS$, leads to reductions in
    diversity that far exceed the empirical estimates of Corbett-Detig et al.
  and reasonable model-based predictions from $\widetilde{N}_e$.}

  \label{suppfig:corbett_detig}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/linked_sel_params.pdf}

  \caption{(A) The estimate of the number of sweeps per basepair, per genome
    ($\nu_\text{BP}$) from Table 2 of \textcite{Elyashiv2016-vt} (the studies
    included are \cite{Li2006-xa,Andolfatto2007-uy,Macpherson2007-qt} and
    \cite{Jensen2008-vz}); the red point is my estimate used in
    this paper. (B) Points are the data from \textcite{Shapiro2007-gp}.  The
    blue line is the non-linear least squares fit to the data, and the green
    dashed line is the sweep model parameterized by the genome-wide average
    sweep coalescent rate $2NS \approx 0.92$ from the classic sweep and
  background selection model of \textcite{Elyashiv2016-vt} ($r_s$ in
Supplementary Table S6).}

  \label{suppfig:linked-sel-params}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/div_popsize_post_phyla.pdf}

  \caption{The posterior distributions for the parameters of the phylogenetic
    mixed-effects model of diversity and population size (this is analogous to
    Figure \ref{fig:figure-2}B) fit separately on chordates ($n=68$), molluscs
    ($n=13$), and arthropods ($n=68$). The phylogenetic mixed-effects model for
    chordates indicated the best-fitting model had no residual variance
    ($\sigma_r^2 = 0$), so an alternate model without this variance component
    was used to ensure proper convergence; this model is shown in green.
    The light blue (green) shaded regions are the 90\% credible intervals, the
    blue (green) lines the posterior averages, the gray shaded regions the OLS
    bootstrap 95\% confidence intervals, and the gray lines the OLS estimate.
    Note that unlike Figure \ref{fig:figure-2}, the OLS estimate uses all taxa,
    not just those present in the phylogeny, since splitting the data by phyla
    reduces sample sizes (OLS with just the subset of taxa in the phylogeny is
    not significant for either chordates and arthropods). The vertical dashed
  gray line indicates zero.}

  \label{suppfig:div_popsize_post_phyla}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/biconmap_diversity.pdf}

  \caption{}

  \label{suppfig:biconmap_diversity}
\end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figures/biconmap.pdf}

  \caption{}

  \label{suppfig:biconmap}
\end{figure}

\end{document}

### 2021-02-08 notebooks/main_analysis.Rmd
New repository with squashed history

---
title: "Main Analyses of the Combined Dataset"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

# Main Analyses of the Combined Dataset

```{r, include=FALSE}
options(warnPartialMatchArgs = FALSE,
        warnPartialMatchDollar = FALSE, 
        warnPartialMatchAttr = FALSE)

knitr::opts_chunk$set(warning = FALSE, message=FALSE)
```

```{r, message=FALSE}

library(nlme)
library(ape)
library(geiger)
library(phytools)
library(broom)
library(xtable)
library(brms)
library(rotl)
library(HDInterval)
library(scales)
library(tidyverse)
library(rstan)
library(loo)
library(viridis)
library(coda)
library(ggrepel)
library(nls.multstart)
source('../R/utilities.r')
source('../R/node_height.r')
set.seed(0)

```

## Overall Plan

This is the main analysis. Publication-quality figures are in
`notebooks/figures` and are each made by their own scripts. This R markdown
notebook uses curated data that's created in `data/` using the
`combined_dataset.r` R script. The sources for this file are all managed by the
`Makefile` in `../data/`.

## Load the Combined Final Merged Dataset

The combined dataset includes data from Leffler et al. (2012), Corbett-Detig
(2015), Stapley et al. (2017), and Romiguier et al. (2014) data. It's all
merged and processed in `data/combined_dataset.r`, which also uses the `taxadb`
and `ritis` to get taxonomical groupings, and the `datelife` package to get a
tree from the Open Tree of Life and calibrate it using datelife.org.

Note that I filter out the parasite/pathogen species because range size
estimates are unreliable (since they depend on host and occurrence data is more
shoddy).

```{r}
d <- read_tsv('../data/combined_data.tsv')

```

Also, I load the Open Tree of Life phylogeny (`otl_tr`) and calibrated datelife
phylogeny (`tr`) created from the `combined_dataset.r` pipeline.

```{r}
load('../data/calibrated_phylogeny.Rdata')
stopifnot(exists('da_tr') && exists('otl_tr') && exists('tr')) 
```

## Curate the main datasets

Our main dataset has a great deal of missing data, since not all taxa have
population sizes, diversity, and map length data. Here, create the main
datasets of the subsequent analyses.

Main datasets (all of metazoan taxa only):

 - `d`: full dataset, with taxa not in phylogeny and missing variables.
 - `dtr`: the subset of the full dataset with taxa in the phylogeny.
 - `da_cc`: all complete cases: map length, diversity, and pop size.
 - `dt_cc`: all complete cases with phylo taxa: map length, diversity, and pop size.
 - `dt_dps`: diversity/population size complete cases with phylo taxa.
 - `da_dps`: diversity/population size complete cases all taxa.

Note: `dt` prefixes are for tree taxa only (`dtr` is used for no suffix, since
`dt` is a function), `da` prefix are for all taxa.

```{r}

# easier to type phyla taxa dataset
dtr <- da_tr
stopifnot(length(tr$tip.label) == nrow(dtr))

# complete data 
da_cc <- d %>% filter(is.finite(log10_popsize), 
                      is.finite(map_length), 
                      is.finite(log10_diversity), 
                      is.finite(log10_body_mass), 
                      is.finite(log10_range))

dt_cc <- dtr %>% filter(is.finite(log10_popsize), 
                        is.finite(map_length), 
                        is.finite(log10_diversity), 
                        is.finite(log10_body_mass), 
                        is.finite(log10_range))
# now, prune the tree
tr_cc <- keep.tip(tr, dt_cc$species)
stopifnot(length(tr_cc$tip.label) == nrow(dt_cc))

# diversity/popsize data
dt_dps <- dtr %>% filter(is.finite(log10_popsize), 
                         is.finite(log10_body_mass), 
                         is.finite(log10_diversity), 
                         is.finite(log10_range))
# now, prune the tree
tr_dps <- keep.tip(tr, dt_dps$species)
stopifnot(length(tr_dps$tip.label) == nrow(dt_dps))

da_dps <- d %>% filter(is.finite(log10_popsize), 
                       is.finite(log10_body_mass), 
                       is.finite(log10_diversity), 
                       is.finite(log10_range))

# map length / popsize data
dt_ml <- dtr %>% filter(is.finite(log10_popsize), 
                        is.finite(log10_body_mass), 
                        is.finite(log10_map_length))
# now, prune the tree
tr_ml <- keep.tip(tr, dt_ml$species)
stopifnot(length(tr_ml$tip.label) == nrow(dt_ml))

da_ml <- d %>% filter(is.finite(log10_popsize), 
                      is.finite(log10_body_mass), 
                      is.finite(log10_map_length))

save(d, dtr, 
     da_cc, dt_cc, tr_cc, 
     dt_dps, tr_dps, da_dps, 
     dt_ml, tr_ml, da_ml, file='../data/main_datasets.Rdata')

```

We lost a few species in matching taxa to the R Open Tree of Life tree; we
quantify that here:

```{r, message=TRUE}

msg <- "Size of animalia in combined dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(d), nrow(dtr)))

msg <- "Size of animalia in complete dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(da_cc), nrow(dt_cc)))

msg <- "Size of animalia in popsize/diversity dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(da_dps), nrow(dt_dps)))

msg <- "Size of animalia in map length/popsize dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(da_ml), nrow(dt_ml)))

```

Finally, we load in the posterior draws of the predicted population sizes.

```{r}

load('../data/post_pred_draws.Rdata')

post_popsizes <- post_popsizes %>% 
                  mutate(log10_mean = map_dbl(log10_popsize, mean),
                         log10_sd = map_dbl(log10_popsize, sd))

```

## Diversity Data Source

Is there good mixing across data sources in the combined diversity date set?

```{r}

ggplot(d, aes(log10_popsize, log10_diversity, color=diversity_data_source)) + geom_point()

```

Yes, looks like so.

## Some Map Length, Body Size, and Chromosome Number EDA

Here, are some quick EDA plots. first, chromosome number and the map length
(mostly from Stapley et al data):

```{r}

d  %>% ggplot(aes(hcn, map_length)) + geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('map length (Morgans)')

```

Body size and map length:

```{r}

d %>% 
  filter_at(vars(hcn, log10_size), ~ !is.na(.)) %>%
  ggplot(aes(log10_size, hcn)) + geom_point() + 
  geom_smooth(method='lm') + ylab('haploid chromosome number') + 
  xlab('log10(size)')

```

Morgans per chromosome by chromosome number:

```{r}
d %>% ggplot(aes(hcn, map_length/hcn)) + 
  geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('Morgans per chromosome (map length / haploid chromosome number)')

```

What do number of Morgans / chromosome across the data look like?

```{r}

d %>% mutate(morgans_per_chrom = map_length/hcn) %>% 
  pull(morgans_per_chrom) %>% hist(50, main='Morgans / chromosome')

```

## Population Size Biomass Analysis

As a quality control on population size estimates, we can look at humans (we
know they have a modern population size of ~8 billion):

```{r}
d %>% filter(species == 'Homo sapiens') %>% 
  select(log10_popsize, size, log10_density, log10_range)
```

This is low, but not terribly so.

What proportion are terrestrial?

```{r}

prop.table(table(d$is_terrestrial))

```

Based on the predictions of Mora et al. 2011

```{r}

nspecies <- c(terrestrial= 7.770e6, marine=2.15e5)
prop.table(nspecies)

```

Now we do a simple biomass analysis -- back of the envelop type stuff.

```{r}

# https://bionumbers.hms.harvard.edu/bionumber.aspx?&id=107256 
# 18-23% of human body is carbon; seems reasonable this holds
# for other animals

# only complete data
xd <- d %>% filter(is.finite(log10_body_mass))

# in Gt
prop_carbon <- signif(sqrt(0.18 * 0.23), 1)  # ~ 0.2
# 1Gt = 1e12 kg
xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * 10^xd$log10_popsize) / 1e12

mass_d <- xd %>% 
  # filter so we don't include NA species in nspecies
  filter(is.finite(carbon_mass), is.finite(log10_popsize)) %>%
  group_by(phylum) %>% 
  summarize(total_carbon_mass=sum(carbon_mass, na.rm=TRUE),
            total_log10_popsize=log10(sum(10^log10_popsize, na.rm=TRUE)),
            nspecies = n()) %>%
  mutate(total = sum(total_carbon_mass), prop_carbon_mass = total_carbon_mass / total)

# chordates are incredibly over-represented in my data set by biomass
baron_d <- tribble(~ phylum, ~ bd_mass,
                   'Arthropoda', 0.2 + 1, # terrestrial + marine
                             # fish + livestock + humans + wild mammals + wild birds
                   'Chordata', 0.7 + 0.1 + 0.06 + 0.007 + 0.002,
                   'Annelida', 0.2, 
                   'Mollusca', 0.2, 
                   'Cnidaria', 0.1,
                   'Nematoda', 0.02) %>% 
           mutate(bd_total_biomass = sum(bd_mass), 
                  bd_prop_mass = bd_mass/bd_total_biomass) %>%
           left_join(mass_d, by='phylum')  %>%
           # not we need to reweight the original proportions as some phyla were
           # dropped
           mutate(prop_carbon_mass = prop_carbon_mass / sum(prop_carbon_mass))

ggplot(baron_d, aes(bd_prop_mass, prop_carbon_mass, color=phylum))  + 
  geom_point(size=5) + geom_abline(slope=1, intercept=0) + 
  xlab('Bar-On et al. 2018 proportion') + 
  ylab('My proportion')

```

Arthropods are under sampled here, and chordates are way over sampled (at least
by biomass).

```{r}

ggplot(baron_d) + 
  geom_point(aes(phylum, total_carbon_mass), size=5)

ggplot(baron_d) + 
  geom_point(aes(phylum, bd_mass), size=5)

baron_d <- baron_d %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)

baron_d %>%
  select(phylum, prop_bd_biomass)

# using the Bar-On weights
baron_d <- baron_d %>% 
              mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
              mutate(reweighted_total_carbon_mass = reweight * total_carbon_mass)

total_sample_carbon_mass <- baron_d %>% pull(reweighted_total_carbon_mass) %>% sum()

# total species here
n <- nrow(xd)
total_animal_species <- sum(nspecies)

print(paste0("fraction of animal species in this data set: ", 
             signif(n/total_animal_species, 2)))

# Bar-On et al. (2018) estimate is 2Gt C
# 1e12 kg = 1Gt
total_animal_cmass <- 2

# if the species here were random
signif(total_animal_cmass * n / total_animal_species, 2)
signif(total_sample_carbon_mass, 2)

0.058 / total_animal_cmass

```

Note: initially it was my plan to see if we could get a sense of what fraction
of the total biomass this dataset contains (which is about 3%). This seems
unreasonably high, even after reweighting as above. The problem is that the
3.4e-5 fraction of animal species is obviously not a random sample based on
size / population, and includes humans, domesticated species, etc. -- this
probably is accentuated by the overabundance of chordates compared arthropods
here.

## Stan Model for Map Length and log10 Population Size

Fit the log-normal model, which seems to be a good fit to the data given the
heteroscedasticity.

```{r}
# no social taxa here; excluded

dt_ml_ns <- dt_ml %>% filter(!social)
tr_ml_ns <- keep.tip(tr_ml, dt_ml_ns$species)

# correlation matrix from the tree
ml_cor_ns <- vcv(tr_ml_ns, cor=TRUE)

# data for the log-normal fit
log10_popsize_seq <- seq(5, 18, length.out=100)
ml_dat <- list(N = nrow(dt_ml_ns), 
              log10_popsize = dt_ml_ns$log10_popsize,
              map_length = dt_ml_ns$map_length, 
              cor = ml_cor_ns,
              x_new = log10_popsize_seq)

PHYLO_MM_LOGNORMAL_CACHE <- '../data/phylo_mm_lognormal_chains.Rdata'

if (!file.exists(PHYLO_MM_LOGNORMAL_CACHE)) {
  ln_fit <- stan('../stan/phylo_mm_lognormal.stan', 
                 data=ml_dat, iter=5000, cores=4)
  save(ln_fit, file=PHYLO_MM_LOGNORMAL_CACHE)
} else {
  load(PHYLO_MM_LOGNORMAL_CACHE)
}

print(ln_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
beta <- extract(ln_fit, pars=c('beta'))

mu_ln_fit <- rstan:::extract(ln_fit, 'mu_rep')$mu_rep

save(ln_fit, log10_popsize_seq, mu_ln_fit,
     file='../data/lognorm_popsize_chains.Rdata')

```

Any serious pathologies? Rhats are low; I'm looking for divergent transitions.

```{r}
pairs(ln_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
```

Doesn't look like it.

What is phylogenetic heritability?

```{r}
lambda_est(ln_fit)

```

What is the credible interval?

```{r}
hdi(beta)

```

What does the posterior predictive distribution look like? Social taxa are
outliers (they have higher levels of recombination) and are shown in blue:

```{r}

mu_rep <- extract(ln_fit, pars='mu_rep')$mu_rep
cols <- c('gray32', 'cornflowerblue')
plot(map_length ~ log10_popsize, dt_ml, pch=19, 
     col=cols[factor(dt_ml$social)])
pred  <- colMeans(mu_rep)
lines(log10_popsize_seq, pred)

```

## Continuous Trait Map

Here I generate a bunch of continuous trait maps, and output data for plots.
First, a convenience function.

```{r}

# this simplifies things, since we're working with dataframes
# ordered by the tip labels, and phytools:::contMap requires
# named vectors
contmap <- function(tree, df, col, ...) {
  stopifnot(all(tree$tip.label == df$species))
  x <- setNames(df[[col]], df$species)
  phytools:::contMap(tree, x, ...)
}

```

```{r}

pscm <- contmap(tr_dps, dt_dps, 'log10_popsize', res=200)
dvcm <- contmap(tr_dps, dt_dps, 'log10_diversity', res=200)

save(pscm, dvcm, tr_dps, dt_dps, file='../data/biconmap_data.Rdata')

```

Now, for the map length / population size dataset.

```{r}

psmlcm <- contmap(tr_ml, dt_ml, 'log10_popsize', res=200)
mlcm <- contmap(tr_ml, dt_ml, 'log10_map_length', res=200)

save(mlcm, psmlcm, tr_ml, dt_ml, file='../data/biconmap_maplen_data.Rdata')

```

## Phylogenetic Mixed-Effects Model of Diversity Population Size Relationship

This is a primary relationship we want to fit, accounting for phylogeny:

```{r}

dps_cor <- vcv(tr_dps, corr=TRUE)
log10_popsize_seq <- seq(4, 16, length.out=100)

dps_dat <- list(N = nrow(dt_dps),
          x = dt_dps$log10_popsize,
          y = dt_dps$log10_diversity,
          x_new=log10_popsize_seq,
          cor=dps_cor)

PHYLO_MM_REGRESSION_CACHE <- '../data/phylo_mm_regression_chains.Rdata'
if (!file.exists(PHYLO_MM_REGRESSION_CACHE)) {
  dps_fit <- stan('../stan/phylo_mm_regression.stan', 
                  seed=0,
                  data=dps_dat, iter=5000, cores=4)
  save(dps_fit, file=PHYLO_MM_REGRESSION_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_CACHE)
}

print(dps_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
pairs(dps_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__'))

phi <- extract(dps_fit, pars='phi')$phi
beta <- extract(dps_fit, pars='beta')$beta
dps_mu_rep <- extract(dps_fit, pars='mu_rep')$mu_rep
plot(dps_dat$x, dps_dat$y, pch=19, xlab='log10 pop size', 
     ylab='log10 diversity')
lines(log10_popsize_seq, colMeans(dps_mu_rep))

dps_ols <- lm(log10_diversity ~ log10_popsize, dt_dps)
summary(dps_ols)

```

The chains mixed well. For quality assurance I compare this to a PGLS, to check
my Stan model is working not too differently than a pre-packaged solution.

```{r}

# compare with pgls
summary(gls(log10_diversity ~ log10_popsize, 
            correlation = corBrownian(phy=tr_dps), 
            data=dt_dps, method = "ML"))

```

This is reasonably close to the phylogenetic mixed-effects model.

Now, let's do a simple bootstrap to get the CIs for the OLS fit.

```{r}

fit_fun <- function(x) {
  lm(log10_diversity ~ log10_popsize, data=x)
}

nboot <- 10000
dps_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

dps_bs_cis <- rbind(ci(dps_bs$intercept, coef(dps_ols)[1]),
                    ci(dps_bs$slope, coef(dps_ols)[2]))
dps_bs_cis
```

Now, save all these data

```{r}

save(dps_fit, dps_ols, dt_dps, dps_bs_cis,
     file='../data/diversity_popsize_chains.Rdata')

```

How many taxa?

```{r}
dps_dat$N
```

Credible intervals:

```{r}
round(hdi(beta), 2)
round(lambda_est(dps_fit), 2)
```

## Diversity Population Size Relationship by Phylum

Next, let's look at phyla separately.

```{r}

min_taxa <- 10
dt_dps_phyla <- dt_dps %>% group_by(phylum) %>%
                  nest() %>% 
                  mutate(n = map_int(data, nrow)) %>%
                  filter(n > min_taxa)

fit_phylo_mm <- function(df, full_tree,
                         stan_file='../stan/phylo_mm_regression.stan', 
                         iter=5000) {
  tr <- keep.tip(tr, df$species)
  cor <- vcv(tr, corr=TRUE)
  log10_popsize_seq <- seq(4, 16, length.out=100)

  dat <- list(N = nrow(df),
              x = df$log10_popsize,
              y = df$log10_diversity,
              x_new=log10_popsize_seq,
              cor=cor)

  fit <- stan(stan_file,
              data=dat, iter=iter, seed=0, cores=4)
  fit
}

# fit all models: 
PHYLO_MM_REGRESSION_BY_PHYLA_CACHE <- '../data/phylo_mm_regression_by_phyla_chainsdf.Rdata'

if (!file.exists(PHYLO_MM_REGRESSION_BY_PHYLA_CACHE)) {
  dt_dps_phyla <- dt_dps_phyla %>% 
                   mutate(fit = map(data, fit_phylo_mm, full_tree=tr_dps))
  save(dt_dps_phyla, file=PHYLO_MM_REGRESSION_BY_PHYLA_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_BY_PHYLA_CACHE)
}

```

Now, we have to ensure convergence worked. For arthropods:

```{r}

dps_pars <- c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__')

pairs(dt_dps_phyla$fit[[3]], pars=dps_pars)

print(dt_dps_phyla$fit[[3]], pars=dps_pars)

```

This looks fine -- all Rhats are below 1.01 now.

For molluscs:

```{r}

pairs(dt_dps_phyla$fit[[2]], pars=dps_pars)

print(dt_dps_phyla$fit[[2]], pars=dps_pars)

```

Again, all Rhats are below 1.01, though clearly there is a high
degree of correlation between slope and intercept (low information
in data to differentiate these). We have some divergent
transitions but this seems to not indicate any pathologies in the
geometry of the posterior.

For chordates (hardest for last):

```{r}

pairs(dt_dps_phyla$fit[[1]], pars=dps_pars)
print(dt_dps_phyla$fit[[1]], pars=dps_pars)

```

Now, something is has gone wrong; Rhats are high and we have many
divergent transitions. The pairs plot shows many pathologies in
the geometry of the posterior surface. The histogram of `sigma_r`
shows that there's a posterior mode with low residual variance
that has a very high log posterior. This seems to indicate that
all the variance could be explained by phylogenetic effects,  not
residual tips.

This requires we fit a special model without `sigma_r` as a term.

```{r}

alt_stan_file <- '../stan/phylo_mm_regression_noresid.stan'

PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE <- '../data/phylo_mm_regression_phyla_alt_chains.Rdata'

if (!file.exists(PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE)) {
  chor_fit <- fit_phylo_mm(dt_dps_phyla$data[[1]], tr_dps, stan_file=alt_stan_file, iter=2000)
  save(chor_fit, file=PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE)
}

dps_pars_alt <- c('phi', 'beta', 'sigma_p', 'lp__')
pairs(chor_fit, pars=dps_pars_alt)
print(chor_fit, pars=dps_pars_alt)

```

This greatly helped in the model fitting process. Note that
parameter estimates have not changed:

```{r}
hdi(rstan:::extract(chor_fit, 'beta')$beta)
hdi(rstan:::extract(dt_dps_phyla$fit[[1]], 'beta')$beta)

```

Note that while it's tempting to compare log posteriors between
these two models as some sort of model comparison approach, this is
[not
appropriate](https://discourse.mc-stan.org/t/model-comparison-lp--/1849/3).

Now, we do simple OLS estimates and bootstraps, and save everything:

```{r}

dt_dps_phyla <- dt_dps_phyla %>% 
                 mutate(ols = map(data, fit_fun)) %>%
                 mutate(intercept = map_dbl(ols, ~ coef(.)[1]),
                        slope = map_dbl(ols, ~ coef(.)[2]))

nboot <- 10000

chor_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps_phyla$data[[1]]))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

arth_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps_phyla$data[[2]]))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

mols_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps_phyla$data[[3]]))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

dt_dps_phyla$boot <- list(chor_bs, arth_bs, mols_bs)

get_cis <- function(boot, slope, intercept) {
  cis <- ci(boot$slope, slope, with_est=FALSE)
  cii <- ci(boot$intercept, intercept, with_est=FALSE)
  tibble(slope_lower=cis[1], slope_upper=cis[2],
         intercept_lower = cii[1], intercept_upper=cii[2])
}

dt_dps_phyla <- dt_dps_phyla  %>%
      mutate(ci_bs = pmap(list(boot, slope, intercept), get_cis)) %>% unnest(ci_bs) %>%
      select(-boot)

save(dt_dps_phyla, chor_fit, 
     file='../data/phyla_diversity_popsize_chains.Rdata')

# just the OLS
ols_phyla <- dt_dps_phyla %>% select(-fit)
save(ols_phyla, file='../data/phyla_diversity_popsize_ols.Rdata')

```

## Phylogenetic Mixed-Effects Model of Diversity and Body Size Relationship

Is the relationship between diversity and population size driven by body mass?
We fit that here to take a look:

```{r}

log10_body_mass_seq <- seq(min(dt_dps$log10_body_mass),
                       max(dt_dps$log10_body_mass), length.out=100)

# body mass version
dbm_dat <- list(N = nrow(dt_dps),
          x = dt_dps$log10_body_mass,
          y = dt_dps$log10_diversity,
          x_new=log10_body_mass_seq,
          cor=dps_cor)

PHYLO_MM_REGRESSION_BODY_SIZE_CACHE <- '../data/phylo_mm_regression_body_size_chains.Rdata'

if (!file.exists(PHYLO_MM_REGRESSION_BODY_SIZE_CACHE)) {
  dbm_fit <- stan('../stan/phylo_mm_regression.stan', 
                  seed=0,
                  data=dbm_dat, iter=5000, cores=4)
  save(dbm_fit, file=PHYLO_MM_REGRESSION_BODY_SIZE_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_BODY_SIZE_CACHE)
}

print(dbm_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
pairs(dbm_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__'))

phi <- extract(dbm_fit, pars='phi')$phi
beta <- extract(dbm_fit, pars='beta')$beta
dbm_mu_rep <- extract(dbm_fit, pars='mu_rep')$mu_rep
plot(dbm_dat$x, dbm_dat$y, pch=19, xlab='log10 body mass', 
     ylab='log10 diversity')
lines(log10_body_mass_seq, colMeans(dbm_mu_rep))

dbm_ols <- lm(log10_diversity ~ log10_body_mass, dt_dps)
summary(dbm_ols)

```

The chains mixed well (Rhats < 1.01 and the pair plots look good). What's the
HDI of beta?

```{r}
hdi(beta)
```

For quality assurance, I run the PGLS and compare parameter estimates. They are
very similar (though it should be noted this model doesn't account for residual
variance):

```{r}
# compare with pgls
summary(gls(log10_diversity ~ log10_body_mass, 
            correlation = corBrownian(phy=tr_dps), 
            data=dt_dps, method = "ML"))

```

## Phylogenetic Mixed-Effects Model of Diversity and Range

Next, we look to see if the diversity population size relationship is driven by
another variable used in population size estimates: range.

```{r}

log10_range_seq <- seq(min(dt_dps$log10_range),
                       max(dt_dps$log10_range), length.out=100)

# body mass version
drn_dat <- list(N = nrow(dt_dps),
          x = dt_dps$log10_range,
          y = dt_dps$log10_diversity,
          x_new=log10_range_seq,
          cor=dps_cor)

PHYLO_MM_REGRESSION_RANGE_CACHE <- '../data/phylo_mm_regression_range_chains.Rdata'
if (!file.exists(PHYLO_MM_REGRESSION_RANGE_CACHE)) {
  drn_fit <- stan('../stan/phylo_mm_regression.stan', 
                  seed=0,
                  data=drn_dat, iter=5000, cores=4)
  save(drn_fit, file=PHYLO_MM_REGRESSION_RANGE_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_RANGE_CACHE)
}

print(drn_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
pairs(drn_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__'))

phi <- extract(drn_fit, pars='phi')$phi
beta <- extract(drn_fit, pars='beta')$beta
drn_mu_rep <- extract(drn_fit, pars='mu_rep')$mu_rep
plot(drn_dat$x, drn_dat$y, pch=19, xlab='log10 range', 
     ylab='log10 diversity')
lines(log10_range_seq, colMeans(drn_mu_rep))

drn_ols <- lm(log10_diversity ~ log10_range, dt_dps)
summary(drn_ols)

```

Again, we compare to the PGLS fit:

```{r}

# compare with pgls
summary(gls(log10_diversity ~ log10_range, 
            correlation = corBrownian(phy=tr_dps), 
            data=dt_dps, method = "ML"))

```

And again the estimates are close, but the PGLS slope estimates are weakened a
bit, likely by allowing for an additional source of variance (the diagonal
residual variance).

Now, we save all these fits for yet another supplementary figure.

```{r}

save(drn_fit, dbm_fit, drn_ols, dbm_ols, 
     log10_body_mass_seq, log10_range_seq,
     file='../data/diversity_range_bodymass_chains.Rdata')

```

## Node Height Tests

Now, let's look at node height.

```{r}

psnh <- node_height(dt_dps$log10_popsize, multi2di(tr_dps))
rnnh <- node_height(dt_dps$log10_range, multi2di(tr_dps))
bmnh <- node_height(dt_dps$log10_body_mass, multi2di(tr_dps))
dvnh <- node_height(dt_dps$diversity, multi2di(tr_dps))

save(psnh, rnnh, bmnh, dvnh, file='../data/nodeheights.Rdata')

```

Just to ensure there's nothing weird going on, let's simulate
Brownian motion 1000 times and see how many times we see a
significant relationship:

```{r}

mtr <- multi2di(tr_dps)
nhsims <- tibble(i=1:5000) %>% 
           mutate(bm_trait = map(i, ~ fastBM(mtr))) %>%
           mutate(nh_rlm = map(bm_trait, ~ node_height(., mtr))) %>%
           mutate(nh_lm = map(bm_trait, ~ node_height(., mtr, method='lm')))

get_lm_pval <- function(x) {
  summary(x$fit)$coefficients[2, 4]
}

get_rlm_pval <- function(x) {
  tval <- summary(x$fit)$coefficients[2, 3]
  df <- summary(x$fit)$df[2]
  dt(abs(tval), df)
}

nhsims %>% mutate(pval = map_dbl(nh_rlm, get_rlm_pval)) %>%
  ggplot(aes(x=pval)) + geom_histogram()

```

This is the distribution of p-values under robust regression. This shows the null-model is a bit wrong (overly conservative); see [David Robinson's great article on this](http://varianceexplained.org/statistics/interpreting-pvalue-histogram/). This is a bit concerning, but note that the p-value histogram under the linear model looks better:

```{r}

nhsims %>% mutate(pval = map_dbl(nh_lm, get_lm_pval)) %>%
  ggplot(aes(x=pval)) + geom_histogram(binwidth=0.02)

```

So the odd `rlm()` p-value distribution seems to be an artifact of
the robust regression.

Overall, simulating Brownian motion shows that there isn't a
strong reason to suspect the shape of this phylogeny affects the
node-height test.

What does a random BM node height test?

```{r}
plot_nh(nhsims$nh_rlm[[2]])
```

```{r}

# no drosophila
dt_dps_nd <- dt_dps %>% filter(genus != 'Drosophila')
tr_dps_nd <- keep.tip(tr_dps, dt_dps_nd$species)

# diversity
dvnh_nd <- node_height(dt_dps_nd$diversity,
                       multi2di(tr_dps_nd))
plot_nh(dvnh_nd)
summary(dvnh_nd$fit)

# popsize
psnh_nd <- node_height(dt_dps_nd$log10_popsize, 
                    multi2di(tr_dps_nd))
plot_nh(psnh_nd)

# 
outliers <- dvnh_nd$d  %>% filter(pic > 0.001) %>% pull(tips) 
not_outliers <- dvnh_nd$d  %>% filter(pic < 0.001) %>% pull(tips)

outliers_ranges <- sapply(outliers, function(x) diff(range(dt_dps_nd$log10_range[dt_dps_nd$species %in% x])))

not_outliers_ranges <- sapply(not_outliers, function(x) diff(range(dt_dps_nd$log10_range[dt_dps_nd$species %in% x])))

tibble(range = c(not_outliers_ranges, outliers_ranges), 
       group = c(rep('outlier', length(not_outliers_ranges)), 
                 rep('not outlier', length(outliers_ranges)))) %>%
  ggplot(aes(x=group, y=range)) + 
geom_boxplot()

  geom_histogram(position='dodge')

```

## Genome Size

A quick EDA plot of log genome size and log population size:

```{r}

d %>% filter(species != "Ambystoma tigrinum", 
             is.finite(log10_popsize),
             is.finite(genome_size)) %>%
  ggplot(aes(log10_popsize, genome_size)) + geom_point() + 
    scale_y_log10()

```

A simple PGLS of log genome size and log population size:

```{r}

# TODO

```

## Simple Hitchhiking Effects Model based on Elyashiv et al. (2016)

Based on the model of Coop and Ralph (2012) and Elyashiv et al. (2016).

Elyashiv et al. 2016's selection coefficient distribution, based on
Supplementary Table S1.

```{r}

# Table S1, with "missing substitutions"; introns/intergenic are all NA
# so not included (same as Table S11)
dm_s <- tribble(~ log10_s, ~ feature, ~ alpha,
                     -1.5,   'exon',       NA,
                     -1.5,   'utr',        NA, 
                     -2.5,   'exon',       0.6,
                     -2.5,   'utr',        NA,
                     -3.5,   'exon',       3.5,
                     -3.5,   'utr',        NA,
                     -4.5,   'exon',       NA,
                     -4.5,   'utr',        5.1,
                     -5.5,   'exon',       36.3,
                     -5.5,   'utr',        42.1) %>% 
         mutate(alpha = alpha/100., s=10^log10_s) # convert from percentage

```

Next are the substitution rates, from Supplementary Text Section A.

```{r}

# Substitution rates
# These are taken from Supplementary Text Section A, p. 5
# Original source: Hu et al. 
# exon is nonsyn
dm_subs <- tribble(~ feature,       ~ prop_data, ~ num_subs,  ~ rate,
                         'exon',            0.7,      64205,  0.0092,
                          'utr',           0.45,     153765,   0.025,
                       'intron',           0.43,     456401,   0.032)

dm_subs <- tribble(~ feature,    ~ num_subs,  ~ rate,
                      'exon',         64205,  0.0092,
                       'utr',        153765,   0.025,
                    'intron',        456401,   0.032)

```

Next, to get a sense of the proportion of base pairs in each annotation
feature, I load in the summary created  in `../data/dmel/`. Note that these are
only for the chromosomes in the Elyashiv et al. study.

```{r}
# annotation 
dm_ann <- read_tsv('../data/dmel/dmel_annotation_summary.tsv')
dm_sl <- read_tsv('../data/dmel/dmel_seqlens.tsv')

# total genome length 
G <- sum(dm_sl$length)

# genic total bp, including CDS, introns, UTRs
Ggenic <- dm_ann %>% filter(feature %in% c('cds', 'intron', 'UTR')) %>% 
  pull(total_bp) %>% sum()

```

For my approximations, I use Drosophila melanogaster values of $\alpha$ and
$m/T$. The main thing is $J_{2,2}$ based on (1) exogenous $N$ and $L$ and (2)
the inferred Elyashiv et al. Drosophila distribution of selection coefficients.

I first average over the features to get a marginalized selection coefficient
distribution. I weight by proportions of UTRs to exonic sequences in the
annotation.

```{r}

# dm_seldist_simple <- tribble(~ log10_s, ~ prop,
#                                   -3.5,    0.04,
#                                   -5.5,     0.4)

dm_seldist<- dm_s %>% left_join(dm_ann)  %>%
  group_by(feature) %>%
  # Total alpha by feature. Sanity check: these match Table S6
  mutate(alpha_feature = sum(alpha, na.rm=TRUE)) %>%
  # calc the selection coefficient distribution by feature 
  mutate(g_feature = alpha / alpha_feature)  %>%
  filter(is.finite(g_feature)) %>%
  ungroup() %>%
  group_by(log10_s, s) %>%
  summarize(g = weighted.mean(g_feature, prop_genome, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(g = g/sum(g))

dm_seldist

```

Now, I estimate the $\alpha$, averaging across annotation features.

```{r}

# alpha is the fraction of substitutions (in genic regions) that are beneficial
dm_alphas <- dm_s %>% left_join(dm_ann) %>% 
               group_by(feature, prop_genome) %>% 
               summarize(alpha = sum(alpha, na.rm=TRUE), 
                         total_bp = unique(total_bp)) %>%
               ungroup() %>%
               mutate(alpha_mean = weighted.mean(alpha, prop_genome))

dm_alpha <- unique(dm_alphas$alpha_mean)
stopifnot(length(dm_alpha) == 1)

dm_alphas

```

These values seem entirely reasonable given Table 1 of Sella et al. (2009).

Next I estimate $m$, the number of substitutions occurring in coding regions.
This uses Elyashiv et al.'s Supplementary Text which reports the substitutions
per feature (input above as `dm_subs`). Again, this is just for the automsomal
chromosomes.

```{r}

# total_bps is the number of bps per feature 
dm_ms <- dm_subs %>% left_join(dm_ann) %>%
          # note, we exclude introns since Elyashiv found 
          # no evidence of adaptive substitutions in introns, 
          # but see note on p. 13 2nd to last paragraph.
          filter(feature %in% c('exon', 'utr')) %>%
          mutate(m = rate * total_bp)

dm_m <- sum(dm_ms$m)

dm_ms %>% select(feature, num_subs, rate, total_bp, m, prop_genome)
dm_m

```

Next, we sanity check this. To get an idea of whether this is reasonable, we
need the divergence times between *D. melanogaster* and *D. simulans*.

```{r}
# parameters
dmel_sim_div <- 4.2e6 # Mya, from Obbard

# assuming Ne of D. mel is 1e6
dm_mu <- 2.8e-9
dm_Ne <- 1e6

dm_pi = da_cc %>% filter(species == "Drosophila melanogaster") %>% pull(diversity)
# 0.2 from syn subs per site in Supp. p. 5
dmel_sim_div2 <- (0.2 / 2) / dm_pi * dm_Ne # Mya, from method in Elyashiv

# I use Obbard's -- both are similar
T <- dmel_sim_div * 10 # gens

dm_m / (2*T * sum(dm_ms$total_bp))

```

This substitution rate is a bit lower than the mutation rate, but given these
are genic regions, this is not too surprising. Overall, nothing too out of the
ordinary here.

Finally, we work to get an estimate of $J_{2,2}$. Here I explore two
approaches, one a simplified version of Elyashiv et al. where I use their
inferred selection coefficient distribution, and another with fixed $s$. First,
some functions:

```{r}

# fixation time, based on diffusion theory (Elyashiv et al. eqn 4)
tau <- function(s, N) {
  gamma <- -digamma(1)
  2*(log(4*N*s) + gamma - 1/(4*N*s)) / s
}

# simple J22 with fixed s, without haldane
J22 <- function(s, N, L){
  t <- tau(s, N)
  # from Mathematica 
  (1 - exp(-2*L*t))/(2*t)
}

# J22 for distribution of s
J22s <- Vectorize(function(s, w, N, L) {
  ts <- tau(s, N)
  # Inf out any negative fixation times (when Ns is small)
  if (any(ts < 0)) warning("some fixation times < 0")
  ts[ts < 0] <- Inf
  sum(w * (1 - exp(-2*L*ts))/(2*ts))
}, c('N', 'L'))

```

What is the $\nu_{BP}$? I also extract some other relevant parameters.

```{r}
# implied nu_bp
dm_nu_bp <- dm_alpha * dm_m / (G*2*T)
dm_nu_bp

dm_log10_Nc <- da_ml %>% filter(species == "Drosophila melanogaster") %>% pull(log10_popsize) 
dm_Nc <- 10^dm_log10_Nc

# some other Drosophila params
dm_L <- da_ml %>% filter(species == "Drosophila melanogaster") %>% pull(map_length)
dm_U <- 1.6 # from Elyashiv
dm_rbp <- mean(c(2.14, 2.77, 2.2, 1.97)) * 1e-2 / 1e6 # from Petrov's website

```

What about $J_{2,2}$? Here, I find that my simple model that marginalizes over
the distribution of selection coefficients seems to estimate a $J_{2,2}$ that
is much weaker (two orders of magnitude) than estimates that have been
previously published and compared to the rate of coalescent events
(Supplementary Table S6). Here is the $J_{2,2}$ from the distribution of
selection coefficients:

```{r}
# J22 
# NOTE: taking L = Inf does not matter since ts is so high.
# What drives this is the 1/fixation_time term.
dm_J22_Elyashiv <- J22s(dm_seldist$s,  dm_seldist$g, dm_Ne, Inf)  
dm_J22_Elyashiv

# 2 N nu_bp J22 / rbp
twoSN_Elyashiv <- 2 * dm_Ne * dm_nu_bp / dm_rbp * dm_J22_Elyashiv
twoSN_Elyashiv

2 * dm_nu_bp / dm_rbp * dm_J22_Elyashiv

```

How does this compare to the value published in the Supplementary Table (S6) of
Elyashiv et al?

```{r}

rs <- 0.92 # from Table S6
# implied by the rs = 0.92 value in Supplementary Table S3
dm_J22_Elyashiv2 <- rs / (2 * dm_Ne * dm_nu_bp / dm_rbp)

# implied J22
signif(rs / (2 * dm_Ne * dm_nu_bp / dm_rbp), 3)

```

Let's compare these to the numbers from Sella et al. 2009.

```{r}

# from Sella et al. 2009
sweep_ests <-
  tribble(            ~ study,                  ~ s,          ~ nu,   ~ nu_gamma,
     'Wiehe and Stephan 1992',             NA_real_,      NA_real_,   1.3e-8,
        'Li and Stephan 2006',      (0.2 + 0.5)/200,  (6+9)/2*1e-11,      NA,
            'Andolfatto 2007',                 1e-5,        7.5e-10,    3e-8,
      'Macpherson et al 2007',                 0.01,        3.6e-12,    1e-7,
          'Jensen et al 2008',              0.2/100,          4e-11,    4e-7) %>%
        mutate(twoSN = nu_gamma * 0.075 / dm_rbp) %>%
        mutate(approx_J22 = twoSN / (2 * dm_Ne * nu / dm_rbp))  %>%
        add_row(study="***", s=NA, nu=dm_nu_bp, approx_J22 = dm_J22_Elyashiv2)
sweep_ests

write_tsv(sweep_ests, "../data/sweep_param_ests.tsv")

```

Below we show all of these, mine is `***`:

```{r}

sweep_ests %>%
  mutate(color=ifelse(study == '***', 'red', 'black')) %>%  
  ggplot() + 
  geom_hline(yintercept=0, color='gray32') + 
  geom_point(aes(nu, 0, color=color), size=4) + 
  geom_text_repel(aes(nu, 0, label=study), point.padding=1, box.padding=3, seed=1) +
  scale_x_log10() +   scale_colour_identity()+
  xlab("beneficial substitutions per bp (ν_bp)") + ylab('')

```

Is this reasonable?  This seems roughly in line with previous findings.

How does our $J_{2,2}$ compare to other published estimates of $J_{2,2}$?

```{r}

# dataset from Graham's notes
dshap_raw <- read_tsv('../data/shapiro_et_al_2007.tsv', col_names=FALSE)

# unsure what all the columns are, but this is what matters
dshap <- tibble(rec = dshap_raw$X3, pi = dshap_raw$X4)

hh_nls <- nls(pi ~ neutral.pi*rec/(rec + alpha), dshap, 
              start=list(neutral.pi=2, alpha=1e-9))

plot(pi ~ rec, dshap, pch=19, xlab='recombination (cM/Mb)', 
     ylab=latex2exp::TeX('$\\pi$'))
rec <- seq(0, 2.5, length.out=100)
y <- predict(hh_nls, data.frame(rec=rec))
lines(rec, y, col='red', lwd=3)

pi0 <- coef(hh_nls)['neutral.pi']
alpha <- coef(hh_nls)['alpha']
lines(rec, pi0/(1 + alpha/rec), col='red', lwd=3)
lines(rec, pi0/(1 + twoSN_Elyashiv/rec), col='blue', lwd=3)

lines(rec, pi0/(1 + rs/rec), col='green', lwd=3, lty=2)

legend(0, 0.06, c('non-linear least squares', 
                  latex2exp::TeX("sel coef $J_{2,2}$"),
                  latex2exp::TeX('Elyashiv et al. $r_S$')),
        fill=c('red', 'blue', 'green'),
        bty='n', border=0, cex=1, ncol=1)

# the implied 2NS / rbp is below -- this is the same 
# as in Coop and Ralph (2012) Table 1
alpha / (100 * 1e6) # units are changed to cM/Mb

# the J22 is
alpha / (100 * 1e6 * 2*dm_Ne * dm_nu_bp)
dm_J22_Elyashiv

```

Overall the selection coefficient distribution approach underestimates
$J_{2,2}$. Looking at Elyashiv et al., this is because the homogeneous sweep
model spaces substitutions too evenly. I use the empirical $J_{2,2}$ which is
stronger and more realistic. Let's look at some comparisons:

```{r}

sweep_ests %>%
   # implied by cs parameter estimate in supp
   add_row(study="Elyashiv et al. 2012", s=NA, nu=dm_nu_bp, twoSN=0.92, 
           approx_J22=0.92 / (2*dm_Ne * dm_nu_bp / dm_rbp)) %>%
   # roughly the homogeneous sweep model
   mutate(one_over_fix= 0.5 / tau(s, dm_Ne), off=one_over_fix/approx_J22)

```

How do the diffusion and deterministic fixation times compare? Could this be an
issue?

```{r}

# how does our diffusion fixation time compare to the deterministic?
crossing(log10_s = -(1:5), log10_N = 3:10) %>%
  mutate(s=10^log10_s, N = 10^log10_N) %>%
  mutate(`simple` = 4/s * log(2 * N), diffusion=tau(s, N)) %>% 
  filter(diffusion > 0) %>%
  ggplot(aes(simple, diffusion, color=as.factor(log10_s), shape=as.factor(log10_N))) + 
  geom_point() + scale_x_log10() + scale_y_log10() + geom_abline(slope=1, intercept=0)

```
No, those look roughly the same.

Now, let's crunch some more parameter estimates, this time
$\gamma_\text{Dmel}$, as well as define some functions for the reductions due
to RHH and BGS.

```{r}

# nu_bp without the 1/G (since we use L)
dm_gamma <- dm_alpha * dm_m / (2*T)
signif(dm_gamma, 2)

# Hudson and Kaplan (1995) equation 8
HK95 <- function(L, U) exp(-U / L)

RHH_S2N <- function(L, N, gamma, J22) {
  S <- gamma / L * J22
  2*N*S
}

RHH_BGS <- function(theta, L, N, B, gamma, J22) {
  S <- gamma / L * J22
  # generally theta is small, so we approx as theta / (1 + theta) ~ theta
  # (Coop and Ralph 2012 do this, e.g. eqns 18, 19). This allows 
  # for me to treat this as a proportion reduction.
  # theta / (theta + 1/B + 2*N*S)
  theta / (1/B + 2*N*S)
}

```

Quickly, what are the relative strengths?

```{r}

x <- 10^da_ml$log10_diversity / 4e-9
bgs <- 1/(HK95(da_ml$map_length, 1.6))
rhh <- RHH_S2N(da_ml$map_length, x, dm_gamma, dm_J22_Elyashiv)
# plot(x, bgs, col='red', pch=19)
# points(x, rhh, col='green', pch=19)
# plot(x, rhh, col='green', pch=19)

plot(log10(x),bgs/(rhh + bgs), xlab='log10 pop size', pch=19)

```

What about the S&C 2016 stuff?

```{r}

fixT <- function(s, Ne, U) (exp(2*s*Ne) - 1) / (2*U*s*Ne)

SC16_R <- function(Ne, U, L, s) {
  T <- fixT(s, Ne, U)
  exp((-2*U / (exp(2*s) + L))*(1 - 1/(U*T))^3)
}

SC16_R(Ne=dm_Ne, U=dm_U, L=dm_L, s=10^{-1.5}) # s from Elyashiv
HK95(U=dm_U, dm_L)

```

Not much of a difference...

## Diversity Changes Due to Linked Selection

First we try to extrapolate what the effect of linked selection would be in the
Corbett-Detig dataset.

```{r}

# load corbett-detig data, which is data with impact_of_sel
dca <- d[!is.na(d$impact_of_sel), ]
# check against real dataset
dc <- read_tsv('../data/corbett_detig_2015_updated.tsv')  %>%
        mutate(data_source = 'Corbett-Detig et al.',
               range_cd = range) %>%
        rename(diversity = obs_pi)
# should be same size if joins were right
stopifnot(sum(dc$kingdom == 'Animalia')  == nrow(dca))

dca_noself <- dca[dca$selfing != 1, ]

```

First, do we see the same diversity / population size relationship as we do
with the Leffler et al diversity data? Yes:

```{r}

ggplot(dca) + 
  geom_smooth(aes(log10_popsize, log10(diversity)), method='lm', se=FALSE) + 
  geom_point(aes(log10_popsize, log10(diversity))) + 
  geom_text_repel(aes(log10_popsize, log10(diversity), label=species))

```

What does impact of selection's relation with our population size look like? We
confirm a finding of Corbett-Detig et al. with our own N estimate, that impact
of selection increases with N:

```{r}

ggplot(dca_noself) + geom_point(aes(log10_popsize, impact_of_sel)) + 
  geom_smooth(aes(log10_popsize, impact_of_sel), method='lm', se=FALSE) + 
  geom_text_repel(aes(log10_popsize, impact_of_sel, label=species))  + 
  xlab("log10 population size") + ylab("impact_of_sel")

```

What's this linear relationship?

```{r}

save(dca, dca_noself, file="../data/corbett_detig_data.Rdata")

plot(dca$log10_popsize, log10(dca$diversity), pch=19)
cd_div_fit <- lm(log10(diversity) ~ log10_popsize, dca_noself)
abline(cd_div_fit)

plot(dca_noself$log10_popsize, dca_noself$impact_of_sel, 
     pch=19)
ios_fit <- lm(impact_of_sel ~ log10_popsize, dca_noself)
abline(ios_fit)
summary(ios_fit)

```

Fit a simple linear model of the impact of selection / popsize relationship.

```{r}
a <- coef(ios_fit)[1]
b <- coef(ios_fit)[2]

```

This is on a subset of the data with population size, diversity,
and map lengths -- about 40 taxa.

```{r}

# here, we use the complete dataset, which we copy here
dml <- da_cc

dml$ios_ls <- predict(ios_fit, 
                         newdata=tibble(log10_popsize=dml$log10_popsize))
dml$Ne_N_CD15 <- 1-dml$ios_ls

# this uses Ne implied by diversity and μ = 1e-9
dml <- dml %>% 
     # estimate from Elyashiv et al.
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     # mutate(J22 = J22s(10^dm_seldist$log10_s,  dm_seldist$g, dm_Ne, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%

     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=diversity/(4e-9), 
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22),
            Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                              B=Ne_N_HK95, gamma=dm_gamma, 
                                              J22=J22 * 10)) %>%
     mutate(Ne_N_HK95_CD15 = Ne_N_HK95 * Ne_N_CD15) %>% 
     mutate(log10_CD15_diversity = log10(diversity / Ne_N_CD15),
            log10_HK95_diversity = log10(diversity / Ne_N_HK95),
            log10_RHH_BGS_diversity = log10(diversity / Ne_N_RHH_BGS),
            log10_RHH_BGS_strongsel_diversity = 
              log10(diversity / Ne_N_RHH_BGS_strongsel),
            log10_diversity = log10(diversity))

# this is the one used in figures -- assumes Ne = Nc
# this got a bit unwieldy as I tried different things... oh well.
dml_full <- da_ml %>%
     # estimate from Elyashiv et al.
     # mutate(J22 = 100*J22s(dm_seldist$s,  dm_seldist$g, 10^log10_popsize, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=Ne_N_HK95, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%

     mutate(Ne_N_RHH_BGS_strongsel_strongBGS = RHH_BGS(1, L=map_length, 
                                                       N=10^log10_popsize,
                                             B=Ne_N_HK95_strongBGS, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%
     # 10-folder greater Nc
     mutate(Ne_N_RHH_BGS2 = RHH_BGS(1, L=map_length, N=10*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%

     # 100-folder greater Nc
     mutate(Ne_N_RHH_BGS3 = RHH_BGS(1, L=map_length, N=100*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=dm_J22_Elyashiv))

save(dml, dml_full, file='../data/linked_sel_div_changes.Rdata')

```

```

What do the reductions using $N_e = 10^6$ (not $N_c$!) for *Drosophila
melanogaster* look like under these parameters for *Drosophila melanogaster*,
under RHH only (no BGS)?

```{r}
# rough approximation, with Ne = 1e6
1-RHH_BGS(1, dm_L, dm_Ne, B=1, gamma=dm_gamma, J22=dm_J22_Elyashiv2)
```

These strong selection reduction, roughly cutting diversity by 66%, is roughly
consistent with Wiehe and Stephan and Elyashiv et al.'s classic sweep only
model (see Figure 6B, CS column).

Using census size, what are the percent reductions in *D. melanogaster*?

```{r}

dml_full %>% filter(species == "Drosophila melanogaster") %>% 
  select(Ne_N_HK95, Ne_N_RHH_BGS)

```

This is a monumental decrease in diversity... it implies we'd be severely
underestimating $\pi_0$.

Ok, now what about the full RHH + BGS model, with *D. melanogaster* $N_e =
10^6$?

```{r}

1-RHH_BGS(1, dm_L, dm_Ne, B=HK95(U=dm_U, L=dm_L), gamma=dm_gamma, J22=dm_J22_Elyashiv2)

```

84% is surprisingly close to Elyashiv et al.'s Figure 6 value of 77% under
BGS+CS, give this simple model.

A draft plot of the effects, red is RHH, blue is BGS. Ultimately I use a
different figure, but this one is still interesting.

```{r}

plot(dml$log10_popsize, dml$log10_diversity, pch=19, 
                  xlim=c(5, 18),
                  ylim=c(-3, 1))

with(dml, segments(log10_popsize, log10_diversity, 
                      log10_popsize, log10_RHH_BGS_diversity, 
                      lty=3, col='red'))
with(dml, segments(log10_popsize, log10_diversity, 
                      log10_popsize, log10_HK95_diversity, 
                      lty=1, col='blue'))

abline(a = log10(4) + -8, b=1)
abline(a = log10(4) + -10, b=1)

```

## IUCN Red List Analysis

Here's a basic analysis of redlist categories and diversity.

```{r}

drl <- da_dps
drl$redlist_cat[is.na(drl$redlist_cat)] <- "NA/DD"
# ordered in terms of endangeredness
iucn_levels <- c("NA/DD", "LC", "NT", "VU", "EN", "CR")
drl$redlist_cat[drl$redlist_cat == 'DD'] <- "NA/DD"
drl$redlist_cat <- factor(drl$redlist_cat, levels=iucn_levels)

drl %>% group_by(redlist_cat) %>% summarize(ntaxa=n())

# use all data, but notice some categories have few species
drl %>% group_by(redlist_cat) %>% summarize(ntaxa=n())

f0_div <- lm(log10_diversity ~ log10_popsize, drl)
f1_div <- lm(log10_diversity ~ redlist_cat + log10_popsize, drl)
f2_div <- lm(log10_diversity ~ redlist_cat, drl)
AIC(f0_div, f1_div, f2_div)
summary(f1_div)
summary(f2_div)

f0_ps <- lm(log10_popsize ~ redlist_cat, drl)
summary(f0_ps)

ggplot(drl, aes(redlist_cat, log10_diversity)) + geom_boxplot()

```

Table for paper:

```{r}

cis <- signif(cbind(coef(f1_div), confint(f1_div)), 2)
xtable(cis)

```

## Session Info

```{r}
sessionInfo()
```

### 2021-02-08 notebooks/stapley_et_al_2017_pcm.Rmd
New repository with squashed history

# Main Analysis of Stapley et al. (2017) Data

```{r, include=FALSE}
options(warnPartialMatchArgs = FALSE,
        warnPartialMatchDollar = FALSE, 
        warnPartialMatchAttr = FALSE)

```

```{r, message=FALSE}

library(ape)
library(phytools)
library(rotl)
library(scales)
library(tidyverse)
library(rstan)
library(loo)
library(coda)
source('../R/utilities.r')

```

## Data Loading

We load the tree from Open Tree of Life downloaded via `rotl` for the taxa in
the Stapley et al. (2017) data. This is for animals, and we've removed all
missing `log10_size` and `map_length` data.

```{r}
load('../data/stapley_et_al_2017_otl.Rdata')
stopifnot(exists('dsatr') && exists('tr'))
```

We lost a few species in matching taxa to the R Open Tree of Life tree; we
quantify that here:

```{r}
total <- suppressMessages(read_tsv('../data/stapley_et_al_2017.tsv')) %>% 
  filter(kingdom %in% 'Animalia') %>%
  nrow()

msg <- "Size of animalia in Stapley et al (2017), total: %d
                            after matching tree: %d"
message(sprintf(msg, total, nrow(dsatr)))
```

## Phylogenetic Tree

Unfortunately, the Open Tree of Life data lacks branch lengths. Like the
original Stapley et al. (2017) analysis, we use unit branch lengths.

```{r, fig.width=6, fig.height=8}
# now, we force an ultrametric tree, just to get the number 
# of edges we need to set to one
utr <- force.ultrametric(tr)

# set all edges to one, as Stapley et al did.
tr$edge.length  <- rep(1, length(utr$edge.length))

# resolve polytomies
trb <- multi2di(tr)
```

What does the tree look like?

```{r}
stopifnot(dsatr$species == tr$tip.label)
nphyla <- length(unique(dsatr$phylum))
phyla_cols <- hue_pal()(nphyla)

# nfamilies <- length(unique(dsatr$family))
# families_cols <- hue_pal()(nfamilies)

# plot(force.ultrametric(trb), cex=0.6, tip.color=phyla_cols[as.factor(dsatr$family)])
plot(force.ultrametric(trb), cex=0.6, tip.color=phyla_cols[as.factor(dsatr$phylum)])

```

## Some Map Length, Body Size, and Chromosome Number EDA

```{r}

dsatr  %>% ggplot(aes(hcn, map_length)) + geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('map length (Morgans)')

dsatr %>% ggplot(aes(hcn, log10_size)) + geom_point() + 
  geom_smooth() + xlab('body size') + 
  ylab('haploid chromosome number')

dsatr  %>% ggplot(aes(log10_size, hcn)) + geom_point() + 
  geom_smooth(method='lm') + ylab('haploid chromosome number') + 
  xlab('log10(size)')

dsatr %>% ggplot(aes(hcn, map_length/hcn)) + 
  geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('Morgans per chromosome (map length / haploid chromosome number)')

```

What do number of Morgans / chromosome across the data look like?

```{r}

dsatr %>% mutate(morgans_per_chrom = map_length/hcn) %>% 
  pull(morgans_per_chrom) %>% hist(50, main='Morgans / chromosome')

```

## The Relationship Between Map Length and Body Size

The relationship we'd like to assess is whether map length is positive
correlated with body size; here body size is a proxy for population density,
where they have a negative relationship (e.g. Damuth (1981)).

For quantitative genetic models of linked selection to decouple diversity from
population size, we need a relationship like: $\log(N) = \alpha / L$ where $L$
is map length in Morgans, and $\alpha$ is an unknown positive constant. Since
$\log(N) \propto -\log(S)$ where $S$ is body size, we have: $\log(S) = -\alpha
/ L$, or $L = -\alpha / \log(S)$.

This is the form of the relationship in the Stapley et al (2017), with my body
size data:

```{r}

with(dsatr, plot(map_length, log10_size, pch=19))

```

This obviously does **not** correct for phylogenetic dependencies in the data. 
Currently, I do not see a way to do all of the following things.

1. Keep the units untransformed, as the theory requires the total map length be
in Morgans. For the sake of evalulating the theory, it does not make sense to
calculate the reduction of diversity on a phylogentically corrected value; the
map length in Morgans of a taxa is known.

2. Ideally, a proper PCM would model the evolution of the traits, map length
and body size. Ideally, all the Metazoan taxa included here would have
well-calibrated phylogeny with branch lengths, and we could actually compare
different trait evolution models (early burst, OU, etc) and control for
phylogenetic non-independence more carefully. However, the current phylogeny of
these taxa from the Open Tree of Life has unit branch lengths. Estimating a
phylgoeny for these Metazoan taxa would be outside the scope of this paper.

3. Model the non-linear nature of the data. Ideally, we would have a way to
test whether there's not only a relationship, but a specific non-linear
relationship between the traits measured. But given the lack of branch lengths
(and as far as I know, methods for this) this is not feasible.

Thus, we have two statistical problems: (1) does the body size / map length
trend exist when we remove the effects of pseudoreplication? (2) does the
general trend fit the non-linear relationship needed to decouple diversity?

My plan is two approach (1) assess the linear relationship with PICs, and (2)
find the non-linear least squared fit on both family, and phylum-level averaged
data. I do this first with non-linear least squares, (3) use Bayesian models
(mostly to address issues with the confidence intervals that occurred with the
non-linear least squares fits).

First, we calculate the family and phylum-level averages of the data, used in
both least-squares and Bayesian model fits.

```{r}

# Side note about what I do below: I also take the log of the mean of groups.
# This makes little difference in practice, but *is* different in theory
# (Jensen's inequality). log10_size2 below is doing the
# *incorrect* thing of taking the mean of the values, then the log. This is
# incorrect for the following reason:
#
# x <- rlnorm(100000, 5, 1)
# hist(log(x), 100, col = "grey", border = "grey")
# abline(v=log(mean(x)), col='red')
# abline(v=mean(log(x)), col='black')
#
# The real mean is 5.

family_mean <- dsatr %>% group_by(family) %>%
                 summarize(log10_size=mean(log10_size, na.rm=TRUE), 
                           log10_size2=log10(mean(size, na.rm=TRUE)),
                           map_length=mean(map_length, na.rm=TRUE),
                           hcn=mean(hcn, na.rm=TRUE),
                           n=n())

phyla_mean <- dsatr %>% group_by(phylum) %>%
                 summarize(log10_size=mean(log10_size, na.rm=TRUE), 
                           log10_size2=log10(mean(size, na.rm=TRUE)),
                           map_length=mean(map_length, na.rm=TRUE), 
                           hcn=mean(hcn, na.rm=TRUE),
                           n=n())

```

### Non-linear Least Squares

First, we use the `nls()` function to fit a few different models, and then use
AIC for model comparisons.

```{r}

## WARNING: I label my models differently here than I do in the Bayesian 
## modeling section, so be cautious in assuming what's what.

# family average
mf0 <- nls(log10_size ~ a + b*map_length, data = family_mean, 
         start = list(a=1, b=-2))

mf1 <- nls(log10_size ~ a + b*map_length^power, data = family_mean, 
         start = list(a=1, b=-2, power = -2))

mf2 <- nls(log10_size ~  b*map_length^power, data = family_mean, 
         start = list(b=-2, power = -2))

# phylum average
mp0 <- nls(log10_size ~ a + b*map_length, data = phyla_mean,
           start = list(a=1, b=-2))

mp1 <- nls(log10_size ~ a + b*map_length^power, data = phyla_mean, 
           start = list(a=1, b=1, power = 1),
           control=list(maxiter = 500))

mp2 <- nls(log10_size ~ b*map_length^power, data = phyla_mean, 
           start = list(b=1, power = 1),
           control=list(maxiter = 500))

summary(mp2)
summary(mf2)

BIC(mp0, mp1, mp2)
BIC(mf0, mf1, mf2)

```

With both phylum and family averages, the best fitting model according to BIC
is the non-linear model without an intercept term.

```{r}
# TODO
mf0hcn <- nls(log10_size ~ a + b*hcn, data = family_mean, 
         start = list(a=1, b=-2))

mf1hcn <- nls(log10_size ~ a + b*hcn^power, data = family_mean, 
         start = list(a=1, b=1, power = 1))

mf2hcn <- nls(log10_size ~  b*hcn^power, data = family_mean, 
         start = list(b=-2, power = -2))

```

Let's bootstrap these model fits.

```{r}

best_model <- function(x, start=NULL) {
  if (is.null(start)) {
    start <- list(b=-2, power=-1)
  }
  nls(log10_size ~  b*map_length^power, data = x,
         start=start, trace = FALSE)
}

bootstrap_sample <- function(x) {
  n <- nrow(x)
  x[sample(n, n, replace=TRUE), ]
}

best_model_safe <- possibly(best_model, NULL)

# x sequences for prediction
sf <- seq(0, 53, length.out=100)
sp <- seq(0, 53, length.out=100)

bs_p <- tibble(rep=1:10000) %>%
  # phylum averages
  mutate(fit = map(rep, ~ best_model_safe(bootstrap_sample(phyla_mean)))) %>%
  mutate(keep = !map_lgl(fit, is.null)) %>% filter(keep)

predict_df <- function(fit, x) {
  tibble(x=x, y=predict(fit, newdata=tibble(map_length=x)))
}

bs_p_ci <- bs_p[sample(nrow(bs_p), 5000), ] %>% 
            mutate(pred = map(fit, predict_df, x=sp)) %>% 
            unnest(pred) %>%
            group_by(x) %>% 
            summarize(lower = quantile(y, probs=0.05/2),
                      upper = quantile(y, probs=1-0.05/2))

bs_f <- tibble(rep=1:10000) %>%
  # family averages
  mutate(fit = map(rep, ~ best_model_safe(bootstrap_sample(family_mean)))) %>%
  mutate(keep = !map_lgl(fit, is.null)) %>% filter(keep)

bs_f_ci <- bs_f[sample(nrow(bs_f), 5000), ] %>% 
            mutate(pred = map(fit, predict_df, x=sf)) %>% 
            unnest(pred) %>%
            group_by(x) %>% 
            summarize(lower = quantile(y, probs=0.05/2),
                      upper = quantile(y, probs=1-0.05/2))

# save the averaged data for plots
save(family_mean, phyla_mean, mp2, mf2, 
     sf, sp,
     bs_f_ci, bs_p_ci, file='../data/body_size_map_length.Rdata')

```

Draft plot (see the `notebooks/figures/` directory for actual figure plots).
Now, I prefer the plot below from the HPD intervals.

```{r}

with(family_mean, plot(map_length, log10_size, pch=19, axes=FALSE, cex=0.7,
                       ylab='', xlab='', xlim=c(0, 50), ylim=c(-4.3, 1.1), col='gray48'))

axis(1, seq(0, 50, 10))
axis(2, seq(-4, 1), seq(-4, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-4, 1))))

with(bs_p_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
                    col=scales::alpha('cornflowerblue', 0.2), lty=0))

with(bs_f_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
                    col=scales::alpha('gray42', 0.2), lty=0))

with(phyla_mean, points(map_length, log10_size, pch=19, col='blue', cex=0.7))

#bs_f %>% mutate(pred = map(fit, predict_df, x=sf)) %>%
#  mutate(x=map(pred, ~ lines(.$x, .$y)))
#
#bs_p %>% mutate(pred = map(fit, predict_df, x=sf)) %>%
#  mutate(x=map(pred, ~ lines(.$x, .$y)))

lines(sf, predict(mf2, newdata=tibble(map_length=sf)), lty='dashed', col='gray32')
lines(sp, predict(mp2, newdata=tibble(map_length=sp)), lty='dashed', col='cornflowerblue')

pivot_ci <- function(x, theta, alpha=0.05) 2*theta - rev(quantile(x, probs=c(alpha/2, 1 - alpha/2)))

bs_f %>% mutate(power = map_dbl(fit, ~ coef(.)[2])) %>% 
  pull(power) %>% pivot_ci(coef(mf2)[2])

```

### Bayesian Models -- Family-level Averages

First model is just a simple regression:

$$ log10(\text{size}) ~ \text{Normal}(\alpha + \beta L, \sigma) $$

```{r}

# Stan models use this range for mu_rep and y_rep, used to calculate
# the HPDI and posterior predictive distribution
x_new <- seq(0.1, 53, length.out=100)

bmf0 <- stan(file='../stan/maplength_log10size_linear_family.stan',
             data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new = x_new),
             seed=50127839,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Let's assess model convergence:

```{r}

bmf0_pars <- c('alpha', 'beta', 'sigma', 'lp__')
print(bmf0, pars=bmf0_pars)
pairs(bmf0, pars=bmf0_pars)

```

The Rhats show great mixing among chains, we have a high number of effective
samples, and there are no divergences.

Now we look at the next model:

$$ log10(\text{size}) ~ \text{Normal}(\beta L^\gamma, \sigma) $$

(log10 here is a transformation; this is not log normal).

```{r}

bmf1 <- stan(file='../stan/maplength_log10size_power_no_intercept_family.stan',
            data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new=x_new),
             seed=59912951,
             iter=10000, cores=4, control=list(adapt_delta=0.95))

bmf1_pars <- c('beta', 'power', 'sigma', 'lp__')
print(bmf1, pars=bmf1_pars)

```

Again, good mixing; let's look at the pairs plot:

```{r}
pairs(bmf1, pars=bmf1_pars)
```

Next, the most complicated model:

$$ log10(\test{size}) ~ \text{Normal}(\alpha + \beta L^\gamma, \sigma) $$

```{r}

bmf2 <- stan(file='../stan/maplength_log10size_power_family.stan',
            data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new = x_new),
             seed=125192,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Now, we check the convergence.

```{r}

bmf2_pars <- c('alpha', 'beta', 'power', 'sigma', 'lp__')
print(bmf2, pars=bmf2_pars)
pairs(bmf2, pars=bmf2_pars)

```

Finally, we estimate the coefficients for a model where we decouple N and L
directly, by setting `power = -1`.

```{r}

bmf3 <- stan(file='../stan/maplength_log10size_inverse_family.stan',
            data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new = x_new),
             seed=7814,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

bmf3_pars <- c('alpha', 'beta', 'sigma', 'lp__')
print(bmf3, pars=bmf3_pars)

```

Now, we need to compare these models.

```{r}

loow <- function(fit) {
  # a loo() wrapper
  log_lik <- extract_log_lik(fit, merge_chains = FALSE)
  r_eff <- relative_eff(exp(log_lik), cores = 2) 
  loo(log_lik, r_eff = r_eff, cores = 2)
}

loow(bmf0)
loow(bmf1)
loow(bmf2)

loo_compare(loow(bmf0), loow(bmf1), loow(bmf2), loow(bmf3))

```

This shows model 2 (mind they index by 1, not zero) does best, but it's not
significantly different from model 1 (see this answer by the `loo` package's
creator arguing at least a few SDs are necessary to say one model is better
than another).

Looking at WAIC:

```{r}

loo_compare(waic(extract_log_lik(bmf0)),
            waic(extract_log_lik(bmf1)),
            waic(extract_log_lik(bmf2)),
            waic(extract_log_lik(bmf3)))

```

So, again model 2 looks to be the best, though the standard error of the
difference is not very large. Still, arguably we should choose the simpler
model that does best, model 1. This indicates we are not over fitting the power
term compared the linear-terms-only model, which is what we want to see.

### Bayesian Models -- Phylum Averages

Next, we fit these same models on the phylum-level averaged data. Note that
it's worth looking at the Stan files in detail. I had to tweak priors quite a
bit, as an early bad prior choice was leading to divergent transitions, and in
some cases, high `Rhat` values. I still use weakly informative priors, option
for pretty diffuse Student's t distribution for the $\alpha$ and $\beta$
parameters, while using a standard Normal for the power, and a tighter prior on
the standard deviation. This last point deserves some justification. These
phyla-level averages are averages of *many* taxa, and thus, the our prior
expectation of what the standard deviation should be decreases. Early on in
testing, I was using the same prior as the family-level averages. The fits were
very different from the non-linear least squares approach, which was
suspicious. Essentially, we were underfitting the data; our expectation should
be averages of more points have less variance. Changing the priors accordingly
lead to fits more consistent (but still different) from those found with
non-linear least squares. See the Stan model files for more details.

```{r}

bmp0 <- stan(file='../stan/maplength_log10size_linear_phylum.stan',
             data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new=x_new),
             seed = 84914,
             iter=10000, cores=4, control=list(adapt_delta=0.99))

bmp1 <- stan(file='../stan/maplength_log10size_power_no_intercept_phylum.stan',
             data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new = x_new),
             # increasing adapt_delta decreases the number of divergent transitions
             seed = 41123,
             iter=10000, cores=4, control=list(adapt_delta=0.999))

bmp2 <- stan(file='../stan/maplength_log10size_power_phylum.stan',
                  data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new=x_new),
             seed = 874891,
             # we increase the number of iterations and warmup to reduce divergent 
             # transitions
             iter=12000, cores=4, 
             warmup = 5000,
             # increasing adapt_delta decreases the number of divergent transitions
             control=list(adapt_delta=0.999))

bmp3 <- stan(file='../stan/maplength_log10size_inverse_phylum.stan',
            data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new = x_new),
             seed=519459,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Next, we let's look at the model output, checking `Rhat` and the number of
effective samples:

```{r}

print(bmp0, pars=bmf0_pars)
print(bmp1, pars=bmf1_pars)
print(bmp2, pars=bmf2_pars)
print(bmp3, pars=bmf3_pars)

```

Next, we look at the pairs plots:

```{r}

pairs(bmp0, pars=bmf0_pars)
pairs(bmp1, pars=bmf1_pars)
pairs(bmp2, pars=bmf2_pars)
pairs(bmp3, pars=bmf2_pars)

```

Again, all of these look reasonable. Notice there are no (or very, very few
divergent transitions). All `Rhat` values are < 1.

Next, let's look at the estimated expected log posteriors:

```{r}
print(loow(bmp0))
print(loow(bmp1))
print(loow(bmp2))
print(loow(bmp3))
```

And we compare them, using `loo`'s compare function and WAIC:

```{r}

loo_compare(loow(bmp0), loow(bmp1), loow(bmp2), loow(bmp3))

loo_compare(waic(extract_log_lik(bmp0)),
            waic(extract_log_lik(bmp1)),
            waic(extract_log_lik(bmp2)),
            waic(extract_log_lik(bmp3)))

```

These indicate that in terms of expected log predictive density, models 1
(again, mind the fact my model numbers are zero-indexed and theirs are
1-indexed), is much better than the linear model.

Next, let's look at the HPD intervals:

```{r}

maplength_bodysize_plot(phyla_mean)
ci_polygon(x_new, bmp1, color='gray', type='hpdi', par='mu_rep')
ci_polygon(x_new, bmp2, color='orange', type='hpdi', par='mu_rep')
ci_polygon(x_new, bmp3, color='red', type='hpdi', par='mu_rep')

```

For models 1 and 2, what do the posterior predictive distributions look like?

```{r}

maplength_bodysize_plot(family_mean)
ci_polygon(x_new, bmf2, color='gray', type='quantile', par='y_rep')
ci_polygon(x_new, bmf1, color='green', type='quantile', par='y_rep')

```

So, we output these models

```{r}

save(family_mean, phyla_mean, 
     bmf1, bmp1,
     bmf2, bmp2,
     bmf3, bmp3,
     x_new,
     file='../data/body_size_map_length_bayesian.Rdata')

```

## Phylogenetic Independent Contrast

Next, I want to establish that the relationship between map length and body
size is significant with phylogenetic corrections. Again, these results would
be greatly improved by having non-unit branch lengths. Given that unit branch
lengths probably over-corrects for phylogenetic signal over these long
evolutionary timeframes, this is a conservative approach.

```{r}
# TODO
plot(trb)

plotTree(force.ultrametric(tr))
# remove polytomies
utrb <- multi2di(tr)

log_ml <- log10(dsatr$map_length)
log_size <- dsatr$log10_size

log_ml_pic = as_tibble(pic(log_ml, utrb, var.contrasts=TRUE))
log_size_pic = as_tibble(pic(log_size, utrb, var.contrasts=TRUE))

plot(sqrt(log_ml_pic$variance), abs(log_ml_pic$contrasts))
plot(sqrt(log_size_pic$variance), abs(log_size_pic$contrasts))

m_pic <- lm(log_size_pic$contrasts ~ log_ml_pic$contrasts)
summary(m_pic)

plot(log_ml_pic$contrasts, log_size_pic$contrasts)
abline(m_pic)

plot(a$contrasts, b$contrasts)
abline(lm(a$contrasts ~ b$contrasts))
m <- lm(a$contrasts ~ b$contrasts)

dsic <- tibble(maplength_contrast=a$contrasts, size_contrast=b$contrasts)

fit_contrast_lm <- function(d) {
  n <- nrow(d)
  lm(size_contrast ~ maplength_contrast, data=d[sample(n, n, replace=TRUE), ])
}

bs <- tibble(rep = 1:5000) %>% 
  mutate(fit = map(rep, ~ fit_contrast_lm(dsic))) %>% 
  mutate(pval = map_dbl(fit, ~ coef(.)[2]))

alpha <- 0.05
quantile(bs$pval, c(alpha/2, 1-alpha/2))

bs %>% ggplot(aes(x=pval)) + geom_histogram()

par(opar)
```

## Diversity and Map Length

First, we merge in all the animal data we have.

```{r}

# load in the Corbett-Detig et al (2015) data
dc <- read_tsv('../data/corbett_detig_2015_updated.tsv') %>% 
         filter(kingdom == 'animal') %>%
         select(species, size, diversity=obs_pi, map_length)

lds <- read_tsv('../data/leffler_et_al_2012_updated.tsv') %>%
         filter(kingdom == 'Animalia') %>%
         select(species, size, diversity)

# The Stapley data has size (that I've added) and map length, but not diversity.
# The Corbett-Detig data has all three. 
# The Leffer data has size (that I've added) and diversity, but not map length.
dsatr2 <- dsatr %>% select(species, size, map_length)

dt <- lds %>% left_join(dsatr2)  %>% 
        filter(!is.na(size), !is.na(diversity), !is.na(map_length)) %>%
        # we join in the Corbett-Detig data
        full_join(dc, by=c('species'))

```

This last join didn't merge like columns. Which is good, because they allow us
to do some quality control (in the good sense, not the Iowa Caucus sense).

```{r}

ggplot(dt, aes(diversity.x, diversity.y)) + geom_point() +
  geom_abline(yintercept=0, slope=1, color='gray')

ggplot(dt, aes(size.x, size.y)) + geom_point() + ylim(0, 0.8) +
  geom_abline(yintercept=0, slope=1, color='gray')

ggplot(dt, aes(map_length.x, map_length.y/100)) + geom_point() + 
  geom_abline(yintercept=0, slope=1, color='gray')

```

The values all look good. I'll average them for the final dataframe:

```{r}

merge_average <- function(x, y) rowMeans(cbind(x, y), na.rm=TRUE)

# final dataframe
dtf <- dt %>% mutate(size = map2_dbl(size.x, size.y, merge_average),
              diversity = map2_dbl(diversity.x, diversity.y, merge_average),
              # the map_length.y / 100 below is converting cM in M
              map_length = map2_dbl(map_length.x, map_length.y/100, merge_average)) %>%
  select(-size.x, -size.y, -diversity.x, -diversity.y, -map_length.x, -map_length.y)

```

Now, we use the SC98 and HK94 reductions in $N_e$ to calculate the shift. This
data creates Figure 2. For SC98, we use $\alpha = 1$, where $\alpha = C^2 /
(1-Z)$. This value is greater than we'd expect under Drosophila. For HK94, we
use the Elyashiv et al data.

$$
\pi = 4 N \mu \exp\{- U / L\}
$$

$$
\pi = \pi_0 \exp\{- U / L\}
$$

Elyashiv et al report the average reduction, which is the ratio of $\bar{\pi} /
\pi_0$. They report under the BG+CS model an average reduction of 77% to 89%.

$$
\pi = \pi_0 \exp\{- U / L\}
$$

```{r}

SC98 <- function(L, alpha) exp(-alpha /L)
HK94 <- function(L, U) exp(-U / L)

dtft <- dtf %>%  
  filter(species != 'Ciona savignyi') %>%
  mutate(Ne_N_SC98 = SC98(map_length, alpha=1))  %>%
  mutate(Ne_N_HK94 = HK94(map_length, U=1.6)) # estimate from Elyashiv et al.

dtft %>%
  filter(species != 'Ciona savignyi') %>%
  mutate(decoupled_diversity = size) %>%
  ggplot(aes(-log10(size), (diversity))) + geom_point()  + 
   # geom_smooth(aes(-log10(size), decoupled_diversity), color='red', method='lm') + 
  geom_segment(aes(-log10(size), (diversity), 
                   xend=-log10(size), yend=(diversity / Ne_N_HK94))) + 
  geom_segment(aes(-log10(size), (diversity), 
                   xend=-log10(size), yend=(diversity / Ne_N_SC98)), color='red') +
  scale_y_log10()

save(dtft, file='../data/diversity_size_theory.Rdata')
```

## Fecundity and body size

Next, I look at the relationship between fecundity and body size. If we imagine
a Poisson number of offspring, the mean level of fecundity is equal to the
variance; this is a conservative model, as overdispersion is very likely.

Note that we violate a key assumption from the start: independence between
points. Each of these points are taxa, so they are phylogenetically
non-independent (Felsenstein, 1985). Knowing this is a serious problem, I still
first fit this relationship as a simple linear model, which illustrates some
pathologies with model fit.

```{r}

dr <- read_tsv('../data/romiguier_et_al_2014_updated.tsv') %>%
        mutate(log10_fecundity = log10(fecundity)) %>%
        #filter(fecundity < 1e5) %>% # optional remove an outlier
        filter(!is.na(log10_fecundity), !is.na(log10_size))

size_fecundity_plot <- function() {
  with(dr %>% filter(fecundity < 8e4),
       plot(log10_size, log10_fecundity, 
            axes=FALSE, 
            pch=19,
            xlim=c(-3.1, 1),
            col='gray42'))
  axis(1, seq(-3, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-3, 1))))
  axis(2, seq(-3, 5), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-3, 5))))
}

size_fecundity_plot()

lmf <- lm(log10_fecundity ~ log10_size, dr)
abline(lmf)

```

We see two things: (1) a great deal of heteroscedasticity, (2) a poor fit
around the right end of the plot (all points are below the line of best fit).
Looking at a plot of the residuals vs fitted:

```{r}
plot(lmf, which=1)
```

This pathology is also reflected in the QQ-plot:

```{r}
plot(lmf, which=2)
```

So instead, I will try to fit this with a Bayesian model using Stan. As before,
I ameliorate non-independence by considering family and phylum-level averages.

```{r}

dr_family_averages  <- dr %>% group_by(family) %>% 
  summarize(log10_size = mean(log10_size, na.rm=TRUE),
            log10_fecundity=mean(log10_fecundity, na.rm=TRUE),
            fecundity=mean(fecundity, na.rm=TRUE),
            n=n())

dr_phylum_averages  <- dr %>% group_by(phylum) %>% 
  summarize(log10_size = mean(log10_size, na.rm=TRUE),
            log10_fecundity=mean(log10_fecundity, na.rm=TRUE),
            fecundity=mean(fecundity, na.rm=TRUE),
            n=n())

```

We first fit a log-normal regression model on the family-level averaged data.

```{r}

lnf_xnew <- seq(-3, 1, length.out=100)

lnf <- stan(file='../stan/fecundity_log10size.stan',
            data=list(N = nrow(dr_family_averages), 
                      log10_fecundity=dr_family_averages$log10_fecundity,
                      log10_size=dr_family_averages$log10_size,
                      x_new = lnf_xnew),
            seed = 861234,
            iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Looking at model diagnostics, we see good number of effective samples, and
`Rhat`, showing good mixing of the chains.

```{r}
lnf_pars <- c('beta', 'sigma', 'lp__')
print(lnf, pars=lnf_pars)
```

And finally, in our pair plot, we see no divergent transitions.

```{r}
pairs(lnf, pars=lnf_pars)
```

Next, I do some posterior predictive checks, visualizing it along side the
HPDI. Overall the model fit is good.

```{r}

size_fecundity_plot()
ci_polygon(lnf_xnew, y=extract(lnf, pars='mu_rep')$mu_rep,
           color='gray', type='hpdi')
ci_polygon(lnf_xnew, y=extract(lnf, pars='y_rep')$y_rep,
           color='red', type='quantile', transp=0.1)
lnf_p <- colMeans(as_tibble(extract(lnf, pars='mu_rep')))
lines(lnf_xnew, lnf_p, col=alpha('black', 0.4), lwd=3)

```

We save these data for prettier publication-quality plots.

```{r}
save(dr,
     dr_family_averages, dr_phylum_averages,
     lnf_xnew, lnf, 
     file='../data/body_size_fecundity.Rdata')
```

## Fecundity and Body Size

Now, we look at the relationship between fecundity, body size, and diversity.

```{r}

dr %>% ggplot() + geom_point(aes(log10_size, log10(piS))) + 
   geom_segment(aes(log10_size, log10(piS), 
                    xend=log10_size, yend=log10(piS*(fecundity))))

```

### 2021-02-08 R/node_height.r

@@ -1,0 +1,89 @@
+
+
+getTips <- function(phy, x) {
+  tips <- seq_along(phy$tip.label)
+  phy$tip.label[intersect(getDescendants(phy, x), tips)]
+}
+
+node_height <- function(x, phy, abs_pics=TRUE, log_pics=FALSE, 
+                        scaled=TRUE, method='rlm') {
+  fit_func <- list('lm'=lm, 'rlm'=MASS:::rlm)[[method]]
+  x <- setNames(x, phy$tip.label)
+  ics <- pic(x, phy, scaled=scaled)
+  if (abs_pics) 
+    ics <- abs(ics)
+  if (log_pics)
+    ics <- log(pics)
+  phy$node.label <- NULL  # otherwise we can't access descendants
+  bt <- branching.times(phy)
+  # this just reverses the x-axis, which I don't like
+  # bt <- abs(bt - max(bt))
+  d <- tibble(bt= bt, pic=ics, nodes=names(bt))
+  d <- d %>% mutate(tips = map(nodes, ~ getTips(phy, .)))
+  fit <- fit_func(ics ~ bt)
+
+  df <- summary(fit)$df[2]
+  tval <- summary(fit)$coefficients[2, 3]
+  pval <- dt(abs(tval), df)
+ 
+  list(d=d, fit = fit, abs_pics=abs_pics, log_pics=log_pics,
+      method=method, pval=pval)
+}
+
+
+# plot node-height tests
+plot_nh <- function(x, title="", ylim=NULL, 
+                    xlab="branching time", twotail=FALSE,
+                    lowess_line=FALSE,
+                    ylab=NULL, alpha=0.05, f=1/4, at=NULL) {
+    lab <- "contrasts"
+  if (x$abs_pics)
+    lab <- "absolute value of contrasts"
+  if (x$log_pics)
+    lab <- sprintf('log(%s)', lab)
+  if (!is.null(ylab))
+    lab <- ylab
+  plot(pic ~ bt, x$d, axes=FALSE,
+       type='n',
+       main='',
+       ylim=ylim,
+       xlab='', ylab='')
+  points(pic ~ bt, x$d, pch=21, 
+         cex=1.3,
+       bg='gray22', col='white',
+       lwd=0.2)
+  if (x$method == 'rlm') {
+    df <- summary(x$fit)$df[2]
+    tval <- summary(x$fit)$coefficients[2, 3]
[…]

### 2021-02-08 R/range_funcs.r

@@ -1,0 +1,296 @@
+library(rgbif)
+library(rnaturalearth)
+library(sf)
+library(tidyverse)
+library(igraph)
+library(alphahull)
+library(maptools)
+
+# WGS84 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
+
+LAND <- rnaturalearth::ne_countries(returnclass = "sf") 
+CONTINENTS <- LAND %>% group_by(continent) %>% summarize() %>% st_union()
+
+sanitize_species <- function(x) {
+  # for the Leffler et al data
+  lapply(strsplit(x, ' +', perl=TRUE), 
+         function(y) {
+           sps <- y[2]
+           if (regexpr('/', sps, fixed=TRUE) != -1) {
+             spss <- strsplit(sps, '/', fixed=TRUE)
+             map(spss, function(z) paste(y[1], z))
+           } else {
+             paste(y[1], y[2])
+           }
+         })
+}
+
+
+# get_occurrences <- function(species) {
+#   key <- name_suggest(q=species, rank='species')$key[1]
+#   if (length(key) == 0) {
+#   }
+#   occ_search(key)
+# }
+
+
+# get this *once*
+countriesSP <- rworldmap:::getMap(resolution='low')
+coords2continent = function(lat, lon) {  
+  # https://stackoverflow.com/questions/21708488/get-country-and-continent-from-longitude-and-latitude-point-in-r
+  points <- data.frame(lon=lon, lat=lat)
+  # converting points to a SpatialPoints object
+  # setting CRS directly to that from rworldmap
+  pointsSP = sp:::SpatialPoints(points, proj4string=sp:::CRS(sp:::proj4string(countriesSP)))  
+
+  # use 'over' to get indices of the Polygons object containing each point 
+  indices = sp:::over(pointsSP, countriesSP)
+
+  #indices$REGION   # returns the continent (7 continent model)
+  indices$continent
+}
+
+
+on_map <- function(occurs, database='world', add=FALSE, 
+                   direct_data=FALSE,
+                   pch=19, cex=0.2, col='firebrick', ...) {
+  if (!direct_data) 
+    d <- occurs$data 
+  else
+    d <- occurs
+  x <- d[, c('decimalLongitude', 'decimalLatitude')]
+  if (!add) {
+    plot.new()
+    maps::map(database)
+  }
[…]

### 2021-02-08 R/utilities.r

@@ -1,0 +1,224 @@
+## utilities.r -- utility functions
+library(HDInterval)
+library(rstan)
+
+wescols <- c("#75BFC1", "#CD2216", "#E8A252", "#D4DAB2", "#6B7181")
+
+
+lambda_est <- function(fit) {
+  sigma_r <- rstan:::extract(fit, pars='sigma_r')$sigma_r
+  sigma_p <- rstan:::extract(fit, pars='sigma_p')$sigma_p
+  l <- sigma_p / (sigma_p + sigma_r)
+  c(mean=mean(l), hdi(l))
+}
+
+# bootstrap samples
+bs_samp  <- function(x) {
+  x[sample(nrow(x), replace=TRUE), ]
+}
+
+
+hpdi <- function(x, alpha) {
+  coda::HPDinterval(coda::as.mcmc(x), prob = 1 - alpha)
+}
+
+ci_polygon <- function(x, fit=NULL, y=NULL, type='quantile',
+                       par = 'y_rep', alpha=0.05, 
+                       color='red', transp=0.4, ...) {
+  xr <- order(x)
+  if (!is.null(fit)) {
+    y <- rstan:::extract(fit, pars=par)[[par]]
+  }
+  stopifnot(!is.null(y))
+  if (type == 'quantile') {
+    cis <- apply(y, 2, function(p) 
+                 quantile(p, probs=c(alpha/2, 1-alpha/2)))
+  } else if (type == 'hpdi') {
+    cis <- apply(y, 2, function(p) hpdi(p, alpha))
+  } else {
+    stop('type must be hpdi or quantile')
+  }
+  polygon(c(x[xr], rev(x[xr])), c(cis[2, xr], rev(cis[1, xr])), 
+          border=0, col=alpha(color, transp), ...)
+  invisible(cis)
+}
+
+maplength_bodysize_plot <- function(x, default_axes=FALSE) {
+  # wraps global data 
+  with(x, plot(map_length, log10_size, pch=19, cex=0.7, col='gray48'))
+ if (!default_axes) {
+    with(x, plot(map_length, log10_size, pch=19, axes=FALSE, cex=0.7,
+                         ylab='', xlab='', 
+                         xlim=c(0, 50), ylim=c(-4.3, 1.1), 
+                         col='gray48'))
[…]

### 2021-02-08 data/anage_data.r

@@ -1,0 +1,99 @@
+library(tidyverse)
+
+ctypes <- cols(
+  HAGRID = col_character(),
+  Kingdom = col_character(),
+  Phylum = col_character(),
+  Class = col_character(),
+  Order = col_character(),
+  Family = col_character(),
+  Genus = col_character(),
+  Species = col_character(),
+  `Common name` = col_character(),
+  `Female maturity (days)` = col_double(),
+  `Male maturity (days)` = col_double(),
+  `Gestation/Incubation (days)` = col_double(),
+  `Weaning (days)` = col_double(),
+  `Litter/Clutch size` = col_double(),
+  `Litters/Clutches per year` = col_double(),
+  `Inter-litter/Interbirth interval` = col_double(),
+  `Birth weight (g)` = col_double(),
+  `Weaning weight (g)` = col_double(),
+  `Adult weight (g)` = col_double(),
+  `Growth rate (1/days)` = col_double(),
+  `Maximum longevity (yrs)` = col_double(),
+  Source = col_character(),
+  `Specimen origin` = col_character(),
+  `Sample size` = col_character(),
+  `Data quality` = col_character(),
+  `IMR (per yr)` = col_double(),
+  `MRDT (yrs)` = col_double(),
+  `Metabolic rate (W)` = col_double(),
+  `Body mass (g)` = col_double(),
+  `Temperature (K)` = col_double(),
+  References = col_number()
+)
+
+
+cnames <- c(
+  'HAGRID',
+  'kingdom',
+  'phylum',
+  'class',
+  'order',
+  'family',
+  'genus',
+  'species',
+  'common_name',
+  'female_maturity_days',
+  'male_maturity_days',
+  'gestation_incubation_days',
+  'weaning_days',
+  'litter_clutch_size',
+  'litters_clutches_per_year',
+  'inter_litter_interbirth_interval',
+  'birth_weight_g',
+  'weaning_weight_g',
+  'adult_weight_g',
+  'growth_rate_recip_days',
+  'maximum_longevity_yrs',
+  'source',
+  'specimen_origin',
+  'sample_size',
+  'data_quality',
+  'imr_per_yr',
+  'mrdt_yrs',
+  'metabolic_rate_w',
+  'body_mass_g',
+  'temperature_k',
+  'references'
+)
+
+da_raw <- read_tsv('./anage_data.tsv', col_types=ctypes)
[…]

### 2021-02-08 data/combined_dataset.r

@@ -1,0 +1,560 @@
+# stapley_leffler_combined.r -- make main dataset, add in range data
+## CACHE FILES
+# combined_dataset_hier.Rdata: taxanomical hierarchy 
+# redlist.Rdata: redlist species, downloaded from web
+# combined_tree.Rdata: tree of life synthetic tree (no branch lengths)
+# datalife.Rdata: synthetic calibrated tree
+# popsize_chains.Rdata: Stan chains (these created by popsize_models.r, 
+#                       which is sourecd here)
+
+# NOTE: the Damuth data are not dry body mass -- I checked a few species.
+# neither are Romiguier et al.
+
+## OUTPUT DATA
+# size_mass_density_lms.Rdata: linear models on Damuth 1987 data
+# calibrated_phylogeny.Rdata: data for metazoan taxa in tree, and calibrated tree
+# redlist_popsizes.tsv: population sizes from redlist data
+# combined_data.tsv: the main dataset, for all metazona taxa (not parpath), even those not in phylogeny
+
+library(ape)
+library(phytools)
+library(datelife)
+library(rotl)
+library(scales)
+library(tidyverse)
+library(rredlist)
+library(ritis)
+map <- purrr::map
+source('../R/utilities.r')
+
+opar <- par(no.readonly=TRUE)
+
+#### Load all in Raw Datasets
+dl <- read_tsv('leffler_et_al_2012_updated.tsv') %>% 
+        # Leffler merged some species; we average them here.
+        # we also merge all site types -- TODO, may want to do it both ways?
+        group_by_at(vars(-size, -diversity, -site_type)) %>% 
+        summarize_all(mean, na.rm=TRUE) %>%
+        mutate(diversity_data_source = 'Leffler et al.')
+
+ds <- read_tsv('stapley_et_al_2017_updated.tsv') %>% 
+        select(-group, -log10_size, -data_source) %>%
+        mutate(maplen_data_source = 'Stapley et al.') %>%
+        mutate(diversity = NA)
+
+dc <- read_tsv('corbett_detig_2015_updated.tsv')  %>%
+        mutate(diversity_data_source = 'Corbett-Detig et al.',
+               maplen_data_source =    'Corbett-Detig et al.',
+               # rename the original Corbett-Detig range
+               range_cd = range) %>%
+        rename(diversity = obs_pi) %>%
+        select(-range, -kingdom)
+
+# Romiguier et al data
+dr <- read_tsv('romiguier_et_al_2014_updated.tsv') %>%
[…]

### 2021-02-08 data/corbett_detig_2015_data.r

@@ -1,0 +1,158 @@
+# data.r -- join all the Corbet-Detig datasets
+library(tidyverse)
+
+maps_raw <- readxl:::read_xls('journal.pbio.1002112.s016.XLS')
+maps <- maps_raw[, c(1:3, 6, 10)]
+colnames(maps) <- c('species', 'num_chr', 'map_length', 'totalMb_covered', 'ave_rec')
+
+div_cord <- readxl:::read_xls('S2_Data.xls')
+
+colnames(div_cord) <- c("species", "abbrv", "kingdom", "size", 
+                        "range", "pred_pi",
+                        "obs_pi", "impact_of_sel", 
+                        "log10_size", "log10_range", 
+                        "neut_model_conf")
+
+# genome size
+gs <- readxl:::read_xls('journal.pbio.1002112.s018.XLS')[c(1, 2, 4, 10)]
+colnames(gs) <- c('species', 'genome_species', 'assembly_size', 'genome_size')
+
+# selfing
+self <- readxl:::read_xls('journal.pbio.1002112.s017.XLS')[c(1, 3, 4)]
+colnames(self) <- c('species', 'selfing', 'self_est')
+
+
+# Below is the Table S5 of McGaugh et al. 2012
+# for reference.
+# This paper is cited as the map source for D. pseudoobscura 
+# in Corbett-Detig et al (2015). The map length of 
+# pseudoobscura in CD2015 is 0.535 (CD2015 Table S13)
+# which seems very low. McGaugh et al say they only use 43%
+# of the chromosome.
+mcg12 <- tribble(~ Mb, ~ cMMb,
+                 0.021916,10.38,
+                 0.019957,4.67,
+                 0.017055,3.64,
+                 0.021615,2.87,
+                 0.024667,6.29,
+                 0.020844,21.25,
+                 0.019964,6.9,
+                 0.019212,3.59,
+                 0.021546,4.11,
+                 0.022377,3.52,
+                 0.021324,13.85,
+                 0.02736,4.31,
+                 0.02314,8.07,
+                 0.012614,2.34,
+                 0.018506,1.59,
+                 0.017728,6.1)
+
+# In looking the D. pseudoobscura data, I cannot
+# recreate CD's estimate and it's a very strong 
+# outlier, so I am ignoring this
[…]

### 2021-02-08 data/dmel/annotation.r

@@ -1,0 +1,98 @@
+library(GenomicRanges)
+library(TxDb.Dmelanogaster.UCSC.dm3.ensGene)
+library(tidyverse)
+
+keep_seqs <- c("chr2L", "chr2R", "chr3R", "chr3L")
+
+tx <- TxDb.Dmelanogaster.UCSC.dm3.ensGene
+reduce <- IRanges:::reduce
+
+sq_filters <- function(x, chroms) x[seqnames(x) %in% chroms]
+
+dm_exs <- exons(tx, filter=list(tx_chrom=keep_seqs))
+dm_introns <- unlist(sq_filters(intronsByTranscript(tx), keep_seqs))
+dm_5UTRs <- unlist(sq_filters(fiveUTRsByTranscript(tx), keep_seqs))
+dm_3UTRs <- unlist(sq_filters(threeUTRsByTranscript(tx), keep_seqs))
+
+seqlens <- seqlengths(tx)
+seqlens <- seqlens[names(seqlens) %in% keep_seqs]
+G <- sum(seqlens)
+
+d <- tribble(~ feature, ~ total_bp, 
+                 'exon', sum(width(reduce(dm_exs, ignore.strand=TRUE))),
+                 'intron', sum(width(reduce(dm_introns, ignore.strand=TRUE))),
+                 'utr', sum(width(reduce(dm_5UTRs))) + 
+                        sum(width(reduce(dm_3UTRs)))) 
+
+# compute non-coding (drop exon; we use CDS)
+noncoding <- G - d %>% pull(total_bp) %>% sum()
+
+d <- d %>% add_row(feature='noncoding', total_bp=noncoding) %>% 
+  mutate(prop_genome = total_bp / G) 
+
+# sanity check --- all the features 1?
+stopifnot(sum(d$prop_genome) == 1)
+
+write_tsv(d, 'dmel_annotation_summary.tsv')
+
+sl <- tibble(chrom = names(seqlens), length=seqlens)
+write_tsv(sl, 'dmel_seqlens.tsv')
+
+
+
+
+# col_types <- "c_ciic___"
+# col_names <- c('seqname', 'source', 'feature', 'start', 'end', 'score', 'filter', 'strand', 'group', 'attribute')
+
+# gff <- read_tsv('./dmel-all-r5.33.gff.gz', comment="#", col_types=col_types, col_names=col_names)
+
[…]

### 2021-02-08 data/leffler_et_al_2012_data.r

@@ -1,0 +1,185 @@
+## leffler_et_al_2012_data.r -- Leffler et al (2012) data
+library(tidyverse)
+library(ritis)
+library(ggrepel)
+source('../R/utilities.r')
+
+# Load Leffler et al (2012) diversity data, with 
+# my hand-collected body length data
+ld <- read_tsv('leffler_et_al_2012.tsv')
+colnames(ld) <- c('species', 'common_name', 'phylum', 'class', 'total_loci',
+                  'mean_diversity_pops', 'thetaW_pi', 'chr_type', 'site_type', 
+                  'sampling', 'reference', 'mating_system', 'habitat', 
+                  'range', 'included_study', 'included_sex_auto', 'size', 
+                  'size_source', 'size_comments')
+
+# # I did not enter body size data in the Leffler data that exists elsewhere; 
+# # we merge that in now
+# dsc <- read_tsv('combined_data.tsv')
+# size_lookup <- as.list(c(dsc$size[!is.na(dsc$size)], ld$size[!is.na(ld$size)]))
+# names(size_lookup) <- c(dsc$species[!is.na(dsc$size)], ld$species[!is.na(ld$size)])
+
+# # before lookup missing size data
+# table(is.na(ld$size))
+# tmp <- sapply(size_lookup[ld$species], function(x) ifelse(is.null(x), NA, x))
+# ld$size <- as.numeric(tmp)
+# # after merge
+# table(is.na(ld$size))
+
+
+## A few subspecies with different names *should* be averaged
+# so we change their names here
+# Combine intestinalis 
+ciona <- regexpr("Ciona intestinalis", ld$species) != -1
+ld$species[ciona] <- "Ciona intestinalis"
+# subtypes A & B are listed as cosmopolitan and broad endemic. I average
+# and call ie cosmopolitan.
+ld$range[ciona] <- "Cosmopolitan"
+
+# Mus subspecies are averaged since the GBIF data doesn't seem to distinguish 
+# between subspecies
+ld[regexpr('Mus musculus', ld$species) != -1, ]$species <- 'Mus musculus'
+
+
+# average the diversity values, looking only at π, autosomes included in the study
[…]

### 2021-02-08 data/popsize_models.r

@@ -1,0 +1,88 @@
+library(tidyverse)
+library(rstan)
+
+# x prefixes are so this can be included in combined_dataset.r
+
+# My combined dataset, animal subset
+xda <- d %>% #read_tsv('../data/combined_data.tsv') %>%
+       filter(kingdom == 'Animalia') %>%
+       filter_at(vars(size, range), ~ is.finite(.)) %>%
+       # put a minimum range of 1km2
+       mutate(range = ifelse(range == 0, 1, range))
+
+# Romiguier et al (2014) data
+xdr <- dr %>%
+        #read_tsv('../data/romiguier_et_al_2014_updated.tsv') %>%
+        filter_at(vars(size, bodymass_g), ~ !is.na(.))
+
+# Damuth 1987 data
+xdd <- dd #read_tsv('../data/damuth1987.tsv')
+
+plot(log10(bodymass_g) ~ log10(size), data=xdr)
+plot(log10(density) ~ log10(mass_g), data=xdd)
+
+dat <- list( # Romiguier et al data
+             N_r = nrow(xdr),
+             body_length_r=xdr$size,
+             body_mass_r=xdr$bodymass_g,
+             # Damuth 1987 data
+             N_d = nrow(xdd),
+             density_d = xdd$density,
+             body_mass_d = xdd$mass_g,
+             # main dataset
+             N = nrow(xda),
+             body_length = xda$size,
+             range_meas = xda$range)
+
+
+species <- xda$species
+
+CACHED_POPSIZE_CHAINS_FILE <- "popsize_chains.Rdata"
+FORCE <- FALSE
+if (FORCE || !file.exists(CACHED_POPSIZE_CHAINS_FILE)) {
+  bayes_fit <- stan('../stan/pred_popsize_missing_centered.stan', data=dat, 
+                    cores=4,
+                    iter=10000)
+  save(bayes_fit, file=CACHED_POPSIZE_CHAINS_FILE)
+} else {
+  load(CACHED_POPSIZE_CHAINS_FILE)
+}
+
+pars <- c('alpha', 'beta', 'sigma_mass', 'gamma', 'theta', 'sigma_density')
+print(bayes_fit, pars=pars)
+# print(fit, pars='pred_density')
+
+popsize <- rstan:::extract(bayes_fit, pars='log10_popsize')$log10_popsize
+mass <- rstan:::extract(bayes_fit, pars='log10_body_mass')$log10_body_mass
[…]

### 2021-02-08 data/romiguier_et_al_2014_data.r

@@ -1,0 +1,125 @@
+library(tidyverse)
+library(readxl)
+library(ritis)
+
+dr_raw <- read_xls('romiguier_et_al_2014_supptable_2.xls')
+
+colnames(dr_raw) <- c("species", "common_name", "family", "order",
+                      "class", "phylum", "number_of_inds", "contigs_number",
+                      "average_length_of_contigs", "SNP_number", "piS",
+                      "piN", "piNpiS", "fit", "invasive_status", 
+                      "maximal_geoographic_distance",
+                      "mean_geographic_distance", "mean_distance_from_equator")
+
+# remove quotes around names
+cleanup_text <- function(x) gsub('"([^"]+)"', '\\1', x)
+
+clean_species <- function(x) {
+  sub('([^_]+)_([^_]+).*', '\\1 \\2', x)
+}
+
+dr <- dr_raw %>% mutate_at(vars(species:phylum), cleanup_text) %>%
+         mutate(species=map_chr(species, clean_species)) %>%
+         mutate(piS=as.numeric(piS), 
+                piN=as.numeric(piN))
+
+
+# Supplementary table 3 data, extracted from PDF using Tabula 
+# and then cleaned up by hand
+s3_files <- list.files('.', pattern='romiguier_et_al_supptable_3_.*\\.csv', 
+              full.names=TRUE)
+
+s3_type <- sub('.*_([^\\.]+)\\.csv', '\\1', s3_files)
+names(s3_files)  <- s3_type
+
+s3d_raw <- Map(function(x, n) {
+                 d <- read_csv(x)
+                 colnames(d)[3:4] <- paste(n, colnames(d)[3:4], sep='_')
+                 d
+              }, s3_files, s3_type)
+
+# join everything
+drs <- reduce(s3d_raw, inner_join, by='species') %>% 
+           # convert to kg
+           mutate(mass = bodymass_g / 1000)  %>%
+           # convert to m
+           mutate(size = size_cm / 100) %>%
+           # convert to m
+           mutate(propagule = propagule_size_cm / 100) %>%
+           mutate(log10_size=log10(size),
+                  log10_mass=log10(mass),
+                  log10_propagule=log10(propagule)) %>%
+           mutate(species=map_chr(species, clean_species)) %>%
+           left_join(dr) %>% 
[…]

### 2021-02-08 data/species_range_fixes.r

@@ -1,0 +1,148 @@
+library(tidyverse)
+
+terrestriality_fixes <- tribble(~ name, ~ is_terrestrial,
+        'Anopheles scanloni', TRUE,
+        'Arabidopsis', TRUE,
+        'Aphanarthrum subglabrum/glabrum', TRUE,
+        'Ciona', FALSE, 
+        'Caenorhabditis', TRUE,
+        'Crassostrea gigas', FALSE,
+        'Cryptomeria japonica', TRUE,
+        # note: daphnia is freshwater, so should be found on continents
+        'Daphnia', TRUE,
+        'Drosophila', TRUE,
+        # occurrences of gray whales are from coasts, so this is 
+        # inferred as a terrestrial, yes, land gray whales... *facepalm* 
+        'Eschrichtius robustus', FALSE,
+        'Ficedula albicollis', TRUE,
+        # Gasterosteus aculeatus / stickelback is classified as terrestrial here --
+        # mostly inland (no changes)
+        'Hoolock leuconedys', TRUE,
+        # chinook is hard-- it's both. We assign as terretrial here
+        # as that gives a better estimate
+        'Oncorhynchus tshawytscha', TRUE,
+        # all urchins are not terrestrial
+        'Strongylocentrotu', FALSE,
+        'Saccharomyces', TRUE,
+        'Bombyx', TRUE,
+        'Agaricus', TRUE,
+        'Agrostis', TRUE,
+        'Agrostis', TRUE,
+        'Amaranthus', TRUE,
+        'Ananas', TRUE,
+        'Bracon', TRUE,
+        'Brassica', TRUE,
+        'Camelina', TRUE,
+        'Capsicum', TRUE,
+        'Carica', TRUE,
+        'Cocos', TRUE,
+        'Coffea', TRUE,
+        'Cucurbita', TRUE,
+        'Cynodon', TRUE,
+        'Daucus', TRUE,
+        # penguin
+        'Eudyptes', TRUE,
+        'Festuca', TRUE,
+        'Gossypium', TRUE,
+        'Hibiscus', TRUE,
+        'Hippocampus', FALSE,
+        'Lactuca', TRUE,
+        'Lineus', FALSE,
+        'Litopenaeus', FALSE,
+        'Lolium', TRUE,
+        'Lupinus', TRUE,
+        'Macropus', TRUE,
+        'Malus', TRUE,
+        'Mangifera', TRUE,
+        'Medicago', TRUE,
[…]

### 2021-02-08 data/species_ranges.r

@@ -1,0 +1,405 @@
+library(rgbif)
+library(rnaturalearth)
+library(sf)
+library(tidyverse)
+library(readxl)
+library(igraph)
+library(units)
+
+source('../R/range_funcs.r')
+
+
+#### Load the data in 
+
+## Leffler data
+message('loading Leffler et al data...')
+dl <- read_tsv('./leffler_et_al_2012_updated.tsv')
+lsp <- sanitize_species(dl$species)
+names(lsp) <- dl$species
+
+# get Leffler species names
+uniq_lsp_names <- unique(names(lsp))
+uniq_lsp <- lsp[uniq_lsp_names]
+
+## Stapley data 
+message('loading Stapley et al data...')
+ds <- read_tsv('./stapley_et_al_2017_updated.tsv')
+# Stapley is all unique species, double check
+stopifnot(length(unique(ds$species)) == nrow(ds))
+
+## Romiguier data 
+# there are some multiple observations, but not combined like Leffler data
+# (hence no need for key)
+message('loading Romiguier et al data...')
+drs <- read_tsv('./romiguier_et_al_2014_updated.tsv')
+drs_sps <- unique(drs$species)
+
+## Combine all data
+# flatten the Leffler data
+species_df <- bind_rows(map2(lsp, names(lsp), function(x, n) {
+                 tibble(key=rep(n, length(x)), species=unlist(x),
+                        dataset='Leffler')
+}))
+
+species_df <- bind_rows(species_df, tibble(key=ds$species, 
+                        species=ds$species, dataset='Stapley'),
+                        tibble(key=drs_sps, species=drs_sps,
+                               dataset='Romiguier'))
+
+uniq_species_df <- species_df %>% distinct(.keep_all=TRUE)
+nrow(uniq_species_df)
+
+#### Species keys and occurence data from GBIF
+safe_occ_search <- safely(occ_search)
+
+OCC_FILE <- 'gbif_species_occ.Rdata'
+if (!file.exists(OCC_FILE)) {
+  message('cached GBIF occurence data et al not found -- downloading from GBIF')
+  message('getting suggested GBIF species names...')
+  uniq_species_df <- uniq_species_df %>% 
+                       mutate(taxon_key = map(species, 
+                                        ~ name_suggest(., rank='species')))
+
+  message('downloading GBIF species occurrences...')
+  n <- nrow(uniq_species_df)
+  uniq_species_df <- uniq_species_df %>% 
[…]

### 2021-02-08 data/stapley_et_al_2017_data.r

@@ -1,0 +1,94 @@
+# stapley_et_al_2017_data.r -- Stapley et al (2017) data
+library(tidyverse)
+
+
+source('../R/utilities.r')
+
+# Stapley et al (2017) data, with my hand-collected body length data
+ds_raw <- read_csv('./stapley_et_al_2017.csv') 
+
+ds <- ds_raw %>% 
+        # convert the genome size to Gb like the Corbett-Detig data
+        mutate(genome_size = genome_size_Mb * 1e6 / 1e9)  %>%
+        # select out certain columns
+        select(species, species_ott, group, genome_size, hcn,  
+               map_length, sexual_system, size,
+               num_markers, parpath) %>%
+        # fix parpath 
+        mutate(parpath=case_when(parpath == 'n' ~ FALSE,
+                                 parpath == 'y' ~ TRUE,
+                                 TRUE ~ NA)) %>%
+        # add metadata
+        mutate(data_source='Stapley et al (2017)') %>%
+        mutate(group=tolower(group)) %>% 
+        # rename group
+        mutate(group=case_when(
+                               group == 'animals' ~ 'animal',
+                               group == 'plants' ~ 'plant',
+                               TRUE ~ group
+                               )) %>%
+       mutate(genus = map_chr(strsplit(species, ' '), 1)) %>%
+       select(-species_ott)
+
+# Stapley data for Apis cerana is wrong...
+# https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0076459
+# the largest linkage block, not total map length reported
+ds$map_length[ds$species == 'Apis cerana'] <- 3942.7
+
+# # grt the taxonomial datas from taxadb
+# ds <- ds %>% mutate(id_raw=map(genus, taxadb::get_ids)) %>%
+#               mutate(id = map_int(id_raw, ~ as.integer(sub('ITIS:', '', .))))
+
+
+# ## Lookup and merge in taxonomy 
+# taxa_file <- 'stapley_et_al_2017_data_hier.Rdata'
+# if (!file.exists(taxa_file)) {
+#   hier <- map(ds$id, possibly(hierarchy_full, NA))
+#   save(hier, file=taxa_file)
+# } else {
+#   load(taxa_file)
+# }
+# ds$taxons <- NA
+# ds$taxons <- hier
+
+# # extract out relevant rank levels
[…]

### 2021-02-08 data/taxonomy_fixes.r

@@ -1,0 +1,383 @@
+## Manual taxonomy fixes
+taxon_levels <- c('kingdom', 'phylum', 'class', 'order', 'family')
+
+taxon_fixes <- list(
+     list(species = "Agaricus bisporus",
+          kingdom = 'Fungi', 
+          phylum='Basidiomycota', 
+          class='Agaricomycetes', 
+          order='Agaricales',
+          family='Agaricaceae'),
+
+     list(species = "Chimonanthus praecox",
+          kingdom = 'Plantae', 
+          phylum=NA, 
+          class=NA, 
+          order='Laurales',
+          family='Calycanthaceae'),
+
+     list(species = "Aphanarthrum subglabrum/glabrum",
+          kingdom = 'Animalia', 
+          phylum='Arthropoda', 
+          class='Insecta', 
+          order='Coleoptera',
+          family='Curculionidae'),
+
+     list(species="Cecidostiba fungosa",
+          kingdom='Animalia',
+          phylum='Arthropoda',
+          class='Insecta',
+          order='Hymenoptera',
+          family='Pteromalidae'),
+
+     list(species="Mesobuthus cyprius/gibbosus",
+          kingdom='Animalia',
+          phylum='Arthropoda',
+          class='Arachnida',
+          order='Scorpiones',
+          family='Buthidae'),
+
+     list(species="Nasonia vitripennis",
+          kingdom='Animalia',
+          phylum='Arthropoda',
+          class='Insecta',
+          order='Hymenoptera',
+          family='Pteromalidae'),
+
+     list(species="Neurospora crassa",
+          kingdom='Animalia',
+          phylum='Ascomycota',
+          class='Sordariomycetes',
+          order='Sordariales',
+          family='Sordariaceae'),
+
+     list(species="Paracoccidioides brasiliensis PS2/PS3/S1",
+          kingdom='Animalia',
+          phylum='Ascomycota',
+          class='Eurotiomycetes',
+          order='Onygenales',
+          family='Ajellomycetaceae'),
+
+     list(species="Paracoccidioides lutzii",
+          kingdom='Animalia',
+          phylum='Ascomycota',
+          class='Eurotiomycetes',
+          order='Onygenales',
+          family='Ajellomycetaceae'),
+
+     list(species="Phlebotomus ariasi",
+          kingdom='Fungi',
+          phylum='Arthropoda',
+          class='Insecta',
+          order='Diptera',
+          family='Psychodidae'),
+
+     list(species="Pieris rapae",
+          kingdom='Animalia',
[…]

### 2021-02-08 notebooks/figures/Nc_figure.r

@@ -1,0 +1,139 @@
+# Nc_figure.r
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+opar <- par(no.readonly=TRUE)
+
+# for stan fit
+load('../../data/diversity_popsize_chains.Rdata')
+load('../../data/main_datasets.Rdata')
+phi <- mean(rstan:::extract(dps_fit, pars='phi')$phi)
+beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
+
+# alias this -- less typing
+d <- da_dps %>% filter(!is.na(log10_popsize)) %>% 
+      group_by(phylum) %>%  
+      mutate(n = n(), phyla_mean = mean(log10_popsize)) %>%
+      filter(n > 5) %>%
+      arrange(phyla_mean, log10_popsize)
+
+d %>% ggplot() + geom_histogram(aes(log10_popsize, fill=phylum), binwidth=0.7, alpha=0.5)
+
+
+d %>% ggplot() + geom_density(aes(log10_popsize, fill=phylum), binwidth=0.7, alpha=0.5)
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+lx <- 4
+ux <- 18
+
+
+output <- TRUE
+# output <- FALSE
+
+if (output)
+  pdf("Nc_figure.pdf", height=7, width=7/1.5)
+
+
+par(mar=c(1, 1, 1, 1), oma=c(0, 0, 0, 0))
+
+
+
+nf <- layout(matrix(c(1, 2), ncol=1), heights=c(2, 8))
+x <- d$log10_popsize
+sps <- d$species
+phy <- d$phylum
+i <- seq_along(x)
+
+## first plot
+bins = seq(lx, ux, length.out=14)
+hd <- d %>% nest() %>% 
+  mutate(hist = map(data, ~ hist(.$log10_popsize, bins, plot=FALSE)))
+
+al <- 1
+
+par(mar=c(1, 1, 1, 1))
+
+h <- hist(d$log10_popsize, 10, plot=FALSE)
+xd <- do.call(rbind, map(hd$hist, 'counts'))
+idx <- c(3, 2, 1, 4)
+xd <- xd[idx, ]
+cols <- phyla_cols[hd$phylum[idx]]
+
+barplot(xd, axes=FALSE, horiz=FALSE,
[…]

### 2021-02-08 notebooks/figures/Ne_popsize.r

@@ -1,0 +1,175 @@
+# Ne_popsize.r -- 
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+
+opar <- par(no.readonly=TRUE)
+
+# for div_fit
+load('../../data/lognorm_fulldata_with_social.Rdata')
+
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+      filter(kingdom == 'Animalia') %>%
+      mutate(log10_diversity = log10(diversity)) %>% 
+      filter_at(vars(log10_diversity, pred_log10_N, log10_popsize), 
+                ~ is.finite(.))
+
+
+# summary(lm(pred_log10_N ~ log10_popsize, d))
+# summary(lm(log10_diversity ~ log10_popsize, d))
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$log10_diversity - log10(4e-8)
+
+plot(x, y)
+
+X <- cbind(x, y)
+center <- colMeans(X)
+sigma <- cov(X)
+
+coords <- ellipse(mu=center, sigma=sigma, alpha=0.4, draw=FALSE)
+
+label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
+# contained <- rep(TRUE, length(x))
+
+d$label <- ifelse(label, d$species, "")
+
+set.seed(2)
+
+xlims <- c(4, 16.5)
+
+REPEL_CACHED_FILE <- 'Ne_repel.tsv'
+if (!file.exists(REPEL_CACHED_FILE)) {
+  p <- ggplot(d) + 
+    geom_point(aes(log10_popsize, log10_diversity - log10(4e-8), 
+                   label=species, color=phylum), size=2) +
+    geom_text_repel(data=d,
+                    mapping=aes(log10_popsize, log10_diversity - log10(4e-8), 
+                                label=label, text=species), 
+                                size=1.5, 
+                                point.padding=0.6, 
+                                force = 2,
[…]

### 2021-02-08 notebooks/figures/biconmap.r

@@ -1,0 +1,46 @@
+library(viridis)
+library(phytools)
+
+load('../../data/biconmap_maplen_data.Rdata')
+
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+pdf('biconmap.pdf', height=8, width=8)
+par(oma=c(0, 0, 0.5, 0))
+vcols <- viridis(100)
+vcols2 <- viridis(100, option='A')
+layout(matrix(1:3, 1, 3), widths=c(0.7, 0.14, 0.7))
+pscm <- psmlcm
+pscm2 <- setMap(pscm, colors=vcols)
+plot(pscm2, fsize=c(0, 0.8), ftype=c("off","reg"), sig=1, lwd=2,
+     outline=FALSE, mar=c(1.1,0.1,1.1,0.01), legend=FALSE)
+add.color.bar.log(500, vcols, prompt=FALSE, outline=FALSE, 
+              subtitle="", digits=0,
+              lims=range(dt_ml$log10_popsize), lwd=4,
+              title='log10(pop size)')
+title(main = 'log10(pop size)')
+# ylim<-c(1-0.12*(length(tr_mlps$tip.label)-1), length(tr_mlps$tip.label))
+
+plot.new()
+plot.window(xlim=c(-0.3,0.3), ylim=c(1, length(tr_ml$tip.label)))
+text(rep(0, length(tr_ml$tip.label)), 1:length(tr_ml$tip.label), 
+     tr_ml$tip.label,
+     font=3, cex=0.5, col=phyla_cols[factor(dt_ml$phylum)])
+mlcm2 <- setMap(mlcm, colors=rev(vcols2))
+plot(mlcm2, fsize=c(0, 0.8), ftype=c("off","reg"),
+    direction="leftwards", sig=1, lwd=2, legend=FALSE,
+    outline=FALSE, mar=c(1.1,0.01,1.1,0.01))
+add.color.bar.log(500, rev(vcols2), prompt=FALSE, outline=FALSE, 
+              subtitle="", x=250, digits=0,
+              lims=range(dt_ml$log10_map_length), lwd=4,
+              title='log10(map length)')
+title(main = 'log10(map length)')
+legend(350, 15, tools:::toTitleCase(unique(dt_ml$phylum)), 
+       fill=phyla_cols,
+       bty='n', border=0, cex=1, ncol=2)
+
+dev.off()
+
+
+

### 2021-02-08 notebooks/figures/biconmap_diversity.r

@@ -1,0 +1,55 @@
+library(viridis)
+library(phytools)
+
+load('../../data/biconmap_data.Rdata')
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+
+# output <- FALSE
+output <- TRUE
+
+if (output)
+  pdf('biconmap_diversity.pdf', height=8, width=8)
+
+par(oma=c(0, 0, 0.5, 0))
+vcols <- viridis(100)
+vcols2 <- viridis(100, option='A')
+layout(matrix(1:3, 1, 3), widths=c(0.7, 0.16, 0.7))
+pscm2 <- setMap(pscm, colors=vcols)
+plot(pscm2, fsize=c(0, 0.8), ftype=c("off","reg"), sig=1, lwd=2,
+     outline=FALSE, mar=c(1.1,0.1,1.1,0.01), legend=FALSE)
+add.color.bar.log(500, vcols, prompt=FALSE, outline=FALSE, 
+              subtitle="", digits=0,
+              lims=range(dt_dps$log10_popsize), lwd=4,
+              title='pop size')
+title(main = 'pop size', font.main = 1, cex.main=1.5)
+# ylim<-c(1-0.12*(length(tr_mlps$tip.label)-1), length(tr_mlps$tip.label))
+
+plot.new()
+plot.window(xlim=c(-0.3,0.3), ylim=c(1, length(tr_dps$tip.label)))
+text(rep(0, length(tr_dps$tip.label)), 
+     1:length(tr_dps$tip.label), 
+     tr_dps$tip.label,
+     font=3, cex=0.5, col=phyla_cols[dt_dps$phylum])
+dvcm2 <- setMap(dvcm, colors=vcols2)
+plot(dvcm2, fsize=c(0, 0.8), ftype=c("off","reg"),
+    direction="leftwards", sig=1, lwd=2, legend=FALSE,
+    outline=FALSE, mar=c(1.1,0.01,1.1,0.01))
+add.color.bar.log(500, vcols2, prompt=FALSE, outline=FALSE, 
+              subtitle="", x=250, digits=0,
+              lims=range(dt_dps$log10_diversity), lwd=4,
+              title='diversity')
+title(main = 'diversity', font.main = 1, cex.main=1.5)
+nms <- tools:::toTitleCase(unique(dt_dps$phylum)) 
+i <- order(nms)
+nms <- nms[i]
+legend(300, 30, nms,
[…]

### 2021-02-08 notebooks/figures/body_mass_length.r

@@ -1,0 +1,50 @@
+library(tidyverse)
+library(scales)
+
+opar <- par(no.readonly=TRUE)
+
+pdf("body_mass_length.pdf", width=5, height=5)
+par(mar=c(5, 5, 2, 1))
+
+load('../../data/popsize_chains.Rdata')
+
+pars <- rstan:::extract(bayes_fit, pars=c('alpha', 'beta'))
+mean_intercept <- mean(pars$alpha)
+mean_slope <- mean(pars$beta)
+levels <- c("Annelida", "Arthropoda", "Chordata", 
+            "Echinodermata", "Mollusca", "Nematoda", "NA")
+d <- read_tsv('../../data/romiguier_et_al_2014_updated.tsv')   %>%
+        mutate(log10_mass_g = log10(bodymass_g)) %>%
+        filter_at(vars(log10_mass_g, log10_size), ~ !is.na(.)) %>%
+        mutate(phylum = ifelse(is.na(phylum), "NA", phylum)) %>% 
+        mutate(phylum = factor(phylum, levels=levels, ordered=TRUE))
+
+ngroups <- length(unique(d$phylum)) - 1
+group_cols <- c(hue_pal()(ngroups), 'gray52')
+
+plot(d$log10_size, d$log10_mass_g, pch=19, 
+     type='n', axes=FALSE, ylim=c(-4, 6),
+     xlim=c(-3, 1),
+     xlab='',
+     ylab='')
+points(d$log10_size, d$log10_mass_g, pch=21, 
+       cex=1.5, lwd=0.4, bg=group_cols[as.factor(d$phylum)], col='white')
+abline(a=mean_intercept, b=mean_slope, lty=2, col='gray52')
+
+xseq <- seq(-3, 1, 1)
+yseq <- seq(-4, 6, 2)
+axis(1, xseq, labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
+     padj = -1,
+     tck=-0.01,
+     cex.axis=0.8, line=0.5)
+axis(2, yseq, labels=latex2exp::TeX(sprintf("$10^{%d}$", yseq)), 
+     las=1,
+     tck=-0.01, hadj=0.5,
+     cex.axis=0.8, line=0.5)
+legend(-0.8, -2, tools:::toTitleCase(levels(as.factor(d$phylum))), fill=group_cols,
+       bty='n', border=0, cex=0.6, ncol=2)
[…]

### 2021-02-08 notebooks/figures/color_scheme.r

@@ -1,0 +1,15 @@
+library(tidyverse)
+library(scales)
+
+# we want consistent phylum colors, which we set here
+
+xx <- read_tsv('../../data/combined_data.tsv')  %>%
+       filter(kingdom == 'Animalia') 
+
+
+nphyla <- length(unique(xx$phylum))
+phyla_cols <- hue_pal()(nphyla)
+names(phyla_cols) <- sort(unique(xx$phylum))
+
+
+ 

### 2021-02-08 notebooks/figures/corbett_detig_fig.r

@@ -1,0 +1,255 @@
+library(viridis)
+library(grDevices)
+library(wesanderson)
+library(shape)
+library(rstan)
+
+opar <- par(no.readonly=TRUE)
+
+# source('../../R/utilities.r')
+source('color_scheme.r')
+
+load('../../data/corbett_detig_data.Rdata')
+load('../../data/linked_sel_div_changes.Rdata')
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("corbett_detig.pdf", width=7.3, height=7.3/1.9)
+
+par(mfrow=c(1, 2), mar=c(3, 3, 2, 1.1))
+# nf <- layout(matrix(c(1, 2), ncol=2))
+# layout.show(nf)
+
+plot(log10_diversity ~ log10_popsize, dca, pch=19, axes=FALSE, 
+     type='n',
+     ylim=c(-3, -1), xlim=c(5.5, 16))
+cd_div_fit <- lm(log10(diversity) ~ log10_popsize, dca_noself)
+abline(cd_div_fit,  col='gray42')
+points(log10_diversity ~ log10_popsize, dca, pch=19, 
+     col=phyla_cols[dca$phylum])
+phyla_cols_s <- phyla_cols[unique(dca_noself$phylum)]
+legend(8, -1, names(phyla_cols_s), 
+       fill=phyla_cols_s, bty='n', border=0, cex=0.6, ncol=2)
+
+xseq <- seq(6, 18, 2)
+axis(1, xseq, 
+     labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
+     padj = -0.9,
+     tck=-0.02,
+     cex.axis=0.8, line=0.3)
+yseq <- seq(-3, -1)
+axis(2, yseq, labels=yseq, cex.axis=0.8, line=0, 
+     tck=-0.02, hadj=0.2,
+     las=1)
+title(ylab="diversity",
+      xlab="approximate population size", line=1.8)
+
+plot(dml_full$log10_popsize, log10(dml_full$Ne_N_RHH_BGS), 
+     axes=FALSE, pch=19,
+     col=wes_palette("Darjeeling1")[3],
+     xlim=c(6, 18))
+
+points(dca_noself$log10_popsize, log10(1-dca_noself$impact_of_sel), pch=19, 
+       col=wes_palette("Darjeeling1")[1])
[…]

### 2021-02-08 notebooks/figures/damuth.r

@@ -1,0 +1,46 @@
+library(tidyverse)
+library(scales)
+
+opar <- par(no.readonly=TRUE)
+
+pdf("damuth.pdf", width=5, height=5)
+par(mar=c(5, 5, 2, 1))
+
+load('../../data/popsize_chains.Rdata')
+
+pars <- rstan:::extract(bayes_fit, pars=c('theta', 'gamma'))
+mean_slope <- mean(pars$theta)
+mean_intercept <- mean(pars$gamma)
+
+d <- read_tsv('../../data/damuth1987.tsv')  %>%
+       mutate(group=factor(group)) %>%
+       mutate(log10_mass_g = log10(mass_g),
+              log10_density = log10(density))
+
+ngroups <- length(unique(d$group))
+group_cols <- hue_pal()(ngroups)
+
+plot(d$log10_mass_g, d$log10_density, pch=19, 
+     type='n', axes=FALSE, xlab='',
+     ylab='')
+points(d$log10_mass_g, d$log10_density, pch=21, 
+       cex=1, lwd=0.4, bg=group_cols[as.factor(d$group)], col='white')
+abline(a=mean_intercept, b=mean_slope, lty=2, col='gray52')
+xseq <- seq(-6, 6, 2)
+yseq <- seq(-2, 12, 2)
+axis(1, xseq, labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
+     padj = -1,
+     tck=-0.01,
+     cex.axis=0.8, line=0.5)
+axis(2, yseq, 
+     las=1,
+     tck=-0.01, hadj=0.5,
+     labels=latex2exp::TeX(sprintf("$10^{%d}$", yseq)), 
+     cex.axis=0.8, line=0.5)
+legend(-6, 0, tools:::toTitleCase(levels(as.factor(d$group))), fill=group_cols,
+       bty='n', border=0, cex=0.6, ncol=2)
+title(xlab=latex2exp::TeX("mass (grams)"), line=2.2)
+title(ylab=latex2exp::TeX("population density (individuals per $km^2$)"),   
+      line=2.4)
+dev.off()
+par(opar)

### 2021-02-08 notebooks/figures/density_mass.r

@@ -1,0 +1,43 @@
+library(tidyverse)
+library(scales)
+
+opar <- par(no.readonly=TRUE)
+
+pdf("popsize_hists", width=5, height=5)
+par(mar=c(5, 5, 2, 1))
+
+load('../../data/post_pred_draws.Rdata')
+
+sps <- c('Homo sapiens', 'Aedes aegypti', 'Drosophila melanogaster',
+         'Balaenoptera bonaerensis', 'Canis latrans/lupus', 'Mus musculus')
+
+d <- post_pred_draws %>% 
+      mutate(mean_popsize = map_dbl(popsize, mean)) %>% 
+      arrange(mean_popsize)
+stopifnot(all(sps %in% d$species))
+
+x <- d %>% filter(species %in% sps)
+
+x %>% unnest() %>% ggplot(aes(x=popsize, fill=species)) +geom_histogram()
+
+nsps <- length(unique(x$species))
+sps_cols <- hue_pal()(nsps)
+
+
+hists <- lapply(x$popsize, function(x) hist(x, breaks=seq(0, 24, length.out=100), freq=TRUE))
+
+lapply(seq_along(hists), function(i) {
+  add = ifelse(i == 1, FALSE, TRUE)
+  plot(hists[[i]], add=add, col=alpha(sps_cols[i], 0.5), 
+       border=0, xlim=c(0, 25), axes=FALSE)
+})
+xseq <- seq(0, 25, 5)
+axis(1, xseq, labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), cex.axis=0.8, line=0.5)
+yseq <- seq(0, 1200, 400)
+axis(2, yseq, cex.axis=0.8, line=0.5)
+
+
+bins <- do.call(rbind, map(hists, 'density'))
+barplot(bins, border=0, col=alpha(sps_cols, 1))
+
+dev.off()

### 2021-02-08 notebooks/figures/div_popsize_post.r

@@ -1,0 +1,48 @@
+# div_popsize_chains.r -- 
+
+library(tidyverse)
+library(scales)
+library(latex2exp)
+library(HDInterval)
+
+
+source('plot_funs.r')
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+opar <- par(no.readonly=TRUE)
+
+# output <- FALSE
+output <- TRUE
+if (output) 
+  pdf('div_popsize_post.pdf', height=1, width=4)
+
+# Stan posteriors, OLS fits, and raw data
+load('../../data/diversity_popsize_chains.Rdata')
+
+
+phi <- rstan:::extract(dps_fit, pars='phi')$phi
+beta <- rstan:::extract(dps_fit, pars='beta')$beta
+sigma_r <- rstan:::extract(dps_fit, pars='sigma_r')$sigma_r 
+sigma_p <- rstan:::extract(dps_fit, pars='sigma_p')$sigma_p
+
+lambda_est(dps_fit)
+
+lambda <- sigma_p / (sigma_p + sigma_r)
+
+layout(matrix(1:3, byrow=TRUE, nrow=1, ncol=3))
+par(mar=c(2, 1, 2, 1), oma=rep(0, 4))
+
+post_plot(phi, main=latex2exp::TeX('intercept ($\\phi$)'), 
+          bs_ci=dps_bs_cis[1, ])
+text(-3.7, 1.23, 'OLS', cex=0.9, xpd=TRUE)
+post_plot(beta, main=latex2exp::TeX('slope ($\\beta$)'), 
+          bs_ci=dps_bs_cis[2, ])
+text(0.125, 20.3, 'OLS', cex=0.9)
+post_plot(lambda, main=latex2exp::TeX('phylo. signal ($\\lambda$)'),
+           xlim=c(0, 1))
+
+if (output) 
+  dev.off()
+
+

### 2021-02-08 notebooks/figures/div_popsize_post_phyla.r

@@ -1,0 +1,103 @@
+# div_popsize_chains_groups.r -- 
+library(tidyverse)
+library(scales)
+library(latex2exp)
+library(HDInterval)
+
+
+source('plot_funs.r')
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+opar <- par(no.readonly=TRUE)
+
+# output <- FALSE
+output <- TRUE
+if (output) 
+  pdf('div_popsize_post_phyla.pdf', height=2, width=4)
+
+load('../../data/phyla_diversity_popsize_chains.Rdata')
+
+# we use the separate Stan fit for chordates
+# chor_fit <- dt_dps_phyla$fit[[1]]
+mols_fit <- dt_dps_phyla$fit[[2]]
+arth_fit <- dt_dps_phyla$fit[[3]]
+
+a_phi <- rstan:::extract(arth_fit, pars='phi')$phi
+a_beta <- rstan:::extract(arth_fit, pars='beta')$beta
+
+a_sigma_r <- rstan:::extract(arth_fit, pars='sigma_r')$sigma_r 
+a_sigma_p <- rstan:::extract(arth_fit, pars='sigma_p')$sigma_p
+
+c_phi <- rstan:::extract(chor_fit, pars='phi')$phi
+c_beta <- rstan:::extract(chor_fit, pars='beta')$beta
+# no residual error
+# c_sigma_r <- rstan:::extract(chor_fit, pars='sigma_r')$sigma_r 
+c_sigma_p <- rstan:::extract(chor_fit, pars='sigma_p')$sigma_p
+
+m_phi <- rstan:::extract(mols_fit, pars='phi')$phi
+m_beta <- rstan:::extract(mols_fit, pars='beta')$beta
+m_sigma_r <- rstan:::extract(mols_fit, pars='sigma_r')$sigma_r 
+m_sigma_p <- rstan:::extract(mols_fit, pars='sigma_p')$sigma_p
+
+
+
+a_lambda <- a_sigma_p / (a_sigma_p + a_sigma_r)
+# c_lambda <- c_sigma_p / (c_sigma_p + c_sigma_r)
+m_lambda <- m_sigma_p / (m_sigma_p + m_sigma_r)
+
+layout(matrix(1:9, byrow=TRUE, nrow=3, ncol=3),
+       height=c(1, 1, 1.1))
+par(mar=c(1.2, 1.4, 1.3, 0.1), oma=c(0.5, 1, 1, 1))
+
[…]

### 2021-02-08 notebooks/figures/diversity_body_size.r

@@ -1,0 +1,82 @@
+# diversity_body_size.r -- create the body size / diversity figure
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(scales)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+
+opar <- par(no.readonly=TRUE)
+
+lds <- read_tsv('../../data/leffler_et_al_2012_updated.tsv') %>%
+         mutate(log10_size=log10(size), 
+                log10_diversity=log10(diversity))
+
+set.seed(2)
+p <- ggplot(lds) + 
+  geom_point(aes(log10_size, log10_diversity, label=species, color=phylum), size=2) +
+  geom_text_repel(aes(log10_size, log10_diversity, label=species, text=species), 
+                  size=1.5, 
+                  point.padding=0.29, 
+                  #box.padding=0.2, 
+                  max.iter=10000)
+
+repel_pos <- extract_ggrepel(p)
+
+# best to restart after messing with grid 
+dev.off()
+
+
+clean_species <- function(x) {
+  cl <- sub('([^/]+)/.*', '\\1', x)
+  sub('([^ ]+) ([^ ]+).*', '\\1 \\2', cl)
+}
+
+ld <- lds %>% filter(!is.na(log10_size), !is.na(log10_diversity)) %>%
+         # simplify species names, remove the species1/species2 in Leffler
+         mutate(species = map_chr(species, clean_species))
+
+
+nphyla <- length(unique(ld$phylum))
+phyla_cols <- hue_pal()(nphyla)
+
+
+pdf("diversity_body_size.pdf", width=7.3, height=7.3/1.5)
+par(mar=c(4, 4, 2, 1))
+x <- ld$log10_size
+y <- ld$log10_diversity
+tx <- repel_pos[, 1]
+ty <- repel_pos[, 2]
+plot(x, y, col=phyla_cols[as.factor(ld$phylum)], type='n', axes=FALSE,
+     ylab='', xlab='', xlim=c(-5.3, 2))
+adjust <- 0
+segments(x, y, tx, ty + adjust*ifelse(ty > y, -0.02, 0.02), col=alpha('gray42', 0.5), lwd=0.5)
+points(x, y, pch=21, cex=1.3, lwd=0.4,
[…]

### 2021-02-08 notebooks/figures/diversity_popsize.r

@@ -1,0 +1,174 @@
+# diversity_popsize.r -- 
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+opar <- par(no.readonly=TRUE)
+
+# for div_fit
+load('../../data/lognorm_fulldata_with_social.Rdata')
+
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+      filter(kingdom == 'Animalia') %>%
+      mutate(log10_diversity = log10(diversity)) %>% 
+      filter_at(vars(log10_diversity, pred_log10_N, log10_popsize), 
+                ~ is.finite(.))
+
+
+# summary(lm(pred_log10_N ~ log10_popsize, d))
+# summary(lm(log10_diversity ~ log10_popsize, d))
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$log10_diversity
+
+plot(x, y)
+
+X <- cbind(x, y)
+center <- colMeans(X)
+sigma <- cov(X)
+
+coords <- ellipse(mu=center, sigma=sigma, alpha=0.35, draw=FALSE)
+
+label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
+# add in all drosophila
+keep_sps <- c("Homo sapiens", "Apis mellifera", "Canis lupus")
+label <- (label | d$family == 'Drosophilidae') | label %in% keep_sps
+# contained <- rep(TRUE, length(x))
+
+d$label <- ifelse(label, d$species, "")
+
+set.seed(2)
+
+xlims <- c(4, 16.5)
+
+REPEL_CACHED_FILE <- 'repel.tsv'
+if (!file.exists(REPEL_CACHED_FILE)) {
+  p <- ggplot(d) + 
+    geom_point(aes(log10_popsize, log10_diversity, 
+                   label=species, color=phylum), size=2) +
+    geom_text_repel(data=d,
+                    mapping=aes(log10_popsize, log10_diversity, 
+                                label=label, text=species), 
[…]

### 2021-02-08 notebooks/figures/diversity_popsize_averages.r

@@ -1,0 +1,105 @@
+# diversity_popsize_averages.r -- 
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+opar <- par(no.readonly=TRUE)
+
+
+load('../../data/main_datasets.Rdata')
+load('../../data/phyla_diversity_popsize_ols.Rdata')
+
+# alias this -- less typing
+d <- da_dps
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$log10_diversity
+
+output <- TRUE
+
+if (output)
+  pdf("diversity_popsize_averages.pdf", width=7, height=7/1.5)
+
+par(mar=c(4, 4, 2, 2))
+
+xlims <- c(4, 18)
+has_label <- d$label != ""
+plot(x, y, col=phyla_cols[d$phylum], type='n', axes=FALSE,
+     ylab='', xlab='', ylim=c(-4.1, -0.9), xlim=xlims) 
+
+# text(6, -0.9, latex2exp:::TeX("$\\mu = 10^{-8}$"), cex=0.5, col='gray52', xpd=TRUE)
+# text(9, -0.9, latex2exp:::TeX("$\\mu = 10^{-10}$"), cex=0.5, col='gray52', xpd=TRUE)
+
+logN <- seq(3, 18, length.out=100)
+y1 <- log10(4e-8) + logN
+y2 <- log10(4e-10) + logN
+# polygon(c(logN, rev(logN)), c(y1, rev(y2)), border=NA, col='gray88')
+
+fits <- ols_phyla$ols
+names(fits) <- ols_phyla$phylum
+
+for (i in seq_along(fits)) {
+  lps <- ols_phyla$data[[i]]$log10_popsize
+  zz <- seq(min(lps), max(lps), length.out=100)
+  yy <- coef(fits[[i]])[1] + coef(fits[[i]])[2]*zz
+  lines(zz, yy, lwd=1.2, lty=2, 
+         col=phyla_cols[names(fits)[i]])
+}
+
+adjust <- 0
+xc <- x[has_label]
+yc <- y[has_label]
+points(x, y, pch=21, cex=1.3, lwd=0.4,
[…]

### 2021-02-08 notebooks/figures/diversity_popsize_full.r

@@ -1,0 +1,206 @@
+# diversity_popsize_full.r -- 
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+opar <- par(no.readonly=TRUE)
+
+# for stan fit
+load('../../data/diversity_popsize_chains.Rdata')
+load('../../data/main_datasets.Rdata')
+phi <- mean(rstan:::extract(dps_fit, pars='phi')$phi)
+beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
+
+# alias this -- less typing
+d <- da_dps
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$log10_diversity
+
+plot(x, y)
+
+X <- cbind(x, y)
+center <- colMeans(X)
+sigma <- cov(X)
+
+coords <- ellipse(mu=center, sigma=sigma, alpha=0.35, draw=FALSE)
+
+label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
+# add in all drosophila
+keep_sps <- c("Homo sapiens", "Apis mellifera", "Canis lupus")
+label <- (label | d$family == 'Drosophilidae') | label %in% keep_sps
+# contained <- rep(TRUE, length(x))
+
+d$label <- ifelse(label, d$species, "")
+
+# simplify some species name (Leffler data has combined them in some cases)
+d$label[d$species == 'Gorilla beringei/gorilla'] <- "Gorilla gorilla"
+d$label[d$species == 'Pongo abelii/pygmaeus'] <- "Pongo abelii"
+d$label[d$species == 'Nomascus gabriellae/leucogenys'] <- "Nomascus gabriellae"
+d$label[d$species == 'Aquila clanga/pomarina'] <- "Aquila clanga"
+
+set.seed(2)
+
+xlims <- c(4, 18)
+
+REPEL_CACHED_FILE <- 'repel_full.tsv'
+# FORCE <- TRUE
+FORCE <- FALSE
+if (FORCE || !file.exists(REPEL_CACHED_FILE)) {
+  p <- ggplot(d) + 
[…]

### 2021-02-08 notebooks/figures/diversity_popsize_linear.r

@@ -1,0 +1,130 @@
+# diversity_popsize_full.r -- 
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+opar <- par(no.readonly=TRUE)
+
+load('../../data/main_datasets.Rdata')
+d <- da_dps
+# summary(lm(pred_log10_N ~ log10_popsize, d))
+# summary(lm(log10_diversity ~ log10_popsize, d))
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$diversity
+
+set.seed(2)
+
+xlims <- c(4, 18)
+
+fit <- lm(y ~ x)
+
+output <- TRUE
+# output <- FALSE
+
+if (output)
+  pdf("diversity_popsize_linear.pdf", width=7, height=7/1.5)
+
+par(mar=c(4, 4, 2, 2))
+
+has_label <- d$label != ""
+plot(x, y, col=phyla_cols[d$phylum], type='n', axes=FALSE,
+     ylab='', xlab='', ylim=c(0, 0.09), xlim=xlims) 
+
+# abline(a=log10(1e-8 * 4), b=1, lty=2, col='cornflowerblue')
+# text(6, -0.9, latex2exp:::TeX("$\\mu = 10^{-8}$"), cex=0.5, col='cornflowerblue', xpd=TRUE)
+
+# logN <- seq(3, 16, length.out=100)
+# y1 <- log10(4e-8) + logN
+# y2 <- log10(4e-10) + logN
+# polygon(c(logN, rev(logN)), c(y1, rev(y2)), border=NA, col='gray88')
+
+endp <- 11
+logN <- seq(3, endp, length.out=100)
+# y1 <- log10(4e-8) + logN
+# y2 <- log10(4e-10) + logN
+# polygon(c(logN, rev(logN)), c(y1, rev(y2)), border=NA, col='gray88')
+
+# two allele
+Kalleles <- 4
+Kboth <- FALSE
+palpha <- 0.4
+if (Kalleles == 2 || Kboth) {
+  y1 <- 10^(log10(4e-8) + logN - log10(2 + 2 * 4e-8 * 10^logN))
[…]

### 2021-02-08 notebooks/figures/diversity_popsize_redlist.r

@@ -1,0 +1,125 @@
+# diversity_popsize_redlist.r -- 
+
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(wesanderson)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+
+opar <- par(no.readonly=TRUE)
+
+
+# for stan fit
+load('../../data/diversity_popsize_chains.Rdata')
+load('../../data/main_datasets.Rdata')
+phi <- mean(rstan:::extract(dps_fit, pars='phi')$phi)
+beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
+
+# alias this -- less typing
+d <- da_dps
+
+iucn_levels <- c("NA/DD", "LC", "NT", "VU", "EN", "CR")
+cols <- c("gray70", "#006666", "aquamarine3", "#cc9900", "#cc6633", "#cc3333")
+d$redlist_cat <- factor(d$redlist_cat, levels=iucn_levels)
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$log10_diversity
+
+xlims <- c(4, 18)
+plot(x, y)
+
+output <- TRUE
+# output <- FALSE
+
+if (output)
+  pdf("diversity_popsize_redlist.pdf", width=7, height=7/1.5)
+
+par(mar=c(4, 4, 3, 4), oma=c(0.5, 0.5, 0.5, 0.5))
+
+d$redlist_cat[is.na(d$redlist_cat)] <- "NA/DD"
+d$redlist_cat[d$redlist_cat == 'DD'] <- "NA/DD" # there's only two, just convert
+#iucn_cols <- setNames(c('gray80', wes_palette("Cavalcanti1")), iucn_levels) 
+iucn_cols <- setNames(cols, iucn_levels) 
+
+plot(x, y, col=iucn_cols[d$redlist_cat], type='n', axes=FALSE,
+     ylab='', xlab='', ylim=c(-4.1, 0), xlim=xlims) 
+
+adjust <- 0
+points(x, y, pch=21, cex=1.3, lwd=0.4,
+       bg=iucn_cols[d$redlist_cat], col='white')
+xseq <- seq(4, 18, 2)
+axis(1, xseq, 
[…]

### 2021-02-08 notebooks/figures/diversity_range_bodymass.r

@@ -1,0 +1,134 @@
+library(tidyverse)
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+load('../../data/main_datasets.Rdata')
+load('../../data/diversity_range_bodymass_chains.Rdata')
+
+d <- da_dps
+
+output <- TRUE
+# output <- FALSE
+
+if (output)
+  pdf("diversity_range_bodymass.pdf", width=7, height=3)
+
+
+par(mar=c(0.5, 0.5, 1, 0.5), oma=rep(0.4, 4))
+opar <- par(no.readonly=TRUE)
+layout(matrix(c(1, 2, 3, 4, 5, 6, 7,
+                8, 8, 8, 9, 10, 10, 10), 
+              byrow=TRUE, ncol=7, nrow=2),
+      height=c(0.3, 1), width=c(0.3, 0.3, 0.3, 0.1, 0.3, 0.3, 0.3))
+
+
+phi <- rstan:::extract(dbm_fit, pars='phi')$phi
+beta <- rstan:::extract(dbm_fit, pars='beta')$beta
+sigma_r <- rstan:::extract(dbm_fit, pars='sigma_r')$sigma_r 
+sigma_p <- rstan:::extract(dbm_fit, pars='sigma_p')$sigma_p
+
+lambda <- sigma_p / (sigma_p + sigma_r)
+
+post_plot(phi, main=latex2exp::TeX('intercept ($\\phi$)'))
+# text(-3.7, 1.23, 'OLS', cex=0.9)
+post_plot(beta, main=latex2exp::TeX('slope ($\\beta$)'))
+# text(0.140, 20.3, 'OLS', cex=0.9)
+post_plot(lambda, main=latex2exp::TeX('phylo. signal ($\\lambda$)'),
+           xlim=c(0, 1))
+
+par(mar=rep(0, 4))
+plot.new()
+
+
+par(mar=c(0.5, 0.5, 1, 0.5))
+phi <- rstan:::extract(drn_fit, pars='phi')$phi
+beta <- rstan:::extract(drn_fit, pars='beta')$beta
+sigma_r <- rstan:::extract(drn_fit, pars='sigma_r')$sigma_r 
+sigma_p <- rstan:::extract(drn_fit, pars='sigma_p')$sigma_p
+
+lambda <- sigma_p / (sigma_p + sigma_r)
+
[…]

### 2021-02-08 notebooks/figures/fecundity_body_size.r

@@ -1,0 +1,106 @@
+# fecundity_body_size.r -- create the fecundity / body size figure
+
+library(tidyverse)
+library(ggrepel)
+library(viridis)
+library(wesanderson)
+library(shape)
+library(rstan)
+source('../../R/utilities.r')
+
+arrow_col <- 'gray32'
+
+opar <- par(no.readonly=TRUE)
+load('../../data/body_size_fecundity.Rdata')
+
+pdf <- TRUE
+#pdf <- FALSE
+if (pdf) pdf("fecundity_body_size.pdf", width=7.3, height=7.3/1.9)
+
+par(mfrow=c(1, 2), mar=c(3.8, 3.6, 2.1, 1.4))
+
+
+with(dr_family_averages %>% filter(fecundity < 8e4),
+     plot(log10_size, log10_fecundity, 
+          axes=FALSE,
+          pch=19,
+          xlab='', ylab='',
+          xlim=c(-3, 1),
+          ylim=c(-3, 5),
+          col='gray42'))
+
+with(dr_phylum_averages %>% filter(fecundity < 8e4),
+     points(log10_size, log10_fecundity, 
+          pch=19,
+          col='cornflowerblue'))
+
+
+# with(dr,
+#      points(log10_size, log10_fecundity, 
+#           pch=19,
+#           cex=0.4,
+#           col='gray42'))
+
+
+axis(1, seq(-3, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-3, 1))))
+axis(2, seq(-3, 6, 2), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-3, 6, 2))))
+title(xlab='body size (meters)', ylab='fecundity', line=2.5)
+
+ci_polygon(lnf_xnew, y=extract(lnf, pars='y_rep')$y_rep,
+           color='gray42', type='quantile', transp=0.1)
+ci_polygon(lnf_xnew, y=extract(lnf, pars='mu_rep')$mu_rep,
+           color='gray42', type='hpdi')
+lnf_p <- colMeans(as_tibble(extract(lnf, pars='mu_rep')))
+lines(lnf_xnew, lnf_p, col=alpha('black', 0.4), lwd=2)
+
+mtext("A", 3, at=-3.22, cex=1.4, xpd=TRUE, font=2)
+
+legend(-3.15, -1.5, inset=0, legend=c('family averages', 'phylum averages'), 
+       fill=c('gray42', 'cornflowerblue'),
[…]

### 2021-02-08 notebooks/figures/figure_3.r

@@ -1,0 +1,277 @@
+library(viridis)
+library(ggrepel)
+library(grDevices)
+library(wesanderson)
+library(scales)
+library(tidyverse)
+library(shape)
+
+opar <- par(no.readonly=TRUE)
+
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+load('../../data/linked_sel_div_changes.Rdata')
+load('../../data/lognorm_popsize_chains.Rdata')
+load('../../data/diversity_popsize_chains.Rdata')
+load('../../data/main_datasets.Rdata')
+
+
+midpoint <- function(x) {
+  out <- lapply(strsplit(x, ','), function(y) {
+           as.numeric(gsub('(\\)|\\]|\\[|\\()', '', y))
+       })
+  rowMeans(do.call(rbind, out))
+}
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("figure_3.pdf", width=7.3, height=7.3/2.3)
+
+
+par(mfrow=c(1, 2), mar=c(3, 3, 2, 1), oma=c(0, 0, 0, 1))
+# nf <- layout(matrix(c(1, 2), ncol=2))
+# layout.show(nf)
+
+# p <- ggplot(dml)  + geom_point(aes(log10_popsize, log10_diversity, size=2, color=map_length)) + 
+#   geom_text_repel(aes(log10_popsize, log10_diversity, color=map_length, label=species)) + scale_color_viridis(option='C', direction=-1)
+# p
+# # repel_pos <- extract_ggrepel(p)
+
+
+x <- log10_popsize_seq
+pred  <- colMeans(mu_ln_fit)
+social <- da_ml$social
+
+da_ml <- da_ml %>% filter(phylum != 'Nematoda')
+
+plot(map_length ~ log10_popsize, da_ml, type='n', 
+     axes=FALSE,
+     ylim=c(0, 60),
+     xlim=c(6, 18), 
+     ylab="", 
+     xlab="")
+ci_polygon(x, fit=ln_fit, par='mu_rep', col=scales::alpha('gray88', 0.4))
+# ci_polygon(x, fit=ln_fit, par='mu_rep', color='gray32')
+# polygon(x=c(5, 8, 8, 5), y=c(60, 60, 63, 63),
[…]

### 2021-02-08 notebooks/figures/figure_3_alt.r

@@ -1,0 +1,265 @@
+library(viridis)
+library(ggrepel)
+library(grDevices)
+library(wesanderson)
+library(scales)
+library(tidyverse)
+library(shape)
+
+opar <- par(no.readonly=TRUE)
+
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+load('../../data/linked_sel_div_changes.Rdata')
+load('../../data/lognorm_popsize_chains.Rdata')
+load('../../data/diversity_popsize_chains.Rdata')
+load('../../data/main_datasets.Rdata')
+
+
+midpoint <- function(x) {
+  out <- lapply(strsplit(x, ','), function(y) {
+           as.numeric(gsub('(\\)|\\]|\\[|\\()', '', y))
+       })
+  rowMeans(do.call(rbind, out))
+}
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("figure_3_alt.pdf", width=7.3, height=7.3/2.3)
+
+par(mfrow=c(1, 2), mar=c(3, 3, 2, 1), oma=c(0, 0, 0, 1))
+
+## subfigure A
+
+
+# I = 1 - pibar/p0
+# pibar / p0 = I - 1
+dml %>% filter(species == 'Drosophila melanogaster') %>%
+  select(impact_of_sel, Ne_N_CD15) %>% 
+  mutate(red = 1-impact_of_sel)
+
+# # if we wantd to use the Elyashiv et al data
+# dmel <- 'Drosophila melanogaster'
+# dml[dml$species == dmel, ]$Ne_N_CD15 <- 0.77
+
+# dml[dml$species == dmel, ]$log10_CD15_diversity <- log10(dml$diversity[dml$species == dmel]  / dml$Ne_N_CD15[dml$species == dmel])
+
+four_alleles <- function(N, mu) {
+  theta <- 4*N*mu
+  theta / (1 + (4*theta)/3)
+}
+
+dml <- dml_full %>% 
+  mutate(Ne_RHH_BGS_strongsel = 10^log10_popsize * Ne_N_RHH_BGS_strongsel) %>%
+  mutate(Ne_RHH_BGS_strongsel_strongBGS = 10^log10_popsize * Ne_N_RHH_BGS_strongsel_strongBGS) %>%
[…]

### 2021-02-08 notebooks/figures/genome_size_maplength.r

@@ -1,0 +1,159 @@
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+
+opar <- par(no.readonly=TRUE)
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+      mutate(log10_genome_size = log10(genome_size),
+             log10_map_length = log10(map_length)) %>% 
+      filter_at(vars(log10_genome_size, map_length), 
+                ~ is.finite(.)) %>% 
+      filter(species != "Ambystoma tigrinum")
+
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_genome_size
+y <- d$log10_map_length
+
+plot(x, y)
+
+X <- cbind(x, y)
+center <- colMeans(X)
+sigma <- cov(X)
+coords <- ellipse(mu=center, sigma=sigma, alpha=0.4, draw=FALSE)
+
+label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
+# contained <- rep(TRUE, length(x))
+
+d$label <- ifelse(label, d$species, "")
+
+set.seed(2)
+
+xlims <- c(-1.2, 1)
+
+REPEL_CACHED_FILE <- 'repel_genome_size_maplength.tsv'
+if (TRUE || !file.exists(REPEL_CACHED_FILE)) {
+  p <- ggplot(d) + 
+    geom_point(aes(log10_genome_size, log10_map_length, 
+                   color=phylum), size=2) +
+    geom_text_repel(data=d,
+                    mapping=aes(log10_genome_size, 
+                                log10_map_length, 
+                                label=label),
+                                size=1.5, 
+                                point.padding=0.6, 
+                                force=2,
+                                min.segment.length=0.4,
+                                seed=1,
+                                box.padding=0.1, 
[…]

### 2021-02-08 notebooks/figures/genome_size_popsize.r

@@ -1,0 +1,162 @@
+## NOTE:
+# if for some reason repel isn't working, the xlim is likely
+# messedup
+library(tidyverse)
+library(plotrix)
+library(ggrepel)
+library(mixtools)
+library(scales)
+library(sp)
+library(grid)
+library(latex2exp)
+
+source('plot_funs.r')
+source('color_scheme.r')
+
+opar <- par(no.readonly=TRUE)
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+      mutate(log10_genome_size = log10(genome_size)) %>% 
+      filter_at(vars(log10_genome_size, log10_popsize), 
+                ~ is.finite(.)) %>% 
+      filter(species != "Ambystoma tigrinum")
+
+
+# what to include
+# x <- d$pred_log10_N  # the simplistic unregularized version
+x <- d$log10_popsize
+y <- d$log10_genome_size
+
+plot(x, y)
+
+X <- cbind(x, y)
+center <- colMeans(X)
+sigma <- cov(X)
+coords <- ellipse(mu=center, sigma=sigma, alpha=0.3,
+                  draw=FALSE)
+
+label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
+# contained <- rep(TRUE, length(x))
+
+d$label <- ifelse(label, d$species, "")
+
+set.seed(2)
+
+xlims <- c(3.9, 18)
+
+REPEL_CACHED_FILE <- 'repel_genome_size_popsize.tsv'
+# FORCE <- TRUE
+FORCE <- FALSE
+if (FORCE || !file.exists(REPEL_CACHED_FILE)) {
+  p <- ggplot(d) + 
+    geom_point(aes(log10_popsize, log10_genome_size, 
+                   label=species, color=phylum), size=2) +
+    geom_text_repel(data=d,
+                    mapping=aes(log10_popsize, log10_genome_size, 
+                                label=label, text=species),
+                                size=1.5, 
+                                point.padding=0.6, 
+                                force=2,
+                                min.segment.length=0.4,
+                                seed=1,
[…]

### 2021-02-08 notebooks/figures/linked_sel_params.r

@@ -1,0 +1,112 @@
+library(scales)
+library(tidyverse)
+library(shape)
+library(ggrepel)
+
+
+opar <- par(no.readonly=TRUE)
+
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+dests <- read_tsv('../../data/sweep_param_ests.tsv')
+
+dshap_raw <- read_tsv('../../data/shapiro_et_al_2007.tsv', col_names=FALSE)
+
+# unsure what all the columns are, but this is what matters
+dshap <- tibble(rec = dshap_raw$X3, pi = dshap_raw$X4)
+
+
+hh_nls <- nls(pi ~ neutral.pi*rec/(rec + alpha), dshap, 
+              start=list(neutral.pi=2, alpha=1e-9))
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("linked_sel_params.pdf", width=7.3, height=7.3/2.3)
+
+par(mfrow=c(1, 2), mar=c(1, 1, 1, 1), oma=c(1, 1, 1, 1))
+
+plot(log10(dests$nu), rep(1, nrow(dests)), xlim=c(-12, -8), 
+    ylab='', xlab='', pch=19, axes=FALSE, 
+    col=ifelse(dests$study == '***', 'red', 'black'))
+
+xlabs <- seq(-12, -8)
+axis(1, xlabs, line=-5.95,
+     cex.axis=0.7,
+     tck=-0.02,
+     padj=-1.2,
+     labels=latex2exp::TeX(sprintf("$10^{%d}$", xlabs)))
+mtext(latex2exp::TeX("sweeps per basepair, per generation ($\\nu_{BP})"),
+      1, line=-3.4)
+mtext("A", 3, at=-12, line=-0, cex=1.4, font=2)
+
+fs <- 0.5
+lwd <- 0.8
+lcol <- 'gray42'
+text(-11.5, 1.05, "Macpherson et al. (2007)", cex=fs)
+segments(-11.5, 1.03, -11.45, 1, lwd=lwd, col=lcol)
+# text(-10.8, 1.1, "***", cex=fs)
+# segments(-10.8, 1.08, -10.61, 1, lwd=lwd, col=lcol)
+text(-9.8, 1.05, "Jensen et al. (2008)", cex=fs)
+segments(-10.4, 1.08, -10.4, 1, lwd=lwd, col=lcol)
+text(-10, 1.1, "Li and Stephan (2006)", cex=fs)
[…]

### 2021-02-08 notebooks/figures/map_length_body_size.r

@@ -1,0 +1,140 @@
+# map_length_body_size.r -- create the body size / map length figure
+
+library(tidyverse)
+library(ggrepel)
+library(viridis)
+library(wesanderson)
+library(shape)
+
+
+load('../../data/body_size_map_length.Rdata')
+opar <- par(no.readonly=TRUE)
+arrow_col <- 'gray32'
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("map_length_body_size.pdf", width=7.3, height=7.3/1.9)
+
+par(mfrow=c(1, 2), mar=c(3.8, 3.6, 2.1, 1.4))
+
+with(family_mean, plot(map_length, log10_size, pch=19, axes=FALSE, cex=0.7,
+                       ylab='', xlab='', xlim=c(0, 52), ylim=c(-4.3, 1.1), col='gray48'))
+
+axis(1, seq(0, 50, 10))
+axis(2, seq(-4, 1), seq(-4, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-4, 1))))
+
+with(bs_p_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
+                    col=scales::alpha('cornflowerblue', 0.2), lty=0))
+
+with(bs_f_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
+                    col=scales::alpha('gray42', 0.2), lty=0))
+
+with(phyla_mean, points(map_length, log10_size, pch=19, col='cornflowerblue', cex=0.7))
+
+lines(sf, predict(mf2, newdata=tibble(map_length=sf)), lty='dashed', col='gray32')
+lines(sp, predict(mp2, newdata=tibble(map_length=sp)), lty='dashed', col='cornflowerblue')
+
+
+legend(27, -3, inset=0, legend=c('family averages', 'phylum averages'), 
+       fill=c('gray42', 'cornflowerblue'),
+			 title='',
+			 adj=c(0, 0.45),
+			 cex=0.7,
+			 bty='n', text.col='gray24',
+			 border=0)
+
+labels <- with(family_mean, identify(map_length, log10_size, family))
+
+text(12.4, -3.83, "Platyhelminthes", col='cornflowerblue', cex=0.4, adj=0, font=3)
+text(4.4, -4.10, "Diplogasteridae", col='gray42', cex=0.4, adj=0, font=3)
+
[…]

### 2021-02-08 notebooks/figures/map_length_body_size_bayesian.r

@@ -1,0 +1,207 @@
+# map_length_body_size.r -- create the body size / map length figure
+
+library(tidyverse)
+library(ggrepel)
+library(viridis)
+library(grDevices)
+library(wesanderson)
+library(shape)
+library(rstan)
+source('../../R/utilities.r')
+
+
+load('../../data/body_size_map_length_bayesian.Rdata')
+opar <- par(no.readonly=TRUE)
+arrow_col <- 'gray32'
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("map_length_body_size_bayesian.pdf", width=7.3, height=7.3/1.9)
+
+par(mfrow=c(1, 2), mar=c(3.8, 3.6, 2.1, 1.4))
+
+with(family_mean, plot(map_length, log10_size, pch=19, axes=FALSE, cex=0.7,
+                       ylab='', xlab='', xlim=c(0, 52), ylim=c(-4.3, 1.1), 
+                       type='n', col='gray48'))
+
+lines(x_new, colMeans(extract(bmf3, 'mu_rep')$mu_rep), 
+      lty='solid', col=scales::alpha('red', 0.8), lwd=1, lend='butt')
+
+with(family_mean, points(map_length, log10_size, pch=19, cex=0.7,
+                        ylab='', xlab='', xlim=c(0, 52), ylim=c(-4.3, 1.1), col='gray48'))
+
+
+axis(1, seq(0, 50, 10))
+axis(2, seq(-4, 1), seq(-4, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-4, 1))))
+
+ci_polygon(x_new, bmf1, color='gray42', type='hpdi', 
+           transp=0.2, par='mu_rep')
+lines(x_new, colMeans(extract(bmf1, 'mu_rep')$mu_rep), 
+      lty='dashed', col='gray42')
+
+with(phyla_mean, points(map_length, log10_size, pch=19, col='cornflowerblue', cex=0.7))
+ci_polygon(x_new, bmp1, color='cornflowerblue', type='hpdi', 
+           transp=0.2, par='mu_rep')
+lines(x_new, colMeans(extract(bmp1, 'mu_rep')$mu_rep), 
+      lty='dashed', col='cornflowerblue')
+
+# PPD
+#ci_polygon(x_new, bmf1, color='gray42', type='quantile', 
+#           transp=0.1, par='y_rep')
[…]

### 2021-02-08 notebooks/figures/map_length_body_size_supp.r

@@ -1,0 +1,56 @@
+# map_length_body_size.r -- create the body size / map length figure
+
+library(tidyverse)
+library(ggrepel)
+library(viridis)
+library(wesanderson)
+library(shape)
+
+
+load('../../data/body_size_map_length.Rdata')
+opar <- par(no.readonly=TRUE)
+arrow_col <- 'gray32'
+
+pdf <- TRUE
+# pdf <- FALSE
+if (pdf) pdf("map_length_body_size_nls.pdf", width=7.3, height=7.3)
+
+
+#par(mfrow=c(1, 1), mar=c(3.8, 3.6, 2.1, 1.4))
+
+with(family_mean, plot(map_length, log10_size, pch=19, axes=FALSE, cex=0.7,
+                       ylab='', xlab='', xlim=c(0, 52), ylim=c(-4.3, 1.1), col='gray48'))
+
+axis(1, seq(0, 50, 10))
+axis(2, seq(-4, 1), seq(-4, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-4, 1))))
+
+with(bs_p_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
+                    col=scales::alpha('cornflowerblue', 0.2), lty=0))
+
+with(bs_f_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
+                    col=scales::alpha('gray42', 0.2), lty=0))
+
+with(phyla_mean, points(map_length, log10_size, pch=19, col='cornflowerblue', cex=0.7))
+
+lines(sf, predict(mf2, newdata=tibble(map_length=sf)), lty='dashed', col='gray32')
+lines(sp, predict(mp2, newdata=tibble(map_length=sp)), lty='dashed', col='cornflowerblue')
+
+
+legend(31, -3.5, inset=0, legend=c('family averages', 'phylum averages'), 
+       fill=c('gray42', 'cornflowerblue'),
+			 title='',
+			 adj=c(0, 0.45),
+			 cex=1.2,
+			 bty='n', text.col='gray24',
+			 border=0)
+
+labels <- with(family_mean, identify(map_length, log10_size, family))
+
+text(12.4, -3.83, "Platyhelminthes", col='cornflowerblue', cex=0.8, adj=0, font=3)
+text(4.4, -4.10, "Diplogasteridae", col='gray42', cex=0.8, adj=0, font=3)
+
[…]

### 2021-02-08 notebooks/figures/node_heights.r

@@ -1,0 +1,101 @@
+library(viridis)
+library(tidyverse)
+library(phytools)
+library(scales)
+
+# note: why not two-tail? Here I don't think it's appropriate since
+# I use the abs. value of the contrasts
+
+load('../../data/nodeheights.Rdata')
+# for EDA at end (to dig around and find what the outlier nodes are)
+load('../../data/main_datasets.Rdata')
+source('color_scheme.r')
+source('../../R/utilities.r')
+source('../../R/node_height.r')
+
+# output <- FALSE
+output <- TRUE
+if (output) 
+  pdf('node_heights.pdf', height=3, width=4)
+
+
+
+layout(matrix(c(1, 2, 3, 
+                1, 4, 5, 
+                1, 6, 6), byrow=TRUE, ncol=3, nrow=3), 
+       width=c(0.25, 1, 1), height=c(1, 1, 0.25))
+
+par(mar=c(1.3, 1.3, 1.2, 1), oma=c(0.1, 0.1, 0.1, 0.1))
+plot.new()
+title(ylab='absolute value of contrasts', line=-1)
+
+plot_nh(dvnh, 'diversity', ylim=c(0, 0.006), 
+        xlab='', ylab='')
+plot_nh(psnh, 'pop size', ylim=c(0, 1.1), 
+        ylab='', 
+        xlab='')
+plot_nh(bmnh, 'body mass', xlab='', ylab='', 
+        ylim=c(0, 0.4))
+plot_nh(rnnh, 'range', ylim=c(0, 1), 
+        ylab='', 
+        xlab='')
+
+# what's excluded?
+dvnh$d  %>% filter(pic > 0.01) %>% pull(tips)
+dvnh$d  %>% filter(pic > 0.002) %>% pull(tips)
+psnh$d  %>% filter(pic > 2) %>% pull(tips)
+rnnh$d  %>% filter(pic > 1) %>% pull(tips)
+tips <- bmnh$d  %>% filter(pic > 0.5) %>% pull(tips)
+
+# these all look reasonable -- just lots of variability
+dt_dps %>% filter(species %in% tips[[1]]) %>%
+  select(species, size, log10_body_mass)
+
+
+# extra eda
+dvnh$d  %>% filter(pic > 0.002) %>% pull(tips)
+
+# now, looking right to left, what are the outliers in the 
[…]

### 2021-02-08 notebooks/figures/plot_funs.r

@@ -1,0 +1,31 @@
+library(ggrepel)
+library(scales)
+library(grid)
+
+
+extract_ggrepel <- function(p) {
+  # source: https://stackoverflow.com/questions/45065567/getting-coordinates-for-the-label-locations-from-ggrepel
+  # Get x and y plot ranges 
+  ggp <- ggplot_build(p)
+  xrg <- ggp$layout$panel_params[[1]]$x.range
+  yrg <- ggp$layout$panel_params[[1]]$y.range
+  print(p)
+  grid.force()
+  nodes <- grid.get("textrepeltree", grep = TRUE)
+  kids <- childNames(nodes)
+
+  # Function: get the x and y positions of a single ggrepel label
+  get.xy.pos.labs <- function(n) {
+    grid.force()
+    grb <- grid.get(n)
+    data.frame(
+      x = xrg[1]+diff(xrg)*convertX(grb$x, "native", valueOnly = TRUE),
+      y = yrg[1]+diff(yrg)*convertY(grb$y, "native", valueOnly = TRUE)
+    )
+  }
+  # Get positions of all ggrepel labels
+  dts <- do.call(rbind, lapply(kids, get.xy.pos.labs))
+  dts
+}
+
+

### 2021-02-08 notebooks/figures/popsize_maplength.r

@@ -1,0 +1,9 @@
+library(tidyverse)
+library(scales)
+
+opar <- par(no.readonly=TRUE)
+
+pdf("popsize_maplength.r", width=5, height=5)
+par(mar=c(5, 5, 2, 1))
+
+

### 2021-02-08 notebooks/figures/popsize_tree.r

@@ -1,0 +1,79 @@
+library(viridis)
+library(tidyverse)
+library(phytools)
+
+load('../../data/biconmap_data.Rdata')
+load('../../data/main_datasets.Rdata')
+source('color_scheme.r')
+source('../../R/utilities.r')
+
+# output <- FALSE
+output <- TRUE
+if (output) 
+  pdf('popsize_tree.pdf', height=5, width=4)
+
+par(oma=c(0.2, 0.2, 0.2, 0.2), mar=c(1, 1, 1, 1))
+vcols <- viridis(100)
+vcols2 <- viridis(100, option='A')
+layout(matrix(c(1, 2, 3, 1, 2, 3), byrow=TRUE, ncol=3, nrow=2),  
+       width=c(1, 0.05, 1))
+pscm2 <- setMap(pscm, colors=vcols)
+plot(pscm2, fsize=c(0, 0.8), ftype=c("off","reg"), sig=1, lwd=1.4,
+     outline=FALSE, mar=c(1.1,0.1,1.1,0.01), legend=FALSE)
+add.color.bar.log(500, vcols, prompt=FALSE, outline=FALSE, 
+              subtitle="",
+              x=10, digits=0,
+              lims=range(dt_dps$log10_popsize), lwd=4,
+              title='pop size')
+title(main = 'pop size', font.main = 1, cex.main=1.5)
+
+plot.new()
+width <- 0.5
+plot.window(xlim=c(0, width), ylim=c(1, length(tr_dps$tip.label)))
+
+# this is dumb... it shouldn't be this hard, but it is.
+tmp <- tibble(species = tr_dps$tip.label) %>%
+  left_join(dt_dps %>% select(species, phylum))  
+
+# get rectangle coords
+tmp2 <- tmp %>%
+  group_by(phylum) %>% summarize(n=n()) %>% mutate(left=0) 
+
+# order by tip phyla
+# get in the same order as data (grrr...)
+order_phy <- unique(tmp$phylum)
+tmp2 <- tmp2[match(order_phy, tmp2$phylum), ] 
+
+tmp2 <- tmp2 %>% mutate(cumsum = cumsum(n), left = lag(cumsum, default=0)) 
+rect_z <- cbind(left=tmp2$left, right=tmp2$cumsum)
+rownames(rect_z) <- tmp2$phylum
+
+
+offset <- 0.5
+for (i in 1:nrow(rect_z)) {
[…]

### 2021-02-08 notebooks/figures/redlist_ranges.r

@@ -1,0 +1,41 @@
+library(tidyverse)
+
+opar <- par(no.readonly=TRUE)
+drl <- read_tsv('../../data/redlist_popsizes.tsv') %>%
+        mutate(log10_range = log10(range), log10_eoo_km2=log10(eoo_km2))
+
+pdf('iucn_redlist_ranges.pdf', width=5, height=5)
+
+fit <- lm(log10_range ~ log10_eoo_km2, data=drl)
+par(mar=c(5, 5, 2, 2))
+plot(log10_range ~ log10_eoo_km2, data=drl, pch=19,
+     axes=FALSE, ylab='',
+     xlab='')
+xlabs <- seq(4, 8)
+abline(b=1, a=0, col='gray42', lty=2)
+abline(fit)
+axis(1, xlabs, labels=latex2exp::TeX(sprintf("$10^{%d}$", xlabs)),
+     padj = -1,
+     tck=-0.01,
+     cex.axis=0.8, line=0.5)
+axis(2, xlabs, labels=latex2exp::TeX(sprintf("$10^{%d}$", xlabs)),
+     las=1,
+     tck=-0.01, hadj=0.5,
+     cex.axis=0.8, line=0.5)
+corr <- cor(drl$log10_eoo_km2, drl$log10_range, use='pairwise.complete')
+corpval <- cor.test(drl$log10_eoo_km2, drl$log10_range, use='pairwise.complete')$p.value
+xpos <- 6.5
+ypos <- 4.5
+text(xpos, ypos + .2, latex2exp::TeX(sprintf('$\\rho = %.2f$', round(corr, 2))), 
+     cex=0.8, adj=0)
+
+mag <- round(log10(corpval)) - 1
+mantissa <- corpval * 10^abs(mag)
+text(xpos, ypos, latex2exp::TeX(sprintf('p-value = $%.2f \\times 10^{%d}$', 
+                                  mantissa, mag)), cex=0.8, adj=0)
+
+title(ylab=latex2exp:::TeX('estimated range ($km^2$)'), line=2.8) 
+title(xlab=latex2exp:::TeX('IUCN Red List estimated extent of occurrence ($km^2$)'),
+      line=2.4) 
+dev.off()
+par(opar)

### 2021-02-08 notebooks/figures/species_ranges.r

@@ -1,0 +1,55 @@
+library(tidyverse)
+library(wesanderson)
+library(RColorBrewer)
+
+range_cat_levels <- c("Island", "Narrow endemic", "Broad endemic", "Cosmopolitan")
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+       filter(is.finite(range), !is.na(range_cat)) %>%
+       mutate(log10_range = log10(range)) %>%
+       mutate(range_cat = ifelse(range_cat == 'Unknown', NA, range_cat)) %>%
+       mutate(range_cat = factor(range_cat, levels=range_cat_levels, ordered=TRUE)) %>%
+       arrange(range_cat, range)  %>% 
+       # space the groups
+       mutate(xpos = 1:nrow(.) + 1.2*as.integer(range_cat)-1)
+
+
+
+opar <- par(no.readonly=TRUE)
+# pdf('range_descriptions.pdf', width=8, height=5)
+pdf('range_categories.pdf', width=5, height=8)
+
+par(mar=c(4, 1, 0, 5))
+cols <- wes_palette('Darjeeling1')[c(5, 2, 3, 1, 4)]
+plot(d$log10_range, d$xpos, pch = 19, col=cols[d$range_cat], 
+     xlab = '',
+     ylab = '', axes=FALSE, xlim=c(2, 9))
+ylabs = seq(2, 8, 2)
+axis(1, ylabs, labels=latex2exp::TeX(sprintf("$10^{%d}$", ylabs)),
+     cex.axis=0.8,
+     line=0.5,
+     tck=-0.01, padj=-1)
+text(y=d$xpos, x=8.8, labels=d$species, 
+     col=cols[d$range_cat], xpd=TRUE, 
+     las=1, srt=0, adj=0, cex=0.3)
+legend(1.9, 126, range_cat_levels, fill=cols, 
+       bty='n', box.lwd=0, border=0, cex=1)
+
+title(xlab=latex2exp::TeX('estimated range ($km^2$)'), line=2.1)
+
+# par(mar=c(8, 5, 2, 1))
+# cols <- wes_palette('Darjeeling1')[c(5, 2, 3, 1, 4)]
+# plot(d$xpos, d$log10_range, pch = 19, col=cols[d$range_cat], 
+#      ylab = latex2exp::TeX('estimated range ($km^2$)'),
[…]

### 2021-02-08 notebooks/figures/species_ranges_long.r

@@ -1,0 +1,55 @@
+library(tidyverse)
+library(wesanderson)
+library(RColorBrewer)
+
+range_cat_levels <- c("Island", "Narrow endemic", "Broad endemic", "Cosmopolitan")
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+       filter(is.finite(range), !is.na(range_cat)) %>%
+       mutate(log10_range = log10(range)) %>%
+       mutate(range_cat = ifelse(range_cat == 'Unknown', NA, range_cat)) %>%
+       mutate(range_cat = factor(range_cat, levels=range_cat_levels, ordered=TRUE)) %>%
+       arrange(range_cat, range)  %>% 
+       # space the groups
+       mutate(xpos = 1:nrow(.) + 0.1*as.integer(range_cat)-1)
+
+
+
+opar <- par(no.readonly=TRUE)
+# pdf('range_descriptions.pdf', width=8, height=5)
+pdf('range_categories_long.pdf', height=5, width=8)
+
+par(mar=c(4, 3, 0, 5))
+cols <- wes_palette('Darjeeling1')[c(5, 2, 3, 1, 4)]
+plot(d$xpos, d$log10_range, pch = 19, col=cols[d$range_cat], 
+     xlab = '', ylim=c(2, 8.9),
+     ylab = '', axes=FALSE) 
+ylabs = seq(2, 8, 2)
+axis(2, ylabs, labels=latex2exp::TeX(sprintf("$10^{%d}$", ylabs)),
+     cex.axis=0.8, las=1,
+     tck=-0.01, hadj=0.5)
+text(x=d$xpos, y=2, labels=d$species, 
+     col=cols[d$range_cat], xpd=TRUE, 
+     las=1,
+     srt=90, adj=1, cex=0.4)
+legend(70, 4, range_cat_levels, fill=cols, 
+       bty='n', box.lwd=0, border=0, cex=1)
+
+title(ylab=latex2exp::TeX('estimated range ($km^2$)'), line=1.7)
+
+# par(mar=c(8, 5, 2, 1))
+# cols <- wes_palette('Darjeeling1')[c(5, 2, 3, 1, 4)]
+# plot(d$xpos, d$log10_range, pch = 19, col=cols[d$range_cat], 
+#      ylab = latex2exp::TeX('estimated range ($km^2$)'),
[…]

### 2021-02-08 notebooks/render_notebook.r

@@ -1,0 +1,3 @@
+library(rmarkdown)
+
+render('main_analysis.Rmd')

### 2021-02-16 manuscript/manuscript.tex
minor corrections to ms, thanks matt!

@@
 
+% Neutral theory predicts that genetic diversity increases with population size,
+% yet observed levels of diversity across species vary only two orders of
+% magnitude while population sizes likely vary over several. The causes of this
+% discrepancy, known as Lewontin's Paradox of Variation, remain unknown. Here I
+% revisit Lewontin's Paradox by quantifying the relationship between pairwise
+% diversity and approximate census size for 172 metazoan species. Using
+% phylogenetic comparative methods, I show this relationship is significant
+% accounting for phylogeny, but has high phylogenetic signal and some lineages
+% experience shifts in diversity deep in the past. I find a negative relationship
+% between recombination map length and census size, suggesting abundant species
+% have less recombination and could experience greater reductions in diversity
+% due to linked selection. However, I show that even using strong selection
+% parameter estimates, models of linked selection are unlikely to explain the
+% observed relationship between diversity and census sizes across species.
+
 \begin{abstract} 
-
-  Under neutral theory, the level of polymorphism in an equilibrium population
+  Under neutral theory, genetic diversity in an equilibrium population
   is expected to increase with population size. However, observed levels of
@@
 broader literature on the empirical relationship between diversity and
-population size \parencite{Frankham1996-yb,Nei1984-zi,Soule1976-he}, and the
+population size
+\parencite{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy}, and the
 ecological and life history correlates of genetic diversity
@@
 
-% Furthermore, Coop's re-analysis finds no evidence
-% of a remaining correlation between recombination and diversity among the
-% highest recombination windows, as would be expected if these $\pi_0$ estimates
-% were still impacted by linked selection.
-
 Here, I revisit Lewontin's Paradox by integrating a variety of data sets and
@@
 Finally, I explore whether the predicted reductions of diversity under
-background selection and recurrent hitchhiking are sufficiently strong enough
+background selection and recurrent hitchhiking are sufficiently strong 
 to resolve Lewontin's Paradox. These predicted reductions in diversity across
@@
 relationship on a log-log scale fits the data well (Figure \ref{fig:figure-1}).
-The OLS slope estimate is significant and implies an increase in diversity of
-0.09 differences per basepair for every order of magnitude census size grows
-(95\% confidence interval $[0.08, 0.12]$; see also the OLS fit per-phyla,
+The OLS slope estimate is significant and implies a 0.09 percent increase in
+differences per basepair for every order of magnitude census size grows (95\%
+confidence interval $[0.08, 0.12]$; see also the OLS fit per-phyla,
 Supplementary Figure \ref{suppfig:figure-1-ave}).
@@
 0.11$), though somewhat attenuated compared to the OLS estimates (Figure
-\ref{fig:figure-1}B). Since the population size estimates are based on range
+\ref{fig:figure-2}B). Since the population size estimates are based on range
 and body mass, they are essentially a composite trait; fitting phylogenetic
@@
 respectively; Figure \ref{fig:figure-2}C).  Considering the constituents of the
-population size estimate, range and body mass, separately, range but not body
-mass shows a significant increase (p-value $1.9 \times 10^{-7}$) in rate
-towards the present.
+population size estimate, range and body mass, separately, the rate of
+evolution of range but not body mass shows a significant increase (p-value $1.9
+\times 10^{-7}$) towards the present.
 
@@
 \parencite{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second difficulty is
-understanding how $R$ varies across taxa, since we lack estimates of critical
-model parameters for most species. Still, I can address a key question: if
-diversity levels were determined by census sizes ($\pi_0 = 4N_c \mu$), are the
-combined effects of background selection and recurrent hitchhiking sufficient
-to reduce diversity to observed levels? Furthermore, does the relationship
-between census size and predicted diversity under linked selection across
-species, $\pi_{BGS+HH} = R \pi_0$, match the observed relationship in Figure
-\ref{fig:figure-1}?
+understanding of how $R$ varies across taxa, since we lack estimates of
+critical model parameters for most species. Still, I can address a key
+question: if diversity levels were determined by census sizes ($\pi_0 = 4N_c
+\mu$), are the combined effects of background selection and recurrent
+hitchhiking sufficient to reduce diversity to observed levels? Furthermore,
+does the relationship between census size and predicted diversity under linked
+selection across species, $\pi_{BGS+HH} = R \pi_0$, match the observed
+relationship in Figure \ref{fig:figure-1}?
 
@@
 
-Parameterizing the model this way, I then set the key parameters that
-determine the impact of recurrent hitchhiking and background selection
-($\gamma$, $J$, and $U$) to high values estimated from \emph{Drosophila
-melanogaster} by \textcite{Elyashiv2016-vt}. My estimate of
-$\gamma_\text{Dmel}$ based Elyashiv et al. implies $\nu_\text{BP,Dmel} \approx
+Parameterizing the model this way, I then set the key parameters that determine
+the impact of recurrent hitchhiking and background selection ($\gamma$, $J$,
+and $U$) to high values estimated from \emph{Drosophila melanogaster} by
+\textcite{Elyashiv2016-vt}. My estimate of $\gamma_\text{Dmel}$ based Elyashiv
+et al. implies a rate of sweeps per basepair of $\nu_\text{BP,Dmel} \approx
 2.34 \times 10^{-11}$, which is close to other estimates from \emph{D.
@@
 calculable from different data and summary statistics
-\parencite{Wang2016-mi,Caballero1994-ao,Caballero2020-wm}. These include
-estimators based on the observed decay in linkage disequilibrium or temporal
-estimators that use the variance in allele frequency change. These alternate
-estimators capture summaries of the effective population size on shorter
-timescales than coalescent-based estimators \parencite{Wang2005-wy}, and thus
-could be used to tease out processes that impact the $N_e$-$N_c$ relationship
-in the more recent past.
+\parencite{Wang2016-mi,Caballero1994-ao,Caballero2020-wm,Galtier2020-fb}. These
+include estimators based on the observed decay in linkage disequilibrium or
+temporal estimators that use the variance in allele frequency change. These
+alternate estimators capture summaries of the effective population size on
+shorter timescales than coalescent-based estimators \parencite{Wang2005-wy},
+and thus could be used to tease out processes that impact the $N_e$-$N_c$
+relationship in the more recent past.
 
@@
 Stapley for kindly providing the recombination map length data, and Yaniv
-Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Sally Otto, Jeff
-Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli, and my
-Kern-Ralph labmates for helpful discussions. Sarah Friedman, Katherine Corn,
+Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Matt Osmond, Sally
+Otto, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli, and
+my Kern-Ralph labmates for helpful discussions. Sarah Friedman, Katherine Corn,
 and Josef Uyeda provided very useful advice about phylogenetic comparative

### 2021-04-29 manuscript/manuscript.tex
new data fixes, bug fixes, etc for revisions

@@
 orders-of-magnitude shortfall between predicted and observed diversity levels.
-Yet, across-taxa surveys indicate that $F_\text{ST}$ is almost never this high
-within species \parencite{Roux2016-lm}.  Still, future work could quantify the
-extent to which spatial processes contribute to Lewontin's Paradox. For
-example, high-$N_c$ taxa usually experience range expansions, likely with
-repeated founder effects and local extinction/recolonization dynamics that
-doubtlessly depress diversity. In particular, with the appropriate data, one
-could estimate the empirical relationship between dispersal distance, range
-size, and coalescent effective population size across taxa.
+This is because the equation for $F_{ST}$ can be rearranged such that $H_S =
+(1-F_{ST})H_T$, where $H_S$ and $H_S$ are the subpopulation and total
+population heterozygosities; if $H_T = 4N_c\mu$, then only $F_{ST} \approx 1$
+can reduce $H_S$ several orders of magnitude. Yet, across-taxa surveys indicate
+that $F_\text{ST}$ is almost never this high within species
+\parencite{Roux2016-lm}.  Still, future work could quantify the extent to which
+spatial processes contribute to Lewontin's Paradox. For example, high-$N_c$
+taxa usually experience range expansions, likely with repeated founder effects
+and local extinction/recolonization dynamics that doubtlessly depress
+diversity. In particular, with the appropriate data, one could estimate the
+empirical relationship between dispersal distance, range size, and coalescent
+effective population size across taxa.
 
@@
 Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Matt Osmond, Sally
-Otto, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli, and
-my Kern-Ralph labmates for helpful discussions. Sarah Friedman, Katherine Corn,
-and Josef Uyeda provided very useful advice about phylogenetic comparative
-methods; yet I take full responsibility for any shortcomings of my analysis.  I
-would like to also thank UO librarian Dean Walton for helping me track down
-some rather difficult to find older papers. This work was supported by an NIH
-Grant (1R01GM117241) awarded to Andrew Kern.
+Otto, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli,
+Margot Wood, and my Kern-Ralph labmates for helpful discussions. Sarah
+Friedman, Katherine Corn, and Josef Uyeda provided very useful advice about
+phylogenetic comparative methods; yet I take full responsibility for any
+shortcomings of my analysis.  I would like to also thank UO librarian Dean
+Walton for helping me track down some rather difficult to find older papers.
+This work was supported by an NIH Grant (1R01GM117241) awarded to Andrew Kern.
 
@@
 
-\section*{Supplementary Table and Figures}
+\section*{Supplementary Information}
+
+\section{Population Size Validation}
 

### 2021-04-29 notebooks/main_analysis.Rmd
new data fixes, bug fixes, etc for revisions

@@
 
+So, marine species are over represented in my dataset.
+
 Now we do a simple biomass analysis -- back of the envelop type stuff.
 
-```{r}
-
-# https://bionumbers.hms.harvard.edu/bionumber.aspx?&id=107256 
-# 18-23% of human body is carbon; seems reasonable this holds
-# for other animals
+A lot of the sources here are from the Bar-on et al. (2018) study.
+
+Across arthropods and chordates, Baron-On et al use a wet weight to carbon
+factor equal to 0.15. Description: "we convert wet weight values to carbon mass
+assuming 70% water content and 50% carbon content out of dry weight."
+
+For molluscs, cnidarians, nematodes, and annelids, they do something slightly
+differently, often basing their conversion on the carbon of a single individual
+(nematode, say). Here I forgo this and just use the conversion factor.
+
+```{r}
 
@@
 
-# in Gt
-prop_carbon <- signif(sqrt(0.18 * 0.23), 1)  # ~ 0.2
+# https://bionumbers.hms.harvard.edu/bionumber.aspx?&id=107256 
+# 18-23% of human body is carbon; his is close to the Bar-On et al. (2018) 
+# estimate
+prop_carbon <- 0.15 # from Bar-On et al.
+
 # 1Gt = 1e12 kg
-xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * 10^xd$log10_popsize) / 1e12
-
+xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * # to kg
+                   10^xd$log10_popsize) / 1e12
+
+# take my data set and find the total carbon biomass of species, grouped by phyla
 mass_d <- xd %>% 
@@
            mutate(bd_total_biomass = sum(bd_mass), 
-                  bd_prop_mass = bd_mass/bd_total_biomass) %>%
+                  bd_prop_mass = bd_mass/bd_total_biomass)
+
+# merge my carbon masses in with Bar-On et al's
+baron_mine <- baron_d %>%
            left_join(mass_d, by='phylum')  %>%
-           # not we need to reweight the original proportions as some phyla were
-           # dropped
+           # note, we need to reweight the original proportions as some phyla were
+           # dropped because they are not in Bar-On's data set (e.g. Porifera, Echinoderms)
            mutate(prop_carbon_mass = prop_carbon_mass / sum(prop_carbon_mass))
 
-ggplot(baron_d, aes(bd_prop_mass, prop_carbon_mass, color=phylum))  + 
+```
+
+How are the proportions?
+
+```{r}
+
+ggplot(baron_mine, aes(bd_prop_mass, prop_carbon_mass, color=phylum))  + 
   geom_point(size=5) + geom_abline(slope=1, intercept=0) + 
@@
 Arthropods are under sampled here, and chordates are way over sampled (at least
-by biomass).
-
-```{r}
-
-ggplot(baron_d) + 
+by biomass)!
+
+Let's look at the biomass totals, in my small data set and Bar-On's:
+
+```{r}
+
+ggplot(baron_mine) + 
   geom_point(aes(phylum, total_carbon_mass), size=5)
 
-ggplot(baron_d) + 
+ggplot(baron_mine) + 
   geom_point(aes(phylum, bd_mass), size=5)
 
-baron_d <- baron_d %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)
-
-baron_d %>%
-  select(phylum, prop_bd_biomass)
+```
+
+This shows the extent to which arthropods are under represented in my dataset
+and chordates are overrepresented.
+
+What fraction of the total biomasses in Bar-On are present in my dataset?
+
+```{r}
+baron_mine <- baron_mine %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)
+
+baron_mine %>% select(phylum, prop_bd_biomass)
+```
+
+4% of chordate biomass in my dataset seems high -- let's look at it.
+
+First the top mammalian species by biomass:
+
+```{r}
+
+xd %>% filter(class == 'Mammalia') %>% arrange(desc(carbon_mass)) %>% 
+  select(species, carbon_mass) %>% 
+  as.data.frame()
+
+```
+
+Baleen whales have a carbon mass of `0.0114e9 / 1e6 = 11.4` megatons! No way.
+
+Actually this isn't wrong --- "In terms of their size and potential to store
+carbon for years or decades, marine vertebrates are the only organisms in the
+ocean comparable to large trees". This is from [this
+paper](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0012444)
+which says ~10 megatons of carbon biomass is what we'd expect, pre-whaling.
+
+```{r}
+
+mamd <- xd %>% filter(class == 'Mammalia') 
+mamd$species  <- factor(mamd$species,
+                        mamd$species[order(mamd$carbon_mass, decreasing=TRUE)], 
+                        ordered=TRUE)
+igno <- c("Balaenoptera bonaerensis", "Homo sapiens", "Eschrichtius robustus")
+mamd %>% arrange(desc(carbon_mass)) %>% pull(species)
+
+mamd %>% 
+  filter(!(species %in% igno)) %>%
+  ggplot(aes(species, carbon_mass))  + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+
+```{r}
+
+xd %>% filter(species == "Aptenodytes patagonicus") %>% select(species, log10_popsize)
+
+xd %>% filter(genus == "Equus") %>% select(species, log10_popsize, log10_range)
+
+```
+
+```{r}
+
+"Homo sapiens"
+"Bos taurus"
+"Ovis aries"
+"Capra hircus"
+
+```
+
+```{r}
 

### 2021-04-29 data/combined_dataset.r

@@ -48,7 +48,22 @@ dc <- read_tsv('corbett_detig_2015_updated.tsv')  %>%
                # rename the original Corbett-Detig range
                range_cd = range) %>%
         rename(diversity = obs_pi) %>%
+        # 
         select(-range, -kingdom)
+
+
+# Corbett-Detig's data has Dmel as 1.5mm which is 60% of what it should be.
+# This was leading the average length across datasets to be too small, 
+# and in turn leading to a much larger density, 10^9.4 rather than 10^9
+# which is what it is for other species 
+# I correct this here manually, based on Wikipedia's value (REVISIONS)
+dc[dc$species == 'Drosophila melanogaster', ]$size <- 2.5e-3
+
+# Corbett-Detig's data has przewalskii's horse's range as 10^7. These strange
+# little horsies have a range that's more about ~300km² forcing it to be NA
+# here should allow the correction in in species_range.r data to go through
+dc[dc$species == 'Equus ferus przewalskii', ]$log10_range <- NA
+dc[dc$species == 'Equus ferus przewalskii', ]$range_cd <- NA
 
 # Romiguier et al data
 dr <- read_tsv('romiguier_et_al_2014_updated.tsv') %>%
@@ -136,7 +151,7 @@ d$parpath <- case_when(is.na(d$parpath) & d$species %in% parpath ~ TRUE,
                        is.na(d$parpath) & !(d$species %in% parpath) ~ FALSE,
                        TRUE ~ d$parpath)
 
-# not used yet, and not complete
+# list of social insects, for recomb plot
 social <- c("Acromyrmex echinatior",
             "Apis mellifera",
             "Apis cerana",
@@ -154,7 +169,6 @@ d$social <- FALSE
 d$social[d$species %in% social] <- TRUE
 
 stopifnot(nrow(d) == length(unique(d$species)))
-
 
 #### Load in Range Data Estimated from Occurrence Data
 ## Load and clean up range size
@@ -185,10 +199,13 @@ d <- d %>%
          unnest(ranges) %>%
          ungroup() %>%
          rowwise() %>%
+         # this averages the ranges from Corbett-Detig and mine
+         # ignoring NAs
          mutate(range = mean(c(range, range_cd), na.rm=TRUE),
                 range = ifelse(is.nan(range), NA, range)) %>%
          select(-range_cd) 
[…]

### 2021-04-29 data/species_range_fixes.r

@@ -3,6 +3,7 @@ library(tidyverse)
 terrestriality_fixes <- tribble(~ name, ~ is_terrestrial,
         'Anopheles scanloni', TRUE,
         'Arabidopsis', TRUE,
+        'Aptenodytes', TRUE,
         'Aphanarthrum subglabrum/glabrum', TRUE,
         'Ciona', FALSE, 
         'Caenorhabditis', TRUE,
@@ -77,6 +78,7 @@ terrestriality_fixes <- tribble(~ name, ~ is_terrestrial,
 )
 
 alpha_fixes <- tribble(~name, ~alpha,
+       'Anoplopoma', c(NA, 10),
        'Crassostrea', c(NA, 10),
        'Eunicella', c(NA, 10),
        'Haliotis', c(NA, 10),

### 2021-04-29 data/species_ranges.r

@@ -11,6 +11,10 @@ source('../R/range_funcs.r')
 
 #### Load the data in 
 
+# we don't load in Corbett-Detig data, which has ranges though some are off.
+# This is merged in combined_dataset.r. I make one correction here though, to
+# Equus ferus przewalskii.
+
 ## Leffler data
 message('loading Leffler et al data...')
 dl <- read_tsv('./leffler_et_al_2012_updated.tsv')
@@ -93,7 +97,6 @@ message('done.')
 #   user="vsbuffalo", pwd="hidden", email="vsbuffalo@gmail.com")
 
 
-
 # combine in alphas
 DEFAULT_ALPHA <- c(8, 70)
 RANGE_DIR <- 'total_range_plots'
@@ -102,6 +105,8 @@ if (!file.exists(RAW_RANGES)) {
   ### Infer terrestriality
   # filter entries without lat/long and infer terrestriality 
   ranges <- uniq_species_df %>% 
+              #filter(species == 'Anoplopoma fimbria') %>%
+              #filter(species == 'Aptenodytes patagonicus') %>%
               mutate(valid_geo = map_lgl(occs, valid_occurs)) %>%
               filter(valid_geo) %>%
               distinct(species, .keep_all=TRUE) %>%
@@ -115,10 +120,10 @@ if (!file.exists(RAW_RANGES)) {
     mutate(is_terrestrial_fixed =
            pmap_lgl(list(genus, species, is_terrestrial),
                     function(g, s, it) {
-                     case_when(g %in% marine_gnr ~ FALSE,
-                               g %in% terrestrial_gnr ~ TRUE,
-                               s %in% marine_sps ~ FALSE,
-                               s %in% terrestrial_sps ~ TRUE,
+                     case_when((g %in% marine_gnr) || 
+                               (s %in% marine_sps) ~ FALSE,
+                               (g %in% terrestrial_gnr) ||
+                               (s %in% terrestrial_sps) ~ TRUE,
                                TRUE ~ it)
                     })) %>%
     mutate(is_terrestrial_orig = is_terrestrial, 
@@ -127,8 +132,6 @@ if (!file.exists(RAW_RANGES)) {
 
   # how many terrestriality fixes?
   table(ranges$is_terrestrial != ranges$is_terrestrial_orig)
-
-
 
   ### Infer Range
   ranges <- ranges %>%  
[…]

### 2021-04-29 notebooks/figures/body_mass_length.r

@@ -43,7 +43,7 @@ axis(2, yseq, labels=latex2exp::TeX(sprintf("$10^{%d}$", yseq)),
      cex.axis=0.8, line=0.5)
 legend(-0.8, -2, tools:::toTitleCase(levels(as.factor(d$phylum))), fill=group_cols,
        bty='n', border=0, cex=0.6, ncol=2)
-title(xlab=latex2exp::TeX("mass length (meters)"), line=2.1)
+title(xlab=latex2exp::TeX("body length (meters)"), line=2.1)
 title(ylab=latex2exp::TeX("body mass (grams)"), line=2.3)
 
 dev.off()

### 2021-05-04 manuscript/manuscript.tex
revision updates

@@
 
+\subsection{Biomass Analysis}
+
+If we have $T$ species, their biomass is $B = \sum_{i=1}^T N_i M_i$ where $N_i$
+is their census size and $M_i$ is their wet body mass. In my dataset, I
+approximate $N_i$ as the product of a species $i$'s range $R_i$ and their
+predicted population density from wet body mass $D_i = \beta_0 + \beta_1 M_i$.
+The biomass in my sample, $B'$ is the sum above with some weights $w_i \in \{0,
+1\}$; we can write:
+
+\begin{align}
+  B' &= \sum_{i=1}^T w_i N_i M_i \\
+    &= \sum_{i=1}^T w_i R_i (\beta_0 + \beta_1 M_i ) M_i \\
+\end{align}
+
 \section{Diversity and IUCN Red List Status}

### 2021-05-04 notebooks/main_analysis.Rmd
revision updates

@@
 
-```
-
-Based on the predictions of Mora et al. 2011
-
-```{r}
-
 nspecies <- c(terrestrial= 7.770e6, marine=2.15e5)
@@
 
-So, marine species are over represented in my dataset.
+Number of species is based on the predictions of Mora et al. 2011. So, marine
+species are over represented in my dataset.
+
+This study doesn't have per-class or phylum data, so I build a quick dataset
+here from a few sources:
+
+```{r}
+
+# sources, 
+# https://www.environment.gov.au/system/files/pages/2ee3f4a1-f130-465b-9c7a-79373680a067/files/nlsaw-2nd-complete.pdf
+nsd <- tribble(~ class, ~ total_species, 
+               "Echinoidea", 940,
+               "Holothuroidea", 1430,
+               "Insecta", 1e6,
+               "Aves",  9990,
+               "Mammalia", 5487,
+               "Clitellata", 8e3, # wikipedia
+               "Amphibia", 6515,
+               "Demospongiae", 8800, # wikipedia
+               "Teleostei", 30e3, # wikipedia
+               "Bivalvia", 10e3,  # https://ucmp.berkeley.edu/taxa/inverts/mollusca/bivalvia.php
+               "Malacostraca", 40e3, # wikipedia
+               "Branchiopoda", 800, # animaldiversity.org
+               "Gastropoda", prod(c(40e3, 100e3))^(1/2), # https://ucmp.berkeley.edu/taxa/inverts/mollusca/gastropoda.php
+               "Chromadorea", 2124, # ITIS
+               "Reptilia", 8734,
+               "Ascidiacea", 2300, # wikipedia
+               "Anthozoa", 6e3, # wikipedia
+               "Euchelicerata", 77000, # wikipedia
+               "Holostei", 8, # wikipedia
+               "Anopla", 600, # wikipedia
+               "Arachnida", 102248,
+               "Ophiuroidea",  2064, # https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0031940#
+               "Polychaeta", 10000, # https://animaldiversity.org/accounts/Polychaeta/
+               "Cephalopoda", 800, # https://ucmp.berkeley.edu/taxa/inverts/mollusca/cephalopoda.php
+             "Maxillopoda", 14e3 # https://www.sciencedirect.com/topics/agricultural-and-biological-sciences/maxillopoda
+)
+
+```
+
+Now, compare to the number of species by class/phylum in my data set:
+
+```{r}
+
+# phyla / class mapping 
+phyla_class <- dt_dps %>% select(phylum, class) %>% group_by(phylum, class) %>% 
+  distinct()
+
+nsd %>% 
+  dt_dps %>% 
+  select(phylum, species) %>% group_by(phylum)
+
+# by phyla
+dnsp_phyla <- dt_dps %>% 
+  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
+  select(phylum, species) %>% group_by(phylum) %>% 
+               summarize(nspecies=n())
+
+# by class
+dnsp_class <- dt_dps %>% 
+  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
+  select(class, species) %>% group_by(class) %>% 
+               summarize(nspecies=n())
+
+dnsp_class %>%
+  ggplot(aes(class, log10(frac), size=log10(total_species))) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+
+By phylum:
+
+```{r}
+
+dnsp_phyla %>%
+  ggplot(aes(phylum, log10(frac), size=log10(total_species))) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
 
@@
 # only complete data
-xd <- d %>% filter(is.finite(log10_body_mass))
+xd <- dt_dps
 
@@
 # 1Gt = 1e12 kg
-xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * # to kg
+xd$carbon_mass <- (prop_carbon*(10^xd$log10_body_mass_pred) / 1e3 * # to kg
                    10^xd$log10_popsize) / 1e12
@@
 
+baron_mine %>% left_join(dnsp_phyla, by='phylum')  %>% select(nspecies.x, nspecies.y)
+  mutate(a=signif(total_carbon_mass / frac, 2)) %>% 
+  select(-bd_total_biomass) %>%
+  select(-total_log10_popsize, -nspecies.y) %>% as.data.frame
+
 ```
@@
 ```{r}
+
 baron_mine <- baron_mine %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)
@@
 baron_mine %>% select(phylum, prop_bd_biomass)
+
 ```
@@
                         ordered=TRUE)
-igno <- c("Balaenoptera bonaerensis", "Homo sapiens", "Eschrichtius robustus")
+#igno <- c("Balaenoptera bonaerensis", "Homo sapiens", "Eschrichtius robustus")
+igno <- c()
 mamd %>% arrange(desc(carbon_mass)) %>% pull(species)
@@
 
-```{r}
-
-xd %>% filter(species == "Aptenodytes patagonicus") %>% select(species, log10_popsize)
-
-xd %>% filter(genus == "Equus") %>% select(species, log10_popsize, log10_range)
-
-```
-
-```{r}
-
-"Homo sapiens"
-"Bos taurus"
-"Ovis aries"
-"Capra hircus"
-
-```
+This seems reasonable... pre-whaling whales, humans, and their domesticates
+dominate. Wolves seem high, but this is a macro-ecological based approach that
+doesn't reflect hunting to low densities.
+
+Let's look at some random species:
+
+```{r}
+
+of_interest = c("Aptenodytes patagonicus", "Equus caballus", "Equus ferus
+                przewalskii", "Homo sapiens", "Bos taurus", "Ovis aries", "Capra
+                hircus", "Canis lupus")
+
+xd %>% filter(species %in% of_interest) %>% select(species, log10_popsize, log10_range)
+
+```
+
+Ok, now let's look at total biomass. Since I have drastically different phylum
+represented here in my taxa sample than on earth, we need to reweight things.
 
@@
 # using the Bar-On weights
-baron_d <- baron_d %>% 
+# prop_carbon_mass is prop in my dataset
+baron_mine <- baron_mine %>% 
               mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
@@
 
-total_sample_carbon_mass <- baron_d %>% pull(reweighted_total_carbon_mass) %>% sum()
+total_sample_carbon_mass <- baron_mine %>% pull(reweighted_total_carbon_mass) %>% sum()
 

### 2021-05-04 data/combined_dataset.r

@@ -205,7 +205,6 @@ d <- d %>%
                 range = ifelse(is.nan(range), NA, range)) %>%
          select(-range_cd) 
 stopifnot(nrow(d) == length(unique(d$species)))
-stop()
 
 #### Get Taxonomical Info
 # get the taxonomial datas from taxadb (via Catalog of Life)

### 2021-05-04 data/species_ranges.r

@@ -379,10 +379,10 @@ ranges_df[ranges_df$species == 'Argopecten irradians', ]$range <- argo_area
 # 
 # since this is a correction to Corbett Detig's data, which hasn't been 
 # merged in yet, I add this row first.
-stop()
 eqfp <- tibble(key = "Equus ferus przewalskii", 
                species = "Equus ferus przewalskii",
                dataset = "Corbett-Detig", is_terrestrial = TRUE, 
+               genus="Equus", 
                is_terrestrial_orig = TRUE, n_occ = NA, range = NA)
 ranges_df <- rbind(ranges_df, eqfp)
 geom_mean <- function(x) prod(x)^(1/length(x))

### 2021-05-04 data/taxonomy_fixes.r

@@ -222,7 +222,7 @@ taxon_fixes <- list(
      list(species="Kryptolebias marmoratus",
           kingdom='Animalia',
           phylum='Chordata',
-          class='Actinopterygii',
+          class='Teleostei',
           order='Cyprinodontiformes',
           family='Rivulidae'),
 

### 2021-05-10 manuscript/manuscript.tex
species range corrections, fig 2 tweaks

--- 2021-04-29 manuscript/manuscript.tex
@@
 \RequirePackage{mathtools}
+\RequirePackage{tabularx}
 \RequirePackage{graphicx}
@@
 
+The approximate census sizes I have estimated here
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[]{figures/species_counts.pdf}
+
+  \caption{}
+
+  \label{suppfig:species_counts}
+\end{figure}
+
+\begin{table}[ht]
+  \resizebox{\textwidth}{!}{%
+\centering
+\begin{tabular}{lr|rr|rrrr|r}
+  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{4}{c}{\bf Present study}  &  \\
+  phylum & total species ($T$) & biomass ($B$) & prop. biomass & biomass ($b$) &  prop. biomass & num. species ($n$) &prop. total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
+  \hline
+  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.77 \times 10^{-4}$ & 0.0100 &  68 & $5.41 \times 10^{-5}$ & 4.26 \\
+  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.68 \times 10^{-2}$ & 0.9720 &  68 & $1.26 \times 10^{-3}$ & 24.50 \\
+  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.24 \times 10^{-5}$ & 0.0005 &   3 & $1.76 \times 10^{-4}$ & 0.35 \\
+  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.47 \times 10^{-4}$ & 0.0162 &  13 & $1.36 \times 10^{-4}$ & 16.40 \\
+  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.21 \times 10^{-5}$ & 0.0012 &   2 & $1.25 \times 10^{-4}$ & 2.57 \\
+  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $4.00 \times 10^{-6}$ & 0.0001 &   1 & $4.00 \times 10^{-5}$ & 5.00 \\
+\end{tabular}}
+
+  \caption{How the total carbon biomass estimates by phylum from
+    \textcite{Bar-On2018-kc} compare to the implied biomass estimates from this
+    study. All biomass estimates are carbon biomass, and the proportions are of
+    total biomass with respect to the study. The proportion of biomass in this
+    study compared to the Bar-On et al. estimates \textcite{Bar-On2018-kc}
+    indicates chordates are over-represented and arthropods are
+    under-represented in the present study. Total species by phylum estimates
+  are from
+\textcite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
+ratio column is the ratio of total biomass implied by the $N_c$ estimates of
+each species in a phylum to the actual biomass of that phylum.}
+
+\end{table}
+
 \section{Diversity and IUCN Red List Status}

### 2021-05-10 notebooks/main_analysis.Rmd
species range corrections, fig 2 tweaks

@@
 
-# sources, 
+# main sources:
 # https://www.environment.gov.au/system/files/pages/2ee3f4a1-f130-465b-9c7a-79373680a067/files/nlsaw-2nd-complete.pdf
+# this is for the classes *in my dataset*
+
+geomean <- function(x) (prod(x))^(1/length(x))
+
 nsd <- tribble(~ class, ~ total_species, 
@@
                "Branchiopoda", 800, # animaldiversity.org
-               "Gastropoda", prod(c(40e3, 100e3))^(1/2), # https://ucmp.berkeley.edu/taxa/inverts/mollusca/gastropoda.php
+               "Gastropoda", geomean(c(40e3, 100e3)), # https://ucmp.berkeley.edu/taxa/inverts/mollusca/gastropoda.php
                "Chromadorea", 2124, # ITIS
@@
 
-nsd %>% 
-  dt_dps %>% 
-  select(phylum, species) %>% group_by(phylum)
-
+# what are the total species by phyla? we join in the mapping from our data
+total_species_by_phyla <- nsd %>% right_join(phyla_class, by='class') %>% 
+  group_by(phylum) %>% summarize(total_species = sum(total_species, na.rm=TRUE))
+
+# these numbers are conditioning on classes we have in this study
+# we actually want all classes on *earth*; for this I use the table 
+# here from wikipedia
+# https://en.wikipedia.org/wiki/Animal#Numbers_and_habitats
+total_species_by_phyla2 <- tribble(~ phylum, ~ total_species,
+                                   "Annelida", 17e3,
+                                   "Arthropoda", 1.257e6,
+                                   "Chordata", geomean(c(45e3, 65e3)),
+                                   "Cnidaria", 16e3,
+                                   "Echinodermata", 7.5e3,
+                                   "Mollusca", geomean(c(85e3, 107e3)),
+                                   "Nematoda", 25e3,
+                                   "Nemertea", 1300,
+                                   "Porifera", 10800)
+
+# and by phyla? again joining in by the mapping in our data
+# conditioning on classes in my dataset
+total_species_by_class <- nsd %>% right_join(phyla_class, by='class') %>% 
+  group_by(class) %>% summarize(total_species = sum(total_species, na.rm=TRUE))
+
+# how different are the number of species when we compare the dataset 
+# further up, conditioning on classes in my dataset, versus the dataset
+# immediately above, from Wikipedia?
+total_species_by_phyla %>% left_join(total_species_by_phyla2, by='phylum') %>%
+  mutate(ratio = total_species.x / total_species.y)
+
+```
+
+Nothing terribly off here; nematodes is the only sizeable difference.
+
+```{r}
+## now join in the total species in our dataset
 # by phyla
 dnsp_phyla <- dt_dps %>% 
+  left_join(total_species_by_phyla2) %>%
   filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
-  select(phylum, species) %>% group_by(phylum) %>% 
-               summarize(nspecies=n())
+  select(phylum, species, total_species) %>% group_by(phylum, total_species) %>% 
+  summarize(nspecies=n()) %>% 
+  mutate(frac = nspecies / total_species)
 
@@
 dnsp_class <- dt_dps %>% 
+  left_join(total_species_by_class) %>%
   filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
-  select(class, species) %>% group_by(class) %>% 
-               summarize(nspecies=n())
+  select(phylum, class, species, total_species) %>% group_by(phylum, class, total_species) %>% 
+  summarize(nspecies=n()) %>%
+  mutate(frac = nspecies / total_species)
 
 dnsp_class %>%
-  ggplot(aes(class, log10(frac), size=log10(total_species))) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  ggplot(aes(class, log10(frac), size=log10(total_species))) + 
+  geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+# this would make a cool supp plot, saving data
+write_tsv(dnsp_class, file='../data/species_by_class.tsv')
 
@@
 dnsp_phyla %>%
-  ggplot(aes(phylum, log10(frac), size=log10(total_species))) + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+  ggplot(aes(phylum, log10(frac), size=log10(total_species))) + 
+  geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 
@@
 
-baron_mine %>% left_join(dnsp_phyla, by='phylum')  %>% select(nspecies.x, nspecies.y)
-  mutate(a=signif(total_carbon_mass / frac, 2)) %>% 
+# join everything
+biomass <- baron_mine %>% left_join(dnsp_phyla, by='phylum')  %>% 
+  # 
+  mutate(factor=signif(total_carbon_mass / (bd_mass * frac), 3)) %>% 
   select(-bd_total_biomass) %>%
-  select(-total_log10_popsize, -nspecies.y) %>% as.data.frame
-
-```
-
-How are the proportions?
+  select(-total_log10_popsize)
+
+stopifnot(all(biomass$nspecies.x == biomass$nspecies.y))
+biomass <- biomass %>% rename(nspecies = nspecies.x) %>%
+  select(-nspecies.y, -total)
+
+write_tsv(biomass, file='../data/biomass.tsv')
+
+bm_tbl <- biomass %>% mutate(total_species = as.integer(total_species)) %>%
+           select(phylum, total_species, bd_mass, bd_prop_mass, 
+                  total_carbon_mass, prop_carbon_mass, nspecies, frac, factor)
+
+print(xtable(bm_tbl, digits=c(1, 2, -2, 2, 4, -2, 4, 2, -2, 2)), 
+      include.rownames=FALSE,
+      tabular.environment='tabular',
+      math.style.exponents=TRUE, type='latex')
+
+```
+
+Things look relatively good here -- nothing is off by several orders of
+magnitude unless there's only 1-3 species sampled.
+
+How are the proportions of biomass by taxa?
 
@@
 ```
-
-Ok, now let's look at total biomass. Since I have drastically different phylum
-represented here in my taxa sample than on earth, we need to reweight things.
-
-```{r}
-
-# using the Bar-On weights
-# prop_carbon_mass is prop in my dataset
-baron_mine <- baron_mine %>% 
-              mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
-              mutate(reweighted_total_carbon_mass = reweight * total_carbon_mass)
-
-total_sample_carbon_mass <- baron_mine %>% pull(reweighted_total_carbon_mass) %>% sum()
-
-# total species here
-n <- nrow(xd)
-total_animal_species <- sum(nspecies)
-
-print(paste0("fraction of animal species in this data set: ", 
-             signif(n/total_animal_species, 2)))
-
-# Bar-On et al. (2018) estimate is 2Gt C
-# 1e12 kg = 1Gt
-total_animal_cmass <- 2
-
-# if the species here were random
-signif(total_animal_cmass * n / total_animal_species, 2)
-signif(total_sample_carbon_mass, 2)
-
-0.058 / total_animal_cmass
-
-```
-
-Note: initially it was my plan to see if we could get a sense of what fraction
-of the total biomass this dataset contains (which is about 3%). This seems
-unreasonably high, even after reweighting as above. The problem is that the
-3.4e-5 fraction of animal species is obviously not a random sample based on
-size / population, and includes humans, domesticated species, etc. -- this
-probably is accentuated by the overabundance of chordates compared arthropods
-here.
 

### 2021-05-10 data/species_ranges.r

@@ -6,6 +6,7 @@ library(readxl)
 library(igraph)
 library(units)
 
+geomean <- function(x) prod(x)^(1/length(x))
 source('../R/range_funcs.r')
 
 
@@ -178,10 +179,15 @@ ranges_df <- ranges %>% select(-occs, -taxon_key, -valid_geo) %>%
                select(-range_area)
 
 ## Manual fixes
+# to ensure we are assining *over*, or correcting a value, 
+# not creating a new entry, we use this
+is_entry <- function(species) stopifnot(any(ranges_df$species == species))
+
 # Human range size is just wrong. Occurrences weren't logged everywhere.
 earth_land_area <- 510.1e6 # km^2
 antartica_land_area <- 13.21e6 # km^2
 human_range <- earth_land_area - antartica_land_area
+is_entry('Homo sapiens')
 ranges_df[ranges_df$species == 'Homo sapiens', ]$range <- human_range
 
 # Nasonia vitripennis 
@@ -189,6 +195,7 @@ ranges_df[ranges_df$species == 'Homo sapiens', ]$range <- human_range
 us_land_area <- 9.834e6 # km^2
 europe_land_area <- 10.18e6  # km^2
 nv_range <- us_land_area + europe_land_area
+is_entry('Nasonia vitripennis')
 ranges_df[ranges_df$species == 'Nasonia vitripennis', ]$range <- nv_range
 
 # Caenorhabditis
@@ -196,6 +203,7 @@ ranges_df[ranges_df$species == 'Nasonia vitripennis', ]$range <- nv_range
 # based on Frezel and Felix (2015)
 western_us <- 863400 # km^2 
 celegans_range <- western_us + europe_land_area
+is_entry('Caenorhabditis elegans')
 ranges_df[ranges_df$species == 'Caenorhabditis elegans', ]$range <- celegans_range
 
 # Pinus balfouriana
@@ -203,6 +211,7 @@ ranges_df[ranges_df$species == 'Caenorhabditis elegans', ]$range <- celegans_ran
 # https://en.wikipedia.org/wiki/Pinus_balfouriana#/media/File:Pinus_balfouriana_range_map_1.png
 ca_area <- 423971 # km^2
 pinus_bal_sel <- (174016 - 172459) / 174016 
+is_entry('Pinus balfouriana')
 ranges_df[ranges_df$species == 'Pinus balfouriana', ]$range <- ca_area* pinus_bal_sel
 
 # Pinus_massoniana
[…]

### 2021-05-10 notebooks/figures/diversity_popsize_full.r

@@ -21,8 +21,11 @@ phi <- mean(rstan:::extract(dps_fit, pars='phi')$phi)
 beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
 
 # alias this -- less typing
-d <- da_dps
+d <- da_dps %>% filter(species != 'Equus ferus przewalskii')
 
+#d = read_tsv('../../data/combined_data.tsv')
+#d %>% filter(class == 'Aves') %>% select(species, log10_range, log10_popsize)
+ 
 # what to include
 # x <- d$pred_log10_N  # the simplistic unregularized version
 x <- d$log10_popsize
@@ -38,11 +41,36 @@ coords <- ellipse(mu=center, sigma=sigma, alpha=0.35, draw=FALSE)
 
 label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
 # add in all drosophila
-keep_sps <- c("Homo sapiens", "Apis mellifera", "Canis lupus")
-label <- (label | d$family == 'Drosophilidae') | label %in% keep_sps
+keep_sps <- c("Homo sapiens", "Apis mellifera", "Canis lupus", "Caenorhabditis elegans")
+drosoph <- c(
+             "Drosophila americana", 
+             "Drosophila ananassae", 
+             #"Drosophila arizonae",
+             #"Drosophila bipectinata", 
+             #"Drosophila buzzatii",  
+             #"Drosophila kikkawai",
+             #"Drosophila malerkotliana", 
+             "Drosophila melanogaster", 
+             "Drosophila miranda",
+             "Drosophila mojavensis", 
+             #"Drosophila novamexicana", 
+             #"Drosophila parabipectinata",
+             "Drosophila persimilis", 
+             "Drosophila pseudoobscura", 
+             "Drosophila sechellia",
+             "Drosophila simulans", 
+             "Drosophila subobscura")
+
+label <- (label | label %in% drosoph) | label %in% keep_sps
 # contained <- rep(TRUE, length(x))
 
 d$label <- ifelse(label, d$species, "")
+
+# ignore a cluster of points a little over Nc=10^8
+# which has human
+#ignore <- c('Aquila clanga', 'Reticulitermes lucifugus', 'Reticulitermes grassei')
[…]

### 2021-05-10 notebooks/figures/plot_funs.r

@@ -14,13 +14,18 @@ extract_ggrepel <- function(p) {
   nodes <- grid.get("textrepeltree", grep = TRUE)
   kids <- childNames(nodes)
 
+  # fix for some upstream change BS
+  #browser() 
+  kids <- kids[regexpr("textrepelgrob", kids) != -1]
+
   # Function: get the x and y positions of a single ggrepel label
   get.xy.pos.labs <- function(n) {
     grid.force()
     grb <- grid.get(n)
     data.frame(
       x = xrg[1]+diff(xrg)*convertX(grb$x, "native", valueOnly = TRUE),
-      y = yrg[1]+diff(yrg)*convertY(grb$y, "native", valueOnly = TRUE)
+      y = yrg[1]+diff(yrg)*convertY(grb$y, "native", valueOnly = TRUE),
+      label = grb$label
     )
   }
   # Get positions of all ggrepel labels

### 2021-05-10 notebooks/figures/species_counts.r

@@ -1,0 +1,42 @@
+# species_counts.r
+
+library(tidyverse)
+source('color_scheme.r')
+
+d <- read_tsv('../../data/species_by_class.tsv') %>% 
+  filter(is.finite(frac))
+opar <- par(no.readonly=TRUE)
+
+size <- 7
+pdf('species_counts.pdf', width=size*0.773, size)
+par(mar=c(5, 0.1, 0.1, 0.6))
+x <- seq_along(d$class)
+plot(log10(d$frac), x, type='n', axes=FALSE, ylab='', xlab='', 
+     xlim=c(-6.5, -0.5))
+axis(1, seq(-1, -5), latex2exp::TeX(sprintf("$10^{%d}$", seq(-1, -5))))
+
+cols <- match(d$phylum, names(phyla_cols))
+points(log10(d$frac), x, pty='n', cex=1.5*(d$nspecies)^0.35, pch=21,
+       col=phyla_cols[cols], lwd=3, bg=scales::alpha(phyla_cols[cols], 0.6))
+text(log10(d$frac) + 0.11*d$nspecies^0.3, x, d$class, 
+     xpd=TRUE, srt=0, cex=0.8, adj=c(0, NA))
+
+text(log10(d$frac), x, d$nspecies, cex=0.6, adj=c(0.5, NA))
+
+z = factor(d$phylum)
+phy <- as.integer(z)
+for (i in 1:max(phy)) {
+  idx <- which(phy == i) 
+  text(-5.5, mean(idx), levels(z)[i], cex=0.8, adj=c(1, NA))
+  segments(-5.4, min(idx) - 0.1, -5.4, max(idx) + 0.1, lwd=3, lend=2, 
+          col=phyla_cols[i])
+}
+
+title(xlab='                   fraction of estimated total species included')
+
+dev.off()
+
+# phyla_cols_s <- phyla_cols[unique(d$phylum)]
+# legend(-5, 15, names(phyla_cols_s), fill=phyla_cols_s,
+#        bty='n', border=0, cex=0.7, ncol=1)
+par(opar)

### 2021-05-11 manuscript/manuscript.tex
more revisions

@@
 are all in the IUCN Red List, but are listed as least concern (though their
-presence in the Red List indicates they are of conservation interest). I also
-investigated the relationship between species' IUCN Red List categories (an
-ordinal scale of how threatened a species is) and both diversity and population
-size, finding that species categorized as more threatened have both smaller
-population sizes and reduced diversity, compared to non-threatened species
-(Figure \ref{fig:figure-1-redlist}) consistent with past work
-\parencite{Spielman2004-mt}. A linear model of diversity regressed on
-population size has lower AIC when the IUCN Red List categories are included,
-and the estimates of the effect of IUCN status are all negative on diversity,
-though not all are significant in part because some categories have three or
-fewer species (Supplementary Table \ref{supptable:tbl-1}).
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=\textwidth]{figures/diversity_popsize_redlist.pdf}
-
-  \caption{A version of Figure \ref{fig:figure-1} with points colored by their
-    IUCN Red List conservation status. Margin boxplots show the diversity and
-    population size ranges (thin lines) and interquartile ranges (thick lines)
-    for each category. NA/DD indicates no IUCN Red List entry, or Red List
-  status Data Deficient; LC is Least Concern, NT is Near Threatened, VU is
-Vulnerable, EN is Endangered, and CR is Critically Endangered.}
-
-  \label{fig:figure-1-redlist}
-\end{figure}
+presence in the Red List indicates they are of conservation interest). In
+Supplementary Material Section \namref{supinfo:div-iucn}, I explore more about
+the relationships between IUCN Red List status, diversity, and population size.
 
@@
 \section{Diversity and IUCN Red List Status}
+\label{supinfo:div-iucn}
+
+I also investigated the relationship between species' IUCN Red List categories
+(an ordinal scale of how threatened a species is) and both diversity and
+population size, finding that species categorized as more threatened have both
+smaller population sizes and reduced diversity, compared to non-threatened
+species (Supplementary Figure \ref{suppfig:figure-1-redlist}) consistent with
+past work \parencite{Spielman2004-mt}. A linear model of diversity regressed on
+population size has lower AIC when the IUCN Red List categories are included,
+and the estimates of the effect of IUCN status are all negative on diversity,
+though not all are significant in part because some categories have three or
+fewer species (Supplementary Table \ref{supptable:tbl-1}).
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=\textwidth]{figures/diversity_popsize_redlist.pdf}
+
+  \caption{A version of Figure \ref{fig:figure-1} with points colored by their
+    IUCN Red List conservation status. Margin boxplots show the diversity and
+    population size ranges (thin lines) and interquartile ranges (thick lines)
+    for each category. NA/DD indicates no IUCN Red List entry, or Red List
+  status Data Deficient; LC is Least Concern, NT is Near Threatened, VU is
+Vulnerable, EN is Endangered, and CR is Critically Endangered.}
+
+  \label{suppfig:figure-1-redlist}
+\end{figure}
+
 \begin{table}[ht]

### 2021-05-11 notebooks/main_analysis.Rmd
more revisions

@@
 # data for the log-normal fit
-log10_popsize_seq <- seq(5, 18, length.out=100)
+log10_popsize_seq <- seq(5, 15.2, length.out=100)
 ml_dat <- list(N = nrow(dt_ml_ns), 
@@
      mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
+     # RHH for comparison
+     mutate(Ne_N_RHH = RHH_BGS(1, L=map_length, N=10^log10_popsize,
+                                   B=1, gamma=dm_gamma, 
+                                   J22=J22)) %>%
      # both BGS and RHH
@@
                                    B=Ne_N_HK95, gamma=dm_gamma, 
-                                   J22=dm_J22_Elyashiv))
+                                   J22=J22))
 
 save(dml, dml_full, file='../data/linked_sel_div_changes.Rdata')
+
+```
+
+What are the relative strengths of BGS and RHH?
+
+```{r}
+
+dml_full %>% mutate(ratio = log10(Ne_N_RHH / Ne_N_HK95)) %>% 
+  ggplot(aes(log10_popsize, ratio)) + geom_point()
 
@@
 plot(dml$log10_popsize, dml$log10_diversity, pch=19, 
-                  xlim=c(5, 18),
+                  xlim=c(5, 16),
                   ylim=c(-3, 1))

### 2021-05-11 data/combined_dataset.r

@@ -208,12 +208,30 @@ stopifnot(nrow(d) == length(unique(d$species)))
 
 #### Get Taxonomical Info
 # get the taxonomial datas from taxadb (via Catalog of Life)
-d <- d %>% mutate(id_raw=map_chr(clean_name(species), taxadb::get_ids)) 
+TAXADB_SPECIES_FILE <- 'taxadb_species_ids.Rdata'
+if (!file.exists(TAXADB_SPECIES_FILE)) {
+  taxadb_species_ids <- map_chr(clean_name(d$species), taxadb::get_ids)
+  save(taxadb_species_ids, file=TAXADB_SPECIES_FILE)
+} else {
+  load(TAXADB_SPECIES_FILE)
+}
+
+# merge in species ids
+d$id_raw <- taxadb_species_ids
+
 # find all missing IDs and get genera-level IDs
 missing_sps <- d$species[is.na(d$id_raw)]
 genera <- map_chr(missing_sps, extract_genus, genus_only=TRUE)
-genera_ids <- map(genera, taxadb::get_ids)
-d$id_raw[is.na(d$id_raw)] <- genera_ids
+
+TAXADB_GENERA_FILE <- 'taxadb_genera_ids.Rdata'
+if (!file.exists(TAXADB_GENERA_FILE)) {
+  taxadb_genera_ids <- map(genera, taxadb::get_ids)
+  save(taxadb_genera_ids, file=TAXADB_GENERA_FILE)
+} else {
+  load(TAXADB_GENERA_FILE)
+}
+
+d$id_raw[is.na(d$id_raw)] <- taxadb_genera_ids
 d <- d %>% mutate(id = map_int(id_raw, ~ as.integer(sub('ITIS:', '', .))))
  
 ## Lookup and merge in taxonomy 
@@ -287,6 +305,7 @@ if (!file.exists(REDLIST_FILE)) {
   load(REDLIST_FILE)
 }
 
+
 # get the occurrences / km2 out
 species <- map(rdl_entries, 'name') 
 species <- map_chr(species, ~ ifelse(is.null(.), NA, .))
@@ -313,6 +332,11 @@ redlist_cat[redlist_cat == 'LR/lc'] <- "LC"
 redlist_cat[redlist_cat == 'LR/nt'] <- "NT" # upgrade to NT
 redlist_cat[redlist_cat == 'LR/nt'] <- "LC"
 redlist_df <- tibble(species, redlist_cat)
+
[…]

### 2021-05-11 data/species_ranges.r

@@ -342,7 +342,7 @@ ret_lucif_rel <- 10064
 austria_area <- 32386
 austria_rel_area <- 4385
 ranges_df[ranges_df$species == 'Reticulitermes grassei', ]$range <- ret_grassei_rel / austria_rel_area * austria_area
-ranges_df[ranges_df$species == 'Reticulitermes lucifugus', ]$range <- log10(ret_lucif_rel / austria_rel_area * austria_area)
+ranges_df[ranges_df$species == 'Reticulitermes lucifugus', ]$range <- ret_lucif_rel / austria_rel_area * austria_area
 
 
 ## Daphnia

### 2021-05-11 notebooks/figures/Nc_figure.r

@@ -21,7 +21,8 @@ phi <- mean(rstan:::extract(dps_fit, pars='phi')$phi)
 beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
 
 # alias this -- less typing
-d <- da_dps %>% filter(!is.na(log10_popsize)) %>% 
+d <- da_dps %>% filter(species != 'Equus ferus przewalskii') %>%
+  filter(!is.na(log10_popsize)) %>% 
       group_by(phylum) %>%  
       mutate(n = n(), phyla_mean = mean(log10_popsize)) %>%
       filter(n > 5) %>%
@@ -35,7 +36,7 @@ d %>% ggplot() + geom_density(aes(log10_popsize, fill=phylum), binwidth=0.7, alp
 # what to include
 # x <- d$pred_log10_N  # the simplistic unregularized version
 lx <- 4
-ux <- 18
+ux <- 16
 
 
 output <- TRUE
@@ -88,7 +89,7 @@ labs[1:3] <- latex2exp::TeX(sprintf("  $10^{%d}$", xseq[1:3]))
 axis(3, xseq, labels=labs, 
      line=1.4,  padj=3,
      las=1, tck=0.01)
-axis(3, xseq, labels=rep("", 8),
+axis(3, xseq, labels=rep("", 7),
      tck=-0.01,
      line=1.4)
 
@@ -130,7 +131,7 @@ labs <- labs[or]
 text(ys, xs, labs, srt=0, pos=sides, cex=0.5, xpd=TRUE)
 
 phyla_cols_s <- phyla_cols[unique(d$phylum)]
-legend(14, 32, names(phyla_cols_s), fill=phyla_cols_s,
+legend(13, 32, names(phyla_cols_s), fill=phyla_cols_s,
        bty='n', border=0, cex=0.7, ncol=1)
 
 

### 2021-05-11 notebooks/figures/div_popsize_post.r

@@ -35,10 +35,10 @@ par(mar=c(2, 1, 2, 1), oma=rep(0, 4))
 
 post_plot(phi, main=latex2exp::TeX('intercept ($\\phi$)'), 
           bs_ci=dps_bs_cis[1, ])
-text(-3.7, 1.23, 'OLS', cex=0.9, xpd=TRUE)
+text(-3.84, 1.23, 'OLS', cex=0.9, xpd=TRUE)
 post_plot(beta, main=latex2exp::TeX('slope ($\\beta$)'), 
           bs_ci=dps_bs_cis[2, ])
-text(0.125, 20.3, 'OLS', cex=0.9)
+text(0.158, 18, 'OLS', cex=0.9)
 post_plot(lambda, main=latex2exp::TeX('phylo. signal ($\\lambda$)'),
            xlim=c(0, 1))
 

### 2021-05-11 notebooks/figures/diversity_popsize_full.r

@@ -24,7 +24,7 @@ beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
 d <- da_dps %>% filter(species != 'Equus ferus przewalskii')
 
 #d = read_tsv('../../data/combined_data.tsv')
-#d %>% filter(class == 'Aves') %>% select(species, log10_range, log10_popsize)
+#d %>% filter(genus == 'Reticulitermes') %>% select(species, log10_range, log10_popsize)
  
 # what to include
 # x <- d$pred_log10_N  # the simplistic unregularized version
@@ -75,10 +75,11 @@ d$label <- ifelse(label, d$species, "")
 # simplify some species name (Leffler data has combined them in some cases)
 d$label[d$species == 'Gorilla beringei/gorilla'] <- "Gorilla gorilla"
 d$label[d$species == 'Pongo abelii/pygmaeus'] <- "Pongo abelii"
+d$label[d$species == 'Varecia variegata variegata'] <- 'Varecia variegata'
 d$label[d$species == 'Nomascus gabriellae/leucogenys'] <- "Nomascus gabriellae"
 d$label[d$species == 'Aquila clanga/pomarina'] <- "Aquila clanga"
 
-set.seed(2)
+set.seed(3)
 
 xlims <- c(4, 16)
 ylims <- c(-4.1, 0)
@@ -96,10 +97,11 @@ if (FORCE || !file.exists(REPEL_CACHED_FILE)) {
                                 label=label, text=species), 
                                 size=2.1,
                                 point.padding=10, 
-                                force = 1.5,
-                                min.segment.length=0.5,
+                                push_force = 1,
+                                pull_force = 2,
+                                min.segment.length=0.1,
                                 seed=1,
-                                box.padding=1e-1, 
+                                box.padding=1.3e-1, 
                                 ylims=ylims,
                                 max.iter=1000000, xlim=xlims) +
     xlim(xlims[1], xlims[2]) + theme(legend.position = "none") + 
@@ -191,7 +193,7 @@ if (Kalleles == 4 || Kboth) {
 # y <- predict(div_fit, newdata=data.frame(log10_popsize=logN))
 # lines(logN, y, lty=2)
 # abline(dps_ols, lty=2, col='gray52')
-z <- seq(xlims[1], xlims[2], length.out=100)
+z <- seq(3.5, xlims[2], length.out=100)
[…]

### 2021-05-11 notebooks/figures/diversity_popsize_redlist.r

@@ -33,7 +33,7 @@ d$redlist_cat <- factor(d$redlist_cat, levels=iucn_levels)
 x <- d$log10_popsize
 y <- d$log10_diversity
 
-xlims <- c(4, 18)
+xlims <- c(4, 16)
 plot(x, y)
 
 output <- TRUE
@@ -55,7 +55,7 @@ plot(x, y, col=iucn_cols[d$redlist_cat], type='n', axes=FALSE,
 adjust <- 0
 points(x, y, pch=21, cex=1.3, lwd=0.4,
        bg=iucn_cols[d$redlist_cat], col='white')
-xseq <- seq(4, 18, 2)
+xseq <- seq(4, 16, 2)
 axis(1, xseq, 
      labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
      padj = -1,
@@ -70,7 +70,7 @@ axis(2, yseq,
 
 y <- -0
 m <- 0.12
-x <- 18
+x <- 16
 n <- 0.25
 for (i in seq_along(iucn_levels)) {
   status <- iucn_levels[[i]]

### 2021-05-11 notebooks/figures/figure_3.r

@@ -44,11 +44,12 @@ pred  <- colMeans(mu_ln_fit)
 social <- da_ml$social
 
 da_ml <- da_ml %>% filter(phylum != 'Nematoda')
+#filter(species != 'Equus ferus przewalskii') %>%
 
 plot(map_length ~ log10_popsize, da_ml, type='n', 
      axes=FALSE,
      ylim=c(0, 60),
-     xlim=c(6, 18), 
+     xlim=c(6, 15), 
      ylab="", 
      xlab="")
 ci_polygon(x, fit=ln_fit, par='mu_rep', col=scales::alpha('gray88', 0.4))
@@ -92,7 +93,7 @@ title(ylab="recombination map length (Morgans)",
 
 
 
-xseq <- seq(6, 18, 2)
+xseq <- seq(6, 14, 2)
 axis(1, xseq, 
      labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
      padj = -0.9,
@@ -161,14 +162,14 @@ Nc <- dml$log10_popsize
 plot(Nc, log10(dml$pi_RHH_BGS_1em9), xlab='', 
      type='n', axes=FALSE,
      ylab='', ylim=c(-3.25, 0),
-     xlim=c(4, 18),
+     xlim=c(4, 15),
      col=phyla_cols[da_ml$phylum], pch=1)
 logN <- seq(0, max(Nc), length.out=100)
 y1 <- (log10(4e-8) + logN - log10(1 + 4/3 * 4e-8 * 10^logN))
 y2 <- (log10(4e-9) + logN - log10(1 + 4/3 * 4e-9 * 10^logN))
 polygon(c(logN, rev(logN)), c(y1, rev(y2)), border=NA, 
         col=scales::alpha('gray88', 0.4))
-lines(x <- seq(10, 18), rep(log10(3/4), length(x)), col='gray88', 
+lines(x <- seq(10, 15), rep(log10(3/4), length(x)), col='gray88', 
       lty=2, lwd=0.1, lend=1)
 
 convex_interval <- function(x, y1, y2, n=100, df=10, 
@@ -232,7 +233,7 @@ text(9.5, -0.6, latex2exp::TeX("$\\pi_{BGS+HH}$"), cex=0.8)
 # dml <- dml %>%  rowwise() %>%
   # mutate(y = mean(log10(pi_RHH_BGS_1em8), log10(pi_RHH_BGS_1em9)))
 # l <- loess(y ~ log10_popsize, dml, span=0.2)
[…]

### 2021-05-11 manuscript/manuscript.tex
updated figures

@@
 
-  \caption{The regression estimates of full IUCN Red List population size model
+  \caption{TODO The regression estimates of full IUCN Red List population size model
     for diversity, $\log_{10}(\pi) = \beta_0 + \beta_{LC} LC + \beta_{NT} NT +

### 2021-05-11 notebooks/figures/corbett_detig_fig.r

@@ -31,7 +31,7 @@ phyla_cols_s <- phyla_cols[unique(dca_noself$phylum)]
 legend(8, -1, names(phyla_cols_s), 
        fill=phyla_cols_s, bty='n', border=0, cex=0.6, ncol=2)
 
-xseq <- seq(6, 18, 2)
+xseq <- seq(6, 16, 2)
 axis(1, xseq, 
      labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
      padj = -0.9,
@@ -47,7 +47,7 @@ title(ylab="diversity",
 plot(dml_full$log10_popsize, log10(dml_full$Ne_N_RHH_BGS), 
      axes=FALSE, pch=19,
      col=wes_palette("Darjeeling1")[3],
-     xlim=c(6, 18))
+     xlim=c(6, 16))
 
 points(dca_noself$log10_popsize, log10(1-dca_noself$impact_of_sel), pch=19, 
        col=wes_palette("Darjeeling1")[1])
@@ -71,7 +71,7 @@ axis(2, yseq,
 #      col=phyla_cols[dca$phylum])
 
 cols = wes_palette('Darjeeling1')[1:3]
-legend(6, -10, 
+legend(6, -8, 
        c( latex2exp::TeX("Corbett-Detig et al."),
          latex2exp::TeX("RHH + BGS, $N = \\pi / 4\\mu$"),
          latex2exp::TeX("RHH + BGS, $N = N_c$")),

### 2021-05-11 notebooks/figures/diversity_popsize_averages.r

@@ -33,7 +33,7 @@ if (output)
 
 par(mar=c(4, 4, 2, 2))
 
-xlims <- c(4, 18)
+xlims <- c(4, 16)
 has_label <- d$label != ""
 plot(x, y, col=phyla_cols[d$phylum], type='n', axes=FALSE,
      ylab='', xlab='', ylim=c(-4.1, -0.9), xlim=xlims) 
@@ -41,7 +41,7 @@ plot(x, y, col=phyla_cols[d$phylum], type='n', axes=FALSE,
 # text(6, -0.9, latex2exp:::TeX("$\\mu = 10^{-8}$"), cex=0.5, col='gray52', xpd=TRUE)
 # text(9, -0.9, latex2exp:::TeX("$\\mu = 10^{-10}$"), cex=0.5, col='gray52', xpd=TRUE)
 
-logN <- seq(3, 18, length.out=100)
+logN <- seq(3, 16, length.out=100)
 y1 <- log10(4e-8) + logN
 y2 <- log10(4e-10) + logN
 # polygon(c(logN, rev(logN)), c(y1, rev(y2)), border=NA, col='gray88')
@@ -62,7 +62,7 @@ xc <- x[has_label]
 yc <- y[has_label]
 points(x, y, pch=21, cex=1.3, lwd=0.4,
        bg=phyla_cols[d$phylum], col='white')
-xseq <- seq(4, 18, 2)
+xseq <- seq(4, 16, 2)
 axis(1, xseq, 
      labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
      padj = -1,
@@ -94,7 +94,7 @@ title(xlab=latex2exp::TeX("approximate population size"), line=2.5, cex.lab=1.2)
 #            max(ld$log10_size[z], na.rm=TRUE), -1.05 + 0.005*i, 
 #            col=phyla_cols[i], lwd=2)
 # }
-phyla_cols_s <- phyla_cols[unique(d$phylum)]
+phyla_cols_s <- phyla_cols[sort(unique(d$phylum))]
 legend(10, -3.6, names(phyla_cols_s), fill=phyla_cols_s,
        bty='n', border=0, cex=0.6, ncol=3)
 

### 2021-05-11 notebooks/figures/diversity_popsize_full.r

@@ -86,8 +86,8 @@ ylims <- c(-4.1, 0)
 
 
 REPEL_CACHED_FILE <- 'repel_full.tsv'
-FORCE <- TRUE
-#FORCE <- FALSE
+#FORCE <- TRUE
+FORCE <- FALSE
 if (FORCE || !file.exists(REPEL_CACHED_FILE)) {
   p <- ggplot(d) + 
     geom_point(aes(log10_popsize, log10_diversity, 

### 2021-05-11 notebooks/figures/diversity_popsize_linear.r

@@ -26,7 +26,7 @@ y <- d$diversity
 
 set.seed(2)
 
-xlims <- c(4, 18)
+xlims <- c(4, 16)
 
 fit <- lm(y ~ x)
 
@@ -88,7 +88,7 @@ if (Kalleles == 4 || Kboth) {
 
 points(x, y, pch=21, cex=1.3, lwd=0.4,
        bg=phyla_cols[d$phylum], col='white')
-xseq <- seq(4, 18, 2)
+xseq <- seq(4, 16, 2)
 axis(1, xseq, 
      labels=latex2exp::TeX(sprintf("$10^{%d}$", xseq)), 
      padj = -1,
@@ -119,8 +119,8 @@ title(xlab=latex2exp::TeX("approximate population size"), line=2.5, cex.lab=1.2)
 #            max(ld$log10_size[z], na.rm=TRUE), -1.05 + 0.005*i, 
 #            col=phyla_cols[i], lwd=2)
 # }
-phyla_cols_s <- phyla_cols[unique(d$phylum)]
-legend(14, 0.075, names(phyla_cols_s), fill=phyla_cols_s,
+phyla_cols_s <- phyla_cols[sort(unique(d$phylum))]
+legend(12.5, 0.075, names(phyla_cols_s), fill=phyla_cols_s,
        bty='n', border=0, cex=0.6, ncol=2)
 
 if (output) dev.off()

### 2021-05-11 notebooks/figures/diversity_range_bodymass.r

@@ -124,7 +124,7 @@ lines(z, coef(drn_ols)[1] + coef(drn_ols)[2]*z, lwd=1.2, lty=2, col='gray42')
 title(ylab="pairwise diversity", line=2.8, cex.lab=1.2)
 title(xlab=latex2exp::TeX("range (km^2)"), line=2, cex.lab=1.2)
 
-legend(2, 0, names(phyla_cols), fill=phyla_cols,
+legend(2.6, 0, names(phyla_cols), fill=phyla_cols,
        bty='n', border=0, cex=0.7, ncol=3)
 
 

### 2021-05-11 notebooks/figures/genome_size_maplength.r

@@ -43,7 +43,8 @@ set.seed(2)
 xlims <- c(-1.2, 1)
 
 REPEL_CACHED_FILE <- 'repel_genome_size_maplength.tsv'
-if (TRUE || !file.exists(REPEL_CACHED_FILE)) {
+FORCE <- TRUE
+if (FORCE || !file.exists(REPEL_CACHED_FILE)) {
   p <- ggplot(d) + 
     geom_point(aes(log10_genome_size, log10_map_length, 
                    color=phylum), size=2) +
@@ -51,12 +52,12 @@ if (TRUE || !file.exists(REPEL_CACHED_FILE)) {
                     mapping=aes(log10_genome_size, 
                                 log10_map_length, 
                                 label=label),
-                                size=1.5, 
-                                point.padding=0.6, 
+                                size=2.1, 
+                                point.padding=10, 
                                 force=2,
-                                min.segment.length=0.4,
+                                min.segment.length=0.1,
                                 seed=1,
-                                box.padding=0.1, 
+                                box.padding=1.3e-1, 
                                 max.iter=100000, xlim=xlims) +
     xlim(xlims[1], xlims[2]) + theme(legend.position = "none")
   repel_pos <- extract_ggrepel(p)
@@ -77,9 +78,15 @@ if (output)
 
 par(mar=c(4, 4, 2, 2))
 
-has_label <- d$label != ""
 tx <- repel_pos$x
 ty <- repel_pos$y
+tl <- repel_pos$label
+
+td <- repel_pos %>% 
+       left_join(d %>% select(label, y=map_length, x=log10_genome_size),
+      by='label', suffix=c('_text', '_point')) %>%
+       as_tibble() %>% filter(is.finite(x_point), is.finite(y_point))
+
 plot(x, y, col=phyla_cols[d$phylum], 
      type='n', axes=FALSE, ylim=c(-0.3, 2),
      ylab='', xlab='', xlim=xlims) 
@@ -108,14 +115,12 @@ if (ave_lines) {
   }
 }
 
-adjust <- 0
-xc <- x[has_label]
-yc <- y[has_label]
-segments(xc, yc, tx, ty, col=alpha('gray42', 0.5), lwd=0.5)
+
+with(td,
+     segments(x_point, y_point, x_text, y_text, col=alpha('gray42', 0.5), lwd=0.5))
[…]

### 2021-05-11 notebooks/figures/genome_size_popsize.r

@@ -19,7 +19,8 @@ d <- read_tsv('../../data/combined_data.tsv') %>%
       mutate(log10_genome_size = log10(genome_size)) %>% 
       filter_at(vars(log10_genome_size, log10_popsize), 
                 ~ is.finite(.)) %>% 
-      filter(species != "Ambystoma tigrinum")
+      filter(species != "Ambystoma tigrinum") %>%
+      filter(species != "Equus ferus przewalskii")
 
 
 # what to include
@@ -32,7 +33,7 @@ plot(x, y)
 X <- cbind(x, y)
 center <- colMeans(X)
 sigma <- cov(X)
-coords <- ellipse(mu=center, sigma=sigma, alpha=0.3,
+coords <- ellipse(mu=center, sigma=sigma, alpha=0.35,
                   draw=FALSE)
 
 label <- !as.logical(point.in.polygon(x, y, coords[, 1], coords[, 2]))
@@ -42,7 +43,7 @@ d$label <- ifelse(label, d$species, "")
 
 set.seed(2)
 
-xlims <- c(3.9, 18)
+xlims <- c(3.9, 16)
 
 REPEL_CACHED_FILE <- 'repel_genome_size_popsize.tsv'
 # FORCE <- TRUE
@@ -54,12 +55,12 @@ if (FORCE || !file.exists(REPEL_CACHED_FILE)) {
     geom_text_repel(data=d,
                     mapping=aes(log10_popsize, log10_genome_size, 
                                 label=label, text=species),
-                                size=1.5, 
-                                point.padding=0.6, 
+                                size=2.1, 
+                                point.padding=10, 
                                 force=2,
-                                min.segment.length=0.4,
+                                min.segment.length=0.1,
                                 seed=1,
-                                box.padding=0.6, 
+                                box.padding=1.3e-1, 
                                 max.iter=100000, xlim=xlims) +
     xlim(xlims[1], xlims[2]) + theme(legend.position = "none")
   repel_pos <- extract_ggrepel(p)
@@ -80,9 +81,16 @@ if (output)
 
 par(mar=c(4, 4, 2, 2))
 
-has_label <- d$label != ""
 tx <- repel_pos$x
[…]

### 2021-05-11 manuscript/manuscript.tex
fixed tables

@@
 \RequirePackage{mathtools}
-\RequirePackage{tabularx}
 \RequirePackage{graphicx}
@@
 presence in the Red List indicates they are of conservation interest). In
-Supplementary Material Section \namref{supinfo:div-iucn}, I explore more about
+Supplementary Material Section \Newnameref{supinfo:div-iucn}, I explore more about
 the relationships between IUCN Red List status, diversity, and population size.
@@
   \hline
-  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.77 \times 10^{-4}$ & 0.0100 &  68 & $5.41 \times 10^{-5}$ & 4.26 \\
-  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.68 \times 10^{-2}$ & 0.9720 &  68 & $1.26 \times 10^{-3}$ & 24.50 \\
-  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.24 \times 10^{-5}$ & 0.0005 &   3 & $1.76 \times 10^{-4}$ & 0.35 \\
-  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.47 \times 10^{-4}$ & 0.0162 &  13 & $1.36 \times 10^{-4}$ & 16.40 \\
-  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.21 \times 10^{-5}$ & 0.0012 &   2 & $1.25 \times 10^{-4}$ & 2.57 \\
-  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $4.00 \times 10^{-6}$ & 0.0001 &   1 & $4.00 \times 10^{-5}$ & 5.00 \\
+  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.78 \times 10^{-4}$ & 0.0100 &  68 & $5.41 \times 10^{-5}$ & 4.29 \\ 
+  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.70 \times 10^{-2}$ & 0.9719 &  68 & $1.26 \times 10^{-3}$ & 24.70 \\ 
+  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & $1.76 \times 10^{-4}$ & 0.35 \\ 
+  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.54 \times 10^{-4}$ & 0.0163 &  13 & $1.36 \times 10^{-4}$ & 16.60 \\ 
+  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.14 \times 10^{-5}$ & 0.0011 &   2 & $1.25 \times 10^{-4}$ & 2.51 \\ 
+  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $3.89 \times 10^{-6}$ & 0.0001 &   1 & $4.00 \times 10^{-5}$ & 4.87 \\ 
 \end{tabular}}
@@
  \hline
-    $\beta_0$ & -2.70 & -3.00 & -2.40 \\ 
-    $\beta_{LC}$ & -0.37 & -0.55 & -0.19 \\ 
-    $\beta_{NT}$ & -0.19 & -0.80 & 0.41 \\ 
-    $\beta_{VU}$ & -0.31 & -0.81 & 0.19 \\ 
-    $\beta_{EN}$ & -0.39 & -0.71 & -0.06 \\ 
+    $\beta_0$ & -2.80 & -3.20 & -2.50 \\ 
+    $\beta_{LC}$ & -0.39 & -0.57 & -0.21 \\ 
+    $\beta_{NT}$ & -0.22 & -0.83 & 0.39 \\ 
+    $\beta_{VU}$ & -0.34 & -0.84 & 0.16 \\ 
+    $\beta_{EN}$ & -0.40 & -0.73 & -0.07 \\ 
     $\beta_{CR}$ & -0.03 & -0.65 & 0.59 \\ 
-    $\beta_{N_c}$ & 0.06 & 0.04 & 0.09 \\ 
+    $\beta_{N_c}$ &  0.08 & 0.05 & 0.11 \\ 
     \hline
   \end{tabular}
-
   \caption{TODO The regression estimates of full IUCN Red List population size model
@@
   203.99$, $AIC_\text{reduced} =212.46$. }
-
 \end{table}
 \label{supptable:tbl-1}
+
 \begin{figure}[!htb]

### 2021-05-11 notebooks/figures/genome_size_maplength.r

@@ -83,7 +83,7 @@ ty <- repel_pos$y
 tl <- repel_pos$label
 
 td <- repel_pos %>% 
-       left_join(d %>% select(label, y=map_length, x=log10_genome_size),
+       left_join(d %>% select(label, y=log10_map_length, x=log10_genome_size),
       by='label', suffix=c('_text', '_point')) %>%
        as_tibble() %>% filter(is.finite(x_point), is.finite(y_point))
 
@@ -102,7 +102,7 @@ dfit <- d %>% group_by(phylum) %>%
      nest() %>%
      mutate(n = map_int(data, nrow)) %>%
      filter(n > 5) %>%
-     mutate(fit = map(data, ~ lm(log10_map_length ~ log10_genome_size, .)))
+     mutate(fit = purrr:::map(data, ~ lm(log10_map_length ~ log10_genome_size, .)))
 
 ave_lines <- TRUE 
 if (ave_lines) {

### 2021-05-13 manuscript/manuscript.tex
table and biomass sections

@@
 
-The approximate census sizes I have estimated here
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/species_counts.pdf}
-
-  \caption{}
-
-  \label{suppfig:species_counts}
-\end{figure}
+I validated the approximate census sizes I have estimated here using a few
+approaches. First, to check broad consistency, I compared the implied biomass
+from my estimates of body mass and census size per species with estimates of
+the total carbon biomass on earth by phylum \parencite{Bar-On2018-kc}. For
+species $i$ with wet body mass $m_i$ and census size $N_i$, the implied biomass
+is $m_i N_i$. For all species in a phylum $S$, this total sample biomass is
+$b_S = \sum_{i \in S} m_i N_i$. I then compare this wet biomass to the carbon
+biomasses by phylum by \textcite{Bar-On2018-kc}. Across animal species, the
+ratio of dry to wet body mass, and carbon body mass dry body mass varies
+little. In their study, Bar-On et al. assume wet body mass has a 70\% water
+content, and 50\% of dry body mass is carbon mass, leading to a wet body mass
+to carbon mass factor of $\nicefrac{1-0.7}{0.5} = 0.15$. I use this factor to
+convert the total wet biomass to carbon biomass per phylum.
 
@@
 \centering
-\begin{tabular}{lr|rr|rrrr|r}
-  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{4}{c}{\bf Present study}  &  \\
-  phylum & total species ($T$) & biomass ($B$) & prop. biomass & biomass ($b$) &  prop. biomass & num. species ($n$) &prop. total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
+\begin{tabular}{lr|rr|rrrrr|r}
+  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{5}{c}{\bf Present study}  &  \\
+  phylum & total species ($T$) & biomass ($B$) & prop.
+biomass & biomass ($b$) &  prop.
+biomass & num.
+species ($n$) & factor overrepresented &prop.
+total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
   \hline
-  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.78 \times 10^{-4}$ & 0.0100 &  68 & $5.41 \times 10^{-5}$ & 4.29 \\ 
-  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.70 \times 10^{-2}$ & 0.9719 &  68 & $1.26 \times 10^{-3}$ & 24.70 \\ 
-  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & $1.76 \times 10^{-4}$ & 0.35 \\ 
-  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.54 \times 10^{-4}$ & 0.0163 &  13 & $1.36 \times 10^{-4}$ & 16.60 \\ 
-  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.14 \times 10^{-5}$ & 0.0011 &   2 & $1.25 \times 10^{-4}$ & 2.51 \\ 
-  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $3.89 \times 10^{-6}$ & 0.0001 &   1 & $4.00 \times 10^{-5}$ & 4.87 \\ 
+  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.78 \times 10^{-4}$ & 0.0100 &  68 & 0.02 & $5.41 \times 10^{-5}$ & 4.29 \\
+  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.70 \times 10^{-2}$ & 0.9719 &  68 & 2.90 & $1.26 \times 10^{-3}$ & 24.70 \\
+  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & 0.01 & $1.76 \times 10^{-4}$ & 0.35 \\
+  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.54 \times 10^{-4}$ & 0.0163 &  13 & 0.21 & $1.36 \times 10^{-4}$ & 16.60 \\
+  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.14 \times 10^{-5}$ & 0.0011 &   2 & 0.03 & $1.25 \times 10^{-4}$ & 2.51 \\
+  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $3.89 \times 10^{-6}$ & 0.0001 &   1 & 0.02 & $4.00 \times 10^{-5}$ & 4.87 \\
 \end{tabular}}
-
   \caption{How the total carbon biomass estimates by phylum from
@@
     study compared to the Bar-On et al. estimates \textcite{Bar-On2018-kc}
-    indicates chordates are over-represented and arthropods are
-    under-represented in the present study. Total species by phylum estimates
-  are from
+    indicates chordates are overrepresented and arthropods are underrepresented
+    in the present study; the factor that each phylum is overrepresented is
+    given in the eighth column. Total species by phylum estimates are from
     \textcite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
-ratio column is the ratio of total biomass implied by the $N_c$ estimates of
-each species in a phylum to the actual biomass of that phylum.}
-
+    ratio column is the ratio of total biomass implied by the $N_c$ estimates
+    of each species in a phylum to the actual biomass of that phylum.}
+    \label{supp:biomass-tbl}
 \end{table}
+
+First, I compared the relative carbon biomass in this study to the relative
+carbon biomass on earth per phylum. This shows that this study's sample is over
+represents chordate biomass (by a factor of $\approx 3$), and under represents
+in arthropod biomass (by a factor of $0.02$) relative to the proportion of
+carbon biomass of these phyla on earth (see column eight of Supplementary Table
+\ref{supp:biomass-tbl}. Second, to check whether the carbon biomass per phylum
+in the sample was broadly consistent with the total on earth by phylum ($B_S$
+for phylum $S$), I calculated the expected sample biomass if species were
+sampled randomly from the total species in a phylum, ($B_S \times
+\nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
+in phylum $S$, $T_S$ is the total number of species in phylum $S$ on earth).
+The fraction of total species on earth included in the sample in this study is
+depicted in Supplementary Figure \ref{suppfig:species_counts}.
+
+Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
+biomass per phylum (Supplementary Table \ref{supp:biomass-tbl}). The
+consistency is quite close for this rough approach and the non-random sample of
+taxa included in this study. The carbon biomass estimates for chordates implied
+by the census size estimates are $\approx 25$-fold higher than expected, but is
+well within reasonable expectations given that the chordate sample includes
+many larger-bodied domesticated species (and is a biased sample in other ways).
+Similarly, the implied arthropod carbon biomass is quite close to what one
+would expect. Overall, these values indicate that the census size estimates
+here do not lead to implied biomasses per phylum that are outside the range of
+plausibility.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[]{figures/species_counts.pdf}
+  \caption{The fraction of total species on earth included in this study's
+  sample, per class. The color of the points represents phylum, and the size of
+the point represents the absolute number of species by class.}
+  \label{suppfig:species_counts}
+\end{figure}
 

### 2021-05-13 notebooks/main_analysis.Rmd
table and biomass sections

@@
 bm_tbl <- biomass %>% mutate(total_species = as.integer(total_species)) %>%
+           mutate(over = prop_carbon_mass / bd_prop_mass) %>%
            select(phylum, total_species, bd_mass, bd_prop_mass, 
-                  total_carbon_mass, prop_carbon_mass, nspecies, frac, factor)
-
-print(xtable(bm_tbl, digits=c(1, 2, -2, 2, 4, -2, 4, 2, -2, 2)), 
+                  total_carbon_mass, prop_carbon_mass, nspecies, over, frac, factor)
+
+print(xtable(bm_tbl, digits=c(1, 2, -2, 2, 4, -2, 4, 2, 2, -2, 2)), 
       include.rownames=FALSE,

### 2021-05-25 manuscript/manuscript.tex
updates to ms, code

@@
 
+% TODO recheck node height and other p-values
+
 % Neutral theory predicts that genetic diversity increases with population size,
@@
 \begin{abstract}
-  Under neutral theory, genetic diversity in an equilibrium population
-  is expected to increase with population size. However, observed levels of
+
+  Under neutral theory, genetic diversity in an equilibrium population is
+  expected to increase with population size. However, observed levels of
   diversity across metazoans vary only two orders of magnitude, while census
@@
   I revisit Lewontin's Paradox and assess whether current models of linked
-  selection are likely to constrain diversity to this extent.  To quantify the
-  discrepancy between pairwise diversity and census population sizes across
+  selection are capable of constraining diversity to this extent. To quantify
+  the discrepancy between pairwise diversity and census population sizes across
   species, I combine genetic data from 172 metazoan taxa with estimates of
@@
   previously-published estimates of genomic diversity and these approximate
-  census sizes to quantify Lewontin's Paradox. While previous across-taxa
-  population genetic studies have avoided accounting for phylogenetic
+  census sizes to quantify Lewontin's Paradox. To account for phylogenetic
   non-independence, I use phylogenetic comparative methods to investigate the
@@
   with large census sizes have shorter map lengths, leading these species to
-  experience greater reductions in diversity due to linked selection. Even
-  after using high estimates of the strength of sweeps and background
+  likely experience greater reductions in diversity due to linked selection.
+  Even after using high estimates of the strength of sweeps and background
   selection, I find linked selection likely cannot explain the shortfall
@@
 expected diversity at neutral sites in a panmictic population of $N_c$ diploids
-is expected to be $\pi \approx 4N_c \mu$, where $\mu$ is per generation
-mutation rate. Given that metazoan germline mutation rates only differ 10-fold
-($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}), and census sizes
-vary over several orders of magnitude, one would expected under neutral theory
-that heterozygosity should also vary over several orders of magnitude. However,
-early allozyme surveys revealed that heterozygosity levels across a wide range
-of species varied just an order of magnitude \parencite[][p.
-208]{Lewontin1974-jb}; this anomaly is known as Lewontin's ``Paradox of
-Variation". With modern sequencing-based estimates of $\pi$ across taxa ranging
-over only three orders of magnitude ($0.01$--$10\%$, \cite{Leffler2012-zj}),
-Lewontin's paradox has persisted unresolved through the genomics era.
+is expected to be $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per
+generation mutation rate. Given that metazoan germline mutation rates only
+differ 10-fold ($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}),
+and census sizes vary over several orders of magnitude, one would expected
+under neutral theory that heterozygosity should also vary over several orders
+of magnitude. However, early allozyme surveys revealed that heterozygosity
+levels across a wide range of species varied just an order of magnitude
+\parencite[][p.  208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox
+of Variation".  With modern sequencing-based estimates of $\pi$ across taxa
+ranging over only three orders of magnitude ($0.01$--$10\%$,
+\cite{Leffler2012-zj}), Lewontin's paradox has persisted unresolved through the
+genomics era.
 
@@
 populations. Undoubtedly, these opposing views represent a false dichotomy, as
-population genomic studies have uncovered complex demographic histories that
-impact diversity within a species (e.g. \cite{Zhao2013-vd,Palkopoulou2015-bg}),
-as well as evidence that selection depresses genome-wide diversity (e.g.
+population genomic studies have uncovered both complex demographic histories
+that impact diversity within a species (e.g.
+\cite{Zhao2013-vd,Palkopoulou2015-bg}) and evidence that selection depresses
+genome-wide diversity (e.g.
 \cite{Elyashiv2016-vt,Begun1992-ey,Aguade1989-jx,McVicker2009-ax}).
@@
 While chance demographic events like bottlenecks and recent expansions have
-long-term impacts on diversity (equilibrium is reached on the order of size of
-the population), characteristics of the breeding structure such as high
-variance ($V_w$) or skew in reproductive success  also suppress diversity below
-the levels predicted by the census size \parencite{Wright1938-tv}. In species
-like marine animals, females are highly fecund, and dispersing larvae face
-extremely low survivorship, leading to high variance in reproductive success
+long-term impacts on diversity (since mutation-drift equilibrium is reached on
+the order of size of the population), characteristics of the breeding structure
+such as high variance ($V_w$) or skew in reproductive success  also suppress
+diversity below the levels predicted by the census size
+\parencite{Wright1938-tv}. For example, in many marine animals,
+females are highly fecund, and dispersing larvae face extremely low
+survivorship, leading to high variance in reproductive success
 \parencite{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
@@
 of diversity. The earliest mathematical model of hitchhiking was proffered as a
-solution to Lewontin's Paradox \parencite{Maynard_Smith1974-zr}.  Since,
+explanation of Lewontin's Paradox \parencite{Maynard_Smith1974-zr}.  Since,
 empirical observations have demonstrated that linked selection shapes patterns
@@
 these hitchhiking models when the rate of coalescence due to selection is much
-greater than the rate of neutral coalescence \parencite[equation
-22]{Coop2012-cd}. Other selection models cannot alone decouple diversity from
-population size, \emph{ceteris paribus}. For example, the reduction in
-diversity predicted under background selection and polygenic fitness variation
-is a proportion reduction in population size, mediated by the total
-recombination map length
+greater than the rate of neutral coalescence \parencite[e.g.][equation
+22]{Coop2012-cd}. Other linked selection models are proportional to population
+size, and thus cannot decouple diversity, all else equal. For example, models
+of background selection and polygenic fitness variation predict diversity is
+proportional to population size, mediated by the total recombination map length
+and the deleterious mutation rate or fitness variation
 \parencite{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.
@@
 across 40 species, and showed the impact of selection increases with two
-proxies of census population size, species range and with body size. These
-authors argued this is evidence that selection could explain Lewontin's
-Paradox; however, in a re-analysis, \textcite{Coop2016-gx} demonstrated that
-the observed scale of these reductions is insufficient to explain the
+proxies of census population size, species range and with body size. They
+argued this is evidence that selection could explain Lewontin's Paradox;
+however, in a re-analysis, \textcite{Coop2016-gx} demonstrated that the
+observed scale of these reductions is insufficient to explain the
 orders-of-magnitude shortfall between observed and expected levels of diversity
@@
 the relationship between census size and previously-published genomic diversity
-estimates across 172 metazoan taxa within nine phyla, which provides a sense of
-the scale of the divergence between $\pi$ and $N_c$ that leads to Lewontin's
+estimates across 172 metazoan taxa within nine phyla, to provide a sense of the
+scale of the divergence between $\pi$ and $N_c$ that leads to Lewontin's
 Paradox.
 
-Past work looking at the relationship between $\pi$ and $N_c$ has largely
-ignored phylogenetic non-independence across taxa
-\parencite{Felsenstein1985-an}. To address this shortcoming, I account for
-phylogenetic non-independence across taxa using a synthetic time-calibrated
-phylogeny and phylogenetic comparative methods (PCMs). Moreover,
-\textcite{Lynch2011-qv} has argued that since coalescent times are much less
-than divergence times, considering phylogenetic non-independence is unnecessary
-for traits like effective population size. Using PCMs, I test this conjecture
-by estimating phylogenetic signal in the diversity census size relationship,
-and investigating how these traits evolve along the phylogeny.
+Past work looking at the relationship between $\pi$ and $N_c$ has been unable
+to account for phylogenetic non-independence across taxa
+\parencite{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
+comparative methods (PCMs) with a synthetic time-calibrated phylogeny to
+account for shared phylogenetic history. Moreover, it is disputed whether
+considering phylogenetic non-independence is necessary in population genetics,
+since coalescent times are much less than divergence times
+\parencite{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by
+estimating the degree of phylogenetic signal in the diversity census size
+relationship, and investigating how these traits evolve along the phylogeny.
 
 Finally, I explore whether the predicted reductions of diversity under
-background selection and recurrent hitchhiking are sufficiently strong 
-to resolve Lewontin's Paradox. These predicted reductions in diversity across
-species are generously estimated using selection strength parameters from
-\emph{Drosophila melanogaster}, a species known to be strongly affected by
-linked selection. Given the effects of linked selection are mediated by
-recombination map length, I investigate how recombination map lengths vary with
-census population size using data from a previously-published survey
-\parencite{Stapley2017-fs}. I find map lengths are typically shorter in
-large--census-size species, increasing the effects of linked selection in these
-species, which might further decouple diversity from census size. Still, I find
-the combined impact of these selection models with available parameter
-estimates falls short in explaining Lewontin's Paradox, and discuss future
-avenues through which the Paradox of Variation could be fully resolved.
+background selection and recurrent hitchhiking are sufficiently strong to
+resolve Lewontin's Paradox. These predicted reductions in diversity across
+species are estimated using strong selection parameters from \emph{Drosophila
+melanogaster}, a species known to be strongly affected by linked selection.
+Given the effects of linked selection are mediated by recombination map length,
+I investigate how recombination map lengths vary with census population size
+using data from a previously-published survey \parencite{Stapley2017-fs}. I
+find map lengths are typically shorter in large--census-size species,
+increasing the effects of linked selection in these species, which might
+further decouple diversity from census size. Still, I find the combined impact
+of these selection models with available parameter estimates falls short in
+explaining Lewontin's Paradox, and discuss future avenues through which the
+Paradox of Variation could be fully resolved.
 
@@
 surveyed the literature for census size estimates
-\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range and body
+\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range or body
 size or qualitative categories as proxies for census size
@@
 population density and body size is driven by energy budgets, and thus reflects
-macroecological equilibria \parencite{Damuth1987-sg}; consequently, population
-sizes for taxa like humans and their domesticated species are underestimated.
-While these methods to estimate census size are crude and approximate, they can
-be efficiently calculated for numerous taxa and are sufficient to estimate the
-scale of Lewontin's Paradox.
+macroecological equilibria \parencite{Damuth1987-sg}. Consequently, population
+sizes are underestimated for taxa like humans and their domesticated species,
+and overestimated for species with anthropogenically reduced densities or
+patchy ranges. For example, the population size of \emph{Lynx lynx} is likely
+around 50,000 \parencite{Iucn2020-ap} which is around two orders of magnitude
+smaller than my estimate. Additionally, the range size estimates do not
+consider whether an area has unsuitable habitat, and thus may be overestimates
+for species with particular niches or patchy habitats. While these methods to
+estimate census size are crude and approximate, they can be efficiently
+calculated for numerous taxa and are sufficient to estimate the scale of
+Lewontin's Paradox.
 
@@
 significant relationship between heterozygosity and the logarithm of population
-size, but these studies relied on heterozygosity measured from allozyme data
-\parencite{Soule1976-he,Frankham1996-yb,Nei1984-zi}. Here, I confirm these
-findings using pairwise diversity estimates from genomic sequence data and the
-estimated census sizes (Figure \ref{fig:figure-1}). The pairwise diversity
-estimates are from three sources: \textcite{Leffler2012-zj},
+size or range size, but these studies relied on heterozygosity measured from
+allozyme data
+\parencite{Soule1976-he,Frankham1996-yb,Nei1984-zi,Leffler2012-zj}. Here, I
+confirm these findings using pairwise diversity estimates from genomic sequence
+data and the estimated census sizes (Figure \ref{fig:figure-1}). The pairwise
+diversity estimates are from three sources: \textcite{Leffler2012-zj},
 \textcite{Corbett-Detig2015-gt}, and \textcite{Romiguier2014-bp}, and are
@@
 presence in the Red List indicates they are of conservation interest). In
-Supplementary Material Section \Newnameref{supinfo:div-iucn}, I explore more about
-the relationships between IUCN Red List status, diversity, and population size.
+Supplementary Material Section \Newnameref{supinfo:div-iucn}, I explore more
+about the relationships between IUCN Red List status, diversity, and population
+size.
 
@@
 
-In quantifying Lewontin's Paradox, I have additionally fit some simple trait
-evolution models that account for phylogenetic non-independence, investigated
+One limitation of using ordinary least squares is that shared phylogenetic
+history can create correlation structure in the residuals, which violates an
+assumption of the regression model
+\parencite{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I
+fit this relationship using a phylogenetic mixed-effects model, investigated
 whether there is a signal of phylogenetic non-independence, estimated the
-continuous trait values on the phylogeny, and assessed how diversity and
+continuous trait values on the phylogeny, and explored how diversity and
 population size evolve. Prior population genetic comparative studies have
@@
 linear relationship between diversity and population size (on a log-log scale)
-accounting for phylogeny, for the 166 taxa with non-missing data and present in
-the synthetic chronogram. As with the non-phylogenetic regression, this
-relationship was positive and significant (95\% credible interval $0.04,
-0.11$), though somewhat attenuated compared to the OLS estimates (Figure
-\ref{fig:figure-2}B). Since the population size estimates are based on range
-and body mass, they are essentially a composite trait; fitting phylogenetic
-mixed-effects models separately on body mass and range indicates these have
-significant negative and positive effects, respectively (Supplementary Figure
-\ref{suppfig:figure-div-range-bodymass}).
+accounting for phylogeny, for the 166 taxa without missing data and present in
+the synthetic chronogram. As with the linear regression, this relationship was
+positive and significant (95\% credible interval $0.04, 0.11$), though somewhat
+attenuated compared to the OLS estimates (Figure \ref{fig:figure-2}B). Since
+the population size estimates are based on range and body mass, they are
+essentially a composite trait; fitting phylogenetic mixed-effects models
+separately on body mass and range indicates these have significant negative and
+positive effects, respectively (Supplementary Figure
+\ref{suppfig:figure-div-range-bodymass}; see also Supplementary Figure
+\ref{suppfig:div_leffler} for the relationship between diversity and the range
+categories of \cite{Leffler2012-zj}).
 
@@
 
-A closer visual inspection of the estimated ancestral continuous values for
-diversity and population size on the phylogeny indicates the high phylogenetic
-signal seems to be driven in part by chordates having low diversity and small
+This high degree of phylogenetic signal suggests
+\citeauthor{Gillespie1991-qa}'s (1991) concern that the $\pi$--$N_c$
+relationship was driven by chordate-arthropod differences may be valid. A
+visual inspection of the estimated ancestral continuous values for diversity
+and population size on the phylogeny indicates the high phylogenetic signal
+seems to be driven in part by chordates having low diversity and small
 population sizes compared to non-chordates (Figure \ref{fig:figure-2}A). This
-suggests \citeauthor{Gillespie1991-qa}'s (1991) earlier critique that the
-$\pi$--$N_c$ relationship was driven by chordate-arthropod differences may be
-valid. This problem resembles Felsenstein's worst-case scenario
+problem resembles Felsenstein's worst-case scenario
 \parencite{Felsenstein1985-an,Uyeda2018-wf}, where a singular event on a
 lineage separating two clades generates a spurious association between two
-traits. To further investigate whether clade-level differences dominated the
-relationship between diversity and population size, I fit phylogenetic
-mixed-effects models to phyla-level subsets of the data for clades with
-sufficient sample sizes (see Methods: \Newnameref{sec:methods-pcm}). This
-analysis shows a significant positive relationship between diversity and
-population size in arthropods, and positive weak relationships in molluscs and
-chordates (Supplementary Figure \ref{suppfig:div_popsize_post_phyla}). Each of
-the 90\% credible intervals for slope overlap, indicating the relationship
-between $\pi$ and $N_c$ is similar across these clades.
-
-One limitation of the phylogenetic mixed-effects models employed here is that
-they assume traits evolve under constant-rate Brownian motion. To test this
-assumption, I performed node-height tests \parencite{Freckleton2006-ww}.
-Node-height tests regress the absolute values of the standardized contrasts
-between lineages against the branching time (since present) of these lineages.
-Under Brownian Motion (BM), standardized contrasts are estimates of the rate of
-character evolution \parencite{Felsenstein1985-an}; if a trait evolves under
-constant rate BM, this relationship should be flat. For both diversity and
-population size, node-height tests indicate a significant increase in the rate
-of evolution towards the present (robust regression p-values 0.028 and 0.00070
-respectively; Figure \ref{fig:figure-2}C).  Considering the constituents of the
-population size estimate, range and body mass, separately, the rate of
-evolution of range but not body mass shows a significant increase (p-value $1.9
-\times 10^{-7}$) towards the present.
+traits.
+
+To investigate whether clade-level differences dominated the relationship
+between diversity and population size, I fit phylogenetic mixed-effects models
+to phyla-level subsets of the data for clades with sufficient sample sizes (see
+Methods: \Newnameref{sec:methods-pcm}). This analysis shows a significant
+positive relationship between diversity and population size in arthropods, and
+positive weak relationships in molluscs and chordates (Supplementary Figure
+\ref{suppfig:div_popsize_post_phyla}). Each of the 90\% credible intervals for
+slope overlap, indicating the relationship between $\pi$ and $N_c$ is similar
+across these clades.
+
+Additionally, I have explored the rate of trait change through time using
+node-height tests \parencite{Freckleton2006-ww}. Node-height tests regress the
+absolute values of the standardized contrasts between lineages against the
+branching time (since present) of these lineages. Under Brownian Motion (BM),
+standardized contrasts are estimates of the rate of character evolution
+\parencite{Felsenstein1985-an}; if a trait evolves under constant rate BM, this
+relationship should be flat. For both diversity and population size,
+node-height tests indicate a significant increase in the rate of evolution
+towards the present (robust regression p-values 0.028 and 0.00070 respectively;
+Figure \ref{fig:figure-2}C). Considering the constituents of the population
+size estimate, range and body mass, separately, the rate of evolution of range
+but not body mass shows a significant increase (p-value $1.9 \times 10^{-7}$)
+towards the present.
 
@@
 experienced a shift on the lineage separating tunicates and vertebrates. At
-nearly 600 Mya, these deep nodes illustrate a counterexample to Lynch's claim
-that expected coalescent times do not share phylogenetic history because they
-are less than divergence times.
+nearly 600 Mya, these deep nodes illustrate that expected coalescent times can
+share phylogenetic history, due to phylogenetic inertia in some combination of
+population size, reproductive system, and mutation rates.
 
@@
 diversity ($\pi$) to the estimated diversity in the absence of linked selection
-($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. There are two difficulties in
-evaluating whether linked selection could resolve Lewontin's Paradox. The first
-difficulty is that $\pi_0$ is unobserved. Previous work has estimated $\pi_0$
-using methods that exploit the spatial heterogeneity in recombination and
-functional density across the genome to fit linked selection models that
-incorporate both hitchhiking and background selection
+($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. Here, $\pi_0$ would reflect demographic
+history and non-heritable variation in reproductive success. There are two
+difficulties in evaluating whether linked selection could resolve Lewontin's
+Paradox. The first difficulty is that $\pi_0$ is unobserved.  Previous work has
+estimated $\pi_0$ using methods that exploit the spatial heterogeneity in
+recombination and functional density across the genome to fit linked selection
+models that incorporate both hitchhiking and background selection
 \parencite{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second difficulty is
@@
 Since we lack estimates of key linked selection parameters across species, I
-generously parameterize the hitchhiking and BGS models using estimates from
+parameterize the hitchhiking and BGS models using estimates from
 \emph{Drosophila melanogaster}, a species known to be strongly affected by
@@
 and $U$) to high values estimated from \emph{Drosophila melanogaster} by
-\textcite{Elyashiv2016-vt}. My estimate of $\gamma_\text{Dmel}$ based Elyashiv
-et al. implies a rate of sweeps per basepair of $\nu_\text{BP,Dmel} \approx
-2.34 \times 10^{-11}$, which is close to other estimates from \emph{D.
-melanogaster} (see Supplementary Figure \ref{suppfig:linked-sel-params}A). The
-rate of deleterious mutations per diploid genome, per generation is
-parameterized using the estimate from Elyashiv et al., $U_\text{Dmel} = 1.6$,
-which is a bit greater than previous estimates based on Bateman-Mukai
-approaches \parencite{Mukai1985-bc,Mukai1988-vs,Charlesworth1987-ab}. Finally,
-the probability that a lineage is trapped in a sweep, $J_\text{Dmel}$, is
+\textcite{Elyashiv2016-vt}. My estimate of the adaptive substitutions per
+generation ($\gamma_\text{Dmel}$) based Elyashiv et al. implies a rate of
+sweeps per basepair of $\nu_\text{BP,Dmel} \approx 2.34 \times 10^{-11}$, which
+is close to other estimates from \emph{D.  melanogaster} (see Supplementary
+Figure \ref{suppfig:linked-sel-params}A). The rate of deleterious mutations per
+diploid genome, per generation is parameterized using the estimate from
+Elyashiv et al., $U_\text{Dmel} = 1.6$, which is a bit greater than previous
+estimates based on Bateman-Mukai approaches
+\parencite{Mukai1985-bc,Mukai1988-vs,Charlesworth1987-ab}. Finally, the
+probability that a lineage is trapped in a sweep, $J_\text{Dmel}$, is
 calculated from the estimated genome-wide average coalescent rate due to sweeps
@@
   \caption{(A) The observed relationship between recombination map length ($L$)
-    and census size ($N_c$) across 131 species with complete data and known
-    phylogeny. Triangle points indicate social taxa excluded from the model
+    and census size ($N_c$) across 136 species with complete data and known
+    phylogeny. Triangle points indicate six social taxa excluded from the model
     fitting since these have adaptively higher recombination map lengths
@@
 magnitude in \emph{Drosophila} (Supplementary Figured
-\ref{suppfig:corbett_detig}B). Overall, while linked selection could decouple
-diversity from census size for high-$N_c$ species, recurrent hitchhiking and
-background selection seem unlikely to explain the observed patterns of
-diversity across species under our understanding of the range of parameter
-estimates.
+\ref{suppfig:corbett_detig}B). The high reductions in $\pi$ predicted here
+(compared to those of \cite{Elyashiv2016-vt}) are a result of using $N_c$,
+rather than $N_e = \nicefrac{\pi_0}{4 \mu}$ in the denominator of Equation
+\eqref{eqn:div}, which leads to a very high rate of sweeps in the population.
+I do not consider selective interference, though the saturation of adaptive
+substitutions per Morgan would only act to limit the reduction in diversity
+\parencite{Weissman2012-aa}, and thus these results are conservative.
+Overall, while linked selection could decouple diversity from census size for
+high-$N_c$ species, recurrent hitchhiking and background selection seem
+unlikely to explain the observed patterns of diversity across species under our
+understanding of the range of parameter estimates.
 
@@
 
-Lewontin's Paradox arises from a comparison of diversity across species, yet
-surprisingly, previous work on this problem has not considered the impact of
-phylogenetic non-independence. I have addressed this limitation, showing that
-diversity does have a significant positive relationship with census size, after
-accounting for shared phylogenetic history among taxa. Additionally, I find a
-high degree of phylogenetic signal, and that arthropods and chordates form
-clusters, showing previous concern about phylogenetic non-independence was
-warranted \parencite{Gillespie1991-qa}. Finally, this high degree of
-phylogenetic signal, as well as evidence of shifts in the rate of evolution of
-genetic diversity on deep timescales in molluscs and chordates, seem to
-contradict Lynch's (2011) claim that since coalescent times are much less than
-divergence times, they are not affected by shared phylogenetic history.
-
-One can reconcile my findings with Lynch's claim by considering what
-evolutionary, ecological, life history, and demographic causal factors
-determine coalescent rates across species, and how these factors evolve across
-deep timescales. Lynch's conjecture that coalescent times should be free of
-phylogenetic signal may be true \emph{were we to condition} on these causal
-factors that could be affected by shared phylogenetic history. In contrast, my
-estimates of phylogenetic signal in diversity are not conditioned on these
+Lewontin's Paradox arises from a comparison of diversity across species, yet it
+has been disputed whether such comparisons require phylogenetic comparative
+methods. Extending previous work that has accounted for phylogeny in particular
+clades \parencite{Leffler2012-zj}, or using taxonomical-level averages
+\parencite{Romiguier2014-bp}, I show that the positive relationship between
+diversity and census size is significant using a mixed-effects model with a
+time-calibrated phylogeny. Additionally, I find a high degree of phylogenetic
+signal, evidence of deep shifts in the rate of evolution of genetic diversity,
+and that arthropods and chordates form clusters. Overall, this suggests that
+previous concerns about phylogenetic non-independence in comparative population
+genetic studies were warranted \parencite{Gillespie1991-qa,Whitney2010-ud}.
+Notably, Lynch (2011) has argued that PCMs for pairwise diversity are
+unnecessary, since mutation rate evolution is fast and thus free of
+phylogenetic inertia, sampling variance should exceed the variance due to
+phylogenetic shared history, and coalescent times are much less than divergence
+times. Since my findings suggest PCMs are necessary in some cases, it is
+worthwhile to address these points.
+
+First, Lynch has correctly pointed out that while coalescent times are much
+less than divergence times and should be free of phylogenetic shared history,
+the factors that determine coalescent times (e.g.  mutation rates and effective
+population size) may not be (2011). In other words, coalescent times are free
+from phylogenetic shared history \emph{were we to condition} on these causal
+factors that could be affected by shared phylogenetic history. My estimates of
+phylogenetic signal in diversity, by contrast, are not conditioned on these
 factors. Importantly, even ``correcting for" phylogeny implicitly favors
@@
 causal factors determine coalescent times across species, as well as how these
-factors evolve across macroevolutionary timescales.
-
-Furthermore, beyond just accounting for phylogenetic non-independence,
+factors evolve across macroevolutionary timescales. Second, it is a
+misconception that a fast rate of trait evolution necessarily reduces
+phylogenetic signal \parencite{Revell2008-in}, and that if either or both
+variables in a regression are free of phylogenetic signal, PCMs are unnecessary
+\parencite{Revell2010-mf,Uyeda2018-wf}. The evidence of high phylogenetic
+signal found in this study suggests PCMs are needed, in part to avoid spurious
+results from phylogenetic pseudoreplication.
+
+Finally, beyond just accounting for phylogenetic non-independence,
 macroevolution and phylogenetic comparative methods are a promising way to
@@
 linked selection poorly fit observed patterns of diversity across species.
-While this suggests these two common modes of linked selection seem unlikely to
-explain across-taxa patterns of diversity, there are three major potential
-limitations of my approach that need further evaluation.
+This result extends the analysis by \textcite{Coop2016-gx} showing that levels
+of $\pi_0$ estimated by \textcite{Corbett-Detig2015-gt} are not decoupled from
+genome-wide average $\pi$, as would occur if linked selection were to explain
+Lewontin's Paradox. Here, my analysis goes a step further and suggests that
+models of recurrent hitchhiking and background selection are not capable of
+explaining the observed relationship between $\pi$ and census size, in part
+because mid-$N_c$ species have sufficiently long recombination map lengths to
+diminish even strong selection. This finding supports the idea the levels of
+diversity across species are primarily determined by past demographic
+fluctuations. Overall, while this suggests these two common modes of linked
+selection seem unlikely to explain across-taxa patterns of diversity, there are
+three major potential limitations of my approach that need further evaluation.
 
@@
 capture the heterogeneity of functionality density and recombination rate in
-real genomes, it is still extraordinary conservative, giving linked selection
-the best possible chance to decouple diversity from census size and explain
-Lewontin's Paradox. This is in part because the strong selection parameter
-estimates from \emph{Drosophila melanogaster} used, but also because I assume
-that the effective population size is the census size. Even then, this
-decoupling only occurs in very high--census-size species, and implies that the
-diversity in the absence of linked selection, $\pi_0$, is currently
-underestimated by several orders of magnitude. Moreover, the study of
-\textcite{Corbett-Detig2015-gt} did consider recombination rate and functional
-density heterogeneity in estimating the reduction due to linked selection
-across species, yet their predicted reductions are orders of magnitude weaker
-than those considered here by assuming that $N_e = N_c$ (Supplementary Figure
-\ref{suppfig:corbett_detig}B).  Overall, even with more realistic models of
-linked selection, current models of linked selection seem fundamentally unable
-to fit the diversity--census-size relationship.
+real genomes, it is still extraordinary conservative, likely overestimating the
+effects of linked selection to see if it could be capable of decoupling
+diversity from census size and explain Lewontin's Paradox. This is in part
+because the strong selection parameter estimates from \emph{Drosophila
+melanogaster} used, but also because I assume that the effective population
+size is the census size.  Even then, this decoupling only occurs in very
+high--census-size species, and implies that the diversity in the absence of
+linked selection, $\pi_0$, is currently underestimated by several orders of
+magnitude. Moreover, the study of \textcite{Corbett-Detig2015-gt} did consider
+recombination rate and functional density heterogeneity in estimating the
+reduction due to linked selection across species, yet their predicted
+reductions are orders of magnitude weaker than those considered here by
+assuming that $N_e = N_c$ (Supplementary Figure \ref{suppfig:corbett_detig}B).
+Overall, even with more realistic models of linked selection, current models of
+linked selection seem fundamentally unable to fit the diversity--census-size
+relationship.
 
@@
 well approximated by quantitative genetic linked selection models (QGLS
-\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}). The reduction in
-diversity under these models is nearly identical to that under background
-selection models, in part because deleterious alleles at mutation-selection
-balance constitute a considerable component of fitness variation (see Appendix
-Section \ref{app:bgs}; \cite{Charlesworth2000-km,Charlesworth2015-am}).
-Finally, I also disregarded the interaction of sweeps and spatial processes.
-For populations spread over wide ranges, limited dispersal slows the spread of
-sweeps, allowing for new beneficial alleles to arise, spread, and compete
-against other segregating beneficial variants
-\parencite{Ralph2015-kl,Ralph2010-ki}. Through limited dispersal should act to
-``soften sweeps" and not impact my findings for the reasons described above,
-future work could investigate how these processes impact diversity in ways not
-captured by hard sweep models.
+\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}). Third, the parameters
+from \textcite{Elyashiv2016-vt} are robust to many types of sweeps that result
+in a substitution (e.g. see p. 19 of their Supplementary Online Materials).
+The reduction in diversity under these models is nearly identical to that under
+background selection models, in part because deleterious alleles at
+mutation-selection balance constitute a considerable component of fitness
+variation (see Appendix Section \ref{app:bgs};
+\cite{Charlesworth2000-km,Charlesworth2015-am}).  Finally, I also disregarded
+the interaction of sweeps and spatial processes.  For populations spread over
+wide ranges, limited dispersal slows the spread of sweeps, allowing for new
+beneficial alleles to arise, spread, and compete against other segregating
+beneficial variants \parencite{Ralph2015-kl,Ralph2010-ki}. Through limited
+dispersal should act to ``soften sweeps" and not impact my findings for the
+reasons described above, future work could investigate how these processes
+impact diversity in ways not captured by hard sweep models.
 
@@
 
-Lewontin's Paradox shows the extent to which the effective population sizes
+Lewontin's Paradox describes the extent to which the effective population sizes
 implied by diversity, $\widetilde{N}_e$, diverge from census population sizes.
@@
 \parencite{Wang2016-mi,Caballero1994-ao,Caballero2020-wm,Galtier2020-fb}. These
-include estimators based on the observed decay in linkage disequilibrium or
-temporal estimators that use the variance in allele frequency change. These
-alternate estimators capture summaries of the effective population size on
-shorter timescales than coalescent-based estimators \parencite{Wang2005-wy},
-and thus could be used to tease out processes that impact the $N_e$-$N_c$
-relationship in the more recent past.
+include estimators based on the site frequency spectrum, observed decay in
+linkage disequilibrium, or temporal estimators that use the variance in allele
+frequency change. These alternate estimators capture summaries of the effective
+population size on shorter timescales than coalescent-based estimators
+\parencite{Wang2005-wy}, and thus could be used to tease apart processes that
+impact the $N_e$-$N_c$ relationship in the more recent past.
 
@@
 \parencite{Frankham1995-xz,Mace1991-ai} and in understanding evolution in
-highly-fecund marine species. Surveys of the $\nicefrac{N_e}{N_c}$ relationship
-across taxa indicate mean $\nicefrac{N_e}{N_c}$ is on order of $\approx 0.1$
-\parencite{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though the
-uncertainty in these estimates is high, and some species with sweepstakes
+highly fecund marine species. Surveys of the short-term $\nicefrac{N_e}{N_c}$
+relationship across taxa indicate mean $\nicefrac{N_e}{N_c}$ is on order of
+$\approx 0.1$ \parencite{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though
+the uncertainty in these estimates is high, and some species with sweepstakes
 reproduction systems like Pacific Oyster (\emph{Crassostrea gigas}) can have
@@
 ecological processes occurring on vastly different timescales. Lewontin's
-Paradox of Variation may persist unresolved for some time because the solution
-requires synthesis and model building at the intersection of all these
-disciplines.
+Paradox of Variation may persist unresolved for some time because the
+explanation requires synthesis and model building at the intersection of all
+these disciplines.
 
@@
 averaged. The total recombination map length data were from both
-\citeauthor{Stapley2017-fs} (2017; 30 taxa), and
-\citeauthor{Corbett-Detig2015-gt} (2015; 11 taxa). Both studies used
+\citeauthor{Stapley2017-fs} (2017; 127 taxa), and
+\citeauthor{Corbett-Detig2015-gt} (2015; 9 taxa). Both studies used
 sex-averaged recombination maps estimated with cross-based approaches; in some
@@
 (\cite{Chamberlain2020-tv,Iucn2020-ap}; Supplementary Figure
-\ref{suppfig:redlist-ranges}). Each census population size is then estimated as
-the product of range and density.
+\ref{suppfig:redlist-ranges}). Each census population size
+is then estimated as the product of range and density.
 
@@
 This calibrated synthetic phylogeny was then subset for the analyses based on
-what species had non-missing trait data. The diversity-population size
+what species had complete trait data. The diversity-population size
 relationship assessed by a linear phylogenetic mixed-effects model implemented
@@
 phylogenetic comparative methods; yet I take full responsibility for any
-shortcomings of my analysis.  I would like to also thank UO librarian Dean
-Walton for helping me track down some rather difficult to find older papers.
-This work was supported by an NIH Grant (1R01GM117241) awarded to Andrew Kern.
+shortcomings of my analysis. Finally, I am indebted to helpful feedback from
+Guy Sella, Matt Pennell, and two other anonymous reviewers. I would like to
+also thank UO librarian Dean Walton for helping me track down some rather
+difficult to find older papers. This work was supported by an NIH Grant
+(1R01GM117241) awarded to Andrew Kern.
 
@@
 \end{align}
-
+%
 This is analogous to the second term of \textcite{Coop2012-cd} equation 17,
@@
 carbon biomass on earth per phylum. This shows that this study's sample is over
-represents chordate biomass (by a factor of $\approx 3$), and under represents
+represents chordate biomass (by a factor of $\mathord{\sim}3$), and under represents
 in arthropod biomass (by a factor of $0.02$) relative to the proportion of
@@
 taxa included in this study. The carbon biomass estimates for chordates implied
-by the census size estimates are $\approx 25$-fold higher than expected, but is
+by the census size estimates are $\mathord{\sim}25$-fold higher than expected, but is
 well within reasonable expectations given that the chordate sample includes
@@
 
+Additionally, note that the body mass based estimates of density for
+\emph{Drosophila} are similar to previously used estimates in surveys of census
+size and diversity. \textcite{Nei1984-zi} suggested a maximum of 5
+\emph{Drosophila} per $\text{m}^2$, including regions of the range that are not
+inhabitable. Across \emph{Drosophila}, the body mass based estimates suggest
+$10^{6.7}-10^{7.6}$ individuals per $\text{km}^2$, or $4.5-36.3$ individuals
+per $\text{m}^2$, which are consistent with this previous estimate.
+\citeauthor{Nei1984-zi}'s estimates of \emph{Drosophila pseudoobscura}'s census
+size are four orders of magnitude smaller than mine, but their approach uses a
+speculated ratio of population sizes of different \emph{Drosophila} species
+rather than range sizes \parencite[p. 81]{Nei1984-zi}.
+
+As another consistency check, I looked at the rank order of mammals by biomass.
+Whale species have the first and third highest biomass with 11.4 and 3.8
+megatons of carbon biomass (for \emph{Balaenoptera bonaerensis} and
+\emph{Eschrichtius robustus}, respectively). While this seems high, a recent
+study shows that across whale species, pre-whaling carbon biomass was at the
+tens of megatons level \parencite[Table 1 and Figure 1]{Pershing2010-fh}.
+Given that my census size estimates represent populations at a macroecological
+equilibrium, they would not reflect reduced density due to whaling or other
+anthropogenic causes. Humans had the second largest biomass, followed by wolf
+species (\emph{Canis lupus} and \emph{C. latrans}); as with whales, the
+population sizes for wolf species represent pre-anthropogenic densities and are
+overestimates compared to current population sizes as expected.
+
+Finally, there are other estimates of approximate population sizes for some
+species that I compared my estimates to. The United Nation's FAOSTAT database
+estimates the total number of horses (\emph{Equus caballus}) on earth as
+$\mathord{\sim}60$ million; the estimate in this study is close to 40 million.
+For other domesticated species like chicken (\emph{Gallus gallus}), estimates
+range from 25 million to 19.6 billion
+\parencite{noauthor_2021-gc,Robinson2014-sx}; the present study's estimate lies
+in the middle at $\mathord{\sim}175$ million. Again, this is a known limitation
+of this method, as the range is estimated from occurrence data and does not
+consider species' niches. This present study's estimate of the number of king
+penguins (\emph{Aptenodytes patagonicus}) is about 3 million; the population
+size was recently estimated as 2.23 million pairs \parencite{Shirihai2008-ft}.
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[]{figures/diversity_leffler_ranges.pdf}
+
+  \caption{Pairwise diversity grouped by the range categories from
+  \textcite{Leffler2012-zj}, with point size indicating the predicted
+   population density. The vertical lines are the range category group means.}
+
+  \label{suppfig:div_leffler}
+\end{figure}
+
 \begin{figure}[!htb]
@@
 \end{figure}
+
+\section{Sensitivity Analysis}
 

### 2021-05-25 notebooks/main_analysis.Rmd
updates to ms, code

@@
 
-Body size and map length:
+Body size and chromosome number:
 
@@
   geom_smooth(method='lm') + ylab('haploid chromosome number') + 
+  xlab('log10(size)')
+
+```
+
+Range and map length:
+
+```{r}
+
+d %>% 
+  filter_at(vars(hcn, log10_size), ~ !is.na(.)) %>%
+  ggplot(aes(log10_range, map_length, color=phylum)) + geom_point() + 
+  geom_smooth(method='lm') + ylab('map length (Morgans)') + 
   xlab('log10(size)')
@@
 
-xd %>% filter(class == 'Mammalia') %>% arrange(desc(carbon_mass)) %>% 
-  select(species, carbon_mass) %>% 
+xd %>% 
+  filter(class == 'Mammalia') %>% 
+  arrange(desc(carbon_mass)) %>% 
+  select(species, carbon_mass, log10_popsize) %>% 
   as.data.frame()

### 2021-05-25 notebooks/figures/diversity_leffler_ranges.r

@@ -1,0 +1,85 @@
+library(tidyverse)
+library(wesanderson)
+library(RColorBrewer)
+
+range_cat_levels <- c("Island", "Narrow endemic", "Broad endemic", "Cosmopolitan")
+
+d <- read_tsv('../../data/combined_data.tsv') %>%
+       filter(is.finite(log10_diversity), is.finite(range), !is.na(range_cat)) %>%
+       mutate(log10_range = log10(range)) %>%
+       mutate(range_cat = ifelse(range_cat == 'Unknown', NA, range_cat)) %>%
+       mutate(range_cat = factor(range_cat, levels=range_cat_levels, ordered=TRUE)) %>%
+       arrange(range_cat, log10_diversity) %>%
+       # space the groups
+       mutate(xpos = seq(1, nrow(.)) + 2.2*as.integer(range_cat)-1) 
+
+means <- d %>% group_by(range_cat) %>% 
+  summarize(xmin=min(xpos), 
+            xmax=max(xpos),
+            log10_diversity = mean(log10_diversity, na.rm=TRUE))
+
+# ggplot(d, aes(xpos, log10_diversity, color=range_cat, size=10^log10_density)) + geom_point()  + facet_wrap(~range_cat)  + geom_text_repel(aes(xpos, log10_diversity, label=species))
+
+opar <- par(no.readonly=TRUE)
+# pdf('range_descriptions.pdf', width=8, height=5)
+pdf('diversity_leffler_ranges.pdf', width=5, height=8)
+
+par(mar=c(4, 1, 1, 5))
+cols <- wes_palette('Darjeeling1')[c(5, 2, 3, 1, 4)]
+
+scaler <- function(x, range=c(1, 3)) {
+  rng <- range(x, na.rm=TRUE)
+  slope = abs(diff(range)) / (rng[2] - rng[1])
+  min(range) + slope*x
+}
+
+plot(d$log10_diversity, d$xpos, pch = 21, col='white',
+     bg=cols[d$range_cat], 
+     lwd=0.7,
+     axes=FALSE,
+     cex=scaler(d$log10_density),
+     xlab = '',
+     ylab = '')
+
+with(means, segments(y0=xmin, y1=xmax, 
+                     x0=log10_diversity, x1=log10_diversity,
[…]

### 2021-05-25 notebooks/figures/species_ranges.r

@@ -13,15 +13,24 @@ d <- read_tsv('../../data/combined_data.tsv') %>%
        # space the groups
        mutate(xpos = 1:nrow(.) + 1.2*as.integer(range_cat)-1)
 
+means <- d %>% group_by(range_cat) %>% 
+  filter(is.finite(log10_range)) %>%
+  summarize(xmin=min(xpos), 
+            xmax=max(xpos),
+            log10_range = mean(log10_range, na.rm=TRUE))
+
 
 
 opar <- par(no.readonly=TRUE)
 # pdf('range_descriptions.pdf', width=8, height=5)
 pdf('range_categories.pdf', width=5, height=8)
 
+
 par(mar=c(4, 1, 0, 5))
 cols <- wes_palette('Darjeeling1')[c(5, 2, 3, 1, 4)]
-plot(d$log10_range, d$xpos, pch = 19, col=cols[d$range_cat], 
+plot(d$log10_range, d$xpos, pch = 21, bg=cols[d$range_cat], 
+     lwd=0.7,
+     col='white',
      xlab = '',
      ylab = '', axes=FALSE, xlim=c(2, 9))
 ylabs = seq(2, 8, 2)
@@ -29,6 +38,14 @@ axis(1, ylabs, labels=latex2exp::TeX(sprintf("$10^{%d}$", ylabs)),
      cex.axis=0.8,
      line=0.5,
      tck=-0.01, padj=-1)
+
+
+with(means, segments(y0=xmin, y1=xmax, 
+                     x0=log10_range, x1=log10_range,
+                     col=cols[range_cat], lwd=1, lty=1))
+
+
+
 text(y=d$xpos, x=8.8, labels=d$species, 
      col=cols[d$range_cat], xpd=TRUE, 
      las=1, srt=0, adj=0, cex=0.3)

### 2021-05-25 manuscript/manuscript.tex
abstract updated

@@
 
-  Under neutral theory, genetic diversity in an equilibrium population is
-  expected to increase with population size. However, observed levels of
-  diversity across metazoans vary only two orders of magnitude, while census
-  population sizes ($N_c$) are expected to vary over several. This unexpectedly
-  narrow range of diversity is a longstanding enigma in evolutionary genetics
-  known as Lewontin's Paradox of Variation (1974). Since Lewontin's
-  observation, it has been argued that selection constrains diversity across
-  species, yet tests of this hypothesis seem to fall short of explaining the
-  orders-of-magnitude reduction in diversity observed in nature. In this work,
-  I revisit Lewontin's Paradox and assess whether current models of linked
-  selection are capable of constraining diversity to this extent. To quantify
-  the discrepancy between pairwise diversity and census population sizes across
-  species, I combine genetic data from 172 metazoan taxa with estimates of
-  census sizes from geographic occurrence data and population densities
-  estimated from body mass.  Next, I fit the relationship between
-  previously-published estimates of genomic diversity and these approximate
-  census sizes to quantify Lewontin's Paradox. To account for phylogenetic
-  non-independence, I use phylogenetic comparative methods to investigate the
-  diversity census size relationship, estimate phylogenetic signal, and explore
-  how diversity changes along the phylogeny. I consider whether the reduction
-  in diversity predicted by models of recurrent hitchhiking and background
-  selection could explain the observed pattern of diversity across species.
-  Since the impact of linked selection is mediated by recombination map length,
-  I also investigate how map lengths vary with census sizes. I find species
-  with large census sizes have shorter map lengths, leading these species to
-  likely experience greater reductions in diversity due to linked selection.
-  Even after using high estimates of the strength of sweeps and background
-  selection, I find linked selection likely cannot explain the shortfall
-  between predicted and observed diversity levels across metazoan species.
-  Furthermore, the predicted diversity under linked selection does not fit the
-  observed diversity--census-size relationship, implying that processes other
-  than background selection and recurrent hitchhiking must be limiting
-  diversity.
+ Neutral theory predicts that genetic diversity increases with population size,
+ yet observed levels of diversity across metazoans vary only two orders of
+ magnitude while population sizes vary over several. This unexpectedly narrow
+ range of diversity is known as Lewontin's Paradox of Variation (1974). While
+ some have suggested selection could constrain diversity, tests of this
+ hypothesis seem to fall short. Here, I revisit Lewontin's Paradox and assess
+ whether current models of linked selection are capable of reducing
+ diversity to this extent. To quantify the discrepancy between pairwise
+ diversity and census population sizes across species, I combine
+ previously-published estimates of pairwise diversity from 172 metazoan taxa
+ with estimates of census sizes. Using phylogenetic comparative methods, I show
+ this relationship is significant accounting for phylogeny, but with high
+ phylogenetic signal and evidence that some lineages experience shifts in the
+ evolutionary rate of diversity deep in the past. Additionally, I find a
+ negative relationship between recombination map length and census size,
+ suggesting abundant species have less recombination and experience greater
+ reductions in diversity due to linked selection. However, I show that even
+ using strong selection parameter estimates, models of linked selection are
+ unlikely to explain the observed relationship between diversity and census
+ sizes across species.
 

### 2021-05-27 manuscript/manuscript.tex
revised ms

@@
  range of diversity is known as Lewontin's Paradox of Variation (1974). While
- some have suggested selection could constrain diversity, tests of this
- hypothesis seem to fall short. Here, I revisit Lewontin's Paradox and assess
- whether current models of linked selection are capable of reducing
- diversity to this extent. To quantify the discrepancy between pairwise
- diversity and census population sizes across species, I combine
- previously-published estimates of pairwise diversity from 172 metazoan taxa
- with estimates of census sizes. Using phylogenetic comparative methods, I show
- this relationship is significant accounting for phylogeny, but with high
- phylogenetic signal and evidence that some lineages experience shifts in the
- evolutionary rate of diversity deep in the past. Additionally, I find a
- negative relationship between recombination map length and census size,
- suggesting abundant species have less recombination and experience greater
- reductions in diversity due to linked selection. However, I show that even
- using strong selection parameter estimates, models of linked selection are
- unlikely to explain the observed relationship between diversity and census
- sizes across species.
+ some have suggested selection constrains diversity, tests of this hypothesis
+ seem to fall short. Here, I revisit Lewontin's Paradox to assess whether
+ current models of linked selection are capable of reducing diversity to this
+ extent. To quantify the discrepancy between pairwise diversity and census
+ population sizes across species, I combine previously-published estimates of
+ pairwise diversity from 172 metazoan taxa with estimates of census sizes.
+ Using phylogenetic comparative methods, I show this relationship is
+ significant accounting for phylogeny, but with high phylogenetic signal and
+ evidence that some lineages experience shifts in the evolutionary rate of
+ diversity deep in the past. Additionally, I find a negative relationship
+ between recombination map length and census size, suggesting abundant species
+ have less recombination and experience greater reductions in diversity due to
+ linked selection. However, I show that even using strong selection parameter
+ estimates, models of linked selection are unlikely to explain the observed
+ relationship between diversity and census sizes across species.
 
@@
 the order of size of the population), characteristics of the breeding structure
-such as high variance ($V_w$) or skew in reproductive success  also suppress
-diversity below the levels predicted by the census size
-\parencite{Wright1938-tv}. For example, in many marine animals,
-females are highly fecund, and dispersing larvae face extremely low
-survivorship, leading to high variance in reproductive success
+such as high variance ($V_w$) or skew in reproductive success continuously
+suppress diversity below the levels predicted by the census size
+\parencite{Wright1938-tv}. For example, in many marine animals, females are
+highly fecund, and dispersing larvae face extremely low survivorship, leading
+to high variance in reproductive success
 \parencite{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
@@
 greater than the rate of neutral coalescence \parencite[e.g.][equation
-22]{Coop2012-cd}. Other linked selection models are proportional to population
-size, and thus cannot decouple diversity, all else equal. For example, models
-of background selection and polygenic fitness variation predict diversity is
-proportional to population size, mediated by the total recombination map length
-and the deleterious mutation rate or fitness variation
+22]{Coop2012-cd}. Under other linked selection models, the resulting effective
+population size is proportional to population size, and thus these models
+cannot decouple diversity, all else equal. For example, models of background
+selection and polygenic fitness variation predict diversity is proportional to
+population size, mediated by the total recombination map length and the
+deleterious mutation rate or fitness variation
 \parencite{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.
@@
 Past work looking at the relationship between $\pi$ and $N_c$ has been unable
-to account for phylogenetic non-independence across taxa
+to fully account for phylogenetic non-independence across taxa
 \parencite{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
@@
 surveyed the literature for census size estimates
-\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range or body
-size or qualitative categories as proxies for census size
+\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size,
+or qualitative categories as proxies for census size
 \parencite{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
@@
 and overestimated for species with anthropogenically reduced densities or
-patchy ranges. For example, the population size of \emph{Lynx lynx} is likely
-around 50,000 \parencite{Iucn2020-ap} which is around two orders of magnitude
-smaller than my estimate. Additionally, the range size estimates do not
-consider whether an area has unsuitable habitat, and thus may be overestimates
-for species with particular niches or patchy habitats. While these methods to
-estimate census size are crude and approximate, they can be efficiently
-calculated for numerous taxa and are sufficient to estimate the scale of
-Lewontin's Paradox.
+fragmented ranges. For example, the population size of \emph{Lynx lynx} is
+likely around 50,000 \parencite{Iucn2020-ap} which is around two orders of
+magnitude smaller than my estimate. Additionally, the range size estimates do
+not consider whether an area has unsuitable habitat, and thus may be
+overestimated for species with particular niches or patchy habitats. While
+these methods to estimate census size are crude and approximate, they can be
+efficiently calculated for numerous taxa and are sufficient to estimate the
+scale of Lewontin's Paradox (see Supplementary Information
+\Newnameref{supp:popsize-valid} for more on validation based on biomass and
+other approaches).
 
@@
     represents the maximum pairwise diversity under the four alleles model. The
-    dark gray dashed line is the OLS regression fit, and the blue
-    dashed line is the regression fit using a phylogenetic mixed-effects model.
-    Points are colored by phylum.}
+    dark gray dashed line is the OLS regression fit, and the blue dashed line
+    is the regression fit using a phylogenetic mixed-effects model.  Points are
+    colored by phylum. The species \emph{Equus ferus przewalskii} ($N_c \approx
+  10^3$ and $\pi = 3.6 \times 10^{-3}$) was an outlier and excluded from this
+figure for visual clarity.}
 
@@
 relationship on a log-log scale fits the data well (Figure \ref{fig:figure-1}).
-The OLS slope estimate is significant and implies a 0.09 percent increase in
+The OLS slope estimate is significant and implies a 13\% percent increase in
 differences per basepair for every order of magnitude census size grows (95\%
-confidence interval $[0.08, 0.12]$; see also the OLS fit per-phyla,
-Supplementary Figure \ref{suppfig:figure-1-ave}).
+confidence interval $[12\%, 14\%]$, adjusted $R^2 = 0.26$; see also the OLS fit
+per-phyla, Supplementary Figure \ref{suppfig:figure-1-ave}).
 
@@
 the synthetic chronogram. As with the linear regression, this relationship was
-positive and significant (95\% credible interval $0.04, 0.11$), though somewhat
+positive and significant (95\% credible interval $0.03, 0.11$), though somewhat
 attenuated compared to the OLS estimates (Figure \ref{fig:figure-2}B). Since
@@
 diversity, the posterior mean of $\lambda = 0.67$ (90\% credible interval
-$[0.59, 0.75]$) indicates that the majority of the variance perhaps might be
+$[0.58, 0.75]$) indicates that the majority of the variance perhaps might be
 due to shared phylogenetic history (Figure \ref{fig:figure-2}B).
@@
 node-height tests indicate a significant increase in the rate of evolution
-towards the present (robust regression p-values 0.028 and 0.00070 respectively;
+towards the present (robust regression p-values 0.023 and 0.00018 respectively;
 Figure \ref{fig:figure-2}C). Considering the constituents of the population
 size estimate, range and body mass, separately, the rate of evolution of range
-but not body mass shows a significant increase (p-value $1.9 \times 10^{-7}$)
+but not body mass shows a significant increase (p-value $1.03 \times 10^{-7}$)
 towards the present.
@@
 diversity ($\pi$) to the estimated diversity in the absence of linked selection
-($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. Here, $\pi_0$ would reflect demographic
-history and non-heritable variation in reproductive success. There are two
-difficulties in evaluating whether linked selection could resolve Lewontin's
-Paradox. The first difficulty is that $\pi_0$ is unobserved.  Previous work has
-estimated $\pi_0$ using methods that exploit the spatial heterogeneity in
-recombination and functional density across the genome to fit linked selection
-models that incorporate both hitchhiking and background selection
-\parencite{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second difficulty is
-understanding of how $R$ varies across taxa, since we lack estimates of
-critical model parameters for most species. Still, I can address a key
-question: if diversity levels were determined by census sizes ($\pi_0 = 4N_c
-\mu$), are the combined effects of background selection and recurrent
+($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. Here, $\pi_0$ would reflect only
+demographic history and non-heritable variation in reproductive success. There
+are two difficulties in evaluating whether linked selection could resolve
+Lewontin's Paradox. The first difficulty is that $\pi_0$ is unobserved.
+Previous work has estimated $\pi_0$ using methods that exploit the spatial
+heterogeneity in recombination and functional density across the genome to fit
+linked selection models that incorporate both hitchhiking and background
+selection \parencite{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
+difficulty is understanding of how $R$ varies across taxa, since we lack
+estimates of critical model parameters for most species. Still, I can address a
+key question: if diversity levels were determined by census sizes ($\pi_0 =
+4N_c \mu$), are the combined effects of background selection and recurrent
 hitchhiking sufficient to reduce diversity to observed levels? Furthermore,
@@
 \ref{supfig:genome_size_maplength}) and genome size and population size
-(Supplementary Figure \ref{supfig:genome_size_popsize}). Overall, the negative
-relationship between map length and census size indicates linked selection is
-expected to be stronger in short map length, high-$N_c$ species.
+(Supplementary Figure \ref{supfig:genome_size_popsize}). These findings are
+consistent with both the hypothesis that non-adaptive processes increase genome
+size in small-$N_e$ species \parencite{Lynch2003-bn} which in turn could
+increase map lengths, as well as the hypothesis that map lengths are adaptively
+longer to more efficiently select against deleterious alleles
+\parencite{Roze2021-ya}. Overall, the negative relationship between map length
+and census size indicates linked selection is expected to be stronger in short
+map length, high-$N_c$ species.
 
@@
 Paradox. First, low to mid-$N_c$ species (census sizes between
-$10^{6}$--$10^{14}$) have sufficiently long map lengths that their diversity
+$10^{4}$--$10^{10}$) have sufficiently long map lengths that their diversity
 levels are only moderately reduced by linked selection, leading to a wide gap
@@
 Second, while hitchhiking and BGS can reduce predicted diversity levels for
-high-$N_c$ species ($N_c > 10^{14}$) to observed levels, this would imply
+high-$N_c$ species ($N_c > 10^{10}$) to observed levels, this would imply
 available estimates of $\pi_0$ are underestimated by several orders of
@@
 how evolutionary, life history, and ecological processes interact to constrain
-diversity across taxa to a narrow range remains a mystery. Since Wright
-(\citeyear{Wright1931-fl,Wright1938-tv}), population geneticists have
-appreciated that various demographic processes shrink effective population
-sizes compared to census sizes, yet it has remained unclear whether these
-neutral processes alone can explain Lewontin's Paradox and across-taxa
-diversity patterns. Alternatively, selective processes that act more strongly
-in larger populations could account for the observed narrow range of diversity.
-A critical first step to discerning the processes that act to transform census
-sizes to diversity levels across species is characterizing the observed
-$\pi$--$N_c$ relationship.
-
-Here, for 172 taxa, I estimated the relationship between genomic estimates of
-pairwise diversity and approximate census population size. Previous surveys
-have used allozyme-based estimates, fewer taxa, or qualitative measures of
-population size. My estimates of census population sizes are quite approximate,
-since they use body size to predict density. An improved estimate might
-consider vagility (as \cite{Soule1976-he} did), though this is harder to do
-systematically across many taxa. Future work might also use other ecological
-information, such as total biomass and estimated numbers of species within
-phyla, to improve census size estimates \parencite{Bar-On2018-kc, Mora2011-wm}.
-Still, it seems more accurate estimates would be unlikely to change the
-qualitative findings here, which resemble those of early surveys
-\parencite{Nei1984-zi,Soule1976-he}.
+diversity across taxa to a narrow range remains a mystery. I revisit Lewontin's
+Paradox by first characterizing the relationship between genomic estimates of
+pairwise diversity and approximate census population size across 172 metazoan
+species. Previous surveys have used allozyme-based estimates, fewer taxa, or
+qualitative measures of population size. My estimates of census population
+sizes are quite approximate, since they use body size to predict density. An
+improved estimate might consider vagility (as \cite{Soule1976-he} did), though
+this is harder to do systematically across many taxa. Future work might also
+use other ecological information, such as total biomass, or species
+distribution modeling to improve census size estimates
+\parencite{Bar-On2018-kc, Mora2011-wm}.  Still, it seems more accurate
+estimates would be unlikely to change the qualitative findings here, which
+resemble those of early surveys \parencite{Nei1984-zi,Soule1976-he}.
 
@@
 because mid-$N_c$ species have sufficiently long recombination map lengths to
-diminish even strong selection. This finding supports the idea the levels of
-diversity across species are primarily determined by past demographic
-fluctuations. Overall, while this suggests these two common modes of linked
-selection seem unlikely to explain across-taxa patterns of diversity, there are
-three major potential limitations of my approach that need further evaluation.
+diminish the effects of even strong selection. This finding supports the idea
+the levels of diversity across species are primarily determined by past
+demographic fluctuations. Overall, while this suggests these two common modes
+of linked selection seem unlikely to explain across-taxa patterns of diversity,
+there are three major potential limitations of my approach that need further
+evaluation.
 
@@
 model simplifications, the predicted reduction in diversity in \emph{Drosophila
-melanogaster} is $85\%$ (using $N_e$, not $N_c$), which is reasonably close to
-the estimated $77\%$ from the more realistic model of Elyashiv et al. that
-accounts for the actual position of substitutions, annotation features, and
-recombination rate heterogeneity (though it should be noted that these both use
-the same parameter estimates). Furthermore, even though my model fails to
+melanogaster} is $85\%$ (when using $N_e$, not $N_c$), which is reasonably
+close to the estimated $77\%$ from the more realistic model of Elyashiv et al.
+that accounts for the actual position of substitutions, annotation features,
+and recombination rate heterogeneity (though it should be noted that these both
+use the same parameter estimates). Furthermore, even though my model fails to
 capture the heterogeneity of functionality density and recombination rate in
@@
 melanogaster} used, but also because I assume that the effective population
-size is the census size.  Even then, this decoupling only occurs in very
-high--census-size species, and implies that the diversity in the absence of
-linked selection, $\pi_0$, is currently underestimated by several orders of
+size is equal to the census size.  Even then, this decoupling only occurs in
+very high--census-size species, and implies that the diversity in the absence
+of linked selection, $\pi_0$, is currently underestimated by several orders of
 magnitude. Moreover, the study of \textcite{Corbett-Detig2015-gt} did consider
@@
 the genome (i.e. due to polygenic fitness variation), levels of diversity are
-well approximated by quantitative genetic linked selection models (QGLS
-\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}). Third, the parameters
-from \textcite{Elyashiv2016-vt} are robust to many types of sweeps that result
-in a substitution (e.g. see p. 19 of their Supplementary Online Materials).
-The reduction in diversity under these models is nearly identical to that under
-background selection models, in part because deleterious alleles at
-mutation-selection balance constitute a considerable component of fitness
-variation (see Appendix Section \ref{app:bgs};
-\cite{Charlesworth2000-km,Charlesworth2015-am}).  Finally, I also disregarded
-the interaction of sweeps and spatial processes.  For populations spread over
-wide ranges, limited dispersal slows the spread of sweeps, allowing for new
-beneficial alleles to arise, spread, and compete against other segregating
-beneficial variants \parencite{Ralph2015-kl,Ralph2010-ki}. Through limited
-dispersal should act to ``soften sweeps" and not impact my findings for the
-reasons described above, future work could investigate how these processes
-impact diversity in ways not captured by hard sweep models.
+well approximated by quantitative genetic linked selection models
+\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}. The reduction in
+diversity under these models is nearly identical to that under background
+selection models, in part because deleterious alleles at mutation-selection
+balance constitute a considerable component of fitness variation (see Appendix
+Section \ref{app:bgs}; \cite{Charlesworth2000-km,Charlesworth2015-am}). Third,
+the parameters from \textcite{Elyashiv2016-vt} are robust to many types of
+sweeps that result in substitution (e.g. see p. 19 of their Supplementary
+Online Materials). Finally, I also disregarded the interaction of sweeps and
+spatial processes.  For populations spread over wide ranges, limited dispersal
+slows the spread of sweeps, allowing for new beneficial alleles to arise,
+spread, and compete against other segregating beneficial variants
+\parencite{Ralph2015-kl,Ralph2010-ki}. Through limited dispersal should act to
+``soften sweeps" and not impact my findings for the reasons described above,
+future work could investigate how these processes impact diversity in ways not
+captured by hard sweep models.
 
@@
 selective events could occur regularly in a species' history yet be
-indistinguishable from demographic bottlenecks from just population genomic
+indistinguishable from demographic bottlenecks with just population genomic
 data.
@@
 $\nicefrac{N_e}{N_c} \approx 10^{-6}$. Estimates of the $\nicefrac{N_e}{N_c}$
-ratio are an important, and under appreciated piece of solving Lewontin's
+ratio are an important, yet under appreciated piece of solving Lewontin's
 Paradox. For example, if $N_e$ is estimated from the allele frequency change
@@
 Looking at Lewontin's Paradox through an macroecological and macroevolutionary
-lens begets interesting questions outside of the realm of population genetics.
-Here, I have found that diversity and $N_c$ have a surprisingly consistent
-relationship without many outliers, despite the wildly disparate ecologies,
-life histories, and evolutionary histories of the taxa included. Furthermore,
-taxa with very large census sizes have surprisingly low diversity. Is this
-explained by macroevolutionary processes, such as different rates of speciation
-for large-$N_c$ taxa? Or, are the levels of diversity we observe today an
-artifact of our timing relative to the last glacial maximum, or the last major
-extinction? Did large-$N_c$ prehistoric animal populations living in other
-geological eras have higher levels of diversity than our present taxa? Or, does
-ecological competition occur on shorter timescales such that strong population
-size contractions transpire and depress diversity, even if a species is
-undisturbed by climatic shifts or mass extinctions? Overall, patterns of
-diversity across taxa are determined by many overlaid evolutionary and
-ecological processes occurring on vastly different timescales. Lewontin's
+lens begets interesting questions outside of the traditional realm of
+population genetics.  Here, I have found that diversity and $N_c$ have a
+surprisingly consistent relationship without many outliers, despite the wildly
+disparate ecologies, life histories, and evolutionary histories of the taxa
+included. Furthermore, taxa with very large census sizes have surprisingly low
+diversity. Is this explained by macroevolutionary processes, such as different
+rates of speciation for large-$N_c$ taxa? Or, are the levels of diversity we
+observe today an artifact of our timing relative to the last glacial maximum,
+or the last major extinction? Did large-$N_c$ prehistoric animal populations
+living in other geological eras have higher levels of diversity than our
+present taxa? Or, does ecological competition occur on shorter timescales such
+that strong population size contractions transpire and depress diversity, even
+if a species is undisturbed by climatic shifts or mass extinctions? Overall,
+patterns of diversity across taxa are determined by many overlaid evolutionary
+and ecological processes occurring on vastly different timescales. Lewontin's
 Paradox of Variation may persist unresolved for some time because the
@@
 Stan model was used to estimate the same relationship between arthropod,
-chordate, and mollusc subsets of the data (Supplementary Figure
-\ref{suppfig:div_popsize_post_phyla}).
+chordate, and mollusc subsets of the data, though a reduced model was used for
+the chordate subset due to identifiability issues leading to poor MCMC
+convergence (Supplementary Figure \ref{suppfig:div_popsize_post_phyla}).
 
@@
 population size is non-linear and heteroscedastic, and was fit using a
-lognormal phylogenetic mixed-effects model. Since social insects have longer
-recombination map lengths \parencite{Wilfert2007-dx}, social taxa were excluded
-when fitting this model.  All \texttt{Rhat} \parencite{Vehtari2019-po} values
-were below 1.01 and the effective number of samples was over 1,000, consistent
-with good mixing; details about the model are available in the GitHub
-repository (\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (Figure
+lognormal phylogenetic mixed-effects model on the 130 species with complete
+data. Since social insects have longer recombination map lengths
+\parencite{Wilfert2007-dx}, social taxa were excluded when fitting this model.
+All \texttt{Rhat} \parencite{Vehtari2019-po} values were below 1.01 and the
+effective number of samples was over 1,000, consistent with good mixing;
+details about the model are available in the GitHub repository
+(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (Figure
 \ref{fig:figure-2}A and Supplementary Material Figure \ref{suppfig:biconmap})
@@
 \end{align}
-
+%
 Using $u$-substitution with $r = \ell r_\text{BP}$ this simplifies to
@@
 \section{Population Size Validation}
+\label{supp:popsize-valid}
 

### 2021-06-01 manuscript/manuscript.tex
minor update
minor changes

@@
 
-\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}
+\title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
+%\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}
 
@@
 using a previously-observed linear relationship that holds across metazoans
-(Supplementary Figure \ref{suppfig:damuth};
+(Figure 1-figure supplement 1 \ref{suppfig:damuth};
 \cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
@@
 confidence interval $[12\%, 14\%]$, adjusted $R^2 = 0.26$; see also the OLS fit
-per-phyla, Supplementary Figure \ref{suppfig:figure-1-ave}).
+per-phyla, Figure 1-figure supplement \ref{suppfig:figure-1-ave}).
 
@@
 This is in part because of the log-log scale, in contrast to previous work
-\parencite{Nei1984-zi,Soule1976-he}; see Supplementary Figure
-\ref{suppfig:figure-1-linear} for a version on a log-linear scale. However, it
-is noteworthy that few taxa have diversity estimates below $10^{-3.5}$
-differences per basepair. Those that do, lynx (\emph{Lynx Lynx}), wolverine
-(\emph{Gulo gulo}), and Massasauga rattlesnake (\emph{Sistrurus catenatus})
-face habitat fragmentation and declining population sizes. These three species
-are all in the IUCN Red List, but are listed as least concern (though their
-presence in the Red List indicates they are of conservation interest). In
-Supplementary Material Section \Newnameref{supinfo:div-iucn}, I explore more
-about the relationships between IUCN Red List status, diversity, and population
-size.
+\parencite{Nei1984-zi,Soule1976-he}; see Figure 1-supplement figure 2 for a
+version on a log-linear scale. However, it is noteworthy that few taxa have
+diversity estimates below $10^{-3.5}$ differences per basepair. Those that do,
+lynx (\emph{Lynx Lynx}), wolverine (\emph{Gulo gulo}), and Massasauga
+rattlesnake (\emph{Sistrurus catenatus}) face habitat fragmentation and
+declining population sizes. These three species are all in the IUCN Red List,
+but are listed as least concern (though their presence in the Red List
+indicates they are of conservation interest). In Supplementary Information
+Section \Newnameref{supinfo:div-iucn}, I explore more about the relationships
+between IUCN Red List status, diversity, and population size.
 
@@
 Finally, an important caveat is the increase in rate towards the tips could be
-caused by measurement noise. Inspecting the lineage pairs that lead to this
+caused by measurement noise, or possibly uncertainty or bias in the divergence
+time estimates deep in the tree. Inspecting the lineage pairs that lead to this
 increase in rate towards the tips indicates these represent plausible rate
@@
 \ref{suppfig:corbett_detig}A), and it is often thought that this is driven by
-higher rates of adaptive substitutions in larger populations, despite equivocal
-evidence \parencite{Galtier2016-dq}. However, there is another mechanism by
-which species with larger population sizes might experience a greater impact of
-linked selection: recombinational map length, $L$, is known to correlate with
-body mass \parencite{Burt1987-tq} and thus varies inversely with population
-size. As this is a critical parameter that determines the genome-wide impact of
-both hitchhiking and background selection, I examine the relationship between
-recombination map length ($L$) and census population size ($N_c$) across taxa,
-using available estimates of map lengths across species
-\parencite{Stapley2017-fs,Corbett-Detig2015-gt}. I find a significant
-non-linear relationship using phylogenetic mixed-effects models (Figure
-\ref{fig:figure-3}A; see Methods: \Newnameref{sec:methods-pcm}). There is also
-a correlation between map length and genome size (Supplementary Figure
-\ref{supfig:genome_size_maplength}) and genome size and population size
+higher rates of adaptive substitutions in larger populations
+\parencite{Ohta1992-yi}, despite equivocal evidence \parencite{Galtier2016-dq}.
+However, there is another mechanism by which species with larger population
+sizes might experience a greater impact of linked selection: recombinational
+map length, $L$, is known to correlate with body mass \parencite{Burt1987-tq}
+and thus varies inversely with population size. As this is a critical parameter
+that determines the genome-wide impact of both hitchhiking and background
+selection, I examine the relationship between recombination map length ($L$)
+and census population size ($N_c$) across taxa, using available estimates of
+map lengths across species \parencite{Stapley2017-fs,Corbett-Detig2015-gt}. I
+find a significant non-linear relationship using phylogenetic mixed-effects
+models (Figure \ref{fig:figure-3}A; see Methods: \Newnameref{sec:methods-pcm}).
+There is also a correlation between map length and genome size (Supplementary
+Figure \ref{supfig:genome_size_maplength}) and genome size and population size
 (Supplementary Figure \ref{supfig:genome_size_popsize}). These findings are
@@
 (\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (Figure
-\ref{fig:figure-2}A and Supplementary Material Figure \ref{suppfig:biconmap})
+\ref{fig:figure-2}A and Supplementary Information Figure \ref{suppfig:biconmap})
 were created using \texttt{phytools} \parencite{Revell2012-zq}. Node-height
@@
 phylogenetic comparative methods; yet I take full responsibility for any
-shortcomings of my analysis. Finally, I am indebted to helpful feedback from
-Guy Sella, Matt Pennell, and two other anonymous reviewers. I would like to
-also thank UO librarian Dean Walton for helping me track down some rather
-difficult to find older papers. This work was supported by an NIH Grant
-(1R01GM117241) awarded to Andrew Kern.
+shortcomings of my analysis. Finally, I am indebted to Guy Sella, Matt Pennell,
+and two other anonymous reviewers for helpful feedback. I would like to also
+thank UO librarian Dean Walton for helping me track down some rather difficult
+to find older papers.  This work was supported by an NIH Grant (1R01GM117241)
+awarded to Andrew Kern.
 
@@
 \renewcommand{\thesection}{S\arabic{section}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
   \hline
-  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.78 \times 10^{-4}$ & 0.0100 &  68 & 0.02 & $5.41 \times 10^{-5}$ & 4.29 \\
-  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.70 \times 10^{-2}$ & 0.9719 &  68 & 2.90 & $1.26 \times 10^{-3}$ & 24.70 \\
+  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.80 \times 10^{-4}$ & 0.0102 &  68 & 0.02 & $5.41 \times 10^{-5}$ & 4.31 \\ 
+  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.67 \times 10^{-2}$ & 0.9715 &  68 & 2.89 & $1.26 \times 10^{-3}$ & 24.40 \\ 
   Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & 0.01 & $1.76 \times 10^{-4}$ & 0.35 \\ 
-  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.54 \times 10^{-4}$ & 0.0163 &  13 & 0.21 & $1.36 \times 10^{-4}$ & 16.60 \\
-  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.14 \times 10^{-5}$ & 0.0011 &   2 & 0.03 & $1.25 \times 10^{-4}$ & 2.51 \\
-  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $3.89 \times 10^{-6}$ & 0.0001 &   1 & 0.02 & $4.00 \times 10^{-5}$ & 4.87 \\
+  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.56 \times 10^{-4}$ & 0.0166 &  13 & 0.21 & $1.36 \times 10^{-4}$ & 16.70 \\ 
+  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.07 \times 10^{-5}$ & 0.0011 &   2 & 0.03 & $1.25 \times 10^{-4}$ & 2.45 \\ 
+  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $4.03 \times 10^{-6}$ & 0.0001 &   1 & 0.02 & $4.00 \times 10^{-5}$ & 5.03 \\
 \end{tabular}}
@@
 First, I compared the relative carbon biomass in this study to the relative
-carbon biomass on earth per phylum. This shows that this study's sample is over
-represents chordate biomass (by a factor of $\mathord{\sim}3$), and under represents
-in arthropod biomass (by a factor of $0.02$) relative to the proportion of
-carbon biomass of these phyla on earth (see column eight of Supplementary Table
-\ref{supp:biomass-tbl}. Second, to check whether the carbon biomass per phylum
-in the sample was broadly consistent with the total on earth by phylum ($B_S$
-for phylum $S$), I calculated the expected sample biomass if species were
-sampled randomly from the total species in a phylum, ($B_S \times
+carbon biomass on earth per phylum. This shows that this study's sample over
+represents chordate biomass (by a factor of $\mathord{\sim}3$), and under
+represents in arthropod biomass (by a factor of $0.02$) relative to the
+proportion of carbon biomass of these phyla on earth (see column eight of
+Supplementary Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
+biomass per phylum in the sample was broadly consistent with the total on earth
+by phylum ($B_S$ for phylum $S$), I calculated the expected sample biomass if
+species were sampled randomly from the total species in a phylum, ($B_S \times
 \nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
@@
 taxa included in this study. The carbon biomass estimates for chordates implied
-by the census size estimates are $\mathord{\sim}25$-fold higher than expected, but is
+by the census size estimates are $\mathord{\sim}24$-fold higher than expected, but is
 well within reasonable expectations given that the chordate sample includes
@@
 As another consistency check, I looked at the rank order of mammals by biomass.
-Whale species have the first and third highest biomass with 11.4 and 3.8
+Whale species have the first and third highest biomass with 11.4 and 3.9
 megatons of carbon biomass (for \emph{Balaenoptera bonaerensis} and
@@
 population sizes for wolf species represent pre-anthropogenic densities and are
-overestimates compared to current population sizes as expected.
+overestimates compared to current population sizes, as expected.
 
@@
   \end{tabular}
-  \caption{TODO The regression estimates of full IUCN Red List population size model
+  \caption{The regression estimates of full IUCN Red List population size model
     for diversity, $\log_{10}(\pi) = \beta_0 + \beta_{LC} LC + \beta_{NT} NT +
@@
     $\log_{10}(\pi) = \beta_0 + \beta_{N_c} \log_{10}(N_c)$, $AIC_\text{full} =
-  203.99$, $AIC_\text{reduced} =212.46$. }
+ 204.9$, $AIC_\text{reduced} = 216.4$. }
 \end{table}
 \label{supptable:tbl-1}
+
+\section{Figure Supplement}
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{}
+\renewcommand{\figurename}{Figure 1-figure supplement 1}
+
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[]{figures/damuth.pdf}
+
+  \caption{The appendix table of \textcite{Damuth1987-sg}; the color indicates
+    Damuth's original group labels. The dashed line was estimated using a
+    lognormal regression model in Stan. References to each measurement are
+  available in \textcite{Damuth1987-sg}.}
+  \label{suppfig:damuth}
+\end{figure}
 
@@
   \label{suppfig:range-cats}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/damuth.pdf}
-
-  \caption{The appendix table of \textcite{Damuth1987-sg}; the color indicates
-    Damuth's original group labels. The dashed line was estimated using a
-    lognormal regression model in Stan. References to each measurement are
-  available in \textcite{Damuth1987-sg}.}
-
-  \label{suppfig:damuth}
 \end{figure}

### 2021-06-04 manuscript/elife/elife-template.tex
more elife template bullshit

--- 2021-06-01 manuscript/manuscript.tex
@@
-\documentclass[11pt]{article}
-\RequirePackage{fullpage}
-\RequirePackage[font=small,labelfont=bf]{caption}
-\RequirePackage{amsmath,amssymb,amsthm}
-\RequirePackage{mathtools}
-\RequirePackage{graphicx}
-\RequirePackage[normalem]{ulem}
-\RequirePackage[hidelinks]{hyperref}
-\RequirePackage{wrapfig}
-\RequirePackage{subcaption}
-\RequirePackage{authblk}
-\RequirePackage{bm}
-\RequirePackage{bbm}
-\RequirePackage{tikz}
-\RequirePackage{wrapfig}
-
-% line numbers:
-\RequirePackage{lineno}
-% \modulolinenumbers[5]
-\linenumbers
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}
-
-% spacing
-\RequirePackage{setspace}
-% \doublespacing
-
-% \RequirePackage[osf]{mathpazo}
-\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
-                date=year,
-                maxbibnames=9,maxnames=5,maxcitenames=2,
-                backend=biber,uniquelist=false,uniquename=false,
-                % style=apa,
-                sorting=nyt,
-                % sorting=,
-                hyperref=true]{biblatex}
-\RequirePackage[colorinlistoftodos]{todonotes}  %disable
-\RequirePackage{color}
-\RequirePackage{nicefrac}
-
-\newcommand{\gc}[1]{{\it \color{red} #1 } }
-\newcommand{\vb}[1]{{\it \color{blue} #1}}
-\newcommand{\vbout}[1]{{\it \color{blue} \sout{#1}}}
-
-% a /nonumber you can turn on/off
-\newcommand{\nnn}{\nonumber}
-%\newcommand{\nnn}{}
-
-\newcommand{\beginsupplement}{%
-        \setcounter{table}{0}
-        \renewcommand{\thetable}{S\arabic{table}}%
-        \setcounter{figure}{0}
-        \renewcommand{\thefigure}{S\arabic{figure}}%
-     }
-
-\newcommand{\Newnameref}[1]{\textit{\nameref{#1}}}
-
-\renewbibmacro{in:}{}
-\AtEveryBibitem{\clearlist{language}}
-
-\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
-\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% ELIFE ARTICLE TEMPLATE
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% PREAMBLE 
+\documentclass[9pt,lineno]{elife}
+% Use the onehalfspacing option for 1.5 line spacing
+% Use the doublespacing option for 2.0 line spacing
+% Please note that these options may affect formatting.
+% Additionally, the use of the \newcommand function should be limited.
+
+\usepackage{lipsum} % Required to insert dummy text
+\usepackage[version=4]{mhchem}
+\usepackage{siunitx}
+\usepackage{nicefrac}
+\usepackage{mathtools}
+\DeclareSIUnit\Molar{M}
 
@@
 
-\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }
-
-\addbibresource{biblio.bib}
-
-\title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
-%\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}
-
-\author{Vince Buffalo}
-\affil[]{\footnotesize University of Oregon, Institute of Ecology and Evolution\\Eugene, Oregon \\ \href{mailto:vsbuffalo@gmail.com}{vsbuffalo@gmail.com}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% ARTICLE SETUP
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to
+Explain the Narrow Range of Diversity Among Species}
+
+\author[1*]{Vince Buffalo}
+\affil[1]{Institute of Ecology and Evolution, University of Oregon, United States}
+
+\corr{vsbuffalo@gmail.com}{VB}
+
+% \contrib[\authfn{1}]{These authors contributed equally to this work}
+% \contrib[\authfn{2}]{These authors also contributed equally to this work}
+
+% \presentadd[\authfn{3}]{Department, Institute, Country}
+% \presentadd[\authfn{4}]{Department, Institute, Country}
+% \presentadd[\authfn{5}]{eLife Sciences editorial Office, eLife Sciences, Cambridge, United Kingdom}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% ARTICLE START
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
+
 \maketitle
-
-% TODO recheck node height and other p-values
-
-% Neutral theory predicts that genetic diversity increases with population size,
-% yet observed levels of diversity across species vary only two orders of
-% magnitude while population sizes likely vary over several. The causes of this
-% discrepancy, known as Lewontin's Paradox of Variation, remain unknown. Here I
-% revisit Lewontin's Paradox by quantifying the relationship between pairwise
-% diversity and approximate census size for 172 metazoan species. Using
-% phylogenetic comparative methods, I show this relationship is significant
-% accounting for phylogeny, but has high phylogenetic signal and some lineages
-% experience shifts in diversity deep in the past. I find a negative relationship
-% between recombination map length and census size, suggesting abundant species
-% have less recombination and could experience greater reductions in diversity
-% due to linked selection. However, I show that even using strong selection
-% parameter estimates, models of linked selection are unlikely to explain the
-% observed relationship between diversity and census sizes across species.
 
@@
  relationship between diversity and census sizes across species.
-
 \end{abstract}
+
+\section{Introduction}
 
@@
 new mutations and their loss by genetic drift
-\parencite{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the
+\citep{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the
 expected diversity at neutral sites in a panmictic population of $N_c$ diploids
@@
 levels across a wide range of species varied just an order of magnitude
-\parencite[][p.  208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox
+\citep[][p.  208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox
 of Variation".  With modern sequencing-based estimates of $\pi$ across taxa
@@
 terms of the neutralist--selectionist controversy
-\parencite{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.
+\citep{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.
 The neutralist view is that beneficial alleles are sufficiently rare and
@@
 diversity are shaped predominantly by genetic drift and mutation
-\parencite{Kimura1984-ia}. Specifically, \emph{non-selective} processes
+\citep{Kimura1984-ia}. Specifically, \emph{non-selective} processes
 decouple the effective population size implied by observed levels of diversity
@@
 
-\subsection*{Possible Explanations of Lewontin's Paradox}
+\subsection{Possible Explanations of Lewontin's Paradox}
 
@@
 population size
-\parencite{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy}, and the
+\citep{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy}, and the
 ecological and life history correlates of genetic diversity
-\parencite{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of
+\citep{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of
 processes stand out as potentially capable of decoupling census sizes from
@@
 than census population sizes, tracing back to early debates between R.A. Fisher
-and Sewall Wright \parencite{Fisher1947-tf,Wright1948-fj}. Possible causes of
+and Sewall Wright \citep{Fisher1947-tf,Wright1948-fj}. Possible causes of
 this divergence between effective and census population sizes include
@@
 glacial maximum severely reduced population sizes
-\parencite{Kimura1984-ia,Ohta1973-pk,Nei1984-zi}, and emphasized that large
+\citep{Kimura1984-ia,Ohta1973-pk,Nei1984-zi}, and emphasized that large
 populations recover to equilibrium diversity levels more slowly
@@
 events, which reduces effective population size
-\parencite{Maruyama1980-xz,Slatkin1977-kd}.
+\citep{Maruyama1980-xz,Slatkin1977-kd}.
 
@@
 suppress diversity below the levels predicted by the census size
-\parencite{Wright1938-tv}. For example, in many marine animals, females are
+\citep{Wright1938-tv}. For example, in many marine animals, females are
 highly fecund, and dispersing larvae face extremely low survivorship, leading
 to high variance in reproductive success
-\parencite{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
+\citep{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
 ``sweepstakes" reproductive systems can lead to remarkably small ratios of
@@
 \nicefrac{1}{V_w}$
-\parencite{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and
+\citep{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and
 require multiple-merger coalescent processes to describe their genealogies
-\parencite{Eldon2006-ui}. Overall, these reproductive systems diminish the
+\citep{Eldon2006-ui}. Overall, these reproductive systems diminish the
 diversity in many species, but seem unlikely to explain Lewontin's Paradox
@@
 of diversity. The earliest mathematical model of hitchhiking was proffered as a
-explanation of Lewontin's Paradox \parencite{Maynard_Smith1974-zr}.  Since,
+explanation of Lewontin's Paradox \citep{Maynard_Smith1974-zr}.  Since,
 empirical observations have demonstrated that linked selection shapes patterns
@@
 and diversity in a variety of species
-\parencite{Aguade1989-jx,Begun1992-ey,Cutter2003-tl,Stephan1998-hh,Cai2009-by}.
+\citep{Aguade1989-jx,Begun1992-ey,Cutter2003-tl,Stephan1998-hh,Cai2009-by}.
 Theoretic work to explain this pattern has considered diversity under a steady
@@
 around genic regions in a variety of species
-\parencite{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now
+\citep{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now
 efforts have shifted towards teasing apart the effects of positive and negative
-selection on genomic diversity \parencite{Elyashiv2016-vt}.
+selection on genomic diversity \citep{Elyashiv2016-vt}.
 
@@
 lineages and forcing them to coalesce
-\parencite{Kaplan1989-sc,Gillespie2000-mh}. In general, decoupling occurs under
+\citep{Kaplan1989-sc,Gillespie2000-mh}. In general, decoupling occurs under
 these hitchhiking models when the rate of coalescence due to selection is much
-greater than the rate of neutral coalescence \parencite[e.g.][equation
+greater than the rate of neutral coalescence \citep[e.g.][equation
 22]{Coop2012-cd}. Under other linked selection models, the resulting effective
@@
 deleterious mutation rate or fitness variation
-\parencite{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.
-
-\subsection*{Recent Approaches Towards Solving Lewontin's Paradox}
-
-Recently, \textcite{Corbett-Detig2015-gt} used population genomic data to
+\citep{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.
+
+\subsection{Recent Approaches Towards Solving Lewontin's Paradox}
+
+Recently, \cite{Corbett-Detig2015-gt} used population genomic data to
 estimate the reduction in diversity due to background selection and hitchhiking
@@
 argued this is evidence that selection could explain Lewontin's Paradox;
-however, in a re-analysis, \textcite{Coop2016-gx} demonstrated that the
+however, in a re-analysis, \cite{Coop2016-gx} demonstrated that the
 observed scale of these reductions is insufficient to explain the
@@
 characteristics related to parental investment, such as propagule size, are
-good predictors diversity in animals \parencite{Romiguier2014-bp,Chen2017-nf}.
+good predictors diversity in animals \citep{Romiguier2014-bp,Chen2017-nf}.
 Nevertheless, while these diversity correlates are important clues, they do not
@@
 to fully account for phylogenetic non-independence across taxa
-\parencite{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
+\citep{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
 comparative methods (PCMs) with a synthetic time-calibrated phylogeny to
@@
 since coalescent times are much less than divergence times
-\parencite{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by
+\citep{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by
 estimating the degree of phylogenetic signal in the diversity census size
@@
 I investigate how recombination map lengths vary with census population size
-using data from a previously-published survey \parencite{Stapley2017-fs}. I
+using data from a previously-published survey \citep{Stapley2017-fs}. I
 find map lengths are typically shorter in large--census-size species,
@@
 
-\section*{Results}
-
-\subsection*{Estimates of Census Population Size}
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=0.6\textwidth]{figures/Nc_figure.pdf}
-  \caption{
-     The distribution of approximate census population sizes estimated by this
+\section{Results}
+
+\subsection{Estimates of Census Population Size}
+
+\begin{figure}
+  \includegraphics[width=0.7\linewidth]{figures/Nc_figure.pdf}
+  \caption{The distribution of approximate census population sizes estimated by this
    study. Some phyla containing few species were excluded for clarity.}
    \label{fig:figure-nc}
+
+   \figsupp[The relationship between body mass and population density found by
+   \cite{Damuth1987-sg}, which is used to predict population densities.]{The
+     source of this data is appendix table of \cite{Damuth1987-sg}; the color
+     indicates Damuth's original group labels. The dashed line was estimated
+     using a lognormal regression model in Stan. References to each measurement
+   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:sf1}
+
+  \figsupp[The fraction of total species on earth included in this study's
+  sample, per class.]{The color of the points represents phylum, and the size of
+the point represents the absolute number of species by
+class.}{\includegraphics{figures/species_counts.pdf}}\label{figsupp:species-counts}
+
+\figsupp[Validation of this paper's range estimates procedure against the IUCN Red List's range estimates.]{
+  The correspondence between the ranges estimated with the alpha hull
+  method applied to GBIF data used in this paper and IUCN Red List's Extent
+  of Occurrence for the subset of species in both datasets. Note that the
+  IUCN Red List contains predominantly endangered species, which leads to
+  ascertainment bias; still, the high correlation between the estimated
+ranges shows the alpha hull method works well.}{\includegraphics[]{figures/iucn_redlist_ranges.pdf}}\label{figsupp:redlist-ranges}
+
+\figsupp[Validation of this paper's range estimates against the categorical labels of \cite{Leffler2012-zj}.]{
+  The estimated ranges using GBIF occurrence data, ordered within and
+  colored by the original range category labels assigned in
+\cite{Leffler2012-zj}.}{\includegraphics[]{figures/range_categories.pdf}}\label{figsupp:range-cats}
+
+  \figsupp[The relationship between body length (meters) and body mass (grams)
+  in the \cite{Romiguier2014-bp} data set.]{The relationship between body
+    length (meters) and body mass (grams) in the \cite{Romiguier2014-bp} data
+    set. This is used to infer body masses for taxa. The gray dashed line is the
+  line of best fit inferred using
+Stan.}{\includegraphics[]{figures/body_mass_length.pdf}}\label{figsupp:body-mass-length}
+
 \end{figure}
@@
 surveyed the literature for census size estimates
-\parencite{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size,
-or qualitative categories as proxies for census size
-\parencite{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
+\citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
+qualitative categories as proxies for census size
+\citep{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
 between genomic estimates of diversity and census population sizes, I first
-approximate census population sizes for 172 metazoan taxa (Figure
-\ref{fig:figure-nc}). My approach predicts population densities from body sizes
-using a previously-observed linear relationship that holds across metazoans
-(Figure 1-figure supplement 1 \ref{suppfig:damuth};
-\cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
-estimate range sizes.  Finally, I estimate population size as the product of
-these predicted densities and range estimates (see Methods:
-\Newnameref{sec:methods-popsize}). Note that the relationship between
+approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). My
+approach predicts population densities from body sizes using a
+previously-observed linear relationship that holds across metazoans (see
+\FIGSUPP[figure-nc]{sf1}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
+geographic occurrence data, I estimate range sizes.  Finally, I estimate
+population size as the product of these predicted densities and range estimates
+(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the relationship between
 population density and body size is driven by energy budgets, and thus reflects
-macroecological equilibria \parencite{Damuth1987-sg}. Consequently, population
+macroecological equilibria \citep{Damuth1987-sg}. Consequently, population
 sizes are underestimated for taxa like humans and their domesticated species,
@@
 fragmented ranges. For example, the population size of \emph{Lynx lynx} is
-likely around 50,000 \parencite{Iucn2020-ap} which is around two orders of
+likely around 50,000 \citep{Iucn2020-ap} which is around two orders of
 magnitude smaller than my estimate. Additionally, the range size estimates do
@@
 efficiently calculated for numerous taxa and are sufficient to estimate the
-scale of Lewontin's Paradox (see Supplementary Information
-\Newnameref{supp:popsize-valid} for more on validation based on biomass and
-other approaches).
-
-\subsection*{Quantifying Lewontin's Paradox}
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=\textwidth]{figures/diversity_popsize_full.pdf}
-
-  \caption{An illustration of Lewontin's Paradox of Variation. Pairwise
+scale of Lewontin's Paradox (see \nameref{supp:popsize-valid} for more on
+validation based on biomass and other approaches).
+
+\subsection{Quantifying Lewontin's Paradox}
+
+\begin{figure}
+  \begin{fullwidth}
+  \includegraphics[width=\linewidth]{figures/diversity_popsize_full.pdf}
+
+  \caption{A visualization of Lewontin's Paradox of Variation. Pairwise
     diversity (data from \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt},
@@
     10^3$ and $\pi = 3.6 \times 10^{-3}$) was an outlier and excluded from this
-figure for visual clarity.}
-
-  \label{fig:figure-1}
+    figure for visual clarity.}\label{fig:pi-Nc}
+
+    \figsupp{A linear-log version of Figure 2. Points are colored by phylum,
+      and the shaded region is the predicted neutral level of diversity
+      assuming $N_e = N_c$ with mutation range ranging between $10^{-10} \le
+    \mu \le 10^{-8}$.}{\includegraphics[width=\textwidth]{figures/diversity_popsize_linear.pdf}}\label{figsupp:popsize-linear}
+
+      \figsupp{A version of Figure 2 with OLS estimates per phylum. Diversity
+      and approximate population size for 172 taxa, colored by phylum; the
+    dashed lines indicate the non-phylogenetic OLS estimates of the
+  relationship between population size and diversity grouped by
+phyla.}{\includegraphics[width=\textwidth]{figures/diversity_popsize_averages.pdf}}\label{figsupp:popsize-ave}
+
+  \figsupp{The posterior distributions and fitted relationship between
+    diversity and both body mass and range size. The relationship between
+    diversity (differences per basepair) and body mass (left) and range (right)
+    across 172 species. The top row are posterior distributions of parameters
+    estimated using the phylogenetic mixed-effects model using 166 taxa in the
+    synthetic phylogeny for the intercept, slope, and phylogenetic signal from
+    the mixed-effects model. The bottom row contain each species as a point,
+    colored by phyla. The gray dashed line is the non-phylogenetic standard
+    regression estimate, and the blue dashed line is the relationship fit by
+    the phylogenetic mixed-effects
+  model.}{\includegraphics[width=\textwidth]{figures/diversity_range_bodymass.pdf}}\label{figsupp:div-range-bodymass}
+
+   \figsupp{Pairwise diversity grouped by the range categories from
+    \cite{Leffler2012-zj}, with point size indicating the predicted population
+    density. The vertical lines are the range category group
+  means.}{\includegraphics{figures/diversity_leffler_ranges.pdf}}\label{figsupp:div-leffler}
+  \end{fullwidth}
 \end{figure}
@@
 size or range size, but these studies relied on heterozygosity measured from
-allozyme data
-\parencite{Soule1976-he,Frankham1996-yb,Nei1984-zi,Leffler2012-zj}. Here, I
-confirm these findings using pairwise diversity estimates from genomic sequence
-data and the estimated census sizes (Figure \ref{fig:figure-1}). The pairwise
-diversity estimates are from three sources: \textcite{Leffler2012-zj},
-\textcite{Corbett-Detig2015-gt}, and \textcite{Romiguier2014-bp}, and are
-predominantly from either synonymous or non-coding DNA (see Methods:
-\Newnameref{sec:methods-diversity}). Overall, an ordinary least squares (OLS)
-relationship on a log-log scale fits the data well (Figure \ref{fig:figure-1}).
-The OLS slope estimate is significant and implies a 13\% percent increase in
-differences per basepair for every order of magnitude census size grows (95\%
-confidence interval $[12\%, 14\%]$, adjusted $R^2 = 0.26$; see also the OLS fit
-per-phyla, Figure 1-figure supplement \ref{suppfig:figure-1-ave}).
+allozyme data \citep{Soule1976-he,Frankham1996-yb,Nei1984-zi,Leffler2012-zj}.
+Here, I confirm these findings using pairwise diversity estimates from genomic
+sequence data and the estimated census sizes (\FIG{pi-Nc}).  The pairwise
+diversity estimates are from three sources: \cite{Leffler2012-zj},
+\cite{Corbett-Detig2015-gt}, and \cite{Romiguier2014-bp}, and are predominantly
+from either synonymous or non-coding DNA (see Methods and Materials:
+\nameref{sec:methods-diversity}). Overall, an ordinary least squares (OLS)
+relationship on a log-log scale fits the data well (\FIG{pi-Nc}, gray dashed
+line). The OLS slope estimate is significant and implies a 13\% percent
+increase in differences per basepair for every order of magnitude census size
+grows (95\% confidence interval $[12\%, 14\%]$, adjusted $R^2 = 0.26$; see also
+the OLS fit per-phyla, \FIGSUPP[pi-Nc]{popsize-ave}).
 
@@
 This is in part because of the log-log scale, in contrast to previous work
-\parencite{Nei1984-zi,Soule1976-he}; see Figure 1-supplement figure 2 for a
+\citep{Nei1984-zi,Soule1976-he}; see \FIGSUPP[pi-Nc]{popsize-linear} for a
 version on a log-linear scale. However, it is noteworthy that few taxa have
@@
 but are listed as least concern (though their presence in the Red List
-indicates they are of conservation interest). In Supplementary Information
-Section \Newnameref{supinfo:div-iucn}, I explore more about the relationships
-between IUCN Red List status, diversity, and population size.
-
-\subsection*{Phylogenetic Non-Independence and the Population Size Diversity Relationship}
+indicates they are of conservation interest). In Appendix \ref{app:div-iucn},
+\nameref{app:div-iucn}, I explore more about the relationships between IUCN Red
+List status, diversity, and population size.
+
+\subsection{Phylogenetic Non-Independence and the Population Size Diversity Relationship}
 
@@
 assumption of the regression model
-\parencite{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I
+\citep{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I
 fit this relationship using a phylogenetic mixed-effects model, investigated
@@
 lacked time-calibrated phylogenies and assumed unit branch lengths
-\parencite{Whitney2010-ud}, a shortcoming that has drawn criticism
-\parencite{Lynch2011-qv}. Here, I use a synthetic time-calibrated phylogeny
-created from the DateLife project \parencite{OMeara2020-ds} to account for
-shared phylogenetic history (see Methods: \Newnameref{sec:methods-pcm}).
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=\textwidth]{figures/diversity_pcm.pdf}
+\citep{Whitney2010-ud}, a shortcoming that has drawn criticism
+\citep{Lynch2011-qv}. Here, I use a synthetic time-calibrated phylogeny
+created from the DateLife project \citep{OMeara2020-ds} to account for
+shared phylogenetic history (see Methods and Materials: \nameref{sec:methods-pcm}).
+
+\begin{figure}
+  \begin{fullwidth}
+  \includegraphics[width=\linewidth]{figures/diversity_pcm.pdf}
 
@@
     cases, these outliers were in the genus \emph{Drosophila}).}
-
-  \label{fig:figure-2}
+    \label{fig:pcm}
+
+  \figsupp{The posterior distributions for the parameters of the phylogenetic
+    mixed-effects model of diversity and population size (this is analogous to
+    Figure \FIG{pcm}B) fit separately on chordates ($n=68$), molluscs
+    ($n=13$), and arthropods ($n=68$). The phylogenetic mixed-effects model for
+    chordates indicated the best-fitting model had no residual variance
+    ($\sigma_r^2 = 0$), so an alternate model without this variance component
+    was used to ensure proper convergence; this model is shown in green.
+    The light blue (green) shaded regions are the 90\% credible intervals, the
+    blue (green) lines the posterior averages, the gray shaded regions the OLS
+    bootstrap 95\% confidence intervals, and the gray lines the OLS estimate.
+    Note that unlike \FIG{pcm}, the OLS estimate uses all taxa,
+    not just those present in the phylogeny, since splitting the data by phyla
+    reduces sample sizes (OLS with just the subset of taxa in the phylogeny is
+    not significant for either chordates and arthropods). The vertical dashed
+  gray line indicates zero.}{\includegraphics[width=\textwidth]{figures/div_popsize_post_phyla.pdf}}\label{figsupp:div-popsize-post-phyla}
+
+  \figsupp{The ancestral continuous trait estimates for diversity and population size with species labels.}{\includegraphics[width=\textwidth]{figures/biconmap_diversity.pdf}}\label{figsupp:anc-species}
+
+  \figsupp{The ancestral continuous trait estimates for recombination map
+  length and diversity and population size with species
+labels.}{\includegraphics[width=\textwidth]{figures/biconmap.pdf}}\label{figsupp:anc-maplen}
+
+  \end{fullwidth}
 \end{figure}
@@
 Using a phylogenetic mixed-effects model
-\parencite{Lynch1991-tp,Hadfield2010-ql,De_Villemereuil2014-kt} implemented in
-Stan \parencite{Carpenter2017-xi,Stan_Development_Team2020-ea}, I estimated the
-linear relationship between diversity and population size (on a log-log scale)
+\citep{Lynch1991-tp,Hadfield2010-ql,De_Villemereuil2014-kt} implemented in Stan
+\citep{Carpenter2017-xi,Stan_Development_Team2020-ea}, I estimated the linear
+relationship between diversity and population size (on a log-log scale)
 accounting for phylogeny, for the 166 taxa without missing data and present in
@@
 positive and significant (95\% credible interval $0.03, 0.11$), though somewhat
-attenuated compared to the OLS estimates (Figure \ref{fig:figure-2}B). Since
+attenuated compared to the OLS estimates (\FIG{pcm}B). Since
 the population size estimates are based on range and body mass, they are
@@
 separately on body mass and range indicates these have significant negative and
-positive effects, respectively (Supplementary Figure
-\ref{suppfig:figure-div-range-bodymass}; see also Supplementary Figure
-\ref{suppfig:div_leffler} for the relationship between diversity and the range
-categories of \cite{Leffler2012-zj}).
+positive effects, respectively (\FIGSUPP[pi-Nc]{div-range-bodymass}; see also
+\FIGSUPP[pi-Nc]{div-leffler} for the relationship between diversity and the
+range categories of \cite{Leffler2012-zj}).
 
@@
 $[0.58, 0.75]$) indicates that the majority of the variance perhaps might be
-due to shared phylogenetic history (Figure \ref{fig:figure-2}B).
+due to shared phylogenetic history (\FIG{pcm}B).
 
@@
 seems to be driven in part by chordates having low diversity and small
-population sizes compared to non-chordates (Figure \ref{fig:figure-2}A). This
+population sizes compared to non-chordates (\FIG{pcm}A). This
 problem resembles Felsenstein's worst-case scenario
-\parencite{Felsenstein1985-an,Uyeda2018-wf}, where a singular event on a
+\citep{Felsenstein1985-an,Uyeda2018-wf}, where a singular event on a
 lineage separating two clades generates a spurious association between two
@@
 to phyla-level subsets of the data for clades with sufficient sample sizes (see
-Methods: \Newnameref{sec:methods-pcm}). This analysis shows a significant
-positive relationship between diversity and population size in arthropods, and
-positive weak relationships in molluscs and chordates (Supplementary Figure
-\ref{suppfig:div_popsize_post_phyla}). Each of the 90\% credible intervals for
-slope overlap, indicating the relationship between $\pi$ and $N_c$ is similar
-across these clades.
+Methods: \nameref{sec:methods-pcm}). This analysis shows a significant positive
+relationship between diversity and population size in arthropods, and positive
+weak relationships in molluscs and chordates
+(\FIGSUPP[pcm]{div-popsize-post-phyla}). Each of the 90\% credible intervals
+for slope overlap, indicating the relationship between $\pi$ and $N_c$ is
+similar across these clades.
 
 Additionally, I have explored the rate of trait change through time using
-node-height tests \parencite{Freckleton2006-ww}. Node-height tests regress the
+node-height tests \citep{Freckleton2006-ww}. Node-height tests regress the
 absolute values of the standardized contrasts between lineages against the
@@
 standardized contrasts are estimates of the rate of character evolution
-\parencite{Felsenstein1985-an}; if a trait evolves under constant rate BM, this
+\citep{Felsenstein1985-an}; if a trait evolves under constant rate BM, this
 relationship should be flat. For both diversity and population size,
@@
 towards the present (robust regression p-values 0.023 and 0.00018 respectively;
-Figure \ref{fig:figure-2}C). Considering the constituents of the population
+\FIG{pcm}C). Considering the constituents of the population
 size estimate, range and body mass, separately, the rate of evolution of range
@@
 Interestingly, the diversity node-height test reveals two rate shifts at deeper
-splits (Figure \ref{fig:figure-2}C, top left) around 570 Mya. These nodes
+splits (\FIG{pcm}C, top left) around 570 Mya. These nodes
 represent the branches between tunicates and vertebrates in chordates, and
@@
 or mutation rate, and there is some evidence of both in this genus \emph{Ciona}
-\parencite{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts
+\citep{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts
 in diversity is mirrored in the population size node-height test (Figure
-\ref{fig:figure-2}C, top right). Rather, it appears a trait impacting diversity
+\FIG{pcm}C, top right). Rather, it appears a trait impacting diversity
 but not census size (e.g. mutation rate or offspring distributions) has
@@
 
-\subsection*{Assessing the Impact of Linked Selection on Diversity Across Taxa}
+\subsection{Assessing the Impact of Linked Selection on Diversity Across Taxa}
 
@@
 that acts to reduce diversity levels from what we would expect given census
-sizes \parencite{Maynard_Smith1974-zr,Gillespie2000-mh,Corbett-Detig2015-gt}.
+sizes \citep{Maynard_Smith1974-zr,Gillespie2000-mh,Corbett-Detig2015-gt}.
 Here, I test this hypothesis by estimating the scale of diversity reductions
@@
 linked selection models that incorporate both hitchhiking and background
-selection \parencite{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
+selection \citep{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
 difficulty is understanding of how $R$ varies across taxa, since we lack
@@
 selection across species, $\pi_{BGS+HH} = R \pi_0$, match the observed
-relationship in Figure \ref{fig:figure-1}?
+relationship in \FIG{pi-Nc}?
 
@@
 \emph{Drosophila melanogaster}, a species known to be strongly affected by
-linked selection \parencite{Sella2009-nx}. Under a generalized model of
-hitchhiking and background selection \parencite{Elyashiv2016-vt,Coop2012-cd}
+linked selection \citep{Sella2009-nx}. Under a generalized model of
+hitchhiking and background selection \citep{Elyashiv2016-vt,Coop2012-cd}
 and assuming $N_e = N_c$, expected diversity is
@@
 deleterious mutation rate, and $L$ is the recombination map length
-\parencite{Hudson1994-oh,Hudson1995-xc,Nordborg1996-nq}. This BGS model is
+\citep{Hudson1994-oh,Hudson1995-xc,Nordborg1996-nq}. This BGS model is
 similar to models of effective population size under polygenic fitness
@@
 and $U$) to high values estimated from \emph{Drosophila melanogaster} by
-\textcite{Elyashiv2016-vt}. My estimate of the adaptive substitutions per
+\cite{Elyashiv2016-vt}. My estimate of the adaptive substitutions per
 generation ($\gamma_\text{Dmel}$) based Elyashiv et al. implies a rate of
 sweeps per basepair of $\nu_\text{BP,Dmel} \approx 2.34 \times 10^{-11}$, which
-is close to other estimates from \emph{D.  melanogaster} (see Supplementary
-Figure \ref{suppfig:linked-sel-params}A). The rate of deleterious mutations per
+is close to other estimates from \emph{D.  melanogaster} (see
+\FIGSUPP[linksel]{linked-sel-params}A). The rate of deleterious mutations per
 diploid genome, per generation is parameterized using the estimate from
@@
 estimates based on Bateman-Mukai approaches
-\parencite{Mukai1985-bc,Mukai1988-vs,Charlesworth1987-ab}. Finally, the
-probability that a lineage is trapped in a sweep, $J_\text{Dmel}$, is
-calculated from the estimated genome-wide average coalescent rate due to sweeps
-from Elyashiv et al. (see Supplementary Figured
-\ref{suppfig:linked-sel-params}B and Methods:
-\Newnameref{sec:methods-reduction} for more details on parameter estimates).
+\citep{Mukai1985-bc,Mukai1988-vs,Charlesworth1987-ab}. Finally, the probability
+that a lineage is trapped in a sweep, $J_\text{Dmel}$, is calculated from the
+estimated genome-wide average coalescent rate due to sweeps from Elyashiv et
+al. (see \FIGSUPP[linksel]{linked-sel-params}B and Methods:
+\nameref{sec:methods-reduction} for more details on parameter estimates).
 Using these \emph{Drosophila} parameters, I then explore how the predicted
@@
 \begin{figure}[t!]
-  \centering
-  \includegraphics[width=\textwidth]{figures/figure_3.pdf}
+  \begin{fullwidth}
+  \includegraphics[width=\linewidth]{figures/figure_3.pdf}
 
@@
     fitting since these have adaptively higher recombination map lengths
-    \parencite{Wilfert2007-dx}. The dark gray line is the estimated
+    \citep{Wilfert2007-dx}. The dark gray line is the estimated
     relationship under a phylogenetic mixed-effects model, and the gray
@@
 under \emph{Drosophila melanogaster} parameters. In both plots, point color
-indicates phylum.}
-
-  \label{fig:figure-3}
+indicates phylum.}\label{fig:linksel}
+
+  \figsupp{The relationship between genome size and approximate census
+    population size. The dashed gray line indicates the OLS fit. Tiger
+    salamander (\emph{Ambystoma tigrinum}) was excluded because of its
+  exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{figures/genome_size_popsize.pdf}}\label{figsupp:genome-size-Nc}
+
+  \figsupp{The relationship between genome size and recombination map length.
+    The dashed gray line indicates the OLS fit for all taxa, and the dashed
+    colored dashed lines indicate the linear relationship fit by phyla. Tiger
+    salamander (\emph{Ambystoma tigrinum}) was excluded because of its
+  exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{figures/genome_size_maplength.pdf}}\label{figsupp:genome-size-maplen}
+
+  \figsupp[The relationship between $N_c$ and diversity in the
+    \cite{Corbett-Detig2015-gt} data, and the relationship between estimated
+    reduction in diversity and census size, for three different approaches.]{(A)
+    The diversity data from \cite{Corbett-Detig2015-gt} and the census
+    population size estimated here for metazoan taxa. (B) The reductions in
+    diversity, $R = \nicefrac{N_e}{N}$, plotted against census size across
+    species. The red points are the reductions estimated by
+    \cite{Corbett-Detig2015-gt}. This confirms
+    \citeauthor{Corbett-Detig2015-gt}'s (2015) finding that the impact of
+    selection ($I = 1-R$) increases with census population size (though, in the
+    original paper size body size and range were used as separate proxy
+    variables for census population size). The green and red points are the
+    predicted reduction in diversity under the recurrent hitchhiking (RHH) and
+    background selection (BGS) model using the \emph{Drosophila melanogaster}
+    parameters as described in the main text.  The reduction in the diversity
+    due to sweeps, from Equation 1, is determined by the term
+    $2NS$.  Green points treat $N$ as the implied effective population size
+    from diversity $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, assuming
+    $\mu = 10^{-9}$. Yellow points treat $N$ as the census size, $N = N_c$.
+    Overall, using the census size, e.g.  $2N_cS$, leads to reductions in
+    diversity that far exceed the empirical estimates of Corbett-Detig et al.
+  and reasonable model-based predictions from
+$\widetilde{N}_e$.}{\includegraphics[width=\textwidth]{figures/corbett_detig.pdf}}\label{figsupp:reductions}
+
+\figsupp{Comparison of the \emph{Drosophila} sweep parameters used in this study with parameters from other studies.
+     (A) The estimate of the number of sweeps per basepair, per genome
+      ($\nu_\text{BP}$) from Table 2 of \cite{Elyashiv2016-vt} (the studies
+      included are \cite{Li2006-xa,Andolfatto2007-uy,Macpherson2007-qt} and
+      \cite{Jensen2008-vz}); the red point is my estimate used in
+      this paper. (B) Points are the data from \cite{Shapiro2007-gp}.  The
+      blue line is the non-linear least squares fit to the data, and the green
+      dashed line is the sweep model parameterized by the genome-wide average
+      sweep coalescent rate $2NS \approx 0.92$ from the classic sweep and
+      background selection model of \cite{Elyashiv2016-vt} ($r_s$ in
+    Supplementary Table S6).}{\includegraphics[width=\textwidth]{figures/linked_sel_params.pdf}}\label{figsupp:linked-sel-params}
+
+  \end{fullwidth}
 \end{figure}
@@
 $N_c$ (\cite{Corbett-Detig2015-gt}; see also Supplementary Figure
-\ref{suppfig:corbett_detig}A), and it is often thought that this is driven by
+\ref{figsupp:corbett_detig}A), and it is often thought that this is driven by
 higher rates of adaptive substitutions in larger populations
-\parencite{Ohta1992-yi}, despite equivocal evidence \parencite{Galtier2016-dq}.
+\citep{Ohta1992-yi}, despite equivocal evidence \citep{Galtier2016-dq}.
 However, there is another mechanism by which species with larger population
 sizes might experience a greater impact of linked selection: recombinational
-map length, $L$, is known to correlate with body mass \parencite{Burt1987-tq}
-and thus varies inversely with population size. As this is a critical parameter
+map length, $L$, is known to correlate with body mass \citep{Burt1987-tq} and
+thus varies inversely with population size. As this is a critical parameter
 that determines the genome-wide impact of both hitchhiking and background
@@
 and census population size ($N_c$) across taxa, using available estimates of
-map lengths across species \parencite{Stapley2017-fs,Corbett-Detig2015-gt}. I
-find a significant non-linear relationship using phylogenetic mixed-effects
-models (Figure \ref{fig:figure-3}A; see Methods: \Newnameref{sec:methods-pcm}).
-There is also a correlation between map length and genome size (Supplementary
-Figure \ref{supfig:genome_size_maplength}) and genome size and population size
-(Supplementary Figure \ref{supfig:genome_size_popsize}). These findings are
-consistent with both the hypothesis that non-adaptive processes increase genome
-size in small-$N_e$ species \parencite{Lynch2003-bn} which in turn could
-increase map lengths, as well as the hypothesis that map lengths are adaptively
-longer to more efficiently select against deleterious alleles
-\parencite{Roze2021-ya}. Overall, the negative relationship between map length
-and census size indicates linked selection is expected to be stronger in short
-map length, high-$N_c$ species.
+map lengths across species \citep{Stapley2017-fs,Corbett-Detig2015-gt}. I find
+a significant non-linear relationship using phylogenetic mixed-effects models
+(Figure \ref{fig:figure-3}A; see Methods and Materials:
+\nameref{sec:methods-pcm}).  There is also a correlation between map length and
+genome size (Supplementary Figure \ref{supfig:genome_size_maplength}) and
+genome size and population size (Supplementary Figure
+\ref{supfig:genome_size_popsize}). These findings are consistent with both the
+hypothesis that non-adaptive processes increase genome size in small-$N_e$
+species \citep{Lynch2003-bn} which in turn could increase map lengths, as well
+as the hypothesis that map lengths are adaptively longer to more efficiently
+select against deleterious alleles \citep{Roze2021-ya}. Overall, the negative
+relationship between map length and census size indicates linked selection is
+expected to be stronger in short map length, high-$N_c$ species.
 
@@
 species within this range than in \emph{Drosophila}, which is incompatible with
-the rate of adaptive substitutions across species \parencite{Galtier2016-dq}
-and mutation rates \parencite{Lynch2010-ki}. Furthermore, linked selection has
+the rate of adaptive substitutions across species \citep{Galtier2016-dq}
+and mutation rates \citep{Lynch2010-ki}. Furthermore, linked selection has
 been quantified in humans, which fall in this census size range, and has been
 found to be relatively weak
-\parencite{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
+\citep{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
 Second, while hitchhiking and BGS can reduce predicted diversity levels for
@@
 magnitude in \emph{Drosophila} (Supplementary Figured
-\ref{suppfig:corbett_detig}B). The high reductions in $\pi$ predicted here
+\ref{figsupp:corbett_detig}B). The high reductions in $\pi$ predicted here
 (compared to those of \cite{Elyashiv2016-vt}) are a result of using $N_c$,
@@
 substitutions per Morgan would only act to limit the reduction in diversity
-\parencite{Weissman2012-aa}, and thus these results are conservative.
+\citep{Weissman2012-aa}, and thus these results are conservative.
 Overall, while linked selection could decouple diversity from census size for
@@
 
-\section*{Discussion}
+\section{Discussion}
 
@@
 distribution modeling to improve census size estimates
-\parencite{Bar-On2018-kc, Mora2011-wm}.  Still, it seems more accurate
+\citep{Bar-On2018-kc,Mora2011-wm}.  Still, it seems more accurate
 estimates would be unlikely to change the qualitative findings here, which
-resemble those of early surveys \parencite{Nei1984-zi,Soule1976-he}.
+resemble those of early surveys \citep{Nei1984-zi,Soule1976-he}.
 
@@
 
-\subsection*{Macroevolution and Across-Taxa Population Genomics}
+\subsection{Macroevolution and Across-Taxa Population Genomics}
 
@@
 methods. Extending previous work that has accounted for phylogeny in particular
-clades \parencite{Leffler2012-zj}, or using taxonomical-level averages
-\parencite{Romiguier2014-bp}, I show that the positive relationship between
+clades \citep{Leffler2012-zj}, or using taxonomical-level averages
+\citep{Romiguier2014-bp}, I show that the positive relationship between
 diversity and census size is significant using a mixed-effects model with a
@@
 previous concerns about phylogenetic non-independence in comparative population
-genetic studies were warranted \parencite{Gillespie1991-qa,Whitney2010-ud}.
+genetic studies were warranted \citep{Gillespie1991-qa,Whitney2010-ud}.
 Notably, Lynch (2011) has argued that PCMs for pairwise diversity are
@@
 certain causal interpretations over others
-\parencite{Westoby1995-wz,Uyeda2018-wf}. Future work could try to untangle what
+\citep{Westoby1995-wz,Uyeda2018-wf}. Future work could try to untangle what
 causal factors determine coalescent times across species, as well as how these
@@
 misconception that a fast rate of trait evolution necessarily reduces
-phylogenetic signal \parencite{Revell2008-in}, and that if either or both
+phylogenetic signal \citep{Revell2008-in}, and that if either or both
 variables in a regression are free of phylogenetic signal, PCMs are unnecessary
-\parencite{Revell2010-mf,Uyeda2018-wf}. The evidence of high phylogenetic
+\citep{Revell2010-mf,Uyeda2018-wf}. The evidence of high phylogenetic
 signal found in this study suggests PCMs are needed, in part to avoid spurious
@@
 
-\subsection*{Spatial and Demographic Processes}
+\subsection{Spatial and Demographic Processes}
 
@@
 which can have a profound impact on population genomic summary statistics
-\parencite{Battey2020-lc}. These issues could systematically bias species-wide
+\citep{Battey2020-lc}. These issues could systematically bias species-wide
 diversity estimates; for example, if diversity estimates from a cosmopolitan
@@
 that $F_\text{ST}$ is almost never this high within species
-\parencite{Roux2016-lm}.  Still, future work could quantify the extent to which
+\citep{Roux2016-lm}.  Still, future work could quantify the extent to which
 spatial processes contribute to Lewontin's Paradox. For example, high-$N_c$
@@
 
-\subsection*{How could selection still explain Lewontin's Paradox?}
+\subsection{How could selection still explain Lewontin's Paradox?}
 
@@
 linked selection poorly fit observed patterns of diversity across species.
-This result extends the analysis by \textcite{Coop2016-gx} showing that levels
-of $\pi_0$ estimated by \textcite{Corbett-Detig2015-gt} are not decoupled from
+This result extends the analysis by \cite{Coop2016-gx} showing that levels
+of $\pi_0$ estimated by \cite{Corbett-Detig2015-gt} are not decoupled from
 genome-wide average $\pi$, as would occur if linked selection were to explain
@@
 selection and recurrent hitchhiking models
-\parencite{Kaplan1989-sc,Hudson1995-xc,Coop2012-cd}, when in reality, there is
+\citep{Kaplan1989-sc,Hudson1995-xc,Coop2012-cd}, when in reality, there is
 genome-wide heterogeneity in functional density, recombination rates, and the
@@
 of linked selection, $\pi_0$, is currently underestimated by several orders of
-magnitude. Moreover, the study of \textcite{Corbett-Detig2015-gt} did consider
+magnitude. Moreover, the study of \cite{Corbett-Detig2015-gt} did consider
 recombination rate and functional density heterogeneity in estimating the
@@
 reductions are orders of magnitude weaker than those considered here by
-assuming that $N_e = N_c$ (Supplementary Figure \ref{suppfig:corbett_detig}B).
+assuming that $N_e = N_c$ (Supplementary Figure \ref{figsupp:corbett_detig}B).
 Overall, even with more realistic models of linked selection, current models of
@@
 models of partial sweeps and hard sweeps, though these imply different rates of
-sweeps \parencite{Coop2012-cd}. Second, in the limit where most fitness
+sweeps \citep{Coop2012-cd}. Second, in the limit where most fitness
 variation is due to weak soft sweeps from standing variation scattered across
@@
 well approximated by quantitative genetic linked selection models
-\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}. The reduction in
+\citep{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}. The reduction in
 diversity under these models is nearly identical to that under background
@@
 Section \ref{app:bgs}; \cite{Charlesworth2000-km,Charlesworth2015-am}). Third,
-the parameters from \textcite{Elyashiv2016-vt} are robust to many types of
+the parameters from \cite{Elyashiv2016-vt} are robust to many types of
 sweeps that result in substitution (e.g. see p. 19 of their Supplementary
@@
 spread, and compete against other segregating beneficial variants
-\parencite{Ralph2015-kl,Ralph2010-ki}. Through limited dispersal should act to
+\citep{Ralph2015-kl,Ralph2010-ki}. Through limited dispersal should act to
 ``soften sweeps" and not impact my findings for the reasons described above,
@@
 selection generally reduces diversity under most conditions
-\parencite{Novak2017-yp}, this could lead seasonality and other sources of
+\citep{Novak2017-yp}, this could lead seasonality and other sources of
 temporal heterogeneity to reduce diversity in large-$N_c$ species with short
@@
 Future work could consider the impact of fluctuating selection on diversity
-under simple models \parencite{Barton2000-zg} if estimates of key parameters
+under simple models \citep{Barton2000-zg} if estimates of key parameters
 governing the rate of such fluctuations were known across taxa. Additionally,
@@
 
-\subsection*{Measures of Effective Population Size, Timescales, and Lewontin's Paradox}
+\subsection{Measures of Effective Population Size, Timescales, and Lewontin's Paradox}
 
@@
 calculable from different data and summary statistics
-\parencite{Wang2016-mi,Caballero1994-ao,Caballero2020-wm,Galtier2020-fb}. These
+\citep{Wang2016-mi,Caballero1994-ao,Caballero2020-wm,Galtier2020-fb}. These
 include estimators based on the site frequency spectrum, observed decay in
@@
 population size on shorter timescales than coalescent-based estimators
-\parencite{Wang2005-wy}, and thus could be used to tease apart processes that
+\citep{Wang2005-wy}, and thus could be used to tease apart processes that
 impact the $N_e$-$N_c$ relationship in the more recent past.
@@
 $\nicefrac{N_e}{N_c}$, which is an important quantity in conservation genetics
-\parencite{Frankham1995-xz,Mace1991-ai} and in understanding evolution in
+\citep{Frankham1995-xz,Mace1991-ai} and in understanding evolution in
 highly fecund marine species. Surveys of the short-term $\nicefrac{N_e}{N_c}$
 relationship across taxa indicate mean $\nicefrac{N_e}{N_c}$ is on order of
-$\approx 0.1$ \parencite{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though
+$\approx 0.1$ \citep{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though
 the uncertainty in these estimates is high, and some species with sweepstakes
@@
 constrains the variance in reproductive success
-\parencite{Wright1938-tv,Nunney1993-ef,Nunney1996-wy}. This implies that apart
+\citep{Wright1938-tv,Nunney1993-ef,Nunney1996-wy}. This implies that apart
 from species with sweepstakes reproductive systems, the variance in
@@
 be predictive of diversity levels across taxa
-\parencite{Romiguier2014-bp,Hallatschek2018-iu,Barry2020-pp}.
+\citep{Romiguier2014-bp,Hallatschek2018-iu,Barry2020-pp}.
 
@@
 resolving Lewontin's Paradox. These estimators, along with short-term estimates
-of the impact of linked selection \parencite{Buffalo2019-qs,Buffalo2020-my},
+of the impact of linked selection \citep{Buffalo2019-qs,Buffalo2020-my},
 can inform us how much diversity is depressed across shorter timescales, free
@@
 are only possible if short-term effective population sizes are orders of
-magnitude larger \parencite{Karasov2010-fb,Barton2010-pi}.
-
-\subsection*{Conclusions}
+magnitude larger \citep{Karasov2010-fb,Barton2010-pi}.
+
+\subsection{Conclusions}
 
@@
 
-\section*{Methods}
-
-\subsection*{Diversity and Map Length Data}
+\section{Methods and Materials}
+
+\subsection{Diversity and Map Length Data}
 \label{sec:methods-diversity}
@@
 surveys. Of the 172 taxa with diversity estimates, 14 are from
-\textcite{Corbett-Detig2015-gt}, 96 are from \textcite{Leffler2012-zj}, and 62
-are from \textcite{Romiguier2014-bp}. The Corbett-Detig et al. data is
+\cite{Corbett-Detig2015-gt}, 96 are from \cite{Leffler2012-zj}, and 62
+are from \cite{Romiguier2014-bp}. The Corbett-Detig et al. data is
 estimated from four-fold degenerate sites, the Romiguier et al. data is
@@
 silent, intronic, and non-coding sites. All types of diversity estimates from
-\textcite{Leffler2012-zj} were included to maximize the taxa in the study,
+\cite{Leffler2012-zj} were included to maximize the taxa in the study,
 since the variability of diversity across functional categories is much less
@@
 
-\subsection*{Macroecological Estimates of Population Size}
+\subsection{Macroecological Estimates of Population Size}
 \label{sec:methods-popsize}
@@
 Then, I re-estimated the relationship between body mass and population density
-using the data in the appendix table of \textcite{Damuth1987-sg}, which
+using the data in the appendix table of \cite{Damuth1987-sg}, which
 includes 696 taxa with body mass and population density measurements across
@@
 arthropods. Though the abundance-body size relationship can be noisy at small
-spatial or phylogenetic scales \parencite[Chapter 5,][]{Gaston2008-pn}, across
+spatial or phylogenetic scales \citep[Chapter 5,][]{Gaston2008-pn}, across
 deeply diverged taxa such as those included in this study and
-\textcite{Damuth1987-sg}, the relationship is linear and homoscedastic (see
-Supplementary Figure \ref{suppfig:damuth}). Using Stan
-\parencite{Stan_Development_Team2020-ea}, I jointly estimated the relationship
-between body mass from body length using the \textcite{Romiguier2014-bp} taxa,
+\cite{Damuth1987-sg}, the relationship is linear and homoscedastic (see
+Supplementary Figure \ref{figsupp:damuth}). Using Stan
+\citep{Stan_Development_Team2020-ea}, I jointly estimated the relationship
+between body mass from body length using the \cite{Romiguier2014-bp} taxa,
 and used this relationship to predict body mass for the taxa in this study.
 These body masses were then used to predict population density simultaneously,
-using the \textcite{Damuth1981-st} relationship. The code of this routine
+using the \cite{Damuth1981-st} relationship. The code of this routine
 (\texttt{pred\_popsize\_missing\_centered.stan}) is available in the GitHub
@@
 To estimate range, I first downloaded occurrence records from Global
-Biodiversity Information Facility \parencite{noauthor_2020-ey} using the
-\texttt{rgbif} R package \parencite{Chamberlain2014-up,Chamberlain2017-uz}.
+Biodiversity Information Facility \citep{noauthor_2020-ey} using the
+\texttt{rgbif} R package \citep{Chamberlain2014-up,Chamberlain2017-uz}.
 Using the occurrence locations, I inferred whether a species was marine or
@@
 overlapped a continent using \texttt{rnaturalearth} and the \texttt{sf}
-packages \parencite{South2017-db,Pebesma2018-bs}. For each taxon, I estimated
+packages \citep{South2017-db,Pebesma2018-bs}. For each taxon, I estimated
 its range by finding the minimum $\alpha$-shape containing these occurrences.
@@
 The range of \emph{C. elegans} was conservatively approximated as the area of
-the Western US and Western Europe based on the map in \textcite{Frezal2015-iw}.
+the Western US and Western Europe based on the map in \cite{Frezal2015-iw}.
 \emph{Drosophila} species ranges are from the Drosophila Speciation Patterns
-website, \parencite{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate
+website, \citep{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate
 these range estimates, I have compared these to the qualitative range
-descriptions \textcite{Leffler2012-zj} (Supplementary Figure
-\ref{suppfig:range-cats}) and compared my $\alpha$-shape method to a subset of
+descriptions \cite{Leffler2012-zj} (Supplementary Figure
+\ref{figsupp:range-cats}) and compared my $\alpha$-shape method to a subset of
 taxa with range estimates from IUCN Red List
 (\cite{Chamberlain2020-tv,Iucn2020-ap}; Supplementary Figure
-\ref{suppfig:redlist-ranges}). Each census population size
+\ref{figsupp:redlist-ranges}). Each census population size
 is then estimated as the product of range and density.
 
-\subsection*{Phylogenetic Comparative Methods}
+\subsection{Population Size Validation}
+\label{supp:popsize-valid}
+
+I validated the approximate census sizes by comparing the implied biomass of
+these estimates to estimates of the total carbon biomass on earth by phylum
+\citep{Bar-On2018-kc}. For species $i$ with wet body mass $m_i$ and census
+size $N_i$, the implied biomass is $m_i N_i$. For all species in a phylum $S$,
+this total sample biomass is $b_S = \sum_{i \in S} m_i N_i$. I then compare
+this wet biomass to the carbon biomasses by phylum by \cite{Bar-On2018-kc}.
+Across animal species, the ratio of dry to wet body mass, and carbon body mass
+dry body mass varies little. In their study, Bar-On et al. assume wet body mass
+has a 70\% water content, and 50\% of dry body mass is carbon mass, leading to
+a wet body mass to carbon mass factor of $\nicefrac{1-0.7}{0.5} = 0.15$. I use
+this factor to convert the total wet biomass to carbon biomass per phylum.
+
+\begin{table}[ht]
+  \resizebox{\textwidth}{!}{%
+\begin{tabular}{lr|rr|rrrrr|r}
+ \toprule
+  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{5}{c}{\bf Present study}  &  \\
+  phylum & total species ($T$) & biomass ($B$) & prop.
+biomass & biomass ($b$) &  prop.
+biomass & num.
+species ($n$) & factor overrepresented &prop.
+total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
+ \midrule
+  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.80 \times 10^{-4}$ & 0.0102 &  68 & 0.02 & $5.41 \times 10^{-5}$ & 4.31 \\ 
+  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.67 \times 10^{-2}$ & 0.9715 &  68 & 2.89 & $1.26 \times 10^{-3}$ & 24.40 \\ 
+  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & 0.01 & $1.76 \times 10^{-4}$ & 0.35 \\ 
+  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.56 \times 10^{-4}$ & 0.0166 &  13 & 0.21 & $1.36 \times 10^{-4}$ & 16.70 \\ 
+  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.07 \times 10^{-5}$ & 0.0011 &   2 & 0.03 & $1.25 \times 10^{-4}$ & 2.45 \\ 
+  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $4.03 \times 10^{-6}$ & 0.0001 &   1 & 0.02 & $4.00 \times 10^{-5}$ & 5.03 \\
+ \bottomrule
+\end{tabular}}
+\caption{How the total carbon biomass estimates by phylum from
+  \cite{Bar-On2018-kc} compare to the implied biomass estimates from this
+  study. All biomass estimates are carbon biomass, and the proportions are of
+  total biomass with respect to the study. The proportion of biomass in this
+  study compared to the Bar-On et al. estimates \cite{Bar-On2018-kc}
+  indicates chordates are overrepresented and arthropods are underrepresented
+  in the present study; the factor that each phylum is overrepresented is
+  given in the eighth column. Total species by phylum estimates are from
+  \cite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
+  ratio column is the ratio of total biomass implied by the $N_c$ estimates
+of each species in a phylum to the actual biomass of that phylum.}
+\label{supp:biomass-tbl}
+\end{table}
+
+First, I compared the relative carbon biomass in this study to the relative
+carbon biomass on earth per phylum. This shows that this study's sample over
+represents chordate biomass (by a factor of $\mathord{\sim}3$), and under
+represents in arthropod biomass (by a factor of $0.02$) relative to the
+proportion of carbon biomass of these phyla on earth (see column eight of
+Supplementary Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
+biomass per phylum in the sample was broadly consistent with the total on earth
+by phylum ($B_S$ for phylum $S$), I calculated the expected sample biomass if
+species were sampled randomly from the total species in a phylum, ($B_S \times
+\nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
+in phylum $S$, $T_S$ is the total number of species in phylum $S$ on earth).
+The fraction of total species on earth included in the sample in this study is
+depicted in Supplementary Figure \ref{figsupp:species_counts}.
+
+Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
+biomass per phylum (Supplementary Table \ref{supp:biomass-tbl}). The
+consistency is quite close for this rough approach and the non-random sample of
+taxa included in this study. The carbon biomass estimates for chordates implied
+by the census size estimates are $\mathord{\sim}24$-fold higher than expected,
+but is well within reasonable expectations given that the chordate sample
+includes many larger-bodied domesticated species (and is a biased sample in
+other ways).  Similarly, the implied arthropod carbon biomass is quite close to
+what one would expect. Overall, these values indicate that the census size
+estimates here do not lead to implied biomasses per phylum that are outside the
+range of plausibility. For other population size consistency checks, see
+Appendix \ref{app:popsize-valid}.
+
+\subsection{Phylogenetic Comparative Methods}
 \label{sec:methods-pcm}
@@
 synthetic calibrated phylogeny was created for 166 species that appear in
-phylogenies in DateLife project \parencite{OMeara2020-ds,Sanchez-Reyes2019-io}.
+phylogenies in DateLife project \citep{OMeara2020-ds,Sanchez-Reyes2019-io}.
 This calibrated synthetic phylogeny was then subset for the analyses based on
@@
 relationship assessed by a linear phylogenetic mixed-effects model implemented
-in Stan \parencite{Stan_Development_Team2020-ea}, according to the methods
+in Stan \citep{Stan_Development_Team2020-ea}, according to the methods
 described in (\cite{De_Villemereuil2014-kt}, see
@@
 the chordate subset due to identifiability issues leading to poor MCMC
-convergence (Supplementary Figure \ref{suppfig:div_popsize_post_phyla}).
+convergence (Supplementary Figure \ref{figsupp:div_popsize_post_phyla}).
 
@@
 data. Since social insects have longer recombination map lengths
-\parencite{Wilfert2007-dx}, social taxa were excluded when fitting this model.
-All \texttt{Rhat} \parencite{Vehtari2019-po} values were below 1.01 and the
-effective number of samples was over 1,000, consistent with good mixing;
-details about the model are available in the GitHub repository
-(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (Figure
-\ref{fig:figure-2}A and Supplementary Information Figure \ref{suppfig:biconmap})
-were created using \texttt{phytools} \parencite{Revell2012-zq}. Node-height
-tests were implemented based on the methods in \texttt{Geiger}
-\parencite{Pennell2014-nx,Harmon2008-tr}, and use robust regression to fit a
-linear relationship between phylogenetic independent contrasts and branching
-times.
-
-\subsection*{Predicted Reductions in Diversity}
+\citep{Wilfert2007-dx}, social taxa were excluded when fitting this model.  All
+\texttt{Rhat} \citep{Vehtari2019-po} values were below 1.01 and the effective
+number of samples was over 1,000, consistent with good mixing; details about
+the model are available in the GitHub repository
+(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (\FIG{pcm}A and
+Supplementary Information Figure \ref{figsupp:biconmap}) were created using
+\texttt{phytools} \citep{Revell2012-zq}. Node-height tests were implemented
+based on the methods in \texttt{Geiger} \citep{Pennell2014-nx,Harmon2008-tr},
+and use robust regression to fit a linear relationship between phylogenetic
+independent contrasts and branching times.
+
+\subsection{Predicted Reductions in Diversity}
 \label{sec:methods-reduction}
@@
 melanogaster}, and the recombination map length estimates from
-\textcite{Stapley2017-fs} and \textcite{Corbett-Detig2015-gt}. The mathematical
+\cite{Stapley2017-fs} and \cite{Corbett-Detig2015-gt}. The mathematical
 details of the simplified sweep model are explained in the Appendix Section
@@
 regions between \emph{D.  melanogaster} and \emph{D. simulans} from
-\textcite{Hu2013-pg}.  Following \textcite{Elyashiv2016-vt}, only substitutions
+\cite{Hu2013-pg}.  Following \cite{Elyashiv2016-vt}, only substitutions
 in UTRs and exons are included, since they found no evidence of sweeps in
@@
 \emph{D. simulans} of $4.2 \times 10^6$ and estimate of ten generations per
-year \parencite{Obbard2012-on}, calculating there are $\gamma_\text{Dmel} =
+year \citep{Obbard2012-on}, calculating there are $\gamma_\text{Dmel} =
 \nicefrac{\alpha m}{2T} = 2.26 \times 10^{-3}$ substitutions per generation.
@@
 close to other previously-published estimates (Supplementary Figure
-\ref{suppfig:linked-sel-params}). Finally, I use $U_\text{Dmel} = 1.6$, from
-\textcite{Elyashiv2016-vt}. With these parameter estimates from \emph{D.
+\ref{figsupp:linked-sel-params}). Finally, I use $U_\text{Dmel} = 1.6$, from
+\cite{Elyashiv2016-vt}. With these parameter estimates from \emph{D.
 melanogaster}, the recombination map lengths across species, and Equation
@@
 
-\section*{Acknowledgments}
+\section{Acknowledgments}
 
@@
 
-\section*{Appendix}
-
-\setcounter{table}{0}
-\renewcommand{\theequation}{A\arabic{equation}}
-\setcounter{section}{0}
-\renewcommand{\thesection}{A\arabic{section}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{A\arabic{figure}}
-
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/diversity_popsize_linear.pdf}
+%   % figure 2-supplement 1
+%   \figsupp{A version of Figure \ref{fig:figure-1} with diversity on a linear,
+%     rather than log, scale. Points are colored by phylum, and the shaded region
+%     is the predicted neutral level of diversity assuming $N_e = N_c$ with mutation
+%   range ranging between $10^{-10} \le \mu \le 10^{-8}$. }
+
+%   \label{suppfig:figure-1-linear}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/diversity_popsize_averages.pdf}
+%   % figure 2-supplement 2
+%   \figsupp{Diversity and approximate population size for 172 taxa, colored by
+%     phylum; the dashed lines indicate the non-phylogenetic OLS estimates of the
+%   relationship between population size and diversity grouped by phyla.}
+
+%   \label{suppfig:figure-1-ave}
+% \end{figure}
+
+% \begin{figure}[!htb] \centering
+%   \includegraphics[width=\textwidth]{figures/diversity_range_bodymass.pdf}
+%   % figure 2-supplement 3
+%   \figsupp{The relationship between diversity (differences per basepair) and
+%     body mass (left) and range (right) across 172 species. The top row are
+%     posterior distributions of parameters estimated using the phylogenetic
+%     mixed-effects model using 166 taxa in the synthetic phylogeny for the
+%     intercept, slope, and phylogenetic signal from the mixed-effects model. The
+%     bottom row contain each species as a point, colored by phyla. The gray
+%     dashed line is the non-phylogenetic standard regression estimate, and the
+%     blue dashed line is the relationship fit by the phylogenetic mixed-effects
+%   model. }
+
+%   \label{suppfig:figure-div-range-bodymass}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/div_popsize_post_phyla.pdf}
+%   % figure 3-supplement 1
+%   \figsupp{The posterior distributions for the parameters of the phylogenetic
+%     mixed-effects model of diversity and population size (this is analogous to
+%     Figure \ref{fig:figure-2}B) fit separately on chordates ($n=68$), molluscs
+%     ($n=13$), and arthropods ($n=68$). The phylogenetic mixed-effects model for
+%     chordates indicated the best-fitting model had no residual variance
+%     ($\sigma_r^2 = 0$), so an alternate model without this variance component
+%     was used to ensure proper convergence; this model is shown in green.
+%     The light blue (green) shaded regions are the 90\% credible intervals, the
+%     blue (green) lines the posterior averages, the gray shaded regions the OLS
+%     bootstrap 95\% confidence intervals, and the gray lines the OLS estimate.
+%     Note that unlike Figure \ref{fig:figure-2}, the OLS estimate uses all taxa,
+%     not just those present in the phylogeny, since splitting the data by phyla
+%     reduces sample sizes (OLS with just the subset of taxa in the phylogeny is
+%     not significant for either chordates and arthropods). The vertical dashed
+%   gray line indicates zero.}
+
+%   \label{suppfig:div_popsize_post_phyla}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/biconmap_diversity.pdf}
+%   % figure 3-supplement 2
+%   \figsupp{}
+%   \label{suppfig:biconmap_diversity}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/biconmap.pdf}
+%   % figure 3-supplement 3
+%   \figsupp{}
+%   \label{suppfig:biconmap}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/genome_size_popsize.pdf}
+%   % figure 4-supplement 1
+%   \figsupp{The relationship between genome size and approximate census
+%     population size. The dashed gray line indicates the OLS fit. Tiger
+%     salamander (\emph{Ambystoma tigrinum}) was excluded because of its
+%   exceptionally large genome size (~30Gbp).}
+
+%   \label{supfig:genome_size_popsize}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/genome_size_maplength.pdf}
+%   % figure 4-supplement 2
+%   \figsupp{The relationship between genome size and recombination map length.
+%     The dashed gray line indicates the OLS fit for all taxa, and the dashed
+%     colored dashed lines indicate the linear relationship fit by phyla. Tiger
+%     salamander (\emph{Ambystoma tigrinum}) was excluded because of its
+%   exceptionally large genome size (~30Gbp).}
+%   \label{supfig:genome_size_maplength}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/corbett_detig.pdf}
+%   % figure 4-supplement 1
+%   \figsupp{(A) The diversity data from \textcite{Corbett-Detig2015-gt} and the
+%     census population size estimated here for metazoan taxa. (B) The
+%     reductions in diversity, $R = \nicefrac{N_e}{N}$, plotted against census
+%     size across species. The red points are the reductions estimated by
+%     \textcite{Corbett-Detig2015-gt}. This confirms
+%     \citeauthor{Corbett-Detig2015-gt}'s (2015) finding that the impact of
+%     selection ($I = 1-R$) increases with census population size (though, in the
+%     original paper size body size and range were used as separate proxy
+%     variables for census population size). The green and red points are the
+%     predicted reduction in diversity under the recurrent hitchhiking (RHH) and
+%     background selection (BGS) model using the \emph{Drosophila melanogaster}
+%     parameters as described in the main text.  The reduction in the diversity
+%     due to sweeps, from Equation \eqref{eqn:div}, is determined by the term
+%     $2NS$.  Green points treat $N$ as the implied effective population size
+%     from diversity $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, assuming
+%     $\mu = 10^{-9}$. Yellow points treat $N$ as the census size, $N = N_c$.
+%     Overall, using the census size, e.g.  $2N_cS$, leads to reductions in
+%     diversity that far exceed the empirical estimates of Corbett-Detig et al.
+%   and reasonable model-based predictions from $\widetilde{N}_e$.}
+%   \label{suppfig:corbett_detig}
+% \end{figure}
+
+% \begin{figure}[!htb]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/linked_sel_params.pdf}
+%   % figure 4-supplement 2
+%   \caption{(A) The estimate of the number of sweeps per basepair, per genome
+%     ($\nu_\text{BP}$) from Table 2 of \textcite{Elyashiv2016-vt} (the studies
+%     included are \cite{Li2006-xa,Andolfatto2007-uy,Macpherson2007-qt} and
+%     \cite{Jensen2008-vz}); the red point is my estimate used in
+%     this paper. (B) Points are the data from \textcite{Shapiro2007-gp}.  The
+%     blue line is the non-linear least squares fit to the data, and the green
+%     dashed line is the sweep model parameterized by the genome-wide average
+%     sweep coalescent rate $2NS \approx 0.92$ from the classic sweep and
+%   background selection model of \textcite{Elyashiv2016-vt} ($r_s$ in
+% Supplementary Table S6).}
+
+%   \label{suppfig:linked-sel-params}
+% \end{figure}
+
+\subsection{Citations}
+\bibliography{biblio.bib}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% APPENDICES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\appendix
+\begin{appendixbox}
 \section{Simplified Sweep Effects Model}
@@
 (cf. equation 1 \cite{Elyashiv2016-vt}, and equation 20 of \cite{Coop2012-cd}).
-The BGS component is given by \textcite{Hudson1995-xc},
+The BGS component is given by \cite{Hudson1995-xc},
 
@@
 Since I use \emph{Drosophila melanogaster} parameter estimates from
-\textcite{Elyashiv2016-vt}, I now reconcile their model's $S$ term with the
+\cite{Elyashiv2016-vt}, I now reconcile their model's $S$ term with the
 simple model above. They estimate $S$ in \emph{Drosophila melanogaster} using a
@@
 
-Note, that we can recover the model of \textcite{Coop2012-cd} from this
+Note, that we can recover the model of \cite{Coop2012-cd} from this
 expression. Suppose there is only one annotation class, and $\alpha$ fraction
@@
 %
-This is analogous to the second term of \textcite{Coop2012-cd} equation 17,
+This is analogous to the second term of \cite{Coop2012-cd} equation 17,
 with $k = i = 2$ and $x = 1$ (e.g.  conditioning on a sweep to fixation). Note
-that there appears to be a factor of two error in \textcite{Elyashiv2016-vt}
-compared to \textcite{Coop2012-cd}; here I include the factor of two.  Then,
+that there appears to be a factor of two error in \cite{Elyashiv2016-vt}
+compared to \cite{Coop2012-cd}; here I include the factor of two.  Then,
 
@@
 close to previous estimates (Supplementary Figure
-\ref{suppfig:linked-sel-params}). For $J$, I use an empirical estimate
+\ref{figsupp:linked-sel-params}). For $J$, I use an empirical estimate
 calculated from the genome-wide average of the rate of coalescent events due to
@@
 \emph{Drosophila} sweep parameters I have used are close to previous estimates
-(Supplementary Figures \ref{suppfig:linked-sel-params} A and B).
-
-%Given that \textcite{Elyashiv2016-vt} provides a distribution of binned
-%selection coefficients $g(s|i_S)$ by annotation class $i_S$ (e.g. UTRs, exons),
-%I average over $i_S$, weighting by the number of basepairs in each annotation
-%category, to calculate the marginal distribution of selection coefficients $g(s)$. 
-%Then, I calculate $J$ by marginalizing over $g(s)$,
-
-%\begin{align}
-%  J = \sum_{s} g(s) \int_{0}^{L} \exp(- 2 r \tau(s, N)) dr.
-%\end{align}
-%%
-%Throughout, I use the expression for fixation time,
-
-%\begin{align}
-%  \tau(s, N) = \frac{2}{s} \left(\log(4 N s) + c - \nicefrac{1}{4Ns}\right)
-%\end{align}
-%%
-%from equation A17 of \textcite{Hermisson2005-hs}, re-parameterized so the
-%fitness of the heterozygote is $1+s$, and $c \approx 0.577$ is Euler's gamma.
-
-%% We can take $L \to \infty$ as
-%% in \textcite{Coop2012-cd}, leaving
-
-%% \begin{align}
-%%   S &= \frac{\alpha m}{T} \frac{1}{L \tau(\bar{s}, N)} 
-%% \end{align}
-
+(Supplementary Figures \ref{figsupp:linked-sel-params} A and B).
+
+\end{appendixbox}
+
+\begin{appendixbox}
 \section{Background Selection and Polygenic Fitness Models}
@@
 
-Under the \textcite{Santiago1998-bs} model, the effective population size is
+Under the \cite{Santiago1998-bs} model, the effective population size is
 $N_e^\text{SC98} = N \exp(\nicefrac{-C^2}{(1-Z)L})$, where $C^2$ is the
@@
 accommodate a variety of modes of selection such as selection on an
-infinitesimal trait \parencite[p.  1016]{Santiago1995-hx}, and the flux of
-either weakly advantageous or deleterious alleles \parencite[p.
+infinitesimal trait \citep[p.  1016]{Santiago1995-hx}, and the flux of
+either weakly advantageous or deleterious alleles \citep[p.
 2109]{Santiago1998-bs}. If the source of fitness variation is entirely the
@@
 (\cite{Crow1970-wm}; \cite{Caballero2020-wm}, p. 167), and $Z = 1 - sh -
-\nicefrac{1}{2N_c}$ \parencite{Santiago1998-bs}. Thus, if $\nicefrac{1}{2N_c}
+\nicefrac{1}{2N_c}$ \citep{Santiago1998-bs}. Thus, if $\nicefrac{1}{2N_c}
 << sh << 1$, then $\nicefrac{C^2}{(1-Z)} \approx U$ and $N_e^\text{SC98}
@@
 alleles across the genome under mutation-selection balance
-\parencite{Charlesworth2015-am,Charlesworth2000-km}.
-
-\printbibliography
-
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{equation}{0}
-\renewcommand{\theequation}{S\arabic{equation}}
-\setcounter{section}{0}
-\renewcommand{\thesection}{S\arabic{section}}
-
-\section*{Supplementary Information}
-
-\section{Population Size Validation}
-\label{supp:popsize-valid}
-
-I validated the approximate census sizes I have estimated here using a few
-approaches. First, to check broad consistency, I compared the implied biomass
-from my estimates of body mass and census size per species with estimates of
-the total carbon biomass on earth by phylum \parencite{Bar-On2018-kc}. For
-species $i$ with wet body mass $m_i$ and census size $N_i$, the implied biomass
-is $m_i N_i$. For all species in a phylum $S$, this total sample biomass is
-$b_S = \sum_{i \in S} m_i N_i$. I then compare this wet biomass to the carbon
-biomasses by phylum by \textcite{Bar-On2018-kc}. Across animal species, the
-ratio of dry to wet body mass, and carbon body mass dry body mass varies
-little. In their study, Bar-On et al. assume wet body mass has a 70\% water
-content, and 50\% of dry body mass is carbon mass, leading to a wet body mass
-to carbon mass factor of $\nicefrac{1-0.7}{0.5} = 0.15$. I use this factor to
-convert the total wet biomass to carbon biomass per phylum.
-
-\begin{table}[ht]
-  \resizebox{\textwidth}{!}{%
-\centering
-\begin{tabular}{lr|rr|rrrrr|r}
-  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{5}{c}{\bf Present study}  &  \\
-  phylum & total species ($T$) & biomass ($B$) & prop.
-biomass & biomass ($b$) &  prop.
-biomass & num.
-species ($n$) & factor overrepresented &prop.
-total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
-  \hline
-  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.80 \times 10^{-4}$ & 0.0102 &  68 & 0.02 & $5.41 \times 10^{-5}$ & 4.31 \\ 
-  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.67 \times 10^{-2}$ & 0.9715 &  68 & 2.89 & $1.26 \times 10^{-3}$ & 24.40 \\ 
-  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.23 \times 10^{-5}$ & 0.0004 &   3 & 0.01 & $1.76 \times 10^{-4}$ & 0.35 \\ 
-  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.56 \times 10^{-4}$ & 0.0166 &  13 & 0.21 & $1.36 \times 10^{-4}$ & 16.70 \\ 
-  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.07 \times 10^{-5}$ & 0.0011 &   2 & 0.03 & $1.25 \times 10^{-4}$ & 2.45 \\ 
-  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $4.03 \times 10^{-6}$ & 0.0001 &   1 & 0.02 & $4.00 \times 10^{-5}$ & 5.03 \\
-\end{tabular}}
-\caption{How the total carbon biomass estimates by phylum from
-  \textcite{Bar-On2018-kc} compare to the implied biomass estimates from this
-  study. All biomass estimates are carbon biomass, and the proportions are of
-  total biomass with respect to the study. The proportion of biomass in this
-  study compared to the Bar-On et al. estimates \textcite{Bar-On2018-kc}
-  indicates chordates are overrepresented and arthropods are underrepresented
-  in the present study; the factor that each phylum is overrepresented is
-  given in the eighth column. Total species by phylum estimates are from
-  \textcite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
-  ratio column is the ratio of total biomass implied by the $N_c$ estimates
-of each species in a phylum to the actual biomass of that phylum.}
-\label{supp:biomass-tbl}
-\end{table}
-
-First, I compared the relative carbon biomass in this study to the relative
-carbon biomass on earth per phylum. This shows that this study's sample over
-represents chordate biomass (by a factor of $\mathord{\sim}3$), and under
-represents in arthropod biomass (by a factor of $0.02$) relative to the
-proportion of carbon biomass of these phyla on earth (see column eight of
-Supplementary Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
-biomass per phylum in the sample was broadly consistent with the total on earth
-by phylum ($B_S$ for phylum $S$), I calculated the expected sample biomass if
-species were sampled randomly from the total species in a phylum, ($B_S \times
-\nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
-in phylum $S$, $T_S$ is the total number of species in phylum $S$ on earth).
-The fraction of total species on earth included in the sample in this study is
-depicted in Supplementary Figure \ref{suppfig:species_counts}.
-
-Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
-biomass per phylum (Supplementary Table \ref{supp:biomass-tbl}). The
-consistency is quite close for this rough approach and the non-random sample of
-taxa included in this study. The carbon biomass estimates for chordates implied
-by the census size estimates are $\mathord{\sim}24$-fold higher than expected, but is
-well within reasonable expectations given that the chordate sample includes
-many larger-bodied domesticated species (and is a biased sample in other ways).
-Similarly, the implied arthropod carbon biomass is quite close to what one
-would expect. Overall, these values indicate that the census size estimates
-here do not lead to implied biomasses per phylum that are outside the range of
-plausibility.
-
-Additionally, note that the body mass based estimates of density for
-\emph{Drosophila} are similar to previously used estimates in surveys of census
-size and diversity. \textcite{Nei1984-zi} suggested a maximum of 5
-\emph{Drosophila} per $\text{m}^2$, including regions of the range that are not
-inhabitable. Across \emph{Drosophila}, the body mass based estimates suggest
-$10^{6.7}-10^{7.6}$ individuals per $\text{km}^2$, or $4.5-36.3$ individuals
-per $\text{m}^2$, which are consistent with this previous estimate.
-\citeauthor{Nei1984-zi}'s estimates of \emph{Drosophila pseudoobscura}'s census
-size are four orders of magnitude smaller than mine, but their approach uses a
-speculated ratio of population sizes of different \emph{Drosophila} species
-rather than range sizes \parencite[p. 81]{Nei1984-zi}.
+\citep{Charlesworth2015-am,Charlesworth2000-km}.
+
+\end{appendixbox}
+
+\begin{appendixbox}
+
+\section{Additional Population Size Validation}
+\label{app:popsize-valid}
+
+In addition to the biomass-based validation described in the main text, I also
+conducted a few other consistency checks. First, note that the body-mass-based
+estimates of density for \emph{Drosophila} are similar to previously used
+estimates in surveys of census size and diversity. \cite{Nei1984-zi}
+suggested a maximum of 5 \emph{Drosophila} per $\text{m}^2$, including regions
+of the range that are not inhabitable. Across \emph{Drosophila}, the body mass
+based estimates suggest $10^{6.7}-10^{7.6}$ individuals per $\text{km}^2$, or
+$4.5-36.3$ individuals per $\text{m}^2$, which are consistent with this
+previous estimate.  \citeauthor{Nei1984-zi}'s estimates of \emph{Drosophila
+pseudoobscura}'s census size are four orders of magnitude smaller than mine,
+but their approach uses a speculated ratio of population sizes of different
+\emph{Drosophila} species rather than range sizes \citep[p.
+81]{Nei1984-zi}.
 
@@
 study shows that across whale species, pre-whaling carbon biomass was at the
-tens of megatons level \parencite[Table 1 and Figure 1]{Pershing2010-fh}.
+tens of megatons level \citep[Table 1 and Figure 1]{Pershing2010-fh}.
 Given that my census size estimates represent populations at a macroecological
@@
 range from 25 million to 19.6 billion
-\parencite{noauthor_2021-gc,Robinson2014-sx}; the present study's estimate lies
+\citep{noauthor_2021-gc,Robinson2014-sx}; the present study's estimate lies
 in the middle at $\mathord{\sim}175$ million. Again, this is a known limitation
@@
 penguins (\emph{Aptenodytes patagonicus}) is about 3 million; the population
-size was recently estimated as 2.23 million pairs \parencite{Shirihai2008-ft}.
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/diversity_leffler_ranges.pdf}
-
-  \caption{Pairwise diversity grouped by the range categories from
-  \textcite{Leffler2012-zj}, with point size indicating the predicted
-   population density. The vertical lines are the range category group means.}
-
-  \label{suppfig:div_leffler}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/species_counts.pdf}
-  \caption{The fraction of total species on earth included in this study's
-  sample, per class. The color of the points represents phylum, and the size of
-the point represents the absolute number of species by class.}
-  \label{suppfig:species_counts}
-\end{figure}
-
-\section{Sensitivity Analysis}
-
+size was recently estimated as 2.23 million pairs \citep{Shirihai2008-ft}.
+
+\end{appendixbox}
+
+\begin{appendixbox}
 \section{Diversity and IUCN Red List Status}
-\label{supinfo:div-iucn}
+\label{app:div-iucn}
 
@@
 smaller population sizes and reduced diversity, compared to non-threatened
-species (Supplementary Figure \ref{suppfig:figure-1-redlist}) consistent with
-past work \parencite{Spielman2004-mt}. A linear model of diversity regressed on
+species (Supplementary Figure \ref{figsupp:figure-1-redlist}) consistent with
+past work \citep{Spielman2004-mt}. A linear model of diversity regressed on
 population size has lower AIC when the IUCN Red List categories are included,
@@
 
-\begin{figure}[t!]
-  \centering
+\end{appendixbox}
+
+\begin{figure}
   \includegraphics[width=\textwidth]{figures/diversity_popsize_redlist.pdf}
@@
 
-  \label{suppfig:figure-1-redlist}
+\label{figsupp:figure-1-redlist}
 \end{figure}
@@
 \begin{table}[ht]
-  \centering
   \begin{tabular}{rrrr}
@@
 
-\section{Figure Supplement}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{}
-\renewcommand{\figurename}{Figure 1-figure supplement 1}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/damuth.pdf}
-
-  \caption{The appendix table of \textcite{Damuth1987-sg}; the color indicates
-    Damuth's original group labels. The dashed line was estimated using a
-    lognormal regression model in Stan. References to each measurement are
-  available in \textcite{Damuth1987-sg}.}
-  \label{suppfig:damuth}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/diversity_popsize_linear.pdf}
-
-  \caption{A version of Figure \ref{fig:figure-1} with diversity on a linear,
-    rather than log, scale. Points are colored by phylum, and the shaded region
-    is the predicted neutral level of diversity assuming $N_e = N_c$ with mutation
-  range ranging between $10^{-10} \le \mu \le 10^{-8}$. }
-
-  \label{suppfig:figure-1-linear}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/diversity_popsize_averages.pdf}
-
-  \caption{Diversity and approximate population size for 172 taxa, colored by
-    phylum; the dashed lines indicate the non-phylogenetic OLS estimates of the
-  relationship between population size and diversity grouped by phyla.}
-
-  \label{suppfig:figure-1-ave}
-\end{figure}
-
-\begin{figure}[!htb] \centering
-  \includegraphics[width=\textwidth]{figures/diversity_range_bodymass.pdf}
-
-  \caption{The relationship between diversity (differences per basepair) and
-    body mass (left) and range (right) across 172 species. The top row are
-    posterior distributions of parameters estimated using the phylogenetic
-    mixed-effects model using 166 taxa in the synthetic phylogeny for the
-    intercept, slope, and phylogenetic signal from the mixed-effects model. The
-    bottom row contain each species as a point, colored by phyla. The gray
-    dashed line is the non-phylogenetic standard regression estimate, and the
-    blue dashed line is the relationship fit by the phylogenetic mixed-effects
-  model. }
-
-  \label{suppfig:figure-div-range-bodymass}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/genome_size_popsize.pdf}
-
-  \caption{The relationship between genome size and approximate census
-    population size. The dashed gray line indicates the OLS fit. Tiger
-    salamander (\emph{Ambystoma tigrinum}) was excluded because of its
-  exceptionally large genome size (~30Gbp).}
-
-  \label{supfig:genome_size_popsize}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/genome_size_maplength.pdf}
-
-  \caption{The relationship between genome size and recombination map length.
-    The dashed gray line indicates the OLS fit for all taxa, and the dashed
-    colored dashed lines indicate the linear relationship fit by phyla. Tiger
-    salamander (\emph{Ambystoma tigrinum}) was excluded because of its
-  exceptionally large genome size (~30Gbp).}
-
-  \label{supfig:genome_size_maplength}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/iucn_redlist_ranges.pdf}
-
-  \caption{The correspondence between the ranges estimated with the alpha hull
-    method applied to GBIF data used in this paper and IUCN Red List's Extent
-    of Occurrence for the subset of species in both datasets. Note that the
-    IUCN Red List contains predominantly endangered species, which leads to
-    ascertainment bias; still, the high correlation between the estimated
-  ranges shows the alpha hull method works well.}
-
-  \label{suppfig:redlist-ranges}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/range_categories.pdf}
-
-  \caption{The estimated ranges using GBIF occurrence data, ordered within and
-    colored by the original range category labels assigned in
-  \textcite{Leffler2012-zj}.}
-
-  \label{suppfig:range-cats}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/body_mass_length.pdf}
-
-  \caption{The relationship between body length (meters) and body mass (grams)
-    in the \textcite{Romiguier2014-bp} data set, used to infer body masses for
-  taxa. The gray dashed line is the line of best fit inferred using Stan.}
-
-  \label{suppfig:body-mass-length}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/corbett_detig.pdf}
-
-  \caption{(A) The diversity data from \textcite{Corbett-Detig2015-gt} and the
-    census population size estimated here for metazoan taxa. (B) The
-    reductions in diversity, $R = \nicefrac{N_e}{N}$, plotted against census
-    size across species. The red points are the reductions estimated by
-    \textcite{Corbett-Detig2015-gt}. This confirms
-    \citeauthor{Corbett-Detig2015-gt}'s (2015) finding that the impact of
-    selection ($I = 1-R$) increases with census population size (though, in the
-    original paper size body size and range were used as separate proxy
-    variables for census population size). The green and red points are the
-    predicted reduction in diversity under the recurrent hitchhiking (RHH) and
-    background selection (BGS) model using the \emph{Drosophila melanogaster}
-    parameters as described in the main text.  The reduction in the diversity
-    due to sweeps, from Equation \eqref{eqn:div}, is determined by the term
-    $2NS$.  Green points treat $N$ as the implied effective population size
-    from diversity $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, assuming
-    $\mu = 10^{-9}$. Yellow points treat $N$ as the census size, $N = N_c$.
-    Overall, using the census size, e.g.  $2N_cS$, leads to reductions in
-    diversity that far exceed the empirical estimates of Corbett-Detig et al.
-  and reasonable model-based predictions from $\widetilde{N}_e$.}
-
-  \label{suppfig:corbett_detig}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/linked_sel_params.pdf}
-
-  \caption{(A) The estimate of the number of sweeps per basepair, per genome
-    ($\nu_\text{BP}$) from Table 2 of \textcite{Elyashiv2016-vt} (the studies
-    included are \cite{Li2006-xa,Andolfatto2007-uy,Macpherson2007-qt} and
-    \cite{Jensen2008-vz}); the red point is my estimate used in
-    this paper. (B) Points are the data from \textcite{Shapiro2007-gp}.  The
-    blue line is the non-linear least squares fit to the data, and the green
-    dashed line is the sweep model parameterized by the genome-wide average
-    sweep coalescent rate $2NS \approx 0.92$ from the classic sweep and
-  background selection model of \textcite{Elyashiv2016-vt} ($r_s$ in
-Supplementary Table S6).}
-
-  \label{suppfig:linked-sel-params}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/div_popsize_post_phyla.pdf}
-
-  \caption{The posterior distributions for the parameters of the phylogenetic
-    mixed-effects model of diversity and population size (this is analogous to
-    Figure \ref{fig:figure-2}B) fit separately on chordates ($n=68$), molluscs
-    ($n=13$), and arthropods ($n=68$). The phylogenetic mixed-effects model for
-    chordates indicated the best-fitting model had no residual variance
-    ($\sigma_r^2 = 0$), so an alternate model without this variance component
-    was used to ensure proper convergence; this model is shown in green.
-    The light blue (green) shaded regions are the 90\% credible intervals, the
-    blue (green) lines the posterior averages, the gray shaded regions the OLS
-    bootstrap 95\% confidence intervals, and the gray lines the OLS estimate.
-    Note that unlike Figure \ref{fig:figure-2}, the OLS estimate uses all taxa,
-    not just those present in the phylogeny, since splitting the data by phyla
-    reduces sample sizes (OLS with just the subset of taxa in the phylogeny is
-    not significant for either chordates and arthropods). The vertical dashed
-  gray line indicates zero.}
-
-  \label{suppfig:div_popsize_post_phyla}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/biconmap_diversity.pdf}
-
-  \caption{}
-
-  \label{suppfig:biconmap_diversity}
-\end{figure}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/biconmap.pdf}
-
-  \caption{}
-
-  \label{suppfig:biconmap}
-\end{figure}
-
 \end{document}

### 2021-06-04 manuscript/elife/elife-template.tex
even more elife formatting bullshit

@@
 new mutations and their loss by genetic drift
-\citep{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the
-expected diversity at neutral sites in a panmictic population of $N_c$ diploids
-is expected to be $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per
+\citep{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the expected
+diversity at neutral sites in a panmictic population of $N_c$ diploids is
+expected to be $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per
 generation mutation rate. Given that metazoan germline mutation rates only
@@
 levels across a wide range of species varied just an order of magnitude
-\citep[][p.  208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox
-of Variation".  With modern sequencing-based estimates of $\pi$ across taxa
+\citep[][p. 208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox of
+Variation". With modern sequencing-based estimates of $\pi$ across taxa
 ranging over only three orders of magnitude ($0.01$--$10\%$,
@@
      using a lognormal regression model in Stan. References to each measurement
-   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:sf1}
+   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:damuth}
 
@@
 Stan.}{\includegraphics[]{figures/body_mass_length.pdf}}\label{figsupp:body-mass-length}
+
+\figdata{The complete diversity and population size dataset for 172 metazoan taxa.}
 
@@
 previously-observed linear relationship that holds across metazoans (see
-\FIGSUPP[figure-nc]{sf1}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
+\FIGSUPP[figure-nc]{damuth}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
 geographic occurrence data, I estimate range sizes.  Finally, I estimate
@@
 variation, and can account for other modes of linked selection
-(\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}, see Appendix Section
-\ref{app:bgs}). The coalescent rate due to sweeps is $S(\gamma, L, J) =
+(\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}, see Appendix 
+\ref{app:bgs}, \nameref{app:bgs}). The coalescent rate due to sweeps is $S(\gamma, L, J) =
 \nicefrac{\gamma}{L} J$, where $\gamma$ is the number of adaptive substitutions
@@
     interval is the 95\% posterior average. (B) Points indicate the observed
-    $\pi$--$N_c$ relationship across taxa shown in Figure \ref{fig:figure-1},
+    $\pi$--$N_c$ relationship across taxa shown in \FIG{pi-Nc},
     and the blue ribbon is the range of predicted diversity were $N_e = N_c$
@@
 Previous work has found that the impact of linked selection increases with
-$N_c$ (\cite{Corbett-Detig2015-gt}; see also Supplementary Figure
-\ref{figsupp:corbett_detig}A), and it is often thought that this is driven by
-higher rates of adaptive substitutions in larger populations
-\citep{Ohta1992-yi}, despite equivocal evidence \citep{Galtier2016-dq}.
-However, there is another mechanism by which species with larger population
-sizes might experience a greater impact of linked selection: recombinational
-map length, $L$, is known to correlate with body mass \citep{Burt1987-tq} and
-thus varies inversely with population size. As this is a critical parameter
-that determines the genome-wide impact of both hitchhiking and background
-selection, I examine the relationship between recombination map length ($L$)
-and census population size ($N_c$) across taxa, using available estimates of
-map lengths across species \citep{Stapley2017-fs,Corbett-Detig2015-gt}. I find
-a significant non-linear relationship using phylogenetic mixed-effects models
-(Figure \ref{fig:figure-3}A; see Methods and Materials:
-\nameref{sec:methods-pcm}).  There is also a correlation between map length and
-genome size (Supplementary Figure \ref{supfig:genome_size_maplength}) and
-genome size and population size (Supplementary Figure
-\ref{supfig:genome_size_popsize}). These findings are consistent with both the
-hypothesis that non-adaptive processes increase genome size in small-$N_e$
-species \citep{Lynch2003-bn} which in turn could increase map lengths, as well
-as the hypothesis that map lengths are adaptively longer to more efficiently
-select against deleterious alleles \citep{Roze2021-ya}. Overall, the negative
-relationship between map length and census size indicates linked selection is
-expected to be stronger in short map length, high-$N_c$ species.
+$N_c$ (\cite{Corbett-Detig2015-gt}; see \FIGSUPP[linksel]{reductions}A), and it
+is often thought that this is driven by higher rates of adaptive substitutions
+in larger populations \citep{Ohta1992-yi}, despite equivocal evidence
+\citep{Galtier2016-dq}.  However, there is another mechanism by which species
+with larger population sizes might experience a greater impact of linked
+selection: recombinational map length, $L$, is known to correlate with body
+mass \citep{Burt1987-tq} and thus varies inversely with population size. As
+this is a critical parameter that determines the genome-wide impact of both
+hitchhiking and background selection, I examine the relationship between
+recombination map length ($L$) and census population size ($N_c$) across taxa,
+using available estimates of map lengths across species
+\citep{Stapley2017-fs,Corbett-Detig2015-gt}. I find a significant non-linear
+relationship using phylogenetic mixed-effects models ( \FIG{linksel}A; see
+Methods and Materials: \nameref{sec:methods-pcm}).  There is also a correlation
+between map length and genome size (\FIGSUPP[linksel]{genome-size-maplen}) and
+genome size and population size (\FIGSUPP[linksel]{genome-size-Nc}). These
+findings are consistent with both the hypothesis that non-adaptive processes
+increase genome size in small-$N_e$ species \citep{Lynch2003-bn} which in turn
+could increase map lengths, as well as the hypothesis that map lengths are
+adaptively longer to more efficiently select against deleterious alleles
+\citep{Roze2021-ya}.  Overall, the negative relationship between map length and
+census size indicates linked selection is expected to be stronger in short map
+length, high-$N_c$ species.
 
@@
 four-alleles model, \cite{Tajima1996-rb}), and compare this to the observed
-relationship between $\pi$ and $N_c$ in Figured \ref{fig:figure-3}B.  Under
-these parameters, linked selection begins to appreciably depress diversity
-around $N_c \approx 10^9$, since $S \approx 10^{-8}$--$10^{-9}$ and linked
-selection dominates drift when $S > \nicefrac{1}{2N}$.  Overall, this reveals
-two problems for the hypothesis that linked selection could solve Lewontin's
+relationship between $\pi$ and $N_c$ in \FIG{linksel}B.  Under these
+parameters, linked selection begins to appreciably depress diversity around
+$N_c \approx 10^9$, since $S \approx 10^{-8}$--$10^{-9}$ and linked selection
+dominates drift when $S > \nicefrac{1}{2N}$.  Overall, this reveals two
+problems for the hypothesis that linked selection could solve Lewontin's
 Paradox. First, low to mid-$N_c$ species (census sizes between
@@
 species within this range than in \emph{Drosophila}, which is incompatible with
-the rate of adaptive substitutions across species \citep{Galtier2016-dq}
-and mutation rates \citep{Lynch2010-ki}. Furthermore, linked selection has
-been quantified in humans, which fall in this census size range, and has been
-found to be relatively weak
+the rate of adaptive substitutions across species \citep{Galtier2016-dq} and
+mutation rates \citep{Lynch2010-ki}. Furthermore, linked selection has been
+quantified in humans, which fall in this census size range, and has been found
+to be relatively weak
 \citep{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
@@
 available estimates of $\pi_0$ are underestimated by several orders of
-magnitude in \emph{Drosophila} (Supplementary Figured
-\ref{figsupp:corbett_detig}B). The high reductions in $\pi$ predicted here
-(compared to those of \cite{Elyashiv2016-vt}) are a result of using $N_c$,
-rather than $N_e = \nicefrac{\pi_0}{4 \mu}$ in the denominator of Equation
-\eqref{eqn:div}, which leads to a very high rate of sweeps in the population.
-I do not consider selective interference, though the saturation of adaptive
-substitutions per Morgan would only act to limit the reduction in diversity
-\citep{Weissman2012-aa}, and thus these results are conservative.
-Overall, while linked selection could decouple diversity from census size for
-high-$N_c$ species, recurrent hitchhiking and background selection seem
-unlikely to explain the observed patterns of diversity across species under our
+magnitude in \emph{Drosophila} (\FIGSUPP[linksel]{reductions}B). The high
+reductions in $\pi$ predicted here (compared to those of
+\cite{Elyashiv2016-vt}) are a result of using $N_c$, rather than $N_e =
+\nicefrac{\pi_0}{4 \mu}$ in the denominator of Equation \eqref{eqn:div}, which
+leads to a very high rate of sweeps in the population.  I do not consider
+selective interference, though the saturation of adaptive substitutions per
+Morgan would only act to limit the reduction in diversity
+\citep{Weissman2012-aa}, and thus these results are conservative.  Overall,
+while linked selection could decouple diversity from census size for high-$N_c$
+species, recurrent hitchhiking and background selection seem unlikely to
+explain the observed patterns of diversity across species under our
 understanding of the range of parameter estimates.
@@
 reductions are orders of magnitude weaker than those considered here by
-assuming that $N_e = N_c$ (Supplementary Figure \ref{figsupp:corbett_detig}B).
+assuming that $N_e = N_c$ (\FIGSUPP[linksel]{reductions}B).
 Overall, even with more realistic models of linked selection, current models of
@@
 surveys. Of the 172 taxa with diversity estimates, 14 are from
-\cite{Corbett-Detig2015-gt}, 96 are from \cite{Leffler2012-zj}, and 62
-are from \cite{Romiguier2014-bp}. The Corbett-Detig et al. data is
-estimated from four-fold degenerate sites, the Romiguier et al. data is
-synonymous sites, and the Leffler et al. data is estimated predominantly from
-silent, intronic, and non-coding sites. All types of diversity estimates from
-\cite{Leffler2012-zj} were included to maximize the taxa in the study,
-since the variability of diversity across functional categories is much less
-than the diversity across taxa. Multiple diversity estimates per taxa were
-averaged. The total recombination map length data were from both
-\citeauthor{Stapley2017-fs} (2017; 127 taxa), and
-\citeauthor{Corbett-Detig2015-gt} (2015; 9 taxa). Both studies used
-sex-averaged recombination maps estimated with cross-based approaches; in some
-cases errors in the original data were found, documented, and corrected. These
-studies also included genome size estimates used to create Supplementary
-Figures \ref{supfig:genome_size_popsize} and
-\ref{supfig:genome_size_maplength}.
+\cite{Corbett-Detig2015-gt}, 96 are from \cite{Leffler2012-zj}, and 62 are from
+\cite{Romiguier2014-bp}. The Corbett-Detig et al. data is estimated from
+four-fold degenerate sites, the Romiguier et al. data is synonymous sites, and
+the Leffler et al. data is estimated predominantly from silent, intronic, and
+non-coding sites. All types of diversity estimates from \cite{Leffler2012-zj}
+were included to maximize the taxa in the study, since the variability of
+diversity across functional categories is much less than the diversity across
+taxa. Multiple diversity estimates per taxa were averaged. The total
+recombination map length data were from both \citeauthor{Stapley2017-fs} (2017;
+127 taxa), and \citeauthor{Corbett-Detig2015-gt} (2015; 9 taxa). Both studies
+used sex-averaged recombination maps estimated with cross-based approaches; in
+some cases errors in the original data were found, documented, and corrected.
+These studies also included genome size estimates used to create
+\FIGSUPP[linksel]{genome-size-maplen} and \FIGSUPP[linksel]{genome-size-Nc}.
 
@@
 \cite{Damuth1987-sg}, the relationship is linear and homoscedastic (see
-Supplementary Figure \ref{figsupp:damuth}). Using Stan
+\FIGSUPP[figure-nc]{damuth}). Using Stan
 \citep{Stan_Development_Team2020-ea}, I jointly estimated the relationship
@@
 Biodiversity Information Facility \citep{noauthor_2020-ey} using the
-\texttt{rgbif} R package \citep{Chamberlain2014-up,Chamberlain2017-uz}.
-Using the occurrence locations, I inferred whether a species was marine or
+\texttt{rgbif} R package \citep{Chamberlain2014-up,Chamberlain2017-uz}.  Using
+the occurrence locations, I inferred whether a species was marine or
 terrestrial, based on whether the majority of their recorded occurrences
 overlapped a continent using \texttt{rnaturalearth} and the \texttt{sf}
-packages \citep{South2017-db,Pebesma2018-bs}. For each taxon, I estimated
-its range by finding the minimum $\alpha$-shape containing these occurrences.
-The $\alpha$ parameters were set more permissive for marine species since
+packages \citep{South2017-db,Pebesma2018-bs}. For each taxon, I estimated its
+range by finding the minimum $\alpha$-shape containing these occurrences.  The
+$\alpha$ parameters were set more permissive for marine species since
 occurrence data for marine taxa were sparser. Then, I intersected the inferred
@@
 \emph{Drosophila} species ranges are from the Drosophila Speciation Patterns
-website, \citep{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate
-these range estimates, I have compared these to the qualitative range
-descriptions \cite{Leffler2012-zj} (Supplementary Figure
-\ref{figsupp:range-cats}) and compared my $\alpha$-shape method to a subset of
-taxa with range estimates from IUCN Red List
-(\cite{Chamberlain2020-tv,Iucn2020-ap}; Supplementary Figure
-\ref{figsupp:redlist-ranges}). Each census population size
-is then estimated as the product of range and density.
+website, \citep{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate these
+range estimates, I have compared these to the qualitative range descriptions
+\cite{Leffler2012-zj} (\FIGSUPP[figure-nc]{range-cats}) and compared my
+$\alpha$-shape method to a subset of taxa with range estimates from IUCN Red
+List (\cite{Chamberlain2020-tv,Iucn2020-ap};
+\FIGSUPP[figure-nc]{redlist-ranges}). Each census population size is then
+estimated as the product of range and density.
 
@@
 proportion of carbon biomass of these phyla on earth (see column eight of
-Supplementary Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
+Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
 biomass per phylum in the sample was broadly consistent with the total on earth
@@
 The fraction of total species on earth included in the sample in this study is
-depicted in Supplementary Figure \ref{figsupp:species_counts}.
+depicted in \FIGSUPP[figure-nc]{species-counts}.
 
 Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
-biomass per phylum (Supplementary Table \ref{supp:biomass-tbl}). The
+biomass per phylum (Table \ref{supp:biomass-tbl}). The
 consistency is quite close for this rough approach and the non-random sample of
@@
 the chordate subset due to identifiability issues leading to poor MCMC
-convergence (Supplementary Figure \ref{figsupp:div_popsize_post_phyla}).
+convergence (\FIGSUPP[pcm]{div-popsize-post-phyla}).
 
@@
 the model are available in the GitHub repository
-(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (\FIG{pcm}A and
-Supplementary Information Figure \ref{figsupp:biconmap}) were created using
+(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (\FIG{pcm}A,
+\FIGSUPP[pcm]{anc-maplen}, and \FIGSUPP[pcm]{anc-species}) were created using
 \texttt{phytools} \citep{Revell2012-zq}. Node-height tests were implemented
@@
 melanogaster}, and the recombination map length estimates from
-\cite{Stapley2017-fs} and \cite{Corbett-Detig2015-gt}. The mathematical
-details of the simplified sweep model are explained in the Appendix Section
+\cite{Stapley2017-fs} and \cite{Corbett-Detig2015-gt}. The mathematical details
+of the simplified sweep model are explained in the Appendix Section
 \ref{app:sweep}. I use estimates of the number of substitutions, $m$, in genic
 regions between \emph{D.  melanogaster} and \emph{D. simulans} from
-\cite{Hu2013-pg}.  Following \cite{Elyashiv2016-vt}, only substitutions
-in UTRs and exons are included, since they found no evidence of sweeps in
-introns. Then, I average over annotation classes to estimate the mean
-proportion of substitutions that are beneficial, $\alpha_\text{Dmel}=0.42$,
-which are consistent with the estimates of Elyashiv et al. and estimates from
+\cite{Hu2013-pg}.  Following \cite{Elyashiv2016-vt}, only substitutions in UTRs
+and exons are included, since they found no evidence of sweeps in introns.
+Then, I average over annotation classes to estimate the mean proportion of
+substitutions that are beneficial, $\alpha_\text{Dmel}=0.42$, which are
+consistent with the estimates of Elyashiv et al. and estimates from
 MacDonald--Kreitman test approaches (see \cite{Eyre-Walker2006-li}, Table 1).
@@
 10^6$. These \emph{Drosophila melanogaster} hitchhiking parameter estimates are
-close to other previously-published estimates (Supplementary Figure
-\ref{figsupp:linked-sel-params}). Finally, I use $U_\text{Dmel} = 1.6$, from
-\cite{Elyashiv2016-vt}. With these parameter estimates from \emph{D.
+close to other previously-published estimates
+(\FIGSUPP[linksel]{linked-sel-params}). Finally, I use $U_\text{Dmel} = 1.6$,
+from \cite{Elyashiv2016-vt}. With these parameter estimates from \emph{D.
 melanogaster}, the recombination map lengths across species, and Equation
@@
 function.
+
+% \subsection{Key Resources Table}
+% \begin{table}[bt]
+%   \caption{}
+%   \begin{tabular}{l l l l l}
+%     \toprule
+%     Reagent type (species) or resource & Designation & Source or reference & Identifiers & Additional information \\
+%     \midrule
+%     \bottomrule
+%   \end{tabular}
+% \end{table}
 
@@
 
-\subsection{Citations}
 \bibliography{biblio.bib}
@@
   \includegraphics[width=\textwidth]{figures/diversity_popsize_redlist.pdf}
-
-  \caption{A version of Figure \ref{fig:figure-1} with points colored by their
+  \captionof{figure}{A version of \FIG{pi-Nc} with points colored by their
     IUCN Red List conservation status. Margin boxplots show the diversity and
@@
 Vulnerable, EN is Endangered, and CR is Critically Endangered.}
-
 \label{figsupp:figure-1-redlist}
@@
 
-\begin{table}[ht]
+\begin{table}
   \begin{tabular}{rrrr}
@@
   \end{tabular}
-  \caption{The regression estimates of full IUCN Red List population size model
+  \captionof{table}{The regression estimates of full IUCN Red List population size model
     for diversity, $\log_{10}(\pi) = \beta_0 + \beta_{LC} LC + \beta_{NT} NT +

### 2021-06-04 notebooks/figures/Nc_figure.r

@@ -28,6 +28,9 @@ d <- da_dps %>% filter(species != 'Equus ferus przewalskii') %>%
       filter(n > 5) %>%
       arrange(phyla_mean, log10_popsize)
 
+d %>% select(species, phylum, log10_popsize) %>%
+  write_tsv(file='popsize_histogram.tsv')
+
 d %>% ggplot() + geom_histogram(aes(log10_popsize, fill=phylum), binwidth=0.7, alpha=0.5)
 
 

### 2021-06-04 notebooks/figures/diversity_popsize_full.r

@@ -21,6 +21,9 @@ phi <- mean(rstan:::extract(dps_fit, pars='phi')$phi)
 beta <- mean(rstan:::extract(dps_fit, pars='beta')$beta)
 
 # alias this -- less typing
+
+da_dps %>% select(species, phylum, log10_popsize, log10_diversity) %>%
+  write_tsv(file='diversity_popsize.tsv')
 d <- da_dps %>% filter(species != 'Equus ferus przewalskii')
 
 #d = read_tsv('../../data/combined_data.tsv')

### 2021-06-04 notebooks/figures/figure_3.r

@@ -44,6 +44,10 @@ pred  <- colMeans(mu_ln_fit)
 social <- da_ml$social
 
 da_ml <- da_ml %>% filter(phylum != 'Nematoda')
+
+da_ml %>% select(species, phylum, log10_popsize, map_length, social) %>%
+  write_tsv(file='popsize_maplen.tsv')
+
 #filter(species != 'Equus ferus przewalskii') %>%
 
 plot(map_length ~ log10_popsize, da_ml, type='n', 
@@ -151,6 +155,10 @@ dml <- dml_full %>%
 # ggplot(dml, aes(log10_popsize, log10(Ne_N_RHH_BGS))) + 
 #   geom_point()  +  
 #   geom_hline(yintercept=0)
+dml %>% select(species, phylum, log10_popsize, log10_diversity, map_length,
+                 pi_RHH_BGS_1em8,
+                 pi_RHH_BGS_1em9) %>%
+  write_tsv(file='diversity_popsize_linkedsel.tsv')
 
 dml %>% filter(species == "Drosophila melanogaster") %>% 
   select(Ne_N_HK95, Ne_N_RHH_BGS, pi_RHH_BGS_1em8, pi_RHH_BGS_1em9, 

### 2021-06-04 manuscript/elife/elife-template.tex
more formatting crap

@@
 
-\figdata{The complete diversity and population size dataset for 172 metazoan taxa.}
+\figdata{The population size estimates for 172 metazoan taxa.}
 
@@
   \end{fullwidth}
+\figdata{The diversity and population size dataset for 172 metazoan taxa.}
 \end{figure}
@@
   \end{fullwidth}
+\figdata{The map length, population size, and linked selection estimates for 136 metazoan taxa.}
 \end{figure}

### 2021-06-04 notebooks/figures/Nc_figure.r

@@ -28,7 +28,7 @@ d <- da_dps %>% filter(species != 'Equus ferus przewalskii') %>%
       filter(n > 5) %>%
       arrange(phyla_mean, log10_popsize)
 
-d %>% select(species, phylum, log10_popsize) %>%
+da_dps %>% select(species, phylum, log10_popsize) %>%
   write_tsv(file='popsize_histogram.tsv')
 
 d %>% ggplot() + geom_histogram(aes(log10_popsize, fill=phylum), binwidth=0.7, alpha=0.5)

### 2021-06-04 notebooks/figures/figure_3.r

@@ -155,7 +155,7 @@ dml <- dml_full %>%
 # ggplot(dml, aes(log10_popsize, log10(Ne_N_RHH_BGS))) + 
 #   geom_point()  +  
 #   geom_hline(yintercept=0)
-dml %>% select(species, phylum, log10_popsize, log10_diversity, map_length,
+dml %>% select(species, phylum, log10_popsize, map_length,
                  pi_RHH_BGS_1em8,
                  pi_RHH_BGS_1em9) %>%
   write_tsv(file='diversity_popsize_linkedsel.tsv')

### 2021-06-05 manuscript/elife/elife-manuscript.tex
more elife formatting, fig data

--- 2021-06-04 manuscript/elife/elife-template.tex
@@
 \begin{figure}
-  \includegraphics[width=0.7\linewidth]{figures/Nc_figure.pdf}
+  \includegraphics[width=0.7\linewidth]{Nc_figure.pdf}
   \caption{The distribution of approximate census population sizes estimated by this
@@
      using a lognormal regression model in Stan. References to each measurement
-   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:damuth}
+   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{damuth.pdf}}\label{figsupp:damuth}
 
@@
 the point represents the absolute number of species by
-class.}{\includegraphics{figures/species_counts.pdf}}\label{figsupp:species-counts}
+class.}{\includegraphics{species_counts.pdf}}\label{figsupp:species-counts}
 
@@
   ascertainment bias; still, the high correlation between the estimated
-ranges shows the alpha hull method works well.}{\includegraphics[]{figures/iucn_redlist_ranges.pdf}}\label{figsupp:redlist-ranges}
+ranges shows the alpha hull method works well.}{\includegraphics[]{iucn_redlist_ranges.pdf}}\label{figsupp:redlist-ranges}
 
@@
   colored by the original range category labels assigned in
-\cite{Leffler2012-zj}.}{\includegraphics[]{figures/range_categories.pdf}}\label{figsupp:range-cats}
+\cite{Leffler2012-zj}.}{\includegraphics[]{range_categories.pdf}}\label{figsupp:range-cats}
 
@@
   line of best fit inferred using
-Stan.}{\includegraphics[]{figures/body_mass_length.pdf}}\label{figsupp:body-mass-length}
+Stan.}{\includegraphics[]{body_mass_length.pdf}}\label{figsupp:body-mass-length}
 
@@
   \begin{fullwidth}
-  \includegraphics[width=\linewidth]{figures/diversity_popsize_full.pdf}
+  \includegraphics[width=\linewidth]{diversity_popsize_full.pdf}
 
@@
       assuming $N_e = N_c$ with mutation range ranging between $10^{-10} \le
-    \mu \le 10^{-8}$.}{\includegraphics[width=\textwidth]{figures/diversity_popsize_linear.pdf}}\label{figsupp:popsize-linear}
+    \mu \le 10^{-8}$.}{\includegraphics[width=\textwidth]{diversity_popsize_linear.pdf}}\label{figsupp:popsize-linear}
 
@@
   relationship between population size and diversity grouped by
-phyla.}{\includegraphics[width=\textwidth]{figures/diversity_popsize_averages.pdf}}\label{figsupp:popsize-ave}
+phyla.}{\includegraphics[width=\textwidth]{diversity_popsize_averages.pdf}}\label{figsupp:popsize-ave}
 
@@
     the phylogenetic mixed-effects
-  model.}{\includegraphics[width=\textwidth]{figures/diversity_range_bodymass.pdf}}\label{figsupp:div-range-bodymass}
+  model.}{\includegraphics[width=\textwidth]{diversity_range_bodymass.pdf}}\label{figsupp:div-range-bodymass}
 
@@
     density. The vertical lines are the range category group
-  means.}{\includegraphics{figures/diversity_leffler_ranges.pdf}}\label{figsupp:div-leffler}
+  means.}{\includegraphics{diversity_leffler_ranges.pdf}}\label{figsupp:div-leffler}
   \end{fullwidth}
@@
   \begin{fullwidth}
-  \includegraphics[width=\linewidth]{figures/diversity_pcm.pdf}
+  \includegraphics[width=\linewidth]{diversity_pcm.pdf}
 
@@
     not significant for either chordates and arthropods). The vertical dashed
-  gray line indicates zero.}{\includegraphics[width=\textwidth]{figures/div_popsize_post_phyla.pdf}}\label{figsupp:div-popsize-post-phyla}
-
-  \figsupp{The ancestral continuous trait estimates for diversity and population size with species labels.}{\includegraphics[width=\textwidth]{figures/biconmap_diversity.pdf}}\label{figsupp:anc-species}
+  gray line indicates zero.}{\includegraphics[width=\textwidth]{div_popsize_post_phyla.pdf}}\label{figsupp:div-popsize-post-phyla}
+
+  \figsupp{The ancestral continuous trait estimates for diversity and population size with species labels.}{\includegraphics[width=\textwidth]{biconmap_diversity.pdf}}\label{figsupp:anc-species}
 
@@
   length and diversity and population size with species
-labels.}{\includegraphics[width=\textwidth]{figures/biconmap.pdf}}\label{figsupp:anc-maplen}
+labels.}{\includegraphics[width=\textwidth]{biconmap.pdf}}\label{figsupp:anc-maplen}
 
@@
   \begin{fullwidth}
-  \includegraphics[width=\linewidth]{figures/figure_3.pdf}
+  \includegraphics[width=\linewidth]{figure_3.pdf}
 
@@
     salamander (\emph{Ambystoma tigrinum}) was excluded because of its
-  exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{figures/genome_size_popsize.pdf}}\label{figsupp:genome-size-Nc}
+  exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{genome_size_popsize.pdf}}\label{figsupp:genome-size-Nc}
 
@@
     salamander (\emph{Ambystoma tigrinum}) was excluded because of its
-  exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{figures/genome_size_maplength.pdf}}\label{figsupp:genome-size-maplen}
+  exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{genome_size_maplength.pdf}}\label{figsupp:genome-size-maplen}
 
@@
   and reasonable model-based predictions from
-$\widetilde{N}_e$.}{\includegraphics[width=\textwidth]{figures/corbett_detig.pdf}}\label{figsupp:reductions}
+$\widetilde{N}_e$.}{\includegraphics[width=\textwidth]{corbett_detig.pdf}}\label{figsupp:reductions}
 
@@
       background selection model of \cite{Elyashiv2016-vt} ($r_s$ in
-    Supplementary Table S6).}{\includegraphics[width=\textwidth]{figures/linked_sel_params.pdf}}\label{figsupp:linked-sel-params}
+    Supplementary Table S6).}{\includegraphics[width=\textwidth]{linked_sel_params.pdf}}\label{figsupp:linked-sel-params}
 
@@
 
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/diversity_popsize_linear.pdf}
-%   % figure 2-supplement 1
-%   \figsupp{A version of Figure \ref{fig:figure-1} with diversity on a linear,
-%     rather than log, scale. Points are colored by phylum, and the shaded region
-%     is the predicted neutral level of diversity assuming $N_e = N_c$ with mutation
-%   range ranging between $10^{-10} \le \mu \le 10^{-8}$. }
-
-%   \label{suppfig:figure-1-linear}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/diversity_popsize_averages.pdf}
-%   % figure 2-supplement 2
-%   \figsupp{Diversity and approximate population size for 172 taxa, colored by
-%     phylum; the dashed lines indicate the non-phylogenetic OLS estimates of the
-%   relationship between population size and diversity grouped by phyla.}
-
-%   \label{suppfig:figure-1-ave}
-% \end{figure}
-
-% \begin{figure}[!htb] \centering
-%   \includegraphics[width=\textwidth]{figures/diversity_range_bodymass.pdf}
-%   % figure 2-supplement 3
-%   \figsupp{The relationship between diversity (differences per basepair) and
-%     body mass (left) and range (right) across 172 species. The top row are
-%     posterior distributions of parameters estimated using the phylogenetic
-%     mixed-effects model using 166 taxa in the synthetic phylogeny for the
-%     intercept, slope, and phylogenetic signal from the mixed-effects model. The
-%     bottom row contain each species as a point, colored by phyla. The gray
-%     dashed line is the non-phylogenetic standard regression estimate, and the
-%     blue dashed line is the relationship fit by the phylogenetic mixed-effects
-%   model. }
-
-%   \label{suppfig:figure-div-range-bodymass}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/div_popsize_post_phyla.pdf}
-%   % figure 3-supplement 1
-%   \figsupp{The posterior distributions for the parameters of the phylogenetic
-%     mixed-effects model of diversity and population size (this is analogous to
-%     Figure \ref{fig:figure-2}B) fit separately on chordates ($n=68$), molluscs
-%     ($n=13$), and arthropods ($n=68$). The phylogenetic mixed-effects model for
-%     chordates indicated the best-fitting model had no residual variance
-%     ($\sigma_r^2 = 0$), so an alternate model without this variance component
-%     was used to ensure proper convergence; this model is shown in green.
-%     The light blue (green) shaded regions are the 90\% credible intervals, the
-%     blue (green) lines the posterior averages, the gray shaded regions the OLS
-%     bootstrap 95\% confidence intervals, and the gray lines the OLS estimate.
-%     Note that unlike Figure \ref{fig:figure-2}, the OLS estimate uses all taxa,
-%     not just those present in the phylogeny, since splitting the data by phyla
-%     reduces sample sizes (OLS with just the subset of taxa in the phylogeny is
-%     not significant for either chordates and arthropods). The vertical dashed
-%   gray line indicates zero.}
-
-%   \label{suppfig:div_popsize_post_phyla}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/biconmap_diversity.pdf}
-%   % figure 3-supplement 2
-%   \figsupp{}
-%   \label{suppfig:biconmap_diversity}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/biconmap.pdf}
-%   % figure 3-supplement 3
-%   \figsupp{}
-%   \label{suppfig:biconmap}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/genome_size_popsize.pdf}
-%   % figure 4-supplement 1
-%   \figsupp{The relationship between genome size and approximate census
-%     population size. The dashed gray line indicates the OLS fit. Tiger
-%     salamander (\emph{Ambystoma tigrinum}) was excluded because of its
-%   exceptionally large genome size (~30Gbp).}
-
-%   \label{supfig:genome_size_popsize}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/genome_size_maplength.pdf}
-%   % figure 4-supplement 2
-%   \figsupp{The relationship between genome size and recombination map length.
-%     The dashed gray line indicates the OLS fit for all taxa, and the dashed
-%     colored dashed lines indicate the linear relationship fit by phyla. Tiger
-%     salamander (\emph{Ambystoma tigrinum}) was excluded because of its
-%   exceptionally large genome size (~30Gbp).}
-%   \label{supfig:genome_size_maplength}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/corbett_detig.pdf}
-%   % figure 4-supplement 1
-%   \figsupp{(A) The diversity data from \textcite{Corbett-Detig2015-gt} and the
-%     census population size estimated here for metazoan taxa. (B) The
-%     reductions in diversity, $R = \nicefrac{N_e}{N}$, plotted against census
-%     size across species. The red points are the reductions estimated by
-%     \textcite{Corbett-Detig2015-gt}. This confirms
-%     \citeauthor{Corbett-Detig2015-gt}'s (2015) finding that the impact of
-%     selection ($I = 1-R$) increases with census population size (though, in the
-%     original paper size body size and range were used as separate proxy
-%     variables for census population size). The green and red points are the
-%     predicted reduction in diversity under the recurrent hitchhiking (RHH) and
-%     background selection (BGS) model using the \emph{Drosophila melanogaster}
-%     parameters as described in the main text.  The reduction in the diversity
-%     due to sweeps, from Equation \eqref{eqn:div}, is determined by the term
-%     $2NS$.  Green points treat $N$ as the implied effective population size
-%     from diversity $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, assuming
-%     $\mu = 10^{-9}$. Yellow points treat $N$ as the census size, $N = N_c$.
-%     Overall, using the census size, e.g.  $2N_cS$, leads to reductions in
-%     diversity that far exceed the empirical estimates of Corbett-Detig et al.
-%   and reasonable model-based predictions from $\widetilde{N}_e$.}
-%   \label{suppfig:corbett_detig}
-% \end{figure}
-
-% \begin{figure}[!htb]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/linked_sel_params.pdf}
-%   % figure 4-supplement 2
-%   \caption{(A) The estimate of the number of sweeps per basepair, per genome
-%     ($\nu_\text{BP}$) from Table 2 of \textcite{Elyashiv2016-vt} (the studies
-%     included are \cite{Li2006-xa,Andolfatto2007-uy,Macpherson2007-qt} and
-%     \cite{Jensen2008-vz}); the red point is my estimate used in
-%     this paper. (B) Points are the data from \textcite{Shapiro2007-gp}.  The
-%     blue line is the non-linear least squares fit to the data, and the green
-%     dashed line is the sweep model parameterized by the genome-wide average
-%     sweep coalescent rate $2NS \approx 0.92$ from the classic sweep and
-%   background selection model of \textcite{Elyashiv2016-vt} ($r_s$ in
-% Supplementary Table S6).}
-
-%   \label{suppfig:linked-sel-params}
-% \end{figure}
-
 \bibliography{biblio.bib}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{figures/diversity_popsize_redlist.pdf}
+  \includegraphics[width=\textwidth]{diversity_popsize_redlist.pdf}
   \captionof{figure}{A version of \FIG{pi-Nc} with points colored by their

### 2021-06-05 notebooks/figures/species_ranges.r

@@ -49,7 +49,7 @@ with(means, segments(y0=xmin, y1=xmax,
 text(y=d$xpos, x=8.8, labels=d$species, 
      col=cols[d$range_cat], xpd=TRUE, 
      las=1, srt=0, adj=0, cex=0.3)
-legend(1.9, 126, range_cat_levels, fill=cols, 
+legend(2.1, 80, range_cat_levels, fill=cols, 
        bty='n', box.lwd=0, border=0, cex=1)
 
 title(xlab=latex2exp::TeX('estimated range ($km^2$)'), line=2.1)

### 2021-06-06 manuscript/manuscript.tex
minor title fix

@@
 
-\title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
-%\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}
+% \title{Natural Selection is Unlikely to Explain Why Species Get a Thin Slice of $\pi$}
+\title{Why do species get a thin slice of $\pi$? Revisiting Lewontin's Paradox of Variation}
 
@@
 \maketitle
-
-% TODO recheck node height and other p-values
 
@@
 using a previously-observed linear relationship that holds across metazoans
-(Figure 1-figure supplement 1 \ref{suppfig:damuth};
+(see Supplementary Figure \ref{suppfig:damuth};
 \cite{Damuth1981-st,Damuth1987-sg}). Then, from geographic occurrence data, I
@@
 This is in part because of the log-log scale, in contrast to previous work
-\parencite{Nei1984-zi,Soule1976-he}; see Figure 1-supplement figure 2 for a
-version on a log-linear scale. However, it is noteworthy that few taxa have
-diversity estimates below $10^{-3.5}$ differences per basepair. Those that do,
-lynx (\emph{Lynx Lynx}), wolverine (\emph{Gulo gulo}), and Massasauga
-rattlesnake (\emph{Sistrurus catenatus}) face habitat fragmentation and
-declining population sizes. These three species are all in the IUCN Red List,
-but are listed as least concern (though their presence in the Red List
-indicates they are of conservation interest). In Supplementary Information
-Section \Newnameref{supinfo:div-iucn}, I explore more about the relationships
-between IUCN Red List status, diversity, and population size.
+\parencite{Nei1984-zi,Soule1976-he}; see Supplementary Figure
+\ref{suppfig:figure-1-linear} for a version on a log-linear scale. However, it
+is noteworthy that few taxa have diversity estimates below $10^{-3.5}$
+differences per basepair. Those that do, lynx (\emph{Lynx Lynx}), wolverine
+(\emph{Gulo gulo}), and Massasauga rattlesnake (\emph{Sistrurus catenatus})
+face habitat fragmentation and declining population sizes. These three species
+are all in the IUCN Red List, but are listed as least concern (though their
+presence in the Red List indicates they are of conservation interest). In
+Supplementary Information Section \Newnameref{supinfo:div-iucn}, I explore more
+about the relationships between IUCN Red List status, diversity, and population
+size.
 
@@
 
-\section{Figure Supplement}
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{}
-\renewcommand{\figurename}{Figure 1-figure supplement 1}
-
 \begin{figure}[!htb]

### 2021-08-04 manuscript/elife/elife-manuscript.tex
minor title fix
edits

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to
-Explain the Narrow Range of Diversity Among Species}
+% \title{Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to Explain the Narrow Range of Diversity Among Species}
+\title{Quantifying the Relationship Between Genetic Diversity and Population Size Suggests Natural Selection Cannot Explain Lewontin's Paradox}
 
@@
  population sizes across species, I combine previously-published estimates of
- pairwise diversity from 172 metazoan taxa with estimates of census sizes.
- Using phylogenetic comparative methods, I show this relationship is
- significant accounting for phylogeny, but with high phylogenetic signal and
- evidence that some lineages experience shifts in the evolutionary rate of
- diversity deep in the past. Additionally, I find a negative relationship
- between recombination map length and census size, suggesting abundant species
- have less recombination and experience greater reductions in diversity due to
- linked selection. However, I show that even using strong selection parameter
- estimates, models of linked selection are unlikely to explain the observed
- relationship between diversity and census sizes across species.
+ pairwise diversity from 172 metazoan taxa with newly derived estimates of
+ census sizes.  Using phylogenetic comparative methods, I show this
+ relationship is significant accounting for phylogeny, but with high
+ phylogenetic signal and evidence that some lineages experience shifts in the
+ evolutionary rate of diversity deep in the past. Additionally, I find a
+ negative relationship between recombination map length and census size,
+ suggesting abundant species have less recombination and experience greater
+ reductions in diversity due to linked selection. However, I show that even
+ assuming strong and abundant selection, models of linked selection are
+ unlikely to explain the observed relationship between diversity and census
+ sizes across species.
+
 \end{abstract}
@@
 A longstanding mystery in evolutionary genetics is that the observed levels of
-genetic variation across sexual species are confined to an unexpectedly narrow
-range. Under neutral theory, the average number of nucleotide differences
-between lineages (pairwise diversity, $\pi$) is determined by the balance of
-new mutations and their loss by genetic drift
-\citep{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, the expected
-diversity at neutral sites in a panmictic population of $N_c$ diploids is
-expected to be $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per
-generation mutation rate. Given that metazoan germline mutation rates only
-differ 10-fold ($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}),
-and census sizes vary over several orders of magnitude, one would expected
-under neutral theory that heterozygosity should also vary over several orders
-of magnitude. However, early allozyme surveys revealed that heterozygosity
-levels across a wide range of species varied just an order of magnitude
-\citep[][p. 208]{Lewontin1974-jb}; this is known as Lewontin's ``Paradox of
-Variation". With modern sequencing-based estimates of $\pi$ across taxa
-ranging over only three orders of magnitude ($0.01$--$10\%$,
-\cite{Leffler2012-zj}), Lewontin's paradox has persisted unresolved through the
-genomics era.
-
-From the beginning, explanations for Lewontin's Paradox have been framed in
-terms of the neutralist--selectionist controversy
-\citep{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.
-The neutralist view is that beneficial alleles are sufficiently rare and
-deleterious alleles removed sufficiently quickly, that levels of genetic
+genetic variation across sexual species span an unexpectedly narrow range.
+Under neutral theory, the average number of nucleotide differences between
+sequences (pairwise diversity, $\pi$) is determined by the balance of new
+mutations and their loss by genetic drift
+\citep{Kimura1964-ul,Malecot1948-zv,Wright1931-fl}. In particular, expected
+pairwise diversity at neutral sites in a panmictic population of $N_c$ diploids
+is $\pi \approx 4N_c \mu$, where $\mu$ is the per basepair per generation
+mutation rate. Given that metazoan germline mutation rates only differ 10-fold
+($10^{-8}$--$10^{-9}$, \cite{Kondrashov2010-fi,Lynch2010-ki}), and census sizes
+vary over several orders of magnitude, under neutral theory one would expect
+that pairwise diversity also vary over several orders of magnitude.  However,
+early allozyme surveys revealed that diversity levels across a wide range of
+species varied just an order of magnitude \citep[][p. 208]{Lewontin1974-jb};
+this is known as Lewontin's ``Paradox of Variation".  With modern
+sequencing-based estimates of $\pi$ across taxa ranging over only three orders
+of magnitude ($0.01$--$10\%$, \cite{Leffler2012-zj}), Lewontin's paradox
+remains unresolved through the genomics era.
+
+Early on, explanations for Lewontin's Paradox have been framed in terms of the
+neutralist--selectionist controversy
+\citep{Lewontin1974-jb,Kimura1984-ia,Gillespie1991-qa,Gillespie2001-mv}.  The
+neutralist view is that beneficial alleles are sufficiently rare and
+deleterious alleles are removed sufficiently quickly, that levels of genetic
 diversity are shaped predominantly by genetic drift and mutation
-\citep{Kimura1984-ia}. Specifically, \emph{non-selective} processes
-decouple the effective population size implied by observed levels of diversity
+\citep{Kimura1984-ia}. Specifically, \emph{non-selective} processes decouple
+the effective population size implied by observed levels of diversity
 $\widehat{\pi}$, $\widetilde{N}_e = \nicefrac{\widehat{\pi}}{4\mu}$, from the
-census size, $N_c$. By contrast, the selectionist view is that the direct and
-indirect effects of linked selection suppress diversity levels across taxa,
-specifically because the impact of linked selection is greater in large
-populations. Undoubtedly, these opposing views represent a false dichotomy, as
-population genomic studies have uncovered both complex demographic histories
-that impact diversity within a species (e.g.
-\cite{Zhao2013-vd,Palkopoulou2015-bg}) and evidence that selection depresses
-genome-wide diversity (e.g.
+census size, $N_c$. By contrast, the selectionist view is that direct selection
+and the indirect effects of selection on linked neutral diversity suppress
+diversity levels across taxa, specifically because the impact of linked
+selection is greater in large populations. Undoubtedly, these opposing views
+represent a false dichotomy, as population genomic studies have uncovered
+evidence for the substantial impact of both demographic history (e.g.
+\cite{Zhao2013-vd,Palkopoulou2015-bg}) and linked selection on genome-wide
+diversity (e.g.
 \cite{Elyashiv2016-vt,Begun1992-ey,Aguade1989-jx,McVicker2009-ax}).
 
-\subsection{Possible Explanations of Lewontin's Paradox}
+\subsection{Possible Resolutions of Lewontin's Paradox}
 
@@
 broader literature on the empirical relationship between diversity and
-population size
-\citep{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy}, and the
-ecological and life history correlates of genetic diversity
-\citep{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of
-processes stand out as potentially capable of decoupling census sizes from
-diversity: non-equilibrium demography, variance and skew in reproductive
-success, and selective processes.
+population size \citep{Frankham1996-yb,Nei1984-zi,Soule1976-he,Leroy2021-gy},
+and the ecological and life history correlates of genetic diversity
+\citep{Nevo1978-wh,Powell1975-lg,Nevo1984-hp}. Three categories of processes
+stand out as potentially capable of decoupling census sizes from diversity:
+non-equilibrium demography, variance and skew in reproductive success, and
+selective processes.
 
@@
 
-While chance demographic events like bottlenecks and recent expansions have
-long-term impacts on diversity (since mutation-drift equilibrium is reached on
-the order of size of the population), characteristics of the breeding structure
-such as high variance ($V_w$) or skew in reproductive success continuously
-suppress diversity below the levels predicted by the census size
-\citep{Wright1938-tv}. For example, in many marine animals, females are
-highly fecund, and dispersing larvae face extremely low survivorship, leading
-to high variance in reproductive success
+While intermittent demographic events like bottlenecks and recent expansions
+have long-term impacts on diversity (since mutation-drift equilibrium is
+reached on the order of size of the population), characteristics of the
+breeding structure such as high variance ($V_w$) or skew in reproductive
+success continuously suppress diversity below the levels predicted by the
+census size \citep{Wright1938-tv}. For example, in many marine animals, females
+are highly fecund, and dispersing larvae face extremely low survivorship,
+leading to high variance in reproductive success
 \citep{Waples2018-kb,Waples2013-wi,Hedgecock2011-ku,Hauser2008-fd}. Such
 ``sweepstakes" reproductive systems can lead to remarkably small ratios of
-effective population size to census population size (e.g. $\nicefrac{N_e}{N_c}$
-can range from $10^{-6}$--$10^{-2}$), since $\nicefrac{N_e}{N} \approx
-\nicefrac{1}{V_w}$
-\citep{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and
-require multiple-merger coalescent processes to describe their genealogies
+effective to census population size (e.g. $\nicefrac{N_e}{N_c}$ can range from
+$10^{-6}$--$10^{-2}$), since $\nicefrac{N_e}{N} \approx \nicefrac{1}{V_w}$
+\citep{Hedgecock1994-gs,Wright1938-tv,Nunney1993-ef,Nunney1996-wy}, and require
+multiple-merger coalescent processes to describe their genealogies
 \citep{Eldon2006-ui}. Overall, these reproductive systems diminish the
-diversity in many species, but seem unlikely to explain Lewontin's Paradox
+diversity in some species, but seem unlikely to explain Lewontin's Paradox
 broadly across metazoans.
@@
 Alternatively, selective processes, and in particular the indirect effects of
-selection on linked neutral variation, could explain the observed narrow range
-of diversity. The earliest mathematical model of hitchhiking was proffered as a
-explanation of Lewontin's Paradox \citep{Maynard_Smith1974-zr}.  Since,
-empirical observations have demonstrated that linked selection shapes patterns
-of genome-wide diversity, as evidenced by the correlation between recombination
-and diversity in a variety of species
+selection on linked neutral variation, could potentially explain the observed
+narrow range of diversity. The earliest mathematical model of hitchhiking was
+proffered as a explanation of Lewontin's Paradox \citep{Maynard_Smith1974-zr}.
+Since, linked selection has been shown to impact diversity levels in a variety
+of species, as evidenced by the correlation between recombination and diversity
 \citep{Aguade1989-jx,Begun1992-ey,Cutter2003-tl,Stephan1998-hh,Cai2009-by}.
-Theoretic work to explain this pattern has considered diversity under a steady
-influx of new beneficial mutations (recurrent hitchhiking;
-\cite{Stephan1992-jc,Stephan1995-ry}), and purifying selection against new
+Theoretic work to explain this pattern has considered the impact of a steady
+influx of beneficial mutations (recurrent hitchhiking;
+\cite{Stephan1992-jc,Stephan1995-ry}), and purifying selection against
 deleterious mutations (background selection, BGS;
@@
 around genic regions in a variety of species
-\citep{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now
-efforts have shifted towards teasing apart the effects of positive and negative
+\citep{McVicker2009-ax,Hernandez2011-gs,Charlesworth1996-px}, and now efforts
+have shifted towards teasing apart the effects of positive and negative
 selection on genomic diversity \citep{Elyashiv2016-vt}.
 
-An important class of theoretic selection models pertaining to Lewontin's
+A class of models that are of particular interest in the context of Lewontin's
 Paradox are recurrent hitchhiking models that decouple diversity from the
-census population size. These models predict diversity when strongly selected
-beneficial mutations regularly enter and sweep through the population, trapping
-lineages and forcing them to coalesce
+census population size. These models predict diversity levels when strongly
+selected beneficial mutations regularly enter and sweep through the population,
+trapping lineages and forcing them to coalesce
 \citep{Kaplan1989-sc,Gillespie2000-mh}. In general, decoupling occurs under
@@
 greater than the rate of neutral coalescence \citep[e.g.][equation
-22]{Coop2012-cd}. Under other linked selection models, the resulting effective
-population size is proportional to population size, and thus these models
-cannot decouple diversity, all else equal. For example, models of background
-selection and polygenic fitness variation predict diversity is proportional to
-population size, mediated by the total recombination map length and the
-deleterious mutation rate or fitness variation
+22]{Coop2012-cd}. In contrast, under other linked selection models, the
+resulting effective population size is proportional to population size; these
+models cannot decouple diversity, all else equal. For example, models of
+background selection and polygenic fitness variation predict diversity is
+proportional to population size, mediated by the total recombination map length
+and the deleterious mutation rate or fitness variation
 \citep{Charlesworth1993-gb,Nicolaisen2012-vs,Nordborg1996-nq,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}.
 
-\subsection{Recent Approaches Towards Solving Lewontin's Paradox}
-
-Recently, \cite{Corbett-Detig2015-gt} used population genomic data to
-estimate the reduction in diversity due to background selection and hitchhiking
-across 40 species, and showed the impact of selection increases with two
-proxies of census population size, species range and with body size. They
-argued this is evidence that selection could explain Lewontin's Paradox;
-however, in a re-analysis, \cite{Coop2016-gx} demonstrated that the
-observed scale of these reductions is insufficient to explain the
-orders-of-magnitude shortfall between observed and expected levels of diversity
-across species. Other recent work has found that certain life history
-characteristics related to parental investment, such as propagule size, are
-good predictors diversity in animals \citep{Romiguier2014-bp,Chen2017-nf}.
-Nevertheless, while these diversity correlates are important clues, they do not
-propose a mechanism by which these traits act to constrain diversity within a
-few orders of magnitude.
-
-Here, I revisit Lewontin's Paradox by integrating a variety of data sets and
-assessing the predicted reductions in diversity under different selection
-models. Prior surveys of genetic diversity either lacked census population size
-estimates, used allozyme-based measures of heterozygosity, or included fewer
-species. To address these shortcomings, I first estimate census sizes by
-combining predictions of population density based on body size with ranges
-estimated from geographic occurrence data. Using these estimates, I quantify
-the relationship between census size and previously-published genomic diversity
-estimates across 172 metazoan taxa within nine phyla, to provide a sense of the
-scale of the divergence between $\pi$ and $N_c$ that leads to Lewontin's
-Paradox.
+\subsection{Recent Approaches Towards Resolving Lewontin's Paradox}
+
+Recently, \cite{Corbett-Detig2015-gt} used population genomic data to estimate
+the reduction in diversity due to background selection and hitchhiking across
+40 species, and showed that the impact of selection increases with two proxies
+of census population size, species range and with body size. Based on this
+evidence, they argued that selection could explain Lewontin's Paradox; however,
+in a re-analysis, \cite{Coop2016-gx} demonstrated that the observed magnitude
+of these reductions is insufficient to explain the orders-of-magnitude
+shortfall between observed and expected levels of diversity across species.
+Other recent work has found that life history characteristics related to
+parental investment, such as propagule size, are good predictors diversity in
+animals \citep{Romiguier2014-bp,Chen2017-nf}.  Nevertheless, while these
+diversity correlates are important clues, they do not propose a mechanism by
+which these traits act to constrain diversity within a few orders of magnitude.
+
+Here, I revisit Lewontin's Paradox by integrating several data sets in order to
+compare the observed reductions between diversity and census size with the
+predicted relationship under different selection models. Prior surveys of
+genetic diversity either lacked census population size estimates, used
+allozyme-based measures of heterozygosity, or included fewer species. To
+address these shortcomings, I first estimate census sizes by combining
+predictions of population density based on body size with ranges estimated from
+geographic occurrence data. Using these estimates, I quantify the relationship
+between census size and previously-published genomic diversity estimates across
+172 metazoan taxa within nine phyla, thus characterizing the relationship
+between $\pi$ and $N_c$ that underlies Lewontin's Paradox.
 
@@
 to fully account for phylogenetic non-independence across taxa
-\citep{Felsenstein1985-an}. To address this shortcoming, I use phylogenetic
-comparative methods (PCMs) with a synthetic time-calibrated phylogeny to
-account for shared phylogenetic history. Moreover, it is disputed whether
-considering phylogenetic non-independence is necessary in population genetics,
-since coalescent times are much less than divergence times
-\citep{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by
-estimating the degree of phylogenetic signal in the diversity census size
-relationship, and investigating how these traits evolve along the phylogeny.
+\citep{Felsenstein1985-an}. To address this, I use phylogenetic comparative
+methods (PCMs) with a synthetic time-calibrated phylogeny to account for shared
+phylogenetic history. Moreover, it is disputed whether considering phylogenetic
+non-independence is necessary in population genetics, given that coalescent
+times within species are much less than divergence times
+\citep{Whitney2010-ud,Lynch2011-qv}. Using PCMs, I address this by estimating
+the degree of phylogenetic signal in the diversity census size relationship,
+and investigating how these traits evolve along the phylogeny.
 
@@
 background selection and recurrent hitchhiking are sufficiently strong to
-resolve Lewontin's Paradox. These predicted reductions in diversity across
-species are estimated using strong selection parameters from \emph{Drosophila
-melanogaster}, a species known to be strongly affected by linked selection.
-Given the effects of linked selection are mediated by recombination map length,
-I investigate how recombination map lengths vary with census population size
-using data from a previously-published survey \citep{Stapley2017-fs}. I
-find map lengths are typically shorter in large--census-size species,
-increasing the effects of linked selection in these species, which might
-further decouple diversity from census size. Still, I find the combined impact
-of these selection models with available parameter estimates falls short in
-explaining Lewontin's Paradox, and discuss future avenues through which the
-Paradox of Variation could be fully resolved.
+resolve Lewontin's Paradox. I do so using selection parameters from
+\emph{Drosophila melanogaster}, a species known to be strongly affected by
+linked selection. Given the effects of linked selection are mediated by
+recombination map length, I also investigate how recombination map lengths vary
+with census population size using data from a previously-published survey
+\citep{Stapley2017-fs}. I find map lengths are typically shorter in
+large--census-size species, increasing the effects of linked selection in these
+species, which could further decouple diversity from census size. Still, I find
+the combined impact of these modes of linked selection fall short in explaining
+Lewontin's Paradox, and discuss future avenues through which the Paradox of
+Variation could be fully resolved.
 
@@
 Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Matt Osmond, Sally
-Otto, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli,
-Margot Wood, and my Kern-Ralph labmates for helpful discussions. Sarah
-Friedman, Katherine Corn, and Josef Uyeda provided very useful advice about
-phylogenetic comparative methods; yet I take full responsibility for any
-shortcomings of my analysis. Finally, I am indebted to Guy Sella, Matt Pennell,
-and two other anonymous reviewers for helpful feedback. I would like to also
-thank UO librarian Dean Walton for helping me track down some rather difficult
-to find older papers.  This work was supported by an NIH Grant (1R01GM117241)
-awarded to Andrew Kern.
+Otto, Molly Przeworski, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina,
+Michael Turelli, Margot Wood, and my Kern-Ralph labmates for helpful
+discussions. Sarah Friedman, Katherine Corn, and Josef Uyeda provided very
+useful advice about phylogenetic comparative methods; yet I take full
+responsibility for any shortcomings of my analysis. Finally, I am indebted to
+Guy Sella, Matt Pennell, and two other anonymous reviewers for helpful
+feedback. I would like to also thank UO librarian Dean Walton for helping me
+track down some rather difficult to find older papers.  This work was supported
+by an NIH Grant (1R01GM117241) awarded to Andrew Kern.
 

### 2021-08-08 manuscript/elife/elife-manuscript.tex
updated figure 3 supp

@@
 
-  \figsupp[The fraction of total species on earth included in this study's
+  \figsupp[The fraction of total species per class on earth included in this study's
   sample, per class.]{The color of the points represents phylum, and the size of
@@
 
-\figsupp[Validation of this paper's range estimates procedure against the IUCN Red List's range estimates.]{
+\figsupp[Comparison of this paper's range estimates procedure against the IUCN Red List's range estimates.]{
   The correspondence between the ranges estimated with the alpha hull
@@
   The estimated ranges using GBIF occurrence data, ordered within and
-  colored by the original range category labels assigned in
+  c olored by the original range category labels assigned in
 \cite{Leffler2012-zj}.}{\includegraphics[]{range_categories.pdf}}\label{figsupp:range-cats}
@@
 
-A major impediment in quantifying Lewontin's Paradox has been estimating census
-population sizes across many taxa, especially for extremely abundant,
-cosmopolitan species that define the upper limit of ranges. Previous work has
-surveyed the literature for census size estimates
+An impediment in resolving Lewontin's Paradox is characterizing the
+relationship between diversity and census population sizes. This is difficult
+because census population sizes are unavailable for many taxa, especially for
+extremely abundant, cosmopolitan species that define the upper limit of ranges.
+Previous work has surveyed the literature for census size estimates
 \citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
@@
 between genomic estimates of diversity and census population sizes, I first
-approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). My
-approach predicts population densities from body sizes using a
-previously-observed linear relationship that holds across metazoans (see
+approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). I
+estimate population densities based on an empirical linear relationship between
+body sizes and density that holds across metazoans (see
 \FIGSUPP[figure-nc]{damuth}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
@@
 population size as the product of these predicted densities and range estimates
-(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the relationship between
-population density and body size is driven by energy budgets, and thus reflects
-macroecological equilibria \citep{Damuth1987-sg}. Consequently, population
-sizes are underestimated for taxa like humans and their domesticated species,
-and overestimated for species with anthropogenically reduced densities or
-fragmented ranges. For example, the population size of \emph{Lynx lynx} is
-likely around 50,000 \citep{Iucn2020-ap} which is around two orders of
-magnitude smaller than my estimate. Additionally, the range size estimates do
-not consider whether an area has unsuitable habitat, and thus may be
-overestimated for species with particular niches or patchy habitats. While
-these methods to estimate census size are crude and approximate, they can be
-efficiently calculated for numerous taxa and are sufficient to estimate the
-scale of Lewontin's Paradox (see \nameref{supp:popsize-valid} for more on
-validation based on biomass and other approaches).
-
-\subsection{Quantifying Lewontin's Paradox}
+(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the
+relationship between population density and body size is driven by energy
+budgets, and thus reflects macroecological equilibria \citep{Damuth1987-sg}.
+Consequently, population sizes are underestimated for taxa like humans and
+their domesticated species, and overestimated for species with
+anthropogenically reduced densities or fragmented ranges. For example, the
+population size of \emph{Lynx lynx} is likely around 50,000 \citep{Iucn2020-ap}
+which is around two orders of magnitude smaller than my estimate. Additionally,
+the range size estimates do not consider whether an area has unsuitable
+habitat, and thus may be overestimated for species with particular niches or
+patchy habitats. While my approach produces approximate and sometimes crude
+estimates, it has the advantage that it can be efficiently calculated for
+numerous taxa, which is sufficient to estimate the magnitude of Lewontin's
+Paradox (see \nameref{supp:popsize-valid} for more on validation based on
+biomass and other approaches).
+
+\subsection{Characterizing the Diversity--Census-size Relationship}
 
@@
     diversity (data from \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt},
-    and \cite{Romiguier2014-bp}), which varies around three orders of
+    and \cite{Romiguier2014-bp}), which varies over three orders of
     magnitude, shows a weak relationship with approximate population size,
@@
 size or range size, but these studies relied on heterozygosity measured from
-allozyme data \citep{Soule1976-he,Frankham1996-yb,Nei1984-zi,Leffler2012-zj}.
-Here, I confirm these findings using pairwise diversity estimates from genomic
-sequence data and the estimated census sizes (\FIG{pi-Nc}).  The pairwise
-diversity estimates are from three sources: \cite{Leffler2012-zj},
-\cite{Corbett-Detig2015-gt}, and \cite{Romiguier2014-bp}, and are predominantly
-from either synonymous or non-coding DNA (see Methods and Materials:
-\nameref{sec:methods-diversity}). Overall, an ordinary least squares (OLS)
-relationship on a log-log scale fits the data well (\FIG{pi-Nc}, gray dashed
-line). The OLS slope estimate is significant and implies a 13\% percent
-increase in differences per basepair for every order of magnitude census size
-grows (95\% confidence interval $[12\%, 14\%]$, adjusted $R^2 = 0.26$; see also
-the OLS fit per-phyla, \FIGSUPP[pi-Nc]{popsize-ave}).
+allozyme data \citep{Soule1976-he,Frankham1996-yb,Nei1984-zi}. I confirm these
+findings using pairwise diversity estimates from genomic sequence data and the
+estimated census sizes (\FIG{pi-Nc}). The pairwise diversity estimates are from
+three sources: \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt}, and
+\cite{Romiguier2014-bp}, and are predominantly from either synonymous or
+non-coding DNA (see Methods and Materials: \nameref{sec:methods-diversity}).
+Overall, an ordinary least squares (OLS) relationship on a log-log scale fits
+the data well (\FIG{pi-Nc}, gray dashed line). The OLS slope estimate is
+significant and implies a 13\% percent increase in differences per basepair for
+every order of magnitude census size grows (95\% confidence interval $[12\%,
+14\%]$, adjusted $R^2 = 0.26$; see also the OLS fit per-phyla,
+\FIGSUPP[pi-Nc]{popsize-ave}).
 
@@
 lynx (\emph{Lynx Lynx}), wolverine (\emph{Gulo gulo}), and Massasauga
-rattlesnake (\emph{Sistrurus catenatus}) face habitat fragmentation and
-declining population sizes. These three species are all in the IUCN Red List,
-but are listed as least concern (though their presence in the Red List
-indicates they are of conservation interest). In Appendix \ref{app:div-iucn},
-\nameref{app:div-iucn}, I explore more about the relationships between IUCN Red
-List status, diversity, and population size.
+rattlesnake (\emph{Sistrurus catenatus}) face habitat loss and declining
+population sizes. These three species are all in the IUCN Red List, but are
+listed as least concern (though their presence in the Red List indicates they
+are of conservation interest). In Appendix \ref{app:div-iucn},
+\nameref{app:div-iucn}, I explore the relationships between IUCN Red List
+status, diversity, and population size.
 
@@
 history can create correlation structure in the residuals, which violates an
-assumption of the regression model
-\citep{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I
-fit this relationship using a phylogenetic mixed-effects model, investigated
-whether there is a signal of phylogenetic non-independence, estimated the
-continuous trait values on the phylogeny, and explored how diversity and
-population size evolve. Prior population genetic comparative studies have
-lacked time-calibrated phylogenies and assumed unit branch lengths
+assumption of the regression model and can lead to bias
+\citep{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I fit
+the diversity--census-size relationship using a phylogenetic mixed-effects
+model, investigated whether there is a signal of phylogenetic non-independence,
+estimated the continuous trait values on the phylogeny, and explored how
+diversity and population size evolve. Prior population genetic comparative
+studies have lacked time-calibrated phylogenies and assumed unit branch lengths
 \citep{Whitney2010-ud}, a shortcoming that has drawn criticism
-\citep{Lynch2011-qv}. Here, I use a synthetic time-calibrated phylogeny
-created from the DateLife project \citep{OMeara2020-ds} to account for
-shared phylogenetic history (see Methods and Materials: \nameref{sec:methods-pcm}).
+\citep{Lynch2011-qv}.  I use a synthetic time-calibrated phylogeny created from
+the DateLife project \citep{OMeara2020-ds} to account for shared phylogenetic
+history (see Methods and Materials: \nameref{sec:methods-pcm}).
 
@@
 accounting for phylogeny, for the 166 taxa without missing data and present in
-the synthetic chronogram. As with the linear regression, this relationship was
-positive and significant (95\% credible interval $0.03, 0.11$), though somewhat
-attenuated compared to the OLS estimates (\FIG{pcm}B). Since
-the population size estimates are based on range and body mass, they are
-essentially a composite trait; fitting phylogenetic mixed-effects models
-separately on body mass and range indicates these have significant negative and
-positive effects, respectively (\FIGSUPP[pi-Nc]{div-range-bodymass}; see also
-\FIGSUPP[pi-Nc]{div-leffler} for the relationship between diversity and the
-range categories of \cite{Leffler2012-zj}).
-
-With the phylogenetic mixed-effects model, I also estimated the variance of the
-phylogenetic effect ($\sigma_p^2$) and the residual variance ($\sigma_r^2$),
-which can be used to estimate a measure of the phylogenetic signal, $\lambda =
-\nicefrac{\sigma_p^2}{(\sigma_p^2 + \sigma_r^2)}$
+the synthetic chronogram. This phylogenetic mixed-effects model can account for
+correlated residuals among more closely related species when estimating the
+relationship between $N_c$ and $\pi$, but does not assume there is phylogenetic
+signal in either of these variables. Since the phylogenetic mixed-effects model
+simultaneously estimates the degree of phylogenetic structure in the residuals
+while fitting the relationship between $N_c$ and $\pi$, this model's estimates
+would match those found by ordinary least squares if the residuals were
+distributed independently. Overall, this approach is conservative, making no
+assumptions about the source of the phylogenetic signal while accounting for
+violations of the regression model due to dependence among the residuals (see
+\cite{Revell2010-mf} for a discussion of this).
+
+As with the linear regression, I find this relationship is positive and
+significant (95\% credible interval $0.03, 0.11$), though somewhat attenuated
+compared to the OLS estimates (\FIG{pcm}B). Since the population size estimates
+are based on range and body mass, they are essentially a composite trait;
+fitting phylogenetic mixed-effects models separately on body mass and range
+indicates these have significant negative and positive effects, respectively
+(\FIGSUPP[pi-Nc]{div-range-bodymass}; see also \FIGSUPP[pi-Nc]{div-leffler} for
+the relationship between diversity and the range categories of
+\cite{Leffler2012-zj}).
+
+Since the phylogenetic mixed-effects model simultaneously estimates the
+variance of the phylogenetic effect ($\sigma_p^2$) and the residual variance
+($\sigma_r^2$), these can be used to estimate the phylogenetic signal, $\lambda
+= \nicefrac{\sigma_p^2}{(\sigma_p^2 + \sigma_r^2)}$
 (\cite{Lynch1991-tp,De_Villemereuil2014-kt}; see \cite{Freckleton2002-ly} for a
-comparison to Pagel's $\lambda$).  If the relationship between diversity and
-population size was free of shared phylogenetic history, $\lambda = 0$ and all
-the variance could be explained by evolution on the tips; this is analogous to
-Lynch's conjecture that coalescent times should be free of phylogenetic signal
+comparison to Pagel's $\lambda$). If the residuals were free of correlations
+due to shared phylogenetic history, then $\lambda = 0$ and all the variance
+could be explained by evolution on the tips; this is analogous to Lynch's
+conjecture that coalescent times should be free of phylogenetic signal
 (\citeyear{Lynch2011-qv}). In the relationship between population size and
 diversity, the posterior mean of $\lambda = 0.67$ (90\% credible interval
-$[0.58, 0.75]$) indicates that the majority of the variance perhaps might be
+$[0.58, 0.75]$) indicates a majority of the variance perhaps might be
 due to shared phylogenetic history (\FIG{pcm}B).
@@
 (\FIGSUPP[pcm]{div-popsize-post-phyla}). Each of the 90\% credible intervals
-for slope overlap, indicating the relationship between $\pi$ and $N_c$ is
+for slope overlap, suggesting the relationship between $\pi$ and $N_c$ is
 similar across these clades.
@@
 Interestingly, the diversity node-height test reveals two rate shifts at deeper
-splits (\FIG{pcm}C, top left) around 570 Mya. These nodes
-represent the branches between tunicates and vertebrates in chordates, and
-cephalopods and pleistomollusca (bivalves and gastropods) in molluscs.  While
-the cephalopod-pleistomollusca split outlier may be an artifact of having a
-single cephalopod (\emph{Sepia officinalis}) in the phylogeny, the
-tunicate-vertebrate split outlier is driven by the low diversity of vertebrates
-and the previously-documented exceptionally high diversity of tunicates (sea
-squirts; \cite{Nydam2010-kg,Small2007-mt}). This deep node representing a rate
-shift in diversity could reflect a change in either effective population size
-or mutation rate, and there is some evidence of both in this genus \emph{Ciona}
-\citep{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts
-in diversity is mirrored in the population size node-height test (Figure
-\FIG{pcm}C, top right). Rather, it appears a trait impacting diversity
-but not census size (e.g. mutation rate or offspring distributions) has
-experienced a shift on the lineage separating tunicates and vertebrates. At
-nearly 600 Mya, these deep nodes illustrate that expected coalescent times can
-share phylogenetic history, due to phylogenetic inertia in some combination of
-population size, reproductive system, and mutation rates.
+splits (\FIG{pcm}C, top left) around 570 Mya. These nodes represent the
+branches between tunicates and vertebrates in chordates, and cephalopods and
+pleistomollusca (bivalves and gastropods) in molluscs.  While the
+cephalopod-pleistomollusca split outlier may be an artifact of having a single
+cephalopod (\emph{Sepia officinalis}) in the phylogeny, the tunicate-vertebrate
+split outlier is driven by the low diversity of vertebrates and the
+previously-documented exceptionally high diversity of tunicates (sea squirts;
+\cite{Nydam2010-kg,Small2007-mt}). This deep node representing a rate shift in
+diversity could reflect a change in either effective population size or
+mutation rate, and there is some evidence of both in this genus \emph{Ciona}
+\citep{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts in
+diversity is mirrored in the population size node-height test (Figure
+\FIG{pcm}C, top right). Rather, it appears a trait impacting diversity but not
+census size (e.g. mutation rate or offspring distributions) has experienced a
+shift on the lineage separating tunicates and vertebrates. At nearly 600 Mya,
+these deep nodes illustrate that expected effective population sizes (and thus
+coalescent times) can share phylogenetic history, due to phylogenetic inertia
+in some combination of population size, reproductive system, and mutation
+rates.
 
 Finally, an important caveat is the increase in rate towards the tips could be
-caused by measurement noise, or possibly uncertainty or bias in the divergence
-time estimates deep in the tree. Inspecting the lineage pairs that lead to this
-increase in rate towards the tips indicates these represent plausible rate
-shifts, e.g.  between cosmopolitan and endemic sister species like
-\emph{Drosophila simulans} and \emph{Drosophila sechellia}; however, ruling out
-measurement noise entirely as an explanation would involve considering the
-uncertainty of diversity and population size estimates.
+caused by measurement noise, or possibly that uncertainty or bias in the
+divergence time estimates deep in the tree. Inspecting the lineage pairs that
+lead to this increase in rate towards the tips indicates these represent
+plausible rate shifts, e.g.  between cosmopolitan and endemic sister species
+like \emph{Drosophila simulans} and \emph{Drosophila sechellia}; however,
+ruling out measurement noise entirely as an explanation would involve
+considering the uncertainty of diversity and population size estimates.
 
@@
 Here, I test this hypothesis by estimating the scale of diversity reductions
-expected under background selection and recurrent hitchhiking, and compare
+expected under background selection and recurrent hitchhiking, and comparing
 these to the observed relationship between $\pi$ and $N_c$.
@@
 selection \citep{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
-difficulty is understanding of how $R$ varies across taxa, since we lack
+difficulty is understanding how $R$ varies across taxa, since we lack
 estimates of critical model parameters for most species. Still, I can address a
@@
   exceptionally large genome size (~30Gbp).}{\includegraphics[width=\textwidth]{genome_size_maplength.pdf}}\label{figsupp:genome-size-maplen}
+
+  \figsupp{The observed $\pi$--$N_c$ relationship (points) across species
+    compared to the predicted diversity (ribbons) under different modes of
+    linked selection and parameters, for a range of mutation rates $\mu =$
+    $10^{-8}--10^{-9}$. In both subplots, the gray ribbon is the expected
+    diversity if $N_e = N_c$. In (A), the predicted impact on diversity for
+    four modes of linked selection are depicted: background selection (purple)
+    and hitchhiking (yellow) individually under the parameters in the
+    \FIG{linksel}B, and strong background selection (red) where
+    $U_\text{strongBGS} = 10 U_\text{Dmel} \approx 16$, and strong recurrent
+    hitchhiking, where $\gamma_\text{strongHH} = 10 \gamma_\text{Dmel} \approx
+    0.23$. (B) The combined strong background selection and strong hitchhiking
+    (orange) compared to the original predicted diversity as in \FIG{linksel}B
+(blue).}{\includegraphics[width=\textwidth]{figure_3_alt.pdf}}\label{figsupp:linked-sel-alt}
 

### 2021-08-08 notebooks/main_analysis.Rmd
revised ms
updated figure 3 supp

@@
 
+What percent increase in pi is that for each order of magnitude change in N?
+
+```{r}
+signif(100 * 10^(coef(dps_ols)[2] - 1), 2)
+ols_cis <- confint(dps_ols)[2, ]
+signif(100 * 10^(ols_cis - 1), 2)
+```
+
 The chains mixed well. For quality assurance I compare this to a PGLS, to check
@@
 
+```
+
+Significance?
+
+```{r}
+sapply(list(popsize=psnh, range=rnnh, bodymass=bmnh, diversity=dvnh), function(x) x$pval)
 ```
@@
      mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
-                                             B=Ne_N_HK95, gamma=dm_gamma, 
-                                             J22=J22 * 10)) %>%
-
+                                             B=Ne_N_HK95, gamma=10*dm_gamma, 
+                                             J22=J22)) %>%
+     mutate(Ne_N_RHH_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
+                                             B=1, gamma=10*dm_gamma, 
+                                             J22=J22)) %>%
      mutate(Ne_N_RHH_BGS_strongsel_strongBGS = RHH_BGS(1, L=map_length, 
                                                        N=10^log10_popsize,
-                                             B=Ne_N_HK95_strongBGS, gamma=dm_gamma, 
-                                             J22=J22 * 10)) %>%
+                                             B=Ne_N_HK95_strongBGS, gamma=10*dm_gamma, 
+                                             J22=J22)) %>%
      # 10-folder greater Nc

### 2021-08-08 notebooks/figures/figure_3_alt.r

@@ -52,11 +52,24 @@ four_alleles <- function(N, mu) {
 
 dml <- dml_full %>% 
   mutate(Ne_RHH_BGS_strongsel = 10^log10_popsize * Ne_N_RHH_BGS_strongsel) %>%
+  mutate(Ne_RHH_strongsel = 10^log10_popsize * Ne_N_RHH_strongsel) %>%
+  mutate(Ne_strongBGS = 10^log10_popsize * Ne_N_HK95_strongBGS) %>%
   mutate(Ne_RHH_BGS_strongsel_strongBGS = 10^log10_popsize * Ne_N_RHH_BGS_strongsel_strongBGS) %>%
   mutate(Ne_RHH_BGS = 10^log10_popsize * Ne_N_RHH_BGS) %>%
   mutate(pi_RHH_BGS_1em9=four_alleles(Ne_RHH_BGS, 1e-9)) %>%
   mutate(pi_RHH_BGS_1em8=four_alleles(Ne_RHH_BGS, 1e-8)) %>%
-  # 10 * J22
+
+  mutate(pi_RHH_1em9=four_alleles(10^log10_popsize * Ne_N_RHH, 1e-9)) %>%
+  mutate(pi_RHH_1em8=four_alleles(10^log10_popsize * Ne_N_RHH, 1e-8)) %>%
+
+  mutate(pi_BGS_1em9=four_alleles(10^log10_popsize * Ne_N_HK95, 1e-9)) %>%
+  mutate(pi_BGS_1em8=four_alleles(10^log10_popsize * Ne_N_HK95, 1e-8)) %>%
+
+  mutate(pi_strongBGS_1em9=four_alleles(Ne_strongBGS, 1e-9)) %>%
+  mutate(pi_strongBGS_1em8=four_alleles(Ne_strongBGS, 1e-8)) %>%
+   mutate(pi_RHH_1em9_strongsel=four_alleles(Ne_RHH_strongsel, 1e-9)) %>%
+  mutate(pi_RHH_1em8_strongsel=four_alleles(Ne_RHH_strongsel, 1e-8)) %>%
+ 
   mutate(pi_RHH_BGS_1em9_strongsel=four_alleles(Ne_RHH_BGS_strongsel, 1e-9)) %>%
   mutate(pi_RHH_BGS_1em8_strongsel=four_alleles(Ne_RHH_BGS_strongsel, 1e-8)) %>%
   # 4 * U and 10 * J22
[…]

### 2021-08-09 manuscript/elife/elife-manuscript.tex
updated ms

@@
 Finally, an important caveat is the increase in rate towards the tips could be
-caused by measurement noise, or possibly that uncertainty or bias in the
-divergence time estimates deep in the tree. Inspecting the lineage pairs that
-lead to this increase in rate towards the tips indicates these represent
-plausible rate shifts, e.g.  between cosmopolitan and endemic sister species
-like \emph{Drosophila simulans} and \emph{Drosophila sechellia}; however,
-ruling out measurement noise entirely as an explanation would involve
-considering the uncertainty of diversity and population size estimates.
+caused by measurement noise, or possibly uncertainty or bias in the divergence
+time estimates deep in the tree. Inspecting the lineage pairs that lead to this
+increase in rate towards the tips indicates these represent plausible rate
+shifts, e.g.  between cosmopolitan and endemic sister species like
+\emph{Drosophila simulans} and \emph{Drosophila sechellia}; however, ruling out
+measurement noise entirely as an explanation would involve modeling the
+uncertainty of diversity and population size estimates.
 
@@
 linked selection models that incorporate both hitchhiking and background
-selection \citep{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
-difficulty is understanding how $R$ varies across taxa, since we lack
-estimates of critical model parameters for most species. Still, I can address a
-key question: if diversity levels were determined by census sizes ($\pi_0 =
-4N_c \mu$), are the combined effects of background selection and recurrent
+selection \citep{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second difficulty
+is understanding how $R$ varies across taxa, since we lack estimates of
+critical model parameters for most species. Still, I can address a key
+question: if diversity levels were determined by census sizes ($\pi_0 = 4N_c
+\mu$), would the combined effects of background selection and recurrent
 hitchhiking sufficient to reduce diversity to observed levels? Furthermore,
@@
 
-Since we lack estimates of key linked selection parameters across species, I
-parameterize the hitchhiking and BGS models using estimates from
-\emph{Drosophila melanogaster}, a species known to be strongly affected by
-linked selection \citep{Sella2009-nx}. Under a generalized model of
-hitchhiking and background selection \citep{Elyashiv2016-vt,Coop2012-cd}
-and assuming $N_e = N_c$, expected diversity is
+Since we lack estimates of selection parameters across species, I parameterize
+the hitchhiking and BGS models using estimates from \emph{Drosophila
+melanogaster}, a species known to be strongly affected by linked selection
+\citep{Sella2009-nx}. Under a generalized model of hitchhiking and background
+selection \citep{Elyashiv2016-vt,Coop2012-cd} and assuming $N_e = N_c$,
+the expected diversity is
 
 \begin{align} 
-  \pi_\text{BGS+HH} \approx \frac{\theta}{\nicefrac{1}{B(U, L)} + 2N_c S(\gamma, L, J)} 
+  \pi_\text{BGS+HH} \approx \frac{\theta}{\nicefrac{1}{B(U, L)} + 2N_c S(\gamma, J, L)} 
   \label{eqn:div} 
@@
 where $\theta = 4N_c\mu$, $B(U,L)$ is the effect of background selection, and
-$S(\gamma, L, J)$ is the rate of coalescence caused by sweeps (c.f.
+$S(\gamma, J, L)$ is the rate of coalescence caused by sweeps (c.f.
 \cite{Elyashiv2016-vt}, equation 1, \cite{Coop2012-cd} equation 20). Under
@@
 (\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}, see Appendix 
-\ref{app:bgs}, \nameref{app:bgs}). The coalescent rate due to sweeps is $S(\gamma, L, J) =
+\ref{app:bgs}, \nameref{app:bgs}). The coalescence rate due to sweeps is $S(\gamma,J,L) =
 \nicefrac{\gamma}{L} J$, where $\gamma$ is the number of adaptive substitutions
@@
 the impact of recurrent hitchhiking and background selection ($\gamma$, $J$,
-and $U$) to high values estimated from \emph{Drosophila melanogaster} by
-\cite{Elyashiv2016-vt}. My estimate of the adaptive substitutions per
-generation ($\gamma_\text{Dmel}$) based Elyashiv et al. implies a rate of
-sweeps per basepair of $\nu_\text{BP,Dmel} \approx 2.34 \times 10^{-11}$, which
-is close to other estimates from \emph{D.  melanogaster} (see
+and $U$) to strong selection values estimated for \emph{Drosophila
+melanogaster} by \cite{Elyashiv2016-vt}. My estimate of the adaptive
+substitutions per generation ($\gamma_\text{Dmel} \approx 2.3 \times 10^{-3}$)
+based Elyashiv et al.  implies a rate of sweeps per basepair of
+$\nu_\text{BP,Dmel} \approx 2.34 \times 10^{-11}$, which is close to other
+estimates from \emph{D.  melanogaster} (see
 \FIGSUPP[linksel]{linked-sel-params}A). The rate of deleterious mutations per
 diploid genome, per generation is parameterized using the estimate from
-Elyashiv et al., $U_\text{Dmel} = 1.6$, which is a bit greater than previous
+Elyashiv et al., $U_\text{Dmel} = 1.6$, which is slightly greater than previous
 estimates based on Bateman-Mukai approaches
 \citep{Mukai1985-bc,Mukai1988-vs,Charlesworth1987-ab}. Finally, the probability
-that a lineage is trapped in a sweep, $J_\text{Dmel}$, is calculated from the
-estimated genome-wide average coalescent rate due to sweeps from Elyashiv et
-al. (see \FIGSUPP[linksel]{linked-sel-params}B and Methods:
-\nameref{sec:methods-reduction} for more details on parameter estimates).
-Using these \emph{Drosophila} parameters, I then explore how the predicted
-range of diversity levels under background selection and recurrent hitchhiking
-varies across species with recombination map length ($L$) and census population
-size ($N_c$).
+that a lineage is trapped in a sweep, $J_\text{Dmel} \approx 4.5 \times
+10^{-4}$, is calculated from the estimated genome-wide average coalescence rate
+due to sweeps from Elyashiv et al. (see \FIGSUPP[linksel]{linked-sel-params}B
+and Methods: \nameref{sec:methods-reduction} for more details on parameter
+estimates).  Using these parameters, I then explore how the predicted range of
+diversity levels varies across species with recombination map length ($L$) and
+census population size ($N_c$).
 
@@
     linked selection and parameters, for a range of mutation rates $\mu =$
-    $10^{-8}--10^{-9}$. In both subplots, the gray ribbon is the expected
+    $10^{-8}$--$10^{-9}$. In both subplots, the gray ribbon is the expected
     diversity if $N_e = N_c$. In (A), the predicted impact on diversity for
     four modes of linked selection are depicted: background selection (purple)
-    and hitchhiking (yellow) individually under the parameters in the
-    \FIG{linksel}B, and strong background selection (red) where
-    $U_\text{strongBGS} = 10 U_\text{Dmel} \approx 16$, and strong recurrent
-    hitchhiking, where $\gamma_\text{strongHH} = 10 \gamma_\text{Dmel} \approx
-    0.23$. (B) The combined strong background selection and strong hitchhiking
+    and hitchhiking (yellow) individually under the \emph{Drosophila
+    melanogaster} parameters as in \FIG{linksel}B, and strong background
+    selection (red) where $U_\text{strongBGS} = 10 U_\text{Dmel} \approx 16$,
+    and strong recurrent hitchhiking, where $\gamma_\text{strongHH} = 10
+    \gamma_\text{Dmel} \approx 0.23$. (B) The predicted diversity under the
+    combined effects of strong background selection and strong hitchhiking
     (orange) compared to the original predicted diversity as in \FIG{linksel}B
-(blue).}{\includegraphics[width=\textwidth]{figure_3_alt.pdf}}\label{figsupp:linked-sel-alt}
+    (blue). Overall, under strong background selection and hitchhiking
+    parameters, predicted diversity would be less than observed for high-$N_c$
+    species, indicating the poor fit to observed data is not sensitive to the choice of
+  \emph{Drosophila melanogaster} parameters.}{\includegraphics[width=\textwidth]{figure_3_alt.pdf}}\label{figsupp:linked-sel-alt}
 
@@
       dashed line is the sweep model parameterized by the genome-wide average
-      sweep coalescent rate $2NS \approx 0.92$ from the classic sweep and
+      sweep coalescence rate $2NS \approx 0.92$ from the classic sweep and
       background selection model of \cite{Elyashiv2016-vt} ($r_s$ in
@@
 with larger population sizes might experience a greater impact of linked
-selection: recombinational map length, $L$, is known to correlate with body
-mass \citep{Burt1987-tq} and thus varies inversely with population size. As
-this is a critical parameter that determines the genome-wide impact of both
-hitchhiking and background selection, I examine the relationship between
-recombination map length ($L$) and census population size ($N_c$) across taxa,
-using available estimates of map lengths across species
+selection: recombination map length, $L$, is known to correlate with body mass
+\citep{Burt1987-tq} and thus varies inversely with population size. As this is
+a critical parameter that determines the genome-wide impact of both hitchhiking
+and background selection, I examine the relationship between recombination map
+length ($L$) and census population size ($N_c$) across taxa, using available
+estimates of map lengths across species
 \citep{Stapley2017-fs,Corbett-Detig2015-gt}. I find a significant non-linear
@@
 genome size and population size (\FIGSUPP[linksel]{genome-size-Nc}). These
-findings are consistent with both the hypothesis that non-adaptive processes
+findings are consistent with the hypothesis that non-adaptive processes
 increase genome size in small-$N_e$ species \citep{Lynch2003-bn} which in turn
@@
 \citep{Roze2021-ya}.  Overall, the negative relationship between map length and
-census size indicates linked selection is expected to be stronger in short map
-length, high-$N_c$ species.
+census size indicates linked selection is expected to be stronger in species
+with short map lengths, which are high-$N_c$ species.
 
 Then, I predict the expected diversity ($\pi_{BGS+HH}$) under background
-selection and hitchhiking, were $N_e= N_c$, and assuming all species had the
+selection and hitchhiking, assuming $N_e= N_c$ and that all species had the
 rate of sweeps and strength of BGS as \emph{D. melanogaster}. Since neutral
@@
 four-alleles model, \cite{Tajima1996-rb}), and compare this to the observed
-relationship between $\pi$ and $N_c$ in \FIG{linksel}B.  Under these
-parameters, linked selection begins to appreciably depress diversity around
-$N_c \approx 10^9$, since $S \approx 10^{-8}$--$10^{-9}$ and linked selection
-dominates drift when $S > \nicefrac{1}{2N}$.  Overall, this reveals two
-problems for the hypothesis that linked selection could solve Lewontin's
-Paradox. First, low to mid-$N_c$ species (census sizes between
-$10^{4}$--$10^{10}$) have sufficiently long map lengths that their diversity
-levels are only moderately reduced by linked selection, leading to a wide gap
-between predicted and observed diversity levels. For this not to be the case,
-the parameters that determine the strength of background selection and
-recurrent hitchhiking would need to be \emph{higher} among these species than
-in \emph{Drosophila melanogaster}. This would require that the rate of adaptive
-mutations or the deleterious mutation rate be orders of magnitude higher for
-species within this range than in \emph{Drosophila}, which is incompatible with
-the rate of adaptive substitutions across species \citep{Galtier2016-dq} and
-mutation rates \citep{Lynch2010-ki}. Furthermore, linked selection has been
-quantified in humans, which fall in this census size range, and has been found
-to be relatively weak
+relationship between $\pi$ and $N_c$ in \FIG{linksel}B.  Under these parameters
+and assumptions, linked selection begins to appreciably constrain diversity for
+$N_c \gtrsim 10^7$, since $S(\gamma_\text{Dmel}, J_\text{Dmel}, L) \approx
+10^{-7}$--$10^{-8}$ and linked selection dominates drift when $S(\gamma, J, L)
+> \nicefrac{1}{2N}$. Overall, this reveals two problems for the hypothesis that
+linked selection could solve Lewontin's Paradox. First, low to mid-$N_c$
+species (census sizes between $10^{4}$--$10^{7}$) have sufficiently long map
+lengths that their diversity levels are only moderately reduced by linked
+selection, leading to a wide gap between predicted and observed diversity
+levels. For this not to be the case, the rate of adaptive mutations or the
+deleterious mutation rate would need to be orders of magnitude higher for
+species within this range than in \emph{Drosophila melanogaster}, which is
+incompatible with the rate of adaptive protein substitutions across species
+\citep{Galtier2016-dq} and overall mutation rates \citep{Lynch2010-ki}.
+Furthermore, linked selection has been quantified in humans, which fall in this
+census size range, and has been found to be relatively weak
 \citep{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
 Second, while hitchhiking and BGS can reduce predicted diversity levels for
-high-$N_c$ species ($N_c > 10^{10}$) to observed levels, this would imply
+high-$N_c$ species ($N_c > 10^{12}$) to observed levels, this would imply
 available estimates of $\pi_0$ are underestimated by several orders of
@@
 species. Previous surveys have used allozyme-based estimates, fewer taxa, or
-qualitative measures of population size. My estimates of census population
-sizes are quite approximate, since they use body size to predict density. An
-improved estimate might consider vagility (as \cite{Soule1976-he} did), though
-this is harder to do systematically across many taxa. Future work might also
-use other ecological information, such as total biomass, or species
-distribution modeling to improve census size estimates
-\citep{Bar-On2018-kc,Mora2011-wm}.  Still, it seems more accurate
-estimates would be unlikely to change the qualitative findings here, which
-resemble those of early surveys \citep{Nei1984-zi,Soule1976-he}.
+proxies of population size. My estimates of census population sizes are rough
+approximates, since they use body size to predict density. An improved estimate
+might account for vagility (as \cite{Soule1976-he} did), though this is harder
+to do systematically across many taxa. Future work might also use other
+ecological information, such as total biomass, or species distribution modeling
+to improve census size estimates \citep{Bar-On2018-kc,Mora2011-wm}.  Still, it
+seems more accurate estimates would be unlikely to change the qualitative
+findings here, which resemble those of early surveys
+\citep{Nei1984-zi,Soule1976-he}.
 
@@
 = \nicefrac{\gamma_\text{Dmel}}{G} = 2.34 \times 10^{-11}$. Finally, I estimate
-$J_\text{Dmel}$ from the estimate of genome-wide average rate of sweeps from
-Elyashiv et al. (Supplementary Table S6) and assuming \emph{Drosophila} $N_e =
-10^6$. These \emph{Drosophila melanogaster} hitchhiking parameter estimates are
-close to other previously-published estimates
-(\FIGSUPP[linksel]{linked-sel-params}). Finally, I use $U_\text{Dmel} = 1.6$,
-from \cite{Elyashiv2016-vt}. With these parameter estimates from \emph{D.
-melanogaster}, the recombination map lengths across species, and Equation
-\eqref{eqn:div}, I estimate $\pi_\text{BGS+HH}$ (assuming $N_c=N_c$) across all
-species. This leads to a range of predicted diversity ranges across species
-corresponding to $\mu=10^{-8}$--$10^{-9}$; to visualize these, I take a convex
-hull of all diversity ranges and smooth this with R's \texttt{smooth.spline}
-function.
+$J_\text{Dmel} \approx 4.5 \times 10^{-4}$ from the estimate of genome-wide
+average rate of sweeps from Elyashiv et al. (Supplementary Table S6) and
+assuming \emph{Drosophila} $N_e = 10^6$. These \emph{Drosophila melanogaster}
+hitchhiking parameter estimates are close to other previously-published
+estimates (\FIGSUPP[linksel]{linked-sel-params}). Finally, I use $U_\text{Dmel}
+= 1.6$, from \cite{Elyashiv2016-vt}. With these parameter estimates from
+\emph{D.  melanogaster}, the recombination map lengths across species, and
+Equation \eqref{eqn:div}, I estimate $\pi_\text{BGS+HH}$ (assuming $N_c=N_c$)
+across all species. This leads to a range of predicted diversity ranges across
+species corresponding to $\mu=10^{-8}$--$10^{-9}$; to visualize these, I take a
+convex hull of all diversity ranges and smooth this with R's
+\texttt{smooth.spline} function.
 
@@
 simultaneously, using substitutions and stratifying by annotation.  For a
-neutral position at site $x$, the coalescent rate due to sweeps is given by
+neutral position at site $x$, the coalescence rate due to sweeps is given by
 Elyashiv et al.'s equation 3,
@@
 \nicefrac{\alpha m}{T}$ as the number of adaptive substitutions per generation
-per entire genome, such that $S(\gamma, L, J) = \nicefrac{\gamma}{L} \; J$ used
+per entire genome, such that $S(\gamma, J, L) = \nicefrac{\gamma}{L} \; J$ used
 in the main text. Using the estimates of $m \approx 4.5 \times 10^{5}$, $\alpha

### 2021-08-09 manuscript/elife/elife-manuscript.tex
more ms reviewer feedback

@@
 could be explained by evolution on the tips; this is analogous to Lynch's
-conjecture that coalescent times should be free of phylogenetic signal
+conjecture that coalescence times should be free of phylogenetic signal
 (\citeyear{Lynch2011-qv}). In the relationship between population size and
@@
 these deep nodes illustrate that expected effective population sizes (and thus
-coalescent times) can share phylogenetic history, due to phylogenetic inertia
+coalescence times) can share phylogenetic history, due to phylogenetic inertia
 in some combination of population size, reproductive system, and mutation
@@
 Morgan would only act to limit the reduction in diversity
-\citep{Weissman2012-aa}, and thus these results are conservative.  Overall,
-while linked selection could decouple diversity from census size for high-$N_c$
-species, recurrent hitchhiking and background selection seem unlikely to
-explain the observed patterns of diversity across species under our
-understanding of the range of parameter estimates.
+\citep{Weissman2012-aa}, and thus these results are conservative.
+
+Finally, the poor fit between observed and predicted levels of diversity across
+species is not remedied by stronger selection parameters. In
+\FIGSUPP[linksel]{linked-sel-alt}B, I increase both selection parameters $U$
+and $\gamma$ ten-fold each, and find the same qualitative pattern: on a log-log
+scale the relationship between $N_c$ and $\pi$ is linear, while the predicted
+diversity under linked selection is non-linear with $N_c$. Under this ten-fold
+higher selection regime, there is more overlap between observed and predicted
+levels of diversity, but diversity is severely \emph{under-predicted} for
+high-$N_c$ species. Additionally, this would imply that selection in
+low-to-mid-$N_c$ species is ten-folder higher than estimated in
+\emph{Drosophila melanogaster}, which is implausible.
+
+Overall, predicted levels of diversity under models of recurrent hitchhiking
+and background selection fit the observed relationship between $N_c$ and $\pi$
+poorly, and this is not overly sensitive to parameter choice. These findings
+extend the argument of \cite{Coop2016-gx}, showing that not only are previous
+empirical estimates of the impact of linked selection unable to explain
+Lewontin's Paradox, but it is unlikely current linked selection models under a
+wide range of selection parameters are capable of explaining the observed
+patterns of diversity across species.
 
@@
 
-One limitation of the dataset in this study is that diversity estimates are
-collated from a variety of sources rather than estimated with a single
-bioinformatic pipeline. This leads to technical noise across diversity
-estimates; perhaps the relationship between $\pi$ and $N_c$ found here could be
-tighter with a standardized bioinformatic pipeline. In addition to this
-technical variation, there might be systematic bioinformatic sources of bias in
-diversity estimates. For example high-diversity sequences may fail to align to
-the reference genome and end up unaccounted for, leading to a downward bias.
-Alternatively, high-diversity sequences might map to the reference genome, but
-adjacent mis-matching SNPs might be mistaken for a short insertion or deletion.
-While these issues might adversely affect the estimates in high-diversity
-species, it is unlikely they will qualitatively change the observed
-$\pi$--$N_c$ relationship.
+One limitation of this study is that diversity estimates are collated from a
+variety of sources rather than estimated with a single bioinformatic pipeline.
+This leads to technical noise across diversity estimates; perhaps the
+relationship between $\pi$ and $N_c$ found here could be tighter with a
+standardized bioinformatic pipeline. In addition, there might be systematic
+bioinformatic sources of bias: for example high-diversity sequences may fail to
+align to the reference genome and end up unaccounted for, leading to a downward
+bias.  Alternatively, a high-diversity sequences might map to the reference
+genome, but adjacent mis-matching SNPs might be mistaken for a short insertion
+or deletion.  While these issues might affect estimates in high-diversity
+species, it is unlikely to change the qualitative relationship between
+$\pi$--$N_c$.
 
@@
 phylogenetic inertia, sampling variance should exceed the variance due to
-phylogenetic shared history, and coalescent times are much less than divergence
-times. Since my findings suggest PCMs are necessary in some cases, it is
-worthwhile to address these points.
-
-First, Lynch has correctly pointed out that while coalescent times are much
+phylogenetic shared history, and coalescence times are much shorter than
+divergence times. Since my findings suggest PCMs are necessary in some cases,
+it is worthwhile to address these points.
+
+First, Lynch has correctly pointed out that while coalescence times are much
 less than divergence times and should be free of phylogenetic shared history,
-the factors that determine coalescent times (e.g.  mutation rates and effective
-population size) may not be (2011). In other words, coalescent times are free
-from phylogenetic shared history \emph{were we to condition} on these causal
-factors that could be affected by shared phylogenetic history. My estimates of
-phylogenetic signal in diversity, by contrast, are not conditioned on these
-factors. Importantly, even ``correcting for" phylogeny implicitly favors
-certain causal interpretations over others
+the factors that determine coalescence times (e.g.  mutation rates and
+effective population size) may not be (2011). In other words, coalescence times
+are free from phylogenetic shared history \emph{were we to condition} on these
+causal factors that could be affected by shared phylogenetic history. My
+estimates of phylogenetic signal in diversity, by contrast, are not conditioned
+on these factors. Importantly, even ``correcting for" phylogeny implicitly
+favors certain causal interpretations over others
 \citep{Westoby1995-wz,Uyeda2018-wf}.  Future work could try to untangle what
-causal factors determine coalescent times across species, as well as how these
+causal factors determine coalescence times across species, as well as how these
 factors evolve across macroevolutionary timescales. Second, it is a
 misconception that a fast rate of trait evolution necessarily reduces
-phylogenetic signal \citep{Revell2008-in}, and that if either or both
-variables in a regression are free of phylogenetic signal, PCMs are unnecessary
-\citep{Revell2010-mf,Uyeda2018-wf}. The evidence of high phylogenetic
-signal found in this study suggests PCMs are needed, in part to avoid spurious
-results from phylogenetic pseudoreplication.
+phylogenetic signal \citep{Revell2008-in}, and that if either or both variables
+in a regression are free of phylogenetic signal, PCMs are unnecessary
+\citep{Revell2010-mf,Uyeda2018-wf}. The evidence of high phylogenetic signal
+found in this study suggests PCMs are necessary when fitting the relationship
+between $N_c$ and $\pi$ in order to account for correlated residuals among
+closely-related species, and to avoid spurious results from phylogenetic
+pseudoreplication.
 
@@
 census population size.
-
-\subsection{Spatial and Demographic Processes}
-
-One limitation of this study is the inability to quantify the impact of spatial
-population genetic processes on the relationship between diversity and census
-population sizes across taxa. The genomic diversity estimates collated in this
-study unfortunately lack details about the sampling process and spatial data,
-which can have a profound impact on population genomic summary statistics
-\citep{Battey2020-lc}. These issues could systematically bias species-wide
-diversity estimates; for example, if diversity estimates from a cosmopolitan
-species were primarily from a single subpopulation, diversity would be an
-underestimate relative to the entire population. However, biased spatial
-sampling alone seems incapable of explaining the $\pi$-$N_c$ divergence in
-high-$N_c$ taxa. In the extreme scenario in which only one subpopulation was
-sampled, $F_\text{ST}$ would need to be close to one for population subdivision
-alone to sufficiently reduce the total population heterozygosity to explain the
-orders-of-magnitude shortfall between predicted and observed diversity levels.
-This is because the equation for $F_{ST}$ can be rearranged such that $H_S =
-(1-F_{ST})H_T$, where $H_S$ and $H_S$ are the subpopulation and total
-population heterozygosities; if $H_T = 4N_c\mu$, then only $F_{ST} \approx 1$
-can reduce $H_S$ several orders of magnitude. Yet, across-taxa surveys indicate
-that $F_\text{ST}$ is almost never this high within species
-\citep{Roux2016-lm}.  Still, future work could quantify the extent to which
-spatial processes contribute to Lewontin's Paradox. For example, high-$N_c$
-taxa usually experience range expansions, likely with repeated founder effects
-and local extinction/recolonization dynamics that doubtlessly depress
-diversity. In particular, with the appropriate data, one could estimate the
-empirical relationship between dispersal distance, range size, and coalescent
-effective population size across taxa.
-
-In this study, I have focused entirely on assessing the role of linked
-selection, rather than demography, in reducing diversity across taxa. In
-contrast to demographic models, models of linked selection have comparatively
-fewer parameters and more readily permit rough estimates of diversity
-reductions across taxa. Still, a full resolution of Lewontin's Paradox would
-require understanding how the demographic processes across taxa with incredibly
-heterogeneous ecologies and life histories transform $N_c$ into $N_e$. With
-population genomic data becoming available for more species, this could involve
-systematically inferring the demographic histories of tens of species and
-looking for correlations in the frequency and size of bottlenecks with $N_c$
-across species.
 
@@
 data.
+
+\subsection{Spatial and Demographic Processes}
+
+One limitation of this study is the inability to quantify the impact of spatial
+population genetic processes on the relationship between diversity and census
+population sizes across taxa. The genomic diversity estimates collated in this
+study unfortunately lack details about the sampling process and spatial data,
+which can have a profound impact on population genomic summary statistics
+\citep{Battey2020-lc}. These issues could systematically bias species-wide
+diversity estimates; for example, if diversity estimates from a cosmopolitan
+species were primarily from a single subpopulation, diversity would be an
+underestimate relative to the entire population. However, biased spatial
+sampling alone seems incapable of explaining the $\pi$-$N_c$ divergence in
+high-$N_c$ taxa. In the extreme scenario in which only one subpopulation was
+sampled, $F_\text{ST}$ would need to be close to one for population subdivision
+alone to sufficiently reduce the total population heterozygosity to explain the
+orders-of-magnitude shortfall between predicted and observed diversity levels.
+This is because the equation for $F_{ST}$ can be rearranged such that $H_S =
+(1-F_{ST})H_T$, where $H_S$ and $H_S$ are the subpopulation and total
+population heterozygosities; if $H_T = 4N_c\mu$, then only $F_{ST} \approx 1$
+can reduce $H_S$ several orders of magnitude. Yet, across-taxa surveys indicate
+that $F_\text{ST}$ is almost never this high within species
+\citep{Roux2016-lm}.  Still, future work could quantify the extent to which
+spatial processes contribute to Lewontin's Paradox. For example, high-$N_c$
+taxa usually experience range expansions, likely with repeated founder effects
+and local extinction/recolonization dynamics that doubtlessly depress
+diversity. In particular, with the appropriate data, one could estimate the
+empirical relationship between dispersal distance, range size, and coalescent
+effective population size across taxa.
+
+In this study, I have focused entirely on assessing the role of linked
+selection, rather than demography, in reducing diversity across taxa. In
+contrast to demographic models, models of linked selection have comparatively
+fewer parameters and more readily permit rough estimates of diversity
+reductions across taxa. Still, a full resolution of Lewontin's Paradox would
+require understanding how the demographic processes across taxa with incredibly
+heterogeneous ecologies and life histories transform $N_c$ into $N_e$. With
+population genomic data becoming available for more species, this could involve
+systematically inferring the demographic histories of tens of species and
+looking for correlations in the frequency and size of bottlenecks with $N_c$
+across species.
 

### 2021-08-10 manuscript/elife/elife-manuscript.tex
more reviewer feedback incorporated

@@
 low-to-mid-$N_c$ species is ten-folder higher than estimated in
-\emph{Drosophila melanogaster}, which is implausible.
-
-Overall, predicted levels of diversity under models of recurrent hitchhiking
-and background selection fit the observed relationship between $N_c$ and $\pi$
-poorly, and this is not overly sensitive to parameter choice. These findings
-extend the argument of \cite{Coop2016-gx}, showing that not only are previous
-empirical estimates of the impact of linked selection unable to explain
-Lewontin's Paradox, but it is unlikely current linked selection models under a
-wide range of selection parameters are capable of explaining the observed
-patterns of diversity across species.
+\emph{Drosophila melanogaster}, which is implausible. Overall, predicted
+levels of diversity under models of recurrent hitchhiking and background
+selection fit the observed relationship between $N_c$ and $\pi$ poorly, and
+this is not overly sensitive to parameter choice.
 
@@
 
-In this study, my goal was not to accurately estimate the levels in diversity
-across species, but rather to give linked selection the best possible chance to
-solve Lewontin's Paradox. Still, I find that even after parameterizing
-hitchhiking and background selection with strong selection parameter estimates
-from \emph{Drosophila melanogaster}, the predicted patterns of diversity under
-linked selection poorly fit observed patterns of diversity across species.
-This result extends the analysis by \cite{Coop2016-gx} showing that levels
-of $\pi_0$ estimated by \cite{Corbett-Detig2015-gt} are not decoupled from
-genome-wide average $\pi$, as would occur if linked selection were to explain
-Lewontin's Paradox. Here, my analysis goes a step further and suggests that
-models of recurrent hitchhiking and background selection are not capable of
-explaining the observed relationship between $\pi$ and census size, in part
-because mid-$N_c$ species have sufficiently long recombination map lengths to
-diminish the effects of even strong selection. This finding supports the idea
-the levels of diversity across species are primarily determined by past
-demographic fluctuations. Overall, while this suggests these two common modes
-of linked selection seem unlikely to explain across-taxa patterns of diversity,
-there are three major potential limitations of my approach that need further
-evaluation.
+Even assuming selection parameters estimated from \emph{Drosophila
+melanogaster}, where the effects of linked selection are thought to be
+especially strong, the predicted patterns of diversity under linked selection
+poorly fit observed patterns of diversity across species. My results support
+the analysis by \cite{Coop2016-gx} showing that levels of $\pi_0$ estimated by
+\cite{Corbett-Detig2015-gt} are not decoupled from genome-wide average $\pi$,
+as would occur if linked selection were to explain Lewontin's Paradox.
+Additionally, my analysis goes a step further, showing that current linked
+selection models under a wide range of selection parameters are incapable of
+explaining the observed relationship between census size and diversity. This is
+in part because mid-$N_c$ species have sufficiently long recombination map
+lengths to diminish the effects of even strong selection. Overall, while this
+suggests these two common modes of linked selection seem unlikely to explain
+patterns of diversity across taxa, there are three major potential limitations
+of my approach that need further evaluation.
 
@@
 capture the heterogeneity of functionality density and recombination rate in
-real genomes, it is still extraordinary conservative, likely overestimating the
-effects of linked selection to see if it could be capable of decoupling
-diversity from census size and explain Lewontin's Paradox. This is in part
-because the strong selection parameter estimates from \emph{Drosophila
-melanogaster} used, but also because I assume that the effective population
-size is equal to the census size.  Even then, this decoupling only occurs in
-very high--census-size species, and implies that the diversity in the absence
-of linked selection, $\pi_0$, is currently underestimated by several orders of
-magnitude. Moreover, the study of \cite{Corbett-Detig2015-gt} did consider
-recombination rate and functional density heterogeneity in estimating the
-reduction due to linked selection across species, yet their predicted
-reductions are orders of magnitude weaker than those considered here by
-assuming that $N_e = N_c$ (\FIGSUPP[linksel]{reductions}B).
-Overall, even with more realistic models of linked selection, current models of
-linked selection seem fundamentally unable to fit the diversity--census-size
-relationship.
+real genomes, it is still conservative, likely overestimating the effects of
+linked selection to see if it could be capable of decoupling diversity from
+census size and explain Lewontin's Paradox. This is in part because the strong
+selection parameter estimates from \emph{Drosophila melanogaster} used, but
+also because I assume that the effective population size is equal to the census
+size. Even then, this decoupling only occurs in very high--census-size species,
+and implies that the diversity in the absence of linked selection, $\pi_0$, is
+currently underestimated by several orders of magnitude. Moreover, the study of
+\cite{Corbett-Detig2015-gt} did consider recombination rate and functional
+density heterogeneity in estimating the reduction due to linked selection
+across species, yet their predicted reductions are orders of magnitude weaker
+than those considered here by assuming that $N_e = N_c$
+(\FIGSUPP[linksel]{reductions}B). Overall, given the effects estimated under
+more realistic inference models are still orders of magnitude weaker than those
+used in this study, current models of linked selection seem fundamentally
+unable to fit the diversity--census-size relationship.
 
@@
 models of partial sweeps and hard sweeps, though these imply different rates of
-sweeps \citep{Coop2012-cd}. Second, in the limit where most fitness
-variation is due to weak soft sweeps from standing variation scattered across
-the genome (i.e. due to polygenic fitness variation), levels of diversity are
-well approximated by quantitative genetic linked selection models
+sweeps \citep{Coop2012-cd}. Second, in the limit where most fitness variation
+is due to weak soft sweeps from standing variation scattered across the genome
+(i.e. due to polygenic fitness variation), levels of diversity are well
+approximated by quantitative genetic linked selection models
 \citep{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}. The reduction in
@@
 Section \ref{app:bgs}; \cite{Charlesworth2000-km,Charlesworth2015-am}). Third,
-the parameters from \cite{Elyashiv2016-vt} are robust to many types of
-sweeps that result in substitution (e.g. see p. 19 of their Supplementary
-Online Materials). Finally, I also disregarded the interaction of sweeps and
-spatial processes.  For populations spread over wide ranges, limited dispersal
-slows the spread of sweeps, allowing for new beneficial alleles to arise,
-spread, and compete against other segregating beneficial variants
+the parameters from \cite{Elyashiv2016-vt} could reflect a mixture of types of
+sweeps (\cite{Elyashiv2016-vt} p. 14 and p. 19 of their Supplementary Online
+Materials). Finally, I also disregarded the interaction of sweeps and spatial
+processes.  For populations spread over wide ranges, limited dispersal slows
+the spread of sweeps, allowing for new beneficial alleles to arise, spread, and
+compete against other segregating beneficial variants
 \citep{Ralph2015-kl,Ralph2010-ki}. Through limited dispersal should act to
@@
 Third, other selective processes, such as fluctuating selection or hard
-selective events, could reduce diversity in ways not captured by the background
-selection and hitchhiking model. Since frequency-independent fluctuating
-selection generally reduces diversity under most conditions
-\citep{Novak2017-yp}, this could lead seasonality and other sources of
-temporal heterogeneity to reduce diversity in large-$N_c$ species with short
-generation times more than longer-lived species with smaller population sizes.
-Future work could consider the impact of fluctuating selection on diversity
-under simple models \citep{Barton2000-zg} if estimates of key parameters
-governing the rate of such fluctuations were known across taxa. Additionally,
-another mode of selection that could severely reduce diversity across taxa, yet
-remains unaccounted for in this study, is periodic hard selective events. These
-selective events could occur regularly in a species' history yet be
-indistinguishable from demographic bottlenecks with just population genomic
-data.
+selective events (i.e. selection resulting in a reduction in the population
+size), could reduce diversity in ways not captured by the background selection
+and hitchhiking models. Since frequency-independent fluctuating selection
+reduces diversity under most conditions \citep{Novak2017-yp}, this could lead
+seasonality and other sources of temporal heterogeneity to reduce diversity in
+large-$N_c$ species with short generation times more than longer-lived species
+with smaller population sizes.  Future work could consider the impact of
+fluctuating selection on diversity under simple models \citep{Barton2000-zg} if
+estimates of key parameters governing the rate of such fluctuations were known
+across taxa. Additionally, another mode of selection that could severely reduce
+diversity across taxa, yet remains unaccounted for in this study, is periodic
+hard selective events. These selective events could occur regularly in a
+species' history yet be indistinguishable from demographic bottlenecks with
+just population genomic data.
 
@@
 One limitation of this study is the inability to quantify the impact of spatial
-population genetic processes on the relationship between diversity and census
-population sizes across taxa. The genomic diversity estimates collated in this
-study unfortunately lack details about the sampling process and spatial data,
-which can have a profound impact on population genomic summary statistics
-\citep{Battey2020-lc}. These issues could systematically bias species-wide
-diversity estimates; for example, if diversity estimates from a cosmopolitan
-species were primarily from a single subpopulation, diversity would be an
-underestimate relative to the entire population. However, biased spatial
-sampling alone seems incapable of explaining the $\pi$-$N_c$ divergence in
-high-$N_c$ taxa. In the extreme scenario in which only one subpopulation was
-sampled, $F_\text{ST}$ would need to be close to one for population subdivision
-alone to sufficiently reduce the total population heterozygosity to explain the
-orders-of-magnitude shortfall between predicted and observed diversity levels.
-This is because the equation for $F_{ST}$ can be rearranged such that $H_S =
-(1-F_{ST})H_T$, where $H_S$ and $H_S$ are the subpopulation and total
-population heterozygosities; if $H_T = 4N_c\mu$, then only $F_{ST} \approx 1$
-can reduce $H_S$ several orders of magnitude. Yet, across-taxa surveys indicate
-that $F_\text{ST}$ is almost never this high within species
-\citep{Roux2016-lm}.  Still, future work could quantify the extent to which
-spatial processes contribute to Lewontin's Paradox. For example, high-$N_c$
-taxa usually experience range expansions, likely with repeated founder effects
-and local extinction/recolonization dynamics that doubtlessly depress
-diversity. In particular, with the appropriate data, one could estimate the
-empirical relationship between dispersal distance, range size, and coalescent
-effective population size across taxa.
+and demographic population genetic processes on the relationship between
+diversity and census population sizes across taxa. The genomic diversity
+estimates collated in this study unfortunately lack details about the sampling
+process and spatial data, which can have a profound impact on population
+genomic summary statistics \citep{Battey2020-lc}. These issues could
+systematically bias species-wide diversity estimates; for example, if diversity
+estimates from a cosmopolitan species were primarily from a single region or
+subpopulation, diversity would be an underestimate relative to the entire
+population. However, biased spatial sampling alone seems incapable of
+explaining the $\pi$-$N_c$ divergence in high-$N_c$ taxa. In the extreme
+scenario in which only one subpopulation was sampled, $F_\text{ST}$ would need
+to be close to one for population subdivision alone to sufficiently reduce the
+total population heterozygosity to explain the orders-of-magnitude shortfall
+between predicted and observed diversity levels.  This can be seen by
+rearranging the expression for $F_{ST}$ as $H_S = (1-F_{ST})H_T$, where $H_S$
+and $H_T$ are the subpopulation and total population heterozygosities; if $H_T
+= 4N_c\mu$, then only $F_{ST} \approx 1$ can reduce $H_S$ several orders of
+magnitude. Yet, across-taxa surveys indicate that $F_\text{ST}$ is almost never
+this high within species \citep{Roux2016-lm}. Future work could quantify the
+extent to which more realistic spatial processes contribute to Lewontin's
+Paradox. For example, high-$N_c$ taxa usually experience range expansions, with
+repeated founder effects and local extinction/recolonization dynamics that
+depress diversity \citep{Slatkin1977-kd}. In particular, with the appropriate
+data, one could estimate the empirical relationship between dispersal distance,
+range size, and coalescent effective population size across taxa.
 
@@
 fewer parameters and more readily permit rough estimates of diversity
-reductions across taxa. Still, a full resolution of Lewontin's Paradox would
-require understanding how the demographic processes across taxa with incredibly
-heterogeneous ecologies and life histories transform $N_c$ into $N_e$. With
-population genomic data becoming available for more species, this could involve
-systematically inferring the demographic histories of tens of species and
-looking for correlations in the frequency and size of bottlenecks with $N_c$
-across species.
+reductions across taxa.  Given that I find that models of linked selection are
+incapable of explaining the observed relationship between $N_c$ and $\pi$, this
+supports the hypothesis the diversity across species are shaped primarily by
+past demographic fluctuations. Still, a full resolution of Lewontin's Paradox
+would require understanding how the demographic processes across taxa with
+incredibly heterogeneous ecologies and life histories transform $N_c$ into
+$N_e$. With population genomic data becoming available for more species, this
+could involve systematically inferring the demographic histories of tens of
+species and looking for correlations in the frequency and size of bottlenecks
+with $N_c$ across species.
 

### 2021-08-11 manuscript/elife/elife-manuscript.tex
more edits

@@
 % \title{Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to Explain the Narrow Range of Diversity Among Species}
+
 \title{Quantifying the Relationship Between Genetic Diversity and Population Size Suggests Natural Selection Cannot Explain Lewontin's Paradox}
@@
 accounting for phylogeny, for the 166 taxa without missing data and present in
-the synthetic chronogram. This phylogenetic mixed-effects model can account for
-correlated residuals among more closely related species when estimating the
-relationship between $N_c$ and $\pi$, but does not assume there is phylogenetic
-signal in either of these variables. Since the phylogenetic mixed-effects model
-simultaneously estimates the degree of phylogenetic structure in the residuals
-while fitting the relationship between $N_c$ and $\pi$, this model's estimates
-would match those found by ordinary least squares if the residuals were
-distributed independently. Overall, this approach is conservative, making no
-assumptions about the source of the phylogenetic signal while accounting for
-violations of the regression model due to dependence among the residuals (see
+the synthetic chronogram. This type of model is needed because closely-related
+species may differ from the average trend between $N_c$ and $\pi$ in similar
+ways due to shared phylogenetic history, similar life history traits, etc., and
+thus do not represent independent observations as is assumed by the standard
+regression model. This is a form of phylogenetic pseudoreplication, and can be
+accounted for with a phylogenetic mixed-effects model. The phylogenetic
+mixed-effects model does not assume that there is phylogenetic structure in
+either $N_c$ or $\pi$ (which itself is not a violation of the standard
+regression model, \cite{Revell2010-mf} and \cite{Uyeda2018-wf}), but rather
+accounts for phylogenetic correlations structure in the residuals if any is
+present. Importantly, phylogenetic mixed-effects models simultaneously estimate
+the degree of phylogenetic structure in the residuals while fitting the
+relationship between $N_c$ and $\pi$. If the residuals are distributed
+independently, the estimated relationship would be similar to that found by
+ordinary least squares, and the estimated phylogenetic signal would be zero.
+Overall, this approach is conservative, making no assumptions about the source
+of the phylogenetic signal while accounting for violations of the regression
+model due to dependence among the residuals if present (see
 \cite{Revell2010-mf} for a discussion of this).
@@
 (\cite{Lynch1991-tp,De_Villemereuil2014-kt}; see \cite{Freckleton2002-ly} for a
-comparison to Pagel's $\lambda$). If the residuals were free of correlations
-due to shared phylogenetic history, then $\lambda = 0$ and all the variance
-could be explained by evolution on the tips; this is analogous to Lynch's
-conjecture that coalescence times should be free of phylogenetic signal
-(\citeyear{Lynch2011-qv}). In the relationship between population size and
-diversity, the posterior mean of $\lambda = 0.67$ (90\% credible interval
-$[0.58, 0.75]$) indicates a majority of the variance perhaps might be
-due to shared phylogenetic history (\FIG{pcm}B).
+comparison to Pagel's $\lambda$). When residuals are free of correlations due
+to shared phylogenetic history, then $\lambda = 0$ and all the variance could
+be explained by evolution or noise on the tips. In the relationship between
+population size and diversity, the posterior mean of $\lambda = 0.67$ (90\%
+credible interval $[0.58, 0.75]$) indicates a majority of the variance perhaps
+might be due to shared phylogenetic history (\FIG{pcm}B).
 
@@
 seems to be driven in part by chordates having low diversity and small
-population sizes compared to non-chordates (\FIG{pcm}A). This
-problem resembles Felsenstein's worst-case scenario
-\citep{Felsenstein1985-an,Uyeda2018-wf}, where a singular event on a
-lineage separating two clades generates a spurious association between two
-traits.
+population sizes compared to non-chordates (\FIG{pcm}A). This problem resembles
+Felsenstein's worst-case scenario \citep{Felsenstein1985-an,Uyeda2018-wf},
+where a singular event on a lineage separating two clades generates a spurious
+association between two traits.
 
@@
 low-to-mid-$N_c$ species is ten-folder higher than estimated in
-\emph{Drosophila melanogaster}, which is implausible. Overall, predicted
-levels of diversity under models of recurrent hitchhiking and background
-selection fit the observed relationship between $N_c$ and $\pi$ poorly, and
-this is not overly sensitive to parameter choice.
+\emph{Drosophila melanogaster}, which is implausible. Overall, this suggests
+that present models of linked selection, even with very strong selection across
+species, are qualitatively incapable of matching the observed relationship
+between $N_c$ and $\pi$ and thus cannot explain Lewontin's Paradox.
 
@@
 causal factors that could be affected by shared phylogenetic history. My
-estimates of phylogenetic signal in diversity, by contrast, are not conditioned
-on these factors. Importantly, even ``correcting for" phylogeny implicitly
-favors certain causal interpretations over others
+estimates of phylogenetic signal in the residuals, by contrast, are not
+conditioned on these factors. Importantly, even ``correcting for" phylogeny
+implicitly favors certain causal interpretations over others
 \citep{Westoby1995-wz,Uyeda2018-wf}.  Future work could try to untangle what
@@
 implied by diversity, $\widetilde{N}_e$, diverge from census population sizes.
-However, there are a variety other effective population size estimates
+However, there are a variety other effective population size estimators
 calculable from different data and summary statistics
@@
 linkage disequilibrium, or temporal estimators that use the variance in allele
-frequency change. These alternate estimators capture summaries of the effective
-population size on shorter timescales than coalescent-based estimators
-\citep{Wang2005-wy}, and thus could be used to tease apart processes that
-impact the $N_e$-$N_c$ relationship in the more recent past.
+frequency change through time. These various estimators capture different
+summaries of effective population size on shorter timescales than
+coalescent-based estimators (see \cite{Wang2005-wy} for a review), and thus
+could be used to tease apart processes that impact the $N_e$-$N_c$ relationship
+in the more recent past.
 
@@
 $\nicefrac{N_e}{N_c}$, which is an important quantity in conservation genetics
-\citep{Frankham1995-xz,Mace1991-ai} and in understanding evolution in
-highly fecund marine species. Surveys of the short-term $\nicefrac{N_e}{N_c}$
+\citep{Frankham1995-xz,Mace1991-ai} and in understanding evolution in highly
+fecund marine species. Surveys of the short-term $\nicefrac{N_e}{N_c}$
 relationship across taxa indicate mean $\nicefrac{N_e}{N_c}$ is on order of
-$\approx 0.1$ \citep{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though
-the uncertainty in these estimates is high, and some species with sweepstakes
+$\approx 0.1$ \citep{Frankham1995-xz,Palstra2008-rt,Palstra2012-vy}, though the
+uncertainty in these estimates is high, and some species with sweepstakes
 reproduction systems like Pacific Oyster (\emph{Crassostrea gigas}) can have
-$\nicefrac{N_e}{N_c} \approx 10^{-6}$. Estimates of the $\nicefrac{N_e}{N_c}$
-ratio are an important, yet under appreciated piece of solving Lewontin's
-Paradox. For example, if $N_e$ is estimated from the allele frequency change
-across a single generation (i.e. \cite{Waples1989-sj}), $\nicefrac{N_e}{N_c}$
-constrains the variance in reproductive success
-\citep{Wright1938-tv,Nunney1993-ef,Nunney1996-wy}. This implies that apart
-from species with sweepstakes reproductive systems, the variance in
-reproductives success each generation (whether heritable or non-heritable) is
-likely insufficient to significantly contribute to constraining
-$\widetilde{N}_e$ for most taxa. Still, further work is needed to characterize
-(1) how $\nicefrac{N_e}{N_c}$ varies with $N_c$ across taxa (though see
+$\nicefrac{N_e}{N_c} \approx 10^{-6}$ \citep{Hedgecock1994-gs}. Estimates of
+the $\nicefrac{N_e}{N_c}$ ratio may be an important, yet under appreciated
+piece of solving Lewontin's Paradox. For example, if $N_e$ is estimated from
+the allele frequency change across a single generation (i.e.
+\cite{Waples1989-sj}), $\nicefrac{N_e}{N_c}$ constrains estimates of the
+variance in reproductive success
+\citep{Wright1938-tv,Nunney1993-ef,Nunney1996-wy}. This implies that apart from
+species with sweepstakes reproductive systems, the variance in reproductive
+success each generation (whether heritable or non-heritable) is likely
+insufficient to significantly contribute to constraining $\widetilde{N}_e$ for
+most taxa. Still, further work is needed to characterize (1) how
+$\nicefrac{N_e}{N_c}$ varies with $N_c$ across taxa (though see
 \cite{Palstra2012-vy}, Figure 2), and (2) the variance of $\nicefrac{N_e}{N_c}$
@@
 resolving Lewontin's Paradox. These estimators, along with short-term estimates
-of the impact of linked selection \citep{Buffalo2019-qs,Buffalo2020-my},
-can inform us how much diversity is depressed across shorter timescales, free
-from the rare strong selective events or severe bottlenecks that impact
+of the impact of linked selection \citep{Buffalo2019-qs,Buffalo2020-my}, can
+inform us how much diversity is depressed by selection on shorter timescales,
+free from the rare strong selective events or severe bottlenecks that impact
 pairwise diversity. It could be that in any one generation, selection
@@
 the generation to generation evolutionary processes acting in a population to
-change allele frequencies. Furthermore, certain observed adaptations are
-inexplicable given implied long-term coalescent effective population sizes, and
-are only possible if short-term effective population sizes are orders of
-magnitude larger \citep{Karasov2010-fb,Barton2010-pi}.
+change allele frequencies. Furthermore, certain observed adaptations occur at a
+pace that is inexplicable given small effective population sizes implied by
+diversity, and are only possible if short-term effective population sizes are
+orders of magnitude larger \citep{Karasov2010-fb,Barton2010-pi}.
 
@@
 population genetics and population ecology into a cohesive discipline of
-population biology. Lewontin's Paradox of Variation remains a critical unsolved
-problem at the nexus of these two different disciplines: across species, we
-fail to understand the processes that connect a central parameter of population
-ecology, census size, to a central parameter of population genetics, effective
-population size. Given that selection seems to fall short in explaining
+population biology. Lewontin's Paradox of Variation remains a major unsolved
+problem at the nexus of these two different disciplines: we fail to understand
+the processes that connect a central parameter of population ecology, census
+size, to a central parameter of population genetics, effective population size
+across species. Given that selection seems to fall short in resolving
 Lewontin's Paradox, a full resolution will require a mechanistic understanding
@@
 population genetics.  Here, I have found that diversity and $N_c$ have a
-surprisingly consistent relationship without many outliers, despite the wildly
-disparate ecologies, life histories, and evolutionary histories of the taxa
-included. Furthermore, taxa with very large census sizes have surprisingly low
-diversity. Is this explained by macroevolutionary processes, such as different
-rates of speciation for large-$N_c$ taxa? Or, are the levels of diversity we
-observe today an artifact of our timing relative to the last glacial maximum,
-or the last major extinction? Did large-$N_c$ prehistoric animal populations
-living in other geological eras have higher levels of diversity than our
-present taxa? Or, does ecological competition occur on shorter timescales such
-that strong population size contractions transpire and depress diversity, even
-if a species is undisturbed by climatic shifts or mass extinctions? Overall,
-patterns of diversity across taxa are determined by many overlaid evolutionary
-and ecological processes occurring on vastly different timescales. Lewontin's
+consistent relationship without many outliers, despite the wildly disparate
+ecologies, life histories, and evolutionary histories of the taxa included.
+Furthermore, taxa with very large census sizes have surprisingly low diversity.
+Is this explained by macroevolutionary processes, such as different rates of
+speciation for large-$N_c$ taxa? Or, are the levels of diversity we observe
+today an artifact of our timing relative to the last glacial maximum, or the
+last major extinction? Did large-$N_c$ prehistoric animal populations living in
+other geological eras have higher levels of diversity than our present taxa?
+Or, does ecological competition occur on shorter timescales such that strong
+population size contractions transpire and depress diversity, even if a species
+is undisturbed by climatic shifts or mass extinctions? Overall, patterns of
+diversity across taxa are determined by many overlaid evolutionary and
+ecological processes occurring on vastly different timescales. Lewontin's
 Paradox of Variation may persist unresolved for some time because the
@@
 %
-(cf. equation 1 \cite{Elyashiv2016-vt}, and equation 20 of \cite{Coop2012-cd}).
+(cf. equation 1 \cite{Elyashiv2016-vt}, equation 4 of \cite{Kim2000-ep}, and equation 20 of \cite{Coop2012-cd}, ).
 The BGS component is given by \cite{Hudson1995-xc},
@@
 %
-(cf. \cite{Coop2012-cd} equation 20) where $J$ is the probability that two
-lineages coalesce down to one, given sweeps occur uniformly along the genome.
-Under this homogeneous sweep model, $J$ is
+(cf. \cite{Coop2012-cd} equation 20) where $\nu_\text{BP}$ and $r_\text{BP}$
+are the substitutions and recombination per basepair respectively, $J$ is the
+probability that two lineages coalesce down to one, given sweeps occur
+uniformly along the genome.  Under this homogeneous sweep model, $J$ is
 
@@
 %
-where $T$ is the number of generations that substitutions accrue, $i_S = 1,
-\ldots, I_S$ is the annotation class (e.g. exons, introns, UTRs), $\alpha(i_S)$
-is the fraction of substitutions in annotation class $i_S$ that are beneficial,
-$a(i_S)$ is the set of all substitutions in annotation class $i_S$, $\tau(s,
-N)$ is the fixation time of a site with additive effect $s$, and $g(s | i_S)$
-is the distribution of selection coefficients for annotation class $i_S$.
+where $T$ is the length of the lineage (in generations) on which substitutions
+accrue, $i_S = 1, \ldots, I_S$ is the annotation class (e.g. exons, introns,
+UTRs), $\alpha(i_S)$ is the fraction of substitutions in annotation class $i_S$
+that are beneficial, $a(i_S)$ is the set of all substitutions in annotation
+class $i_S$, $\tau(s, N)$ is the fixation time of a site with additive effect
+$s$, and $g(s | i_S)$ is the distribution of selection coefficients for
+annotation class $i_S$.
 
